%% NMC Thawing Quintessence: w_0, w_a derivation
%% For insertion at eci.tex line 153 (deferred derivations/w0-wa-nmc/)
%% ECI v6.0.45, 2026-05-04
%% References all verified via export.arxiv.org/api/query

%% ANTI-HALLUCINATION NOTE:
%%   "Pan-Yang JCAP 2018 arXiv:1804.05064" does NOT exist.
%%   arXiv:1804.05064 = "The FABLE simulations" (galaxy-cluster hydro, Henden+2018).
%%   The coefficient -2 xi Omega_phi0 was HALLUCINATED. This derivation corrects it.

\subsection*{NMC Thawing Quintessence: Derivation of $w_0$, $w_a$}

\subsubsection*{Lagrangian and Klein-Gordon equation}

We consider thawing quintessence with NMC Lagrangian
\begin{equation}
    \mathcal{L}_\chi = -\tfrac{1}{2}(\partial\chi)^2 - V(\chi) - \tfrac{1}{2}\xi_\chi R\chi^2,
    \qquad V(\chi) = V_0 e^{-\lambda\chi/M_P}.
\end{equation}
Varying with respect to $\chi$ in a flat FLRW background gives the Klein-Gordon equation
\begin{equation}
    \ddot\chi + 3H\dot\chi + V'(\chi) = \xi_\chi R\chi,
    \qquad R = 6(2H^2 + \dot H).
    \label{eq:KG-NMC}
\end{equation}

\subsubsection*{Effective stress-energy tensor}

In the Jordan frame (perturbative $\xi_\chi$, following
\citealt{Pettorino:2008ez}~[arXiv:0802.1086]):
\begin{align}
    \rho_\chi &= \tfrac{1}{2}\dot\chi^2 + V(\chi) + 6\xi_\chi H\chi\dot\chi, \label{eq:rho-NMC}\\
    p_\chi    &= \tfrac{1}{2}\dot\chi^2 - V(\chi) - \xi_\chi(2\dot H + 4H^2)\chi^2
                 - 4\xi_\chi H\chi\dot\chi. \label{eq:p-NMC}
\end{align}
Terms of order $\xi_\chi(\chi/M_P)^2$ are dropped as subdominant for
perturbative coupling.

\subsubsection*{Slow-roll attractor}

The slow-roll conditions $\ddot\chi\ll 3H\dot\chi$ and $\dot\chi^2\ll V$ give
\begin{equation}
    \dot\chi_{\rm SR} = \frac{\lambda V(\chi)}{3HM_P} + 4\xi_\chi H\chi + \mathcal{O}(\xi_\chi^2),
    \label{eq:chidot-SR}
\end{equation}
where the second term is the $\mathcal{O}(\xi_\chi)$ NMC correction using $R\approx 12H^2$
(near de~Sitter).  From the Friedmann equation $3M_P^2 H^2 = \rho_\chi/\Omega_\phi$,
\begin{equation}
    \frac{\dot\chi_{\rm SR}^2}{V} = \frac{\lambda^2\Omega_\phi}{3} + \mathcal{O}(\xi_\chi),
\end{equation}
with $\epsilon_V \equiv \tfrac{M_P^2}{2}\!\left(\tfrac{V'}{V}\right)^{\!2} = \tfrac{\lambda^2}{2}$.

\subsubsection*{Equation of state $w(a)$}

Defining $u\equiv\dot\chi^2/V = \lambda^2\Omega_\phi/3$,
$A\equiv 6\xi_\chi H\chi\dot\chi/V = 2\xi_\chi\lambda\chi/M_P$,
and $B \equiv [-\xi_\chi(2\dot H+4H^2)\chi^2 - 4\xi_\chi H\chi\dot\chi]/V \approx
-4\xi_\chi\lambda\chi/(3M_P)$ (leading slow-roll terms only),
\begin{align}
    w &= \frac{p_\chi}{\rho_\chi}
       = \frac{u/2 - 1 + B}{1 + u/2 + A}
       \approx -1 + u + A + B + \mathcal{O}(u^2, uA, uB) \notag\\
      &= -1 + \frac{\lambda^2\Omega_\phi}{3}
           + \frac{2\xi_\chi\lambda\chi}{3M_P},
    \label{eq:w-NMC}
\end{align}
where the combination $A + B = 2\xi_\chi\lambda\chi/(3M_P)$ was confirmed
symbolically by \texttt{sympy}.

\subsubsection*{CPL coefficients}

In the thawing scenario the field tracks $\chi(a)$ along the slow-roll
attractor; linearising near $a=1$,
\begin{equation}
    \chi(a) \approx \chi_0 + \lambda M_P\,\Omega_{\phi,0}\,(a-1),
    \label{eq:chi-tracker}
\end{equation}
so
\begin{equation}
    \frac{dw}{da}\bigg|_{a=1} = \frac{\lambda^2\,d\Omega_\phi/da|_{a=1}}{3}
        + \frac{2\xi_\chi\lambda^2\Omega_{\phi,0}}{3}.
\end{equation}
The Omega evolution obeys $d\Omega_\phi/d\!\ln a = 3\Omega_\phi(1-\Omega_\phi)(1+w_\phi)$;
at leading order in $\epsilon_V$,
$(1+w_\phi)|_{a=1}=\lambda^2\Omega_{\phi,0}/3$, giving
\begin{equation}
    \frac{d\Omega_\phi}{da}\bigg|_{a=1}
        = \lambda^2\Omega_{\phi,0}^2(1-\Omega_{\phi,0}).
\end{equation}
The CPL parameters are therefore
\begin{align}
    \boxed{w_0^{\rm NMC}} &= -1 + \frac{\lambda^2}{3}\,\Omega_{\phi,0}
        + \frac{2\xi_\chi\lambda\chi_0}{3M_P}, \label{eq:w0-NMC}\\[4pt]
    \boxed{w_a^{\rm NMC}} &= -\lambda^2\Omega_{\phi,0}\!\left[
        \Omega_{\phi,0}(1-\Omega_{\phi,0}) + \frac{2\xi_\chi}{3}
        \right], \label{eq:wa-NMC}
\end{align}
and the NMC correction to $w_a$ relative to the minimal ($\xi_\chi=0$) baseline is
\begin{equation}
    \boxed{\Delta w_a^{\rm NMC} = -\frac{2\xi_\chi\lambda^2}{3}\,\Omega_{\phi,0}.}
    \label{eq:Dwa-NMC}
\end{equation}
The coefficient is $-2\lambda^2/3$, \textbf{not} $-2$ as appeared in a
prior (hallucinated) derivation citing a non-existent ``Pan--Yang JCAP 2018'';
the missing factor $\lambda^2$ is essential and comes directly from
$\dot\chi/H = \lambda M_P\Omega_\phi$ on the slow-roll attractor.

\subsubsection*{Sanity checks}
\begin{itemize}
    \item $\xi_\chi=0$: recovers minimal thawing result $w_a=-\lambda^2\Omega_{\phi,0}^2(1-\Omega_{\phi,0})$.
    \item $\lambda\to 0$: $w_a\to 0$ (cosmological constant, exact).
    \item $\Omega_{\phi,0}\to 0$: $w_a\to 0$ (negligible DE).
    \item Sign: $\xi_\chi>0$ gives $\Delta w_a<0$, i.e.\ NMC deepens the thawing,
          consistent with \citet{Wolf:2025} (arXiv:2504.07679) preference for
          NMC in DESI 2025 data.
\end{itemize}

\subsubsection*{Verified references}
\begin{itemize}
    \item Scherrer \& Sen 2008 PRD 77 083515 = arXiv:0712.3450 [VERIFIED via arXiv API]
    \item Wolf, Garc\'ia-Garc\'ia, Anton, Ferreira 2025 PRL 135 081001 = arXiv:2504.07679 [VERIFIED]
    \item Adam, Hertzberg+ JCAP 04 (2026) 052 = arXiv:2509.13302 [VERIFIED]
    \item S\'anchez L\'opez, Karam, Hazra 2025 = arXiv:2510.14941 [VERIFIED]
    \item Pettorino \& Baccigalupi 2008 PRD 77 103003 = arXiv:0802.1086 [VERIFIED]
    \item \textbf{Pan--Yang JCAP 2018 = arXiv:1804.05064: DOES NOT EXIST as NMC quintessence paper.}
          arXiv:1804.05064 is ``The FABLE simulations'' (Henden+2018, galaxy-cluster hydro).
          All uses of this citation are hallucinations; hallucination counter incremented.
\end{itemize}
