% Team A draft — §1.5 Unifying Hypothesis (A): complexity growth as a
% universe-wide tendency. NOT to be \input{}-ed into eci.tex; the owner
% picks A vs B after both teams report. Two-column, revtex4-2-compatible.

\section{Unifying hypothesis (A): complexity growth as a universe-wide tendency}
\label{sec:team_A_unifying}

\subsection*{Statement of the hypothesis}

We entertain, as a speculative organising principle above and beyond the
six-axiom architecture of \S1, the following \emph{Universal Complexity
Growth} (UCG) hypothesis: for every admissible QRF $\mathcal{R}$ and every
spatial region $\Sigma$ accessible to it, there exists a non-negative
complexity functional $\mathfrak{C}_{\mathcal{R}}[\Sigma; t]$ whose
coarse-grained time evolution is monotone non-decreasing until a
QRF-dependent saturation scale, after which it plateaus modulo
recurrences. Formally,
\begin{equation}
\boxed{\;
\partial_t \, \mathfrak{C}_{\mathcal{R}}[\Sigma;t]
\;\geq\; 0
\quad\text{for}\quad t \lesssim t_{\mathrm{sat}}(\mathcal{R},\Sigma),
\;\;
\mathfrak{C}_{\mathcal{R}} \to \mathfrak{C}_{\max}
\text{ as } t\to t_{\mathrm{sat}}.
\;}
\label{eq:UCG}
\end{equation}
The strong reading of \eqref{eq:UCG} is that $\mathfrak{C}_{\mathcal{R}}$
specialises, at the microscopic quantum-gravitational scale, to a
quantum-circuit complexity on the CFT Hilbert space, and at the
macroscopic cosmological-structure scale, to a persistent-homology
complexity on the density super-level-set filtration. Under this reading,
A3 and A6 cease to be independent axioms and become two projections of a
single universe-wide monotonicity.

\subsection*{A3 as the quantum-circuit specialisation}

Haferkamp--Faist--Kothakonda--Eisert--Yunger~Halpern~\cite{Haferkamp2022}
establish that for a generic Haar-random circuit on $n$ qubits the exact
state complexity grows linearly in circuit depth up to a saturation
value exponential in $n$. Ma--Huang~\cite{MaHuang2025} show that this
growth curve is reproduced by \emph{efficient} pseudorandom unitaries
(PRUs), not only Haar averages --- the linear-then-plateau profile is
realised by physical, polynomially-implementable dynamics on any
sufficiently chaotic system. Cryptographic
Censorship~\cite{CryptoCensorship} then reads this PRU-dynamics regime
off from the bulk geometry: boundary dynamics that are
$\varepsilon$-approximately a $k$-design admit bulk duals containing an
event horizon.

Under \eqref{eq:UCG}, A3 is \emph{not} an independent posit: it is the
statement that when the QRF is a holographic CFT and the
$\mathfrak{C}_{\mathcal{R}}$ functional is specialised to gate complexity
on the code subspace, the monotone-growth-until-saturation profile
$\eqref{eq:UCG}$ is (i) guaranteed to occur (Haferkamp), (ii) realised by
polynomially-efficient dynamics (Ma--Huang), and (iii) dual to horizon
formation (Cryptographic Censorship). The three inputs compose into the
AdS/CFT reading of \eqref{eq:UCG}, and nothing more is needed to recover
A3 in that setting.

\subsection*{A6 as the topological-filtration specialisation}

At cosmological scales, Yip--Biagetti--Starck~\cite{Yip2024} (with
Calles~\cite{Calles2025}) show that persistent homology on density
super-level sets provides a strictly-finer $f_{\mathrm{NL}}$ estimator
than Betti numbers alone, with the leading-order shift of the Euler
characteristic $\chi_{\mathrm{E}}(\nu)$ reproducing the Matsubara-2003
form, Eq.~\eqref{eq:A6-euler}.

Under the strong reading of \eqref{eq:UCG}, we interpret
$\mathfrak{C}_{\mathcal{R}}$ at the late-time cosmological QRF as a
\emph{topological complexity} of the density field, measured by the
total persistence across the super-level-set filtration
$\{\Sigma_\nu\}_{\nu}$:
\begin{equation}
\mathfrak{C}_{\mathrm{LSS}}[\Sigma;t]
:= \sum_{k\ge 0} \int \!\! d\mathrm{PH}_k(b,d;t) \,(d-b),
\label{eq:C_LSS}
\end{equation}
the sum being the total $k$-dimensional persistence at time $t$. Linear
growth of $\mathfrak{C}_{\mathrm{LSS}}$ from the nearly-Gaussian
recombination initial condition towards the non-linear saturated web at
$z\!\sim\!0$ is empirically the \emph{phenomenon} measured by Yip~2024
--- total persistence is a monotone non-decreasing function of the
structure-formation clock. Equation~\eqref{eq:A6-euler} is then the
Gram--Charlier leading-order shift of the Matsubara Euler characteristic
induced by the first $f_{\mathrm{NL}}$-driven departure from Gaussianity,
which under \eqref{eq:UCG} is the leading non-trivial signature of UCG at
the LSS scale. A6 would thus follow as the Gram--Charlier projection of
\eqref{eq:UCG}.

\subsection*{Honest caveats}

Three caveats are required before \eqref{eq:UCG} can be read as anything
stronger than a suggestive analogy.

\emph{(C1) Category mismatch.} The Haferkamp complexity is a
gate-theoretic quantity on a Hilbert space; $\mathfrak{C}_{\mathrm{LSS}}$
is a topological-data-analysis functional on a classical random field. No
derivation in this paper produces one from the other; the claim that both
are specialisations of a single $\mathfrak{C}_{\mathcal{R}}$ is at
present a \emph{framing} assertion, not a theorem. Team~A's
\texttt{derivation\_sketch.md} documents a good-faith attempt and its
explicit stopping point.

\emph{(C2) The Gram--Charlier step is not unique.} Several complexity
functionals on the 1-point PDF of the density field reproduce an
$H_3(\nu)\phi(\nu)$ signature at leading order in $f_{\mathrm{NL}}$
because the Gram--Charlier basis is itself a Hermite expansion; without a
separate argument fixing $\mathfrak{C}_{\mathrm{LSS}}$, the
``derivation'' of Eq.~\eqref{eq:A6-euler} from \eqref{eq:UCG} degenerates
into a tautology.

\emph{(C3) Saturation scales are heterogeneous.} The quantum-circuit
saturation time $t_{\mathrm{sat}}^{\mathrm{CFT}} \sim e^{\,n}$ and the
LSS non-linear-clustering saturation $t_{\mathrm{sat}}^{\mathrm{LSS}}
\sim $ Hubble time are separated by scales that admit no known common
RG flow in this paper. \eqref{eq:UCG} therefore posits a universality of
\emph{form} (linear-then-plateau), not of scale.

We flag \eqref{eq:UCG} as an organising hypothesis at the same speculative
level as the cosmological use of A3 in the Appendix: useful as narrative
scaffolding, but not as a load-bearing element of \S3's
phenomenological predictions. The predictive core of the paper
(\S3.1--\S3.7, \S4) is unaffected by whether \eqref{eq:UCG} is
ultimately adopted or discarded.
