%% bianchi_ix_modular_shadow.tex
%% M45 draft v0.1 (2026-05-06)
%% Conjecture-paper for Comm. Math. Phys.
%% "Type-II_infty modular-shadow algebra for Bianchi IX vacuum sector
%%  via Manin-Marcolli symbolic dynamics"
%%
%% Author: Kevin Remondiere (independent), Tarbes, France
%% Status: DRAFT — 6 [TBD: prove] markers explicit; conjecture paper, not theorem.
%% All 14 arXiv IDs WebFetch-verified 2026-05-06.
%% Hallu count entering: 86. Held.

\documentclass[12pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=2.5cm}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}{Conjecture}
\newtheorem{remark}{Remark}
\newtheorem{definition}[theorem]{Definition}

\title{\textbf{A type-II$_\infty$ modular-shadow algebra for the\\
       Bianchi IX vacuum sector via\\
       Manin--Marcolli symbolic dynamics}\\[6pt]
       \large (conjecture paper, with falsifier protocol)}

\author{%
  K\'evin Remondi\`ere\\[2pt]
  \small Independent researcher, Tarbes, France\\[2pt]
  \small \textit{Draft for Communications in Mathematical Physics}
}

\date{Preprint: \today\quad(M45 v0.1)}

\begin{document}

\maketitle

\begin{abstract}
We propose a type-II$_\infty$ modular-shadow algebra
$\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}$ for the vacuum sector
of Bianchi~IX cosmology in the Belinski--Khalatnikov--Lifshitz (BKL)
mixmaster phase, obtained as the Speranza (2025)
crossed product of a wedge-complement type-III$_1$ field algebra
$\mathcal{N}$ by the Connes--Takesaki modular automorphism flow.
Combining Manin--Marcolli's (2015) symbolic-dynamics
representation of the BKL Kasner bounce sequence as the geodesic flow
of the modular surface $X(2) = \Gamma(2)\backslash\mathbb{H}$ with the
generic crossed-product framework of Witten (2022)
and Speranza (2025), we
\emph{conjecture} that the modular flow rate on
$\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}$ equals
$\lambda_{\mathrm{mod}} = 2\pi\,\lambda_{\mathrm{BKL}}$, where
$\lambda_{\mathrm{BKL}} = h_{\mathrm{KS}}(\sigma_{\mathrm{Gauss}})$ is
the Kolmogorov--Sinai entropy of the Gauss-shift symbolic encoding of
BKL Kasner bounces.
This in turn predicts a Maldacena--Shenker--Stanford-style modular
Krylov-complexity bound
$\mathcal{C}_K^{\mathrm{mod}}(s) \le A\,\cosh(2\pi\,\lambda_{\mathrm{BKL}}\,s)$
on $\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}$.
The construction is presented as four explicit conjectures
(Conjectures M45.1.A--M45.1.D) and a one-sided numerical falsifier
based on direct computation of $\lambda_{\mathrm{BKL}}$ via the Gauss
shift and comparison with Krylov-rate growth in a discretised
Bianchi~IX wave-function evolution.
We do \emph{not} prove any of the four conjectures; six explicit
``[TBD: prove]'' markers are flagged.
The proposal is compatible with
the Heinzle--Uggla (2009) Bianchi~IX attractor theorem
and with our companion work on the Modular Shadow MSS bound for the
isotropic FLRW sector.
\end{abstract}

% =====================================================================
\section{Introduction}\label{sec:intro}
% =====================================================================

The Maldacena--Shenker--Stanford (MSS) bound on chaos~\cite{MSS2016}
states that the Lyapunov rate of any large-$N$ thermal quantum system
satisfies $\lambda_L \le 2\pi/\beta$.
Its modular-Krylov refinement
of Caputa--Mag\'{a}n--Patramanis--Tonni~\cite{Caputa2023} and the
universal operator-growth inequality of
Parker--Cao--Avdoshkin--Scaffidi--Altman~\cite{Parker2019} give a
sharper, large-$N$-free statement on the operator-Hilbert space of
modular evolution.

In parallel, the program initiated by Witten~\cite{Witten2018,Witten2022}
and Chandrasekaran--Longo--Penington--Witten~\cite{CLPW2022} has
established that gravitational observable algebras with quantum
reference frames are type-II$_\infty$ crossed products
$\mathcal{A}_{\mathrm{II}_\infty} = \mathcal{A}_{\mathrm{III}_1}
\rtimes_{\sigma^\phi} \mathbb{R}$, restoring a semifinite trace.
Speranza~\cite{Speranza2025} extended this to a generic
\emph{intrinsic} cosmological observer construction, applicable to
backgrounds without a global timelike Killing vector.
Kudler-Flam--Leutheusser--Satishchandran~\cite{KLS2024} concurrently
worked out the FLRW (isotropic) case in detail.

The Bianchi~IX vacuum, with its anisotropic SO(3)-symmetric metric and
three independent scale factors, is the canonical \emph{anisotropic}
analog of FLRW.
In the BKL approach to its initial-singularity behaviour, generic
solutions exhibit a chaotic infinite sequence of Kasner epochs
separated by bounces governed by a hyperbolic billiard in Misner
variables~\cite{DamourHenneauxNicolai2002,CornishLevin1996}.
Manin--Marcolli~\cite{ManinMarcolli2015} proved a remarkable
correspondence:
\begin{equation}
  \text{Mixmaster BKL bounce sequence}
  \quad \longleftrightarrow \quad
  \text{geodesic flow on } X(2) = \Gamma(2)\backslash\mathbb{H},
  \label{eq:MM-corr}
\end{equation}
realised by a two-sided continued-fraction shift, with Gauss measure
as the unique invariant probability of maximal entropy.
Fan--Fathizadeh--Marcolli~\cite{FFM2015} further showed that the
Connes spectral action on Bianchi~IX gravitational instantons admits
modular-form expansions of weights $2$, $14$, $18$.

De~Clerck--Hartnoll--Santos~\cite{HartnollDCS2024} studied AdS
Mixmaster chaos via the $\Gamma(2)$ arithmetic substrate
(arXiv:2312.11622), focusing on Hamiltonian-quantization eigenvalues
of the interior geometry.
Our framework provides the natural type-II$_\infty$ crossed-product
$+$ Krylov-complexity completion of that programme: we identify the
Speranza modular flow with the BKL Gauss-shift rate, and conjecture
a sharp $\lambda_K \le \lambda_{\mathrm{BKL}}$ bound.

\paragraph{Aim of this paper.}
We propose a \emph{conjectural} synthesis of the two pictures: combine
the Speranza/CLPW crossed-product construction with the
Manin--Marcolli geodesic-flow representation of BKL chaos, identify
the resulting type-II$_\infty$ algebra
$\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}$ as the
\emph{modular-shadow algebra} of the Bianchi~IX vacuum sector, and
\emph{conjecture} that its modular flow rate equals $2\pi$ times the
Kolmogorov--Sinai entropy $\lambda_{\mathrm{BKL}}$ of the Gauss shift.
This in turn predicts an MSS-style Krylov-complexity bound at
$\beta_{\mathrm{eff}} = 2\pi/\lambda_{\mathrm{BKL}}$.

\paragraph{What this paper does \emph{not} prove.}
None of the four sub-conjectures
(Conjectures~\ref{conj:wedge}--\ref{conj:bound}) is proved.
Six explicit ``[TBD: prove]'' markers are flagged in the
construction sketch (\S\ref{sec:construction}).
This is a \emph{programmatic} note: it identifies a precise, testable
target, sketches the construction, and presents a one-sided numerical
falsifier (\S\ref{sec:falsifier}).

\paragraph{Honesty bar.}
The Manin--Marcolli correspondence~\eqref{eq:MM-corr} is a theorem;
the type-II$_\infty$ crossed-product framework of
Witten/Speranza/CLPW is established; the MSS bound and its
Krylov refinement are established.
The \emph{novelty} here lies entirely in identifying the rate of the
Speranza modular flow with $2\pi\lambda_{\mathrm{BKL}}$.
That identification is open.

\paragraph{Organization.}
\S\ref{sec:prelim} reviews the three preliminary frameworks.
\S\ref{sec:conjecture} states the central Conjecture~M45.1.
\S\ref{sec:construction} sketches the four-step construction.
\S\ref{sec:bound} formulates the MSS-style Krylov bound.
\S\ref{sec:lemma} relates to our Bianchi~IX Lemma~A.1 (paper~\#5).
\S\ref{sec:falsifier} presents a numerical falsifier protocol.
\S\ref{sec:outlook} lists open problems.

% =====================================================================
\section{Preliminaries: three frameworks}\label{sec:prelim}
% =====================================================================

\subsection{Manin--Marcolli BKL $\leftrightarrow$ X(2) geodesic flow}\label{ssec:MM}

Misner variables for Bianchi~IX vacuum reduce the dynamics to the
geodesic flow of an indefinite metric on a hyperbolic billiard table
in $\mathbb{H}^2$~\cite{DamourHenneauxNicolai2002}.
Cornish--Levin~\cite{CornishLevin1996} demonstrated, via observer-independent
fractal methods, the existence of a strange repellor in the
mixmaster dynamics.
Manin--Marcolli~\cite{ManinMarcolli2015} showed that the BKL
Kasner-bounce sequence is conjugate to the two-sided shift
$\sigma:\Sigma_2\to\Sigma_2$ associated with the continued-fraction
expansion, equivalently the geodesic flow on the modular surface
$X(2) = \Gamma(2)\backslash\mathbb{H}$ along its imaginary axis.
The unique invariant probability of maximal entropy for $\sigma$ is
the Gauss measure
\begin{equation}
  d\mu_{\mathrm{Gauss}}(x)
  = \frac{1}{\log 2}\,\frac{dx}{1+x},\qquad x\in[0,1],
  \label{eq:gauss-measure}
\end{equation}
with topological entropy $h_{\mathrm{top}}(\sigma) = \pi^2/6$ (the
``Lochs--Khinchin'' constant; see~\cite{ManinMarcolli2015} \S2.4).

\begin{definition}[BKL Lyapunov rate]\label{def:lambda-BKL}
We denote by
\begin{equation}
  \lambda_{\mathrm{BKL}}
  \;:=\; h_{\mathrm{KS}}(\sigma,\mu_{\mathrm{Gauss}})
  \;=\; \frac{\pi^2}{6\log 2}
  \;\approx\; 2.3731,
  \label{eq:lambda-BKL}
\end{equation}
the Kolmogorov--Sinai entropy of the Gauss shift with respect to
$\mu_{\mathrm{Gauss}}$ (Manin--Marcolli~\cite{ManinMarcolli2015},
Theorem~2.4 and~\cite{FFM2015}).
\end{definition}

\subsection{Witten II$_\infty$ crossed product}\label{ssec:Witten}

Let $\mathcal{N}$ be the type-III$_1$ von~Neumann algebra of
gauge-invariant local field operators on a wedge-complement region of
some background spacetime, in the cyclic-and-separating GNS
representation of a KMS state $\phi$ at inverse temperature $\beta$.
Tomita--Takesaki gives the modular automorphism group
$\sigma_t^\phi:\mathcal{N}\to\mathcal{N}$.
Witten~\cite{Witten2022} and CLPW~\cite{CLPW2022}
showed that the crossed product
$\mathcal{N}\rtimes_{\sigma^\phi}\mathbb{R}$
is type-II$_\infty$, semifinite, with trace dual to the modular flow.
The Connes--Takesaki flow of weights identifies the natural rate
of $\sigma_t^\phi$ with $2\pi/\beta$ (in normalised units).

\subsection{Speranza intrinsic observer}\label{ssec:Speranza}

For backgrounds without a global timelike Killing vector
(de~Sitter static patch, FLRW, Bianchi),
Speranza~\cite{Speranza2025} constructs an \emph{intrinsic
cosmological observer}: a worldline $\gamma$ equipped with a
quantum reference clock, and an associated crossed-product algebra
\begin{equation}
  \mathcal{A}_\gamma^{\mathrm{II}_\infty}
  \;=\; \mathcal{A}_{\gamma^c}^{\mathrm{III}_1}
        \rtimes_{\sigma^{\phi_\gamma}}\mathbb{R},
  \label{eq:speranza}
\end{equation}
where $\mathcal{A}_{\gamma^c}^{\mathrm{III}_1}$ is the type-III$_1$
algebra of fields on the wedge complement of $\gamma$, and
$\phi_\gamma$ is the geometric (vacuum-like) state for which a local
modular automorphism is well-defined.
Kudler-Flam--Leutheusser--Satishchandran~\cite{KLS2024} worked this
out explicitly for FLRW backgrounds.
For Bianchi~IX, the construction has not yet been carried out; this
gap is precisely what Conjectures~M45.1.B and M45.1.C below address.

% =====================================================================
\section{Central conjecture}\label{sec:conjecture}
% =====================================================================

Let $M_{\mathrm{BIX}}$ denote a Bianchi~IX vacuum spacetime in
mixmaster (BKL) phase, $\gamma$ an intrinsic Speranza observer with
worldline parametrised by Misner volume time $\tau$, and
$\mathcal{A}_{\mathrm{obs}} := \mathcal{A}_{\gamma^c}^{\mathrm{III}_1}$
the type-III$_1$ algebra of bounded gauge-invariant local field
observables on the wedge complement of $\gamma$ in
$M_{\mathrm{BIX}}$.

\begin{conjecture}[Bianchi~IX modular-shadow algebra,
Conjecture~M45.1]\label{conj:central}
There exists a faithful normal representation
$\pi:\mathcal{A}_{\mathrm{obs}}\to B(\mathcal{H}_{\mathrm{BIX}})$
on a Hilbert space $\mathcal{H}_{\mathrm{BIX}}$ unitarily equivalent
to $L^2(X(2),\mu_{\mathrm{Gauss}})\otimes\ell^2$, such that:
\begin{enumerate}
\item[(i)] the Speranza crossed product
$\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}
:= \mathcal{A}_{\mathrm{obs}}\rtimes_{\sigma^{\phi_\gamma}}\mathbb{R}$
is a type-II$_\infty$ factor with semifinite trace $\tau_{\mathrm{II}}$;
\item[(ii)] the modular automorphism $\sigma_t^{\phi_\gamma}$ is
implemented (up to inner automorphism) by the Manin--Marcolli geodesic
flow $g_t$ on $L^2(X(2),\mu_{\mathrm{Gauss}})$;
\item[(iii)] consequently, the natural modular flow rate is
$\lambda_{\mathrm{mod}} = 2\pi\,\lambda_{\mathrm{BKL}}$ with
$\lambda_{\mathrm{BKL}}$ from Definition~\ref{def:lambda-BKL}.
\end{enumerate}
\end{conjecture}

We split this into four sub-conjectures (\S\ref{sec:construction})
and derive the falsifiable Krylov bound
(Conjecture~\ref{conj:bound}, \S\ref{sec:bound}).

% =====================================================================
\section{Sketch of construction}\label{sec:construction}
% =====================================================================

% ============================================================
\subsection{Step A: Wedge algebra setup}\label{ssec:wedge}
% ============================================================
% M78 detailed sketch merged here (2026-05-06); replaces 20-line stub.
% Sub-markers TBD-1.1/1.2/1.3 expand on the original TBD-1.

\subsubsection*{Setup}

Let $M_{\mathrm{BIX}} = \mathbb{R}\times S^3$ carry a Bianchi~IX vacuum metric
\begin{equation}
  ds^2 = -N^2(\tau)\,d\tau^2
         + \tfrac{1}{4}\,e^{-2\Omega(\tau)}\,
           \bigl(e^{2\beta(\tau)}\bigr)_{ij}\,\sigma^i\otimes\sigma^j,
  \label{eq:BIX-metric}
\end{equation}
where $\tau$ is Misner volume time ($\Omega = -\frac{1}{3}\ln V$,
monotone in the BKL backward direction),
$\{\sigma^i\}$ are the left-invariant 1-forms on $S^3\cong\mathrm{SU}(2)$,
and $\beta = \beta_+\,\mathrm{diag}(1,1,-2) + \beta_-\,\mathrm{diag}(1,-1,0)$
is the Misner anisotropy matrix.
The spatial sections are homogeneous under the SO(3) left-action.
In the BKL (Mixmaster) phase approaching the singularity
$\tau\to+\infty$, the metric~\eqref{eq:BIX-metric} undergoes an infinite
sequence of Kasner epochs separated by bounces encoded by the
Gauss-shift map $\sigma_G(x)=\{1/x\}$~\cite{ManinMarcolli2015}.

Fix an intrinsic observer $\gamma:\mathbb{R}\to M_{\mathrm{BIX}}$
with worldline parametrised by $\tau$.
Define the \emph{wedge complement}
\begin{equation}
  W_\gamma = \{\,p\in M_{\mathrm{BIX}}\;:\;
     p \text{ is spacelike-separated from every point on }\gamma\,\}
\end{equation}
(well-defined locally, since the homogeneous SO(3) foliation provides a
canonical spacelike splitting~\cite{HeinzleUggla2009}).
Consider a free real scalar field $\phi$ on $M_{\mathrm{BIX}}$
satisfying the covariant Klein--Gordon equation $\Box_g\phi = m^2\phi$;
the graviton case is analogous but algebraically more involved.
Let $\mathcal{A}(W_\gamma)$ be the von~Neumann algebra of bounded
operators generated by the Weyl operators $W(f) = e^{i\phi(f)}$
with $\mathrm{supp}(f)\subset W_\gamma$, in the GNS representation
of a reference state~$\omega_{\mathrm{BIX}}$ (specified below).

\begin{conjecture}[Wedge algebra type, M45.1.A --- TBD-1]
\label{conj:wedge}
$\mathcal{A}(W_\gamma)$ is the unique hyperfinite type-\textrm{III}$_1$ factor
$\mathcal{R}_\infty$.
\end{conjecture}

\paragraph{Why the standard arguments do not apply.}
The classical route to type-III$_1$ for wedge algebras
runs through the \emph{Bisognano--Wichmann theorem}~\cite{BW76}:
in Minkowski space, the modular Hamiltonian of the Rindler wedge in the
Minkowski vacuum is the \emph{boost generator}~$K$, a global Killing
vector field.
By Tomita--Takesaki theory, this forces the modular automorphism
group $\sigma_t^{\omega}(A) = e^{itK}Ae^{-itK}$ to be a continuous
one-parameter group of outer automorphisms, and the associated Connes
spectrum $S(\mathcal{A}) = [0,\infty)$, which characterises
type-III$_1$~\cite{Witten2022}.
For the de~Sitter static patch,
Chandrasekaran--Longo--Penington--Witten~\cite{CLPW2022}
show the algebra is the unique hyperfinite III$_1$ factor
$\mathcal{R}_\infty$, again by identifying the
modular Hamiltonian with the de~Sitter isometry generator
(a Killing vector of the static patch).

\emph{Neither argument transfers to Bianchi~IX.}
The metric~\eqref{eq:BIX-metric} admits \emph{no global timelike Killing
vector}: the three spatial Killing vectors $\{\xi_i\}$ generating the
SO(3) symmetry are spacelike, and the Misner volume~$\tau$ is not a
Killing coordinate (it appears only as a synchronous gauge choice).
Consequently, the modular Hamiltonian of $\mathcal{A}(W_\gamma)$
is \emph{not} identified with any known geometric symmetry generator,
and the Bisognano--Wichmann route is unavailable.

We note also that the FLRW case handled by
Kudler-Flam--Leutheusser--Satishchandran~\cite{KLS2024}
follows a \emph{different} path: the FLRW algebra is shown to be
type-II$_\infty$ \emph{directly} (via the inflaton zero-mode and
the KMS condition with respect to conformal-time translations),
not via a prior type-III$_1$ identification.
The Bianchi~IX situation is harder: the anisotropic $\beta$-dynamics
breaks the FLRW conformal-time symmetry, so the KLS shortcut is also
unavailable.

\paragraph{Three partial strategies toward TBD-1.}

\vspace{4pt}
\noindent\textbf{Strategy~(i): Haag--Kastler axioms $+$ SO(3) homogeneity.}
If the net $O\mapsto\mathcal{A}(O)$ satisfies the Haag--Kastler
axioms (isotony, locality, covariance under the residual SO(3) isometry)
and the Bianchi~IX quantum field theory satisfies the
\emph{Buchholz--Junglas split property} for the homogeneous spatial foliation
(i.e., for any two causally complete regions $O_1\Subset O_2$,
there exists an intermediate type-I factor),
then standard AQFT theorems~\cite{Buchholz1986,BV1996} imply
that the local algebras are type-III factors.
The additional step from type-III to type-III$_1$ requires showing
that the Connes $S$-invariant $S(\mathcal{A}(W_\gamma))=[0,\infty)$;
we argue this below via the BKL ergodicity.

\noindent\textbf{[TBD-1.1: prove]} The split property for free
scalar fields on Bianchi~IX has not been established.
The proof would require uniform energy bounds on the quantum field
relative to the BIX homogeneous foliation ---
a non-trivial task given the time-dependent metric~\eqref{eq:BIX-metric}
and the absence of a static time.

\vspace{4pt}
\noindent\textbf{Strategy~(ii): Connes $S$-invariant via BKL ergodicity.}
For any factor $\mathcal{N}$ with faithful normal weight $\phi$,
the Connes $S$-invariant is defined as
\begin{equation}
  S(\mathcal{N}) \;=\; \bigcap_{\phi}\,
  \mathrm{Sp}(\Delta_\phi),
  \label{eq:S-invariant}
\end{equation}
where $\Delta_\phi$ is the modular operator of $\phi$ and
the intersection runs over all faithful normal semifinite weights.
$\mathcal{N}$ is type-III$_1$ if and only if $S(\mathcal{N})=[0,\infty)$,
type-III$_\lambda$ (for $\lambda\in(0,1)$) if $S(\mathcal{N})=
\{\lambda^n:n\in\mathbb{Z}\}$, and type-III$_0$ if $S(\mathcal{N})=\{0,1\}$.

For the reference state $\omega_{\mathrm{BIX}}$,
the modular operator $\Delta_{\omega_{\mathrm{BIX}}}$ encodes the
``temperature'' seen by the observer along $\gamma$.
The Kasner-bounce sequence in the BKL phase is encoded by the
Gauss-shift $\sigma_G$ with Kolmogorov--Sinai entropy
$\lambda_{\mathrm{BKL}} = \pi^2/(6\log 2) > 0$
(Definition~\ref{def:lambda-BKL}).
The positive KS entropy means the Gauss-shift is
\emph{strongly mixing} and \emph{Bernoulli} (Adler--Weiss theorem),
so the associated time-evolution has \emph{absolutely continuous
spectrum} on $L^2([0,1],\mu_{\mathrm{Gauss}})$.
If this continuous spectrum is reflected in the modular spectrum
$\mathrm{Sp}(\Delta_{\omega_{\mathrm{BIX}}})$
via the Manin--Marcolli geodesic-flow representation~\cite{ManinMarcolli2015},
one expects $S(\mathcal{A}(W_\gamma))=[0,\infty)$, i.e., type-III$_1$.

\noindent\textbf{[TBD-1.2: prove]} The link between BKL Gauss-shift KS entropy
(a classical dynamical systems result) and the quantum modular spectrum
$\mathrm{Sp}(\Delta_{\omega_{\mathrm{BIX}}})$ is conjectural.
Making it precise requires:
\begin{enumerate}
  \item[(a)] Identifying $\omega_{\mathrm{BIX}}$ as a KMS state
     (or quasi-free state) for some BIX-adapted dynamics;
  \item[(b)] Showing the modular operator $\Delta_{\omega_{\mathrm{BIX}}}$
     has continuous spectrum via the Manin--Marcolli embedding
     $\mathcal{A}(W_\gamma)\hookrightarrow
     B(L^2(X(2),\mu_{\mathrm{Gauss}}))$
     (which is itself Conjecture~M45.1.C);
  \item[(c)] Ruling out III$_\lambda$ ($\lambda\in(0,1)$) ---
     i.e., that the spectrum is the full $[0,\infty)$ and not
     a discrete subgroup $\{\lambda^n\}$.
\end{enumerate}
For step~(c), the aperiodicity of the Gauss shift (its spectrum is purely
continuous on $L^2$ with no point masses at $\{\lambda^n\}$) is encouraging
but does not automatically imply III$_1$ for the quantum algebra.

\vspace{4pt}
\noindent\textbf{Strategy~(iii): Mixmaster scrambling as surrogate for geometric modular flow.}
In de~Sitter, the de~Sitter temperature $T_{dS} = H/2\pi$ controls the
modular flow rate, with $T_{dS}$ arising from the Killing horizon structure.
For Bianchi~IX, there is no Killing horizon, but the Mixmaster oscillations
produce an effective ``scrambling temperature'' at rate $\lambda_{\mathrm{BKL}}$.
One may \emph{conjecture} (heuristically) that the modular flow of
$\mathcal{A}(W_\gamma)$ in $\omega_{\mathrm{BIX}}$ is
\emph{ergodic relative to the Gauss-measure foliation of $X(2)$}, and
that this ergodicity --- in the quantum context --- forces
$S(\mathcal{A}(W_\gamma)) = [0,\infty)$.

\noindent\textbf{[TBD-1.3: prove]} This heuristic requires:
identifying a \emph{geometric} realisation of the modular Hamiltonian
$K_{\mathrm{BIX}}$ as a differential operator on
$L^2(X(2),\mu_{\mathrm{Gauss}})$;
and showing $K_{\mathrm{BIX}}$ has purely continuous spectrum.
A possible route is the Bost--Connes arithmetic algebra associated with
$\Gamma(2)$ (Connes--Marcolli, \emph{Noncommutative Geometry, Quantum
Fields and Motives}, Ch.~III), where the modular Hamiltonian is related
to the geodesic flow generator on $X(2)$ and the spectrum is known to be
continuous.
This route depends on Conjecture~M45.1.C and is therefore circular
without an independent proof of Step~C.

\subsubsection*{Summary of TBD-1 sub-markers}

\begin{center}
\small
\begin{tabular}{lll}
\toprule
Marker & Dependency & Status \\
\midrule
TBD-1.1 & Split property for free scalar on BIX homogeneous foliation
        & Open (AQFT/curved spacetime) \\
TBD-1.2 & Gauss-shift continuous spectrum $\Rightarrow$ $S(\mathcal{A})=[0,\infty)$
        & Open (requires steps (a)--(c) above) \\
TBD-1.3 & BIX modular Hamiltonian = geodesic flow generator on $X(2)$
        & Open (circular with Conj.~M45.1.C) \\
\bottomrule
\end{tabular}
\end{center}

\noindent The most tractable near-term route is
Strategy~(i) $+$ partial~(ii):
establish the Haag--Kastler net and split property on BIX homogeneous
slices (using SO(3) covariance and energy bounds for the free scalar
in a time-dependent homogeneous background), obtain the type-III label,
and then use the Gauss-shift ergodicity as heuristic input to argue
III$_1$ rather than III$_\lambda$.
A rigorous proof of type-III$_1$ for any non-Killing-symmetric cosmological
background would be a significant result in AQFT on curved spacetime
in its own right, independent of the applications to crossed products
pursued here.

\subsection{Step B: Crossed product}\label{ssec:crossed}

Given Conjecture~\ref{conj:wedge}, Tomita--Takesaki provides the
modular automorphism $\sigma_t^{\phi_\gamma}$ for any KMS state
$\phi_\gamma$, e.g., the Hartle--Hawking-analogue state on
$M_{\mathrm{BIX}}$ (whose existence is itself non-trivial in
Bianchi~IX; see~\cite{HeinzleUggla2009}).

\begin{conjecture}[Speranza transfer, M45.1.B]\label{conj:speranza}
The Speranza intrinsic-observer crossed-product
construction~\eqref{eq:speranza} applies to $M_{\mathrm{BIX}}$ with
$\tau$-foliation, yielding a type-II$_\infty$ factor
$\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}
= \mathcal{A}_{\mathrm{obs}}\rtimes_{\sigma^{\phi_\gamma}}\mathbb{R}$
with semifinite trace $\tau_{\mathrm{II}}$.
\end{conjecture}

\noindent\textbf{[TBD-2: prove]} Speranza's flow-of-weights argument
uses cosmological time. Bianchi~IX has no global cosmological time:
only the Misner volume $\tau = -\frac{1}{3}\log V$ plays a comparable
role, and $\tau$ is monotonic only in the BKL backward direction.
The argument must handle the non-globality of $\tau$.

\subsection{Step C: Modular-flow $=$ Manin--Marcolli geodesic flow}\label{ssec:mm-flow}

This is the load-bearing identification.

\begin{conjecture}[Flow identification, M45.1.C]\label{conj:flow}
On $\mathcal{H}_{\mathrm{BIX}} \cong L^2(X(2),\mu_{\mathrm{Gauss}})\otimes\ell^2$,
the modular flow $\sigma_t^{\phi_\gamma}$ of Conjecture~\ref{conj:speranza}
acts (up to inner automorphism on the $\ell^2$ factor) as
\begin{equation}
  \sigma_t^{\phi_\gamma}(f)
  \;=\; f \circ g_{-c\,t},\qquad
  f \in L^2(X(2),\mu_{\mathrm{Gauss}}),
  \label{eq:flow-id}
\end{equation}
where $g_t$ is the Manin--Marcolli geodesic flow on $X(2)$ and
$c > 0$ is a normalisation constant identified below.
\end{conjecture}

\noindent\textbf{[TBD-3: prove]} The identification of the
Speranza-Connes-Takesaki flow with the Manin--Marcolli geodesic flow
is the deepest open step. A possible route: realise
$\mathcal{A}_{\mathrm{obs}}$ as a sub-algebra of the Bost--Connes type
arithmetic algebra associated with $\Gamma(2)$ (Connes--Marcolli
\emph{Noncommutative Geometry, Quantum Fields and Motives}, Ch.~III),
where the modular flow has an intrinsic geometric realisation by the
geodesic flow on $X(2)$.

\subsection{Step D: Modular-flow rate}\label{ssec:rate}

Given Conjecture~\ref{conj:flow}, the rate of $\sigma_t^{\phi_\gamma}$
relative to the natural Gauss-measure parametrisation is
$\lambda_{\mathrm{mod}} = 2\pi/\beta_{\mathrm{eff}}$ where
$\beta_{\mathrm{eff}}$ is the inverse Hawking-like temperature of the
Bianchi~IX intrinsic observer.

\begin{conjecture}[Rate identification, M45.1.D]\label{conj:rate}
The normalisation constant $c$ in~\eqref{eq:flow-id} is fixed by
\begin{equation}
  \lambda_{\mathrm{mod}} \;=\; 2\pi\,\lambda_{\mathrm{BKL}}
  \;=\; \frac{\pi^3}{3\log 2}
  \;\approx\; 14.907\quad\text{(natural BKL units)},
  \label{eq:rate}
\end{equation}
where $\lambda_{\mathrm{BKL}} = h_{\mathrm{KS}}(\sigma_{\mathrm{Gauss}})
= \pi^2/(6\log 2)$ from Definition~\ref{def:lambda-BKL}.
\end{conjecture}

\noindent\textbf{[TBD-4: prove]} The factor $2\pi$ comes from the
Bisognano--Wichmann normalisation $\beta_{\mathrm{BW}}=2\pi$ in the
absence of explicit boost generator.

% =====================================================================
\section{Modular Krylov complexity bound}\label{sec:bound}
% =====================================================================

Combining Conjectures~M45.1.A--M45.1.D with the modular-Krylov
formalism of Caputa--Mag\'{a}n--Patramanis--Tonni~\cite{Caputa2023} and
the universal Lanczos-cosh bound of
Parker--Cao--Avdoshkin--Scaffidi--Altman~\cite{Parker2019} yields a
falsifiable target.

\begin{conjecture}[Bianchi~IX modular Krylov bound]\label{conj:bound}
For any operator $\mathcal{O}\in\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}$
with finite modular two-point function
$(\mathcal{O},\mathcal{O})_{\phi_\gamma}<\infty$, the modular Krylov
complexity satisfies
\begin{equation}
  \mathcal{C}_K^{\mathrm{mod}}(\mathcal{O};s)
  \;\le\;
  A_{\mathcal{O}}\;\cosh\!\bigl(2\pi\,\lambda_{\mathrm{BKL}}\,s\bigr),
  \label{eq:Krylov-bound}
\end{equation}
with $A_{\mathcal{O}}$ a state-dependent prefactor and
$\lambda_{\mathrm{BKL}} = \pi^2/(6\log 2)$.
The corresponding modular Lyapunov exponent is bounded by
$\lambda_K^{\mathrm{mod}} \le 2\pi\,\lambda_{\mathrm{BKL}}$.
\end{conjecture}

\noindent\textbf{[TBD-5: prove]} The Parker--Cao argument
in~\cite{Parker2019} requires linearly growing Lanczos coefficients
$b_n \sim \alpha n$. Their argument transposes from
Hamiltonian to modular evolution (this is the
Caputa~et~al.~\cite{Caputa2023} step) but the
\emph{slope} $\alpha = \pi\lambda_{\mathrm{BKL}}$ is conjectural.

\paragraph{Counter-example caveat.}
Bhattacharjee--Cao--Nandy--Pathak~\cite{BCNP2022} exhibit
saddle-dominated scrambling systems where Krylov complexity grows
exponentially without underlying chaos. In Bianchi~IX, the BKL
stochasticity is genuine (Gauss-shift entropy is positive); but the
\emph{converse} statement --- saturation of~\eqref{eq:Krylov-bound}
implies BKL chaos --- is \emph{not} claimed here.

\paragraph{Effective temperature.}
Define the effective inverse temperature of the Bianchi~IX intrinsic
observer by
\begin{equation}
  \beta_{\mathrm{eff}}
  \;:=\; \frac{2\pi}{\lambda_{\mathrm{BKL}}}
  \;=\; \frac{12\log 2}{\pi}
  \;\approx\; 2.6479.
  \label{eq:beta-eff}
\end{equation}
This is a \emph{prediction}: the intrinsic Hawking-like temperature
of an observer in the BKL phase of Bianchi~IX is
$T_{\mathrm{eff}} = 1/\beta_{\mathrm{eff}} \approx 0.3777$ in natural
BKL units.
The von~Neumann coupling constant of
$\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}$
(in the Murray--von Neumann sense) is conjecturally
proportional to $\lambda_{\mathrm{BKL}}^{-1}$.

\noindent\textbf{[TBD-6: prove]} The proportionality constant in the
von~Neumann coupling identification is undetermined and constitutes
the main content of the rate conjecture beyond the symbolic-dynamics
input.

% =====================================================================
\section{Connection to Lemma~A.1 (paper~\#5)}\label{sec:lemma}
% =====================================================================

In~\cite{ECI-paper5} we proved (Lemma~A.1) that the set of
configurations in Bianchi~IX vacuum with eigenvalue crossings of the
spatial Ricci tensor has Lebesgue measure zero in the Misner
configuration space.
This is consistent with Conjecture~M45.1: eigenvalue crossings
correspond to non-generic Kasner-bounce orbits, which form a
measure-zero subset of $X(2)$ under $\mu_{\mathrm{Gauss}}$ (the
periodic orbits of $\sigma$, dense but countable).

\paragraph{Compatibility check.}
The Fan--Fathizadeh--Marcolli~\cite{FFM2015} expansion of the
Connes spectral action on Bianchi~IX in modular forms of weight
$2,14,18$ should restrict on
$\mathcal{M}_{\mathrm{BIX}}^{\mathrm{II}_\infty}$ to a graded
trace-class component; this is a non-trivial consistency check on
Conjecture~M45.1.

% =====================================================================
\section{Falsifier protocol}\label{sec:falsifier}
% =====================================================================

The construction above is falsifiable.
A concrete numerical protocol (developed in detail in the companion
note \texttt{falsifier\_protocol.md}) consists of:

\begin{enumerate}
\item[(F1)] \emph{Numerical verification of $\lambda_{\mathrm{BKL}}$.}
Iterate the Gauss shift $\sigma:[0,1]\to[0,1]$,
$\sigma(x)=\{1/x\}$, on $N=10^6$ trajectories of length $T=10^4$,
sample the Lyapunov rate via finite-time Birkhoff averages, and
verify $\lambda_{\mathrm{BKL}} = \pi^2/(6\log 2) \approx 2.3731$ to
four significant figures. (This is a textbook exercise; serves as
sanity check.)

\item[(F2)] \emph{Discretised Bianchi~IX Krylov rate.}
Take a truncation of the WDW (Wheeler--DeWitt) Hamiltonian on
Bianchi~IX in Misner variables to $n_{\mathrm{rank}}=10^3$ basis
modes, compute the Lanczos coefficients $b_n$ of a generic local
field operator under the truncated modular evolution, and extract the
Krylov-complexity growth rate $\lambda_K^{\mathrm{trunc}}$.

\item[(F3)] \emph{Rate comparison.}
Conjecture~\ref{conj:bound} predicts
$\lambda_K^{\mathrm{trunc}} \le 2\pi\,\lambda_{\mathrm{BKL}}
\approx 14.907$, with saturation in the BKL-chaotic regime and
violation only in the (measure-zero) periodic-Kasner orbits.
A numerical Lyapunov rate exceeding $14.907$ by more than the truncation
error would falsify Conjecture~M45.1.

\item[(F4)] \emph{Lemma~A.1 compatibility.}
Verify that the periodic-orbit subset of $X(2)$ (of $\sigma$-measure
zero) maps to the eigenvalue-crossing set of paper~\#5
Lemma~A.1 under the Misner-to-modular-surface change of variables.
\end{enumerate}

% =====================================================================
\section{Outlook}\label{sec:outlook}
% =====================================================================

\begin{itemize}
\item \emph{Lean formalization.} Step~A (\S\ref{ssec:wedge}) and the
Gauss-shift entropy computation~\eqref{eq:lambda-BKL} are tractable
in mathlib (continued fractions and ergodic theory both exist).
Steps B, C, D require von~Neumann algebras (sparse in mathlib as of
2026).
\item \emph{Extension to Bianchi class A (types VIII, VI$_{-1/9}$).}
The Manin--Marcolli framework
extends~\cite{ManinMarcolli2015} to all class~A Bianchi types; the
construction here should also extend.
\item \emph{Genus-$g\ge 2$.} The companion paper~\#4 (Modular Shadow
LMP v2.5) noted that the genus-$g\ge 2$ extension fails for the FLRW
case. The Bianchi~IX setting is intrinsically tied to $X(2)$ (genus
zero) and so the question is moot here; but extension to higher level
$X(N)$ is interesting.
\item \emph{Co-authors targeted.} Matilde Marcolli (Caltech, for the
symbolic-dynamics half) and Antony Speranza (Maryland, for the
intrinsic-observer construction); see the companion
\texttt{outreach\_strategy.md}.
\end{itemize}

% =====================================================================
\section*{Discipline statement}
% =====================================================================

\begin{itemize}
\item All 14 arXiv references in the bibliography below were
WebFetch-verified on 2026-05-06 against \texttt{arxiv.org} directly.
\item Six \textbf{[TBD: prove]} markers are flagged in
\S\ref{sec:construction}--\ref{sec:bound}; this paper proves
\emph{none} of the four conjectures M45.1.A--M45.1.D.
\item No large-language-model output was treated as authoritative
without independent live re-verification (Mistral strict-banned per
project policy).
\item This is a \emph{conjecture paper}: the Manin--Marcolli
correspondence~\eqref{eq:MM-corr} is a theorem, the
Speranza/Witten/CLPW crossed-product framework is established, but
the identification $\sigma_t^{\phi_\gamma} \leftrightarrow g_t$
(Conjecture~\ref{conj:flow}) is the load-bearing open step.
\end{itemize}

% =====================================================================
\begin{thebibliography}{99}
% =====================================================================

\bibitem{MSS2016}
J.~Maldacena, S.~H.~Shenker, D.~Stanford,
\emph{A bound on chaos},
JHEP \textbf{08} (2016) 106, arXiv:1503.01409.

\bibitem{Parker2019}
D.~E.~Parker, X.~Cao, A.~Avdoshkin, T.~Scaffidi, E.~Altman,
\emph{A Universal Operator Growth Hypothesis},
Phys.~Rev.~X \textbf{9}, 041017 (2019), arXiv:1812.08657.

\bibitem{Caputa2023}
P.~Caputa, J.~M.~Mag\'{a}n, D.~Patramanis, E.~Tonni,
\emph{Krylov complexity of modular Hamiltonian evolution},
Phys.~Rev.~D \textbf{109}, 086013 (2024), arXiv:2306.14732.

\bibitem{Witten2018}
E.~Witten,
\emph{Notes on Some Entanglement Properties of Quantum Field Theory},
Rev.~Mod.~Phys. \textbf{90}, 045003 (2018), arXiv:1803.04993.

\bibitem{Witten2022}
E.~Witten,
\emph{Gravity and the Crossed Product},
JHEP \textbf{10} (2022) 008, arXiv:2112.12828.

\bibitem{CLPW2022}
V.~Chandrasekaran, R.~Longo, G.~Penington, E.~Witten,
\emph{An Algebra of Observables for de Sitter Space},
JHEP \textbf{02} (2023) 082, arXiv:2206.10780.

\bibitem{KLS2024}
J.~Kudler-Flam, S.~Leutheusser, G.~Satishchandran,
\emph{Algebraic Observational Cosmology},
arXiv:2406.01669 (2024).

\bibitem{Speranza2025}
A.~J.~Speranza,
\emph{An intrinsic cosmological observer},
arXiv:2504.07630 (2025).

\bibitem{ManinMarcolli2015}
Yu.~Manin, M.~Marcolli,
\emph{Symbolic Dynamics, Modular Curves, and Bianchi IX Cosmologies},
Ann.~Fac.~Sci.~Toulouse Math. \textbf{25} (2016), arXiv:1504.04005.

\bibitem{FFM2015}
W.~Fan, F.~Fathizadeh, M.~Marcolli,
\emph{Modular forms in the spectral action of Bianchi IX
gravitational instantons},
JHEP \textbf{04} (2020) 087, arXiv:1511.05321.

\bibitem{DamourHenneauxNicolai2002}
T.~Damour, M.~Henneaux, H.~Nicolai,
\emph{Cosmological Billiards},
Class.~Quant.~Grav. \textbf{20} (2003) R145, arXiv:hep-th/0212256.

\bibitem{CornishLevin1996}
N.~Cornish, J.~Levin,
\emph{The Mixmaster Universe: A Chaotic Farey Tale},
Phys.~Rev.~D \textbf{55} (1997) 7489, arXiv:gr-qc/9612066.

\bibitem{HeinzleUggla2009}
J.~M.~Heinzle, C.~Uggla,
\emph{A new proof of the Bianchi type IX attractor theorem},
Class.~Quant.~Grav. \textbf{26} (2009) 075015, arXiv:0901.0806.

\bibitem{BCNP2022}
B.~Bhattacharjee, X.~Cao, P.~Nandy, T.~Pathak,
\emph{Krylov complexity in saddle-dominated scrambling},
JHEP \textbf{05} (2022) 174, arXiv:2203.03534.

\bibitem{ECI-paper5}
K.~Remondi\`ere,
\emph{Eigenvalue-crossing measure-zero lemma for Bianchi~IX vacuum
configurations} (paper~\#5, in preparation, 2026).

%%  R2 Bisognano-Wichmann (no arXiv; journal-verified)
\bibitem{BW76}
J.~J.~Bisognano, E.~H.~Wichmann,
\emph{On the duality condition for a Hermitian scalar field},
J.~Math.~Phys. \textbf{16} (1975) 985;
\emph{On the duality condition for quantum fields},
J.~Math.~Phys. \textbf{17} (1976) 303.

%%  R3 Buchholz-Junglas split property (no arXiv; Commun. Math. Phys.)
\bibitem{Buchholz1986}
D.~Buchholz, E.~H.~Wichmann,
\emph{Causal independence and the energy-level density of states in local
quantum field theory},
Commun.~Math.~Phys. \textbf{106} (1986) 321--344.

%%  R4 Buchholz-Verch (split property, type-III review)
\bibitem{BV1996}
D.~Buchholz, R.~Verch,
\emph{Scaling algebras and renormalization group in algebraic quantum field
theory},
Rev.~Math.~Phys. \textbf{7} (1995) 1195--1240.

%%  R1 Hartnoll-DCS reference (M59, 2026-05-06; arXiv live-verified; JHEP M48-verified)
\bibitem{HartnollDCS2024}
M.~De~Clerck, S.~A.~Hartnoll, and J.~E.~Santos,
\emph{Mixmaster chaos in an AdS black hole interior},
JHEP \textbf{07} (2024) 202, arXiv:2312.11622.

\end{thebibliography}

\end{document}
