% frw_note.tex --- standalone math.OA note
% Type II_infty for the gravitational algebra in radiation-dominated FRW
% via conformal pullback (no Killing vector required).
%
% Author: Kevin Remondiere
% Date: 2026-05-02
% Compile: pdflatex frw_note ; bibtex frw_note ; pdflatex frw_note ; pdflatex frw_note

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{cite}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Open Question}

% Macros
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\HH}{\mathcal{H}}
\renewcommand{\AA}{\mathcal{A}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\OO}{\mathcal{O}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\Ad}{\operatorname{Ad}}
\newcommand{\BW}{\mathrm{BW}}
\newcommand{\HL}{\mathrm{HL}}
\newcommand{\FRW}{\mathrm{FRW}}
\newcommand{\Mink}{\mathrm{M}}
\newcommand{\Tr}{\operatorname{Tr}}
\DeclareMathOperator{\Sp}{Sp}

\title{Type II$_\infty$ for the gravitational algebra of a comoving observer in radiation-dominated FRW via conformal pullback\thanks{Standalone technical note. The result was first recorded in commit \texttt{aa9a3ee} of \url{https://github.com/AIdevsmartdata/crossed-cosmos} (2026-05-02).}}
\author{Kevin Remondi\`ere\\
\small Independent researcher\\
\small \href{mailto:kevin.remondiere@gmail.com}{kevin.remondiere@gmail.com}\\
\small ORCID: \href{https://orcid.org/0009-0008-2443-7166}{0009-0008-2443-7166}}
\date{2026-05-02}

\begin{document}
\maketitle

\begin{abstract}
We observe that, for the conformally coupled massless scalar field on a radiation-dominated FRW spacetime ($a(\eta)=a_0 \eta$), the local algebra $\AA(\OO)_\FRW$ associated to a causal diamond of a comoving observer with $\eta_i>0$ (away from the Big Bang) is \emph{unitarily} equivalent, as a W$^*$-algebra equipped with its vacuum state, to the corresponding Minkowski local algebra $\AA(M_\OO)_\Mink$ on the conformally identified diamond. The unitary is the second-quantised version of the standard field rescaling $\varphi \mapsto a(\eta)\,\varphi$; we verify symbolically (sympy) the off-shell conformal-covariance identity for the wave operator and the cancellation of $\partial_\eta a$ derivative terms in the symplectic form on a Cauchy slice. By Tomita--Takesaki uniqueness, the modular automorphism group of $(\AA(\OO)_\FRW,\Omega_\FRW)$ is therefore unitarily intertwined with the Hislop--Longo modular flow of the Minkowski diamond in the vacuum, hence has Connes spectrum $\Gamma(\sigma)=\R$. Consequently the crossed product $\AA(\OO)_\FRW \rtimes_\sigma \R$ is type II$_\infty$. This sidesteps the stationarity hypothesis (Killing vector field) of Chandrasekaran--Longo--Penington--Witten 2022 and gives the first explicit non-stationary cosmological example of the CLPW type II$_\infty$ classification. The argument extends verbatim to matter-dominated FRW ($a(\eta)\propto \eta^2$, Theorem~\ref{thm:matter}) and to slow-roll inflation modelled by exact de Sitter ($a(\eta) = -1/(H\eta)$, Theorem~\ref{thm:dS}), provided the diamond is bounded away from $\eta = 0$ in each case so that $a(\eta)$ is smooth and strictly positive on the closure. The argument is strictly limited to the conformally coupled massless scalar on a conformally flat background: generic FRW backgrounds (with non-conformal matter content), fields with $\xi \neq 1/6$ or $m \neq 0$, and wedge-type regions extending to $\eta=0$ are explicitly outside its scope.
\end{abstract}

\section{Introduction}\label{sec:intro}

The crossed-product construction of Witten~\cite{Witten2022} and Chandrasekaran--Longo--Penington--Witten (CLPW)~\cite{CLPW2023} produces, from the type III$_1$ local algebra of a free quantum field theory and a suitable modular automorphism group, a type II von Neumann algebra of \emph{gravitationally dressed} observables. The key ingredient is a one-parameter group of automorphisms with full Connes spectrum $\Gamma(\sigma)=\R$, which under crossed-product yields a type II$_\infty$ factor (semifinite, with a well-defined trace up to scaling). For static patches of de Sitter space CLPW~\cite{CLPW2023} obtain instead a type II$_1$ factor by including the constraint that the modular Hamiltonian of the static patch is bounded below. Faulkner--Speranza~\cite{FaulknerSperanza2024} extend the crossed-product framework to general Killing horizons and prove a non-perturbative generalised second law in this setting.

In all of these constructions, the modular automorphism is identified with the flow of a Killing vector field of the background spacetime: the boost in Minkowski (Bisognano--Wichmann)~\cite{BisognanoWichmann1975, BisognanoWichmann1976}, the de Sitter boost in the static patch~\cite{CLPW2023}, the Killing horizon generator in~\cite{FaulknerSperanza2024}. In other words, all known cases of the CLPW type-II classification rely, at the present level of the literature, on the existence of a stationary Killing vector field whose flow geometrically implements the modular automorphism.

The purpose of this note is to record one example where this stationarity hypothesis can be removed. We consider a radiation-dominated FRW spacetime, which has \emph{no} timelike Killing vector field along a generic comoving worldline, but \emph{is} conformally flat. For the conformally coupled massless scalar field, the conformal isometry to Minkowski space lifts to a unitarily implemented W$^*$-algebraic isomorphism of local algebras, intertwining the FRW vacuum (conformal vacuum) and the Minkowski vacuum. This reduces the FRW modular flow to the Minkowski Hislop--Longo~\cite{HislopLongo1982} modular flow on the conformally identified diamond. Tomita--Takesaki uniqueness then yields the type II$_\infty$ classification of the crossed product, with no Killing vector required.

\medskip

\noindent\textbf{Statement (informal, see Theorem~\ref{thm:main}).} \emph{Let $\varphi$ be the conformally coupled massless scalar field on radiation-dominated FRW with $a(\eta)=a_0\eta$ on $\eta>0$, in the conformal vacuum. Let $\OO=D_\OO$ be a causal diamond of a comoving observer with $0<\eta_i<\eta_f<\infty$, and $\AA(\OO)$ the associated local von Neumann algebra. Then $\AA(\OO)\rtimes_\sigma \R$ is a type II$_\infty$ factor.}

\medskip

\noindent\textbf{Scope and limitations.} The argument uses conformal flatness of the FRW background and conformal invariance of the field equation $(\Box_g - R/6)\varphi=0$ in $d=4$. It applies as written to any conformally flat FRW background with a strictly positive scale factor on the closed diamond, including matter-dominated FRW with $a(\eta)\propto \eta^2$ and slow-roll inflation modelled by exact de Sitter $a(\eta)=-1/(H\eta)$. We restrict throughout to \emph{causal diamonds} $D_\OO$ with $\eta_i>0$ (resp.\ $\eta_f<0$ for de Sitter): the wedge case is excluded because a Rindler-type wedge in conformal coordinates extends to $\eta\to 0$ and unavoidably crosses the singularity (or, in the de Sitter case, the conformal boundary). The argument does \emph{not} apply to generic FRW backgrounds beyond conformal flatness, nor to massive or non-conformal scalar fields, nor to interacting fields. We discuss these limitations and the corresponding open problems in Section~\ref{sec:open}.

\section{Preliminaries}\label{sec:prelim}

\subsection{Conformal flatness of radiation-dominated FRW}\label{sec:conformal}

A spatially flat FRW spacetime in conformal time $\eta$ has metric
\begin{equation}
\label{eq:frw-metric}
\mathrm{d}s^2_\FRW = a(\eta)^2 \big(-\mathrm{d}\eta^2 + \mathrm{d}\mathbf{x}^2\big),
\qquad \mathbf{x}\in\R^3.
\end{equation}
For radiation-dominated matter content ($a^4\rho = \mathrm{const}$, $p=\rho/3$), the Friedmann equations integrate to
\begin{equation}
\label{eq:rd-scale}
a(\eta) = a_0\,\eta, \qquad \eta\in(0,\infty).
\end{equation}
The Big Bang is at $\eta=0$. The metric \eqref{eq:frw-metric} is, by inspection, $a(\eta)^2$ times the Minkowski metric $\eta_{\mu\nu} = \mathrm{diag}(-1,+1,+1,+1)$. This is the standard fact that any spatially flat FRW spacetime is conformally flat in conformal coordinates~\cite[\S5.1, \S5.3]{Wald1994}.

For matter-dominated FRW ($a^3\rho = \mathrm{const}$, $p=0$), the same coordinate change gives $a(\eta) \propto \eta^2$, again conformally flat. Generic mixed FRW backgrounds with non-conformal matter content (e.g.\ a cosmological constant + dust) are likewise conformally flat as Lorentzian manifolds, since spatial-flatness plus FRW symmetry forces the Weyl tensor to vanish; the conformal factor $a(\eta)$ is then determined by integrating the Friedmann equation.

\subsection{Conformally coupled massless scalar}\label{sec:cc}

The conformally coupled massless scalar in $d=4$ has classical action (Wald sign convention~\cite[\S4.6]{Wald1994})
\begin{equation}
\label{eq:S-cc}
S[\varphi] = -\tfrac{1}{2}\int \mathrm{d}^4 x\, \sqrt{-g}\,\Big[g^{\mu\nu}\partial_\mu\varphi\,\partial_\nu\varphi - \tfrac{1}{6}\,R\,\varphi^2\Big],
\end{equation}
with $\xi=1/6$ the conformal coupling. Its equation of motion is $(\Box_g - R/6)\varphi = 0$. Under a Weyl rescaling $g_{\mu\nu} = \Omega^2 \tilde g_{\mu\nu}$ with $\Omega=a(\eta)$ and $\tilde g_{\mu\nu}$ the Minkowski metric, the conformally rescaled field
\begin{equation}
\label{eq:rescale}
\tilde\varphi \,:=\, a(\eta)\,\varphi
\end{equation}
satisfies the flat wave equation $\Box_{\tilde g}\,\tilde\varphi = 0$ on Minkowski space, as a consequence of the off-shell conformal-covariance identity
\begin{equation}\label{eq:cc-identity}
\Box_{\tilde g}\big(a\varphi\big) \;=\; a^3\,\big(\Box_g - R/6\big)\varphi
\end{equation}
which is the standard conformal covariance of the $d=4$ conformally coupled scalar~\cite[\S4.6]{Wald1994}. We re-verify \eqref{eq:cc-identity} symbolically (sympy) in Section~\ref{sec:symplectic}.

\subsection{Local algebras and Bunch--Davies vacuum}\label{sec:algebras}

We work in the algebraic framework. To each open relatively compact region $\OO$ of the FRW spacetime with causally complete closure we associate the (smeared) field algebra
\begin{equation}
\label{eq:algebra}
\AA(\OO)_\FRW \,:=\, \pi_\Omega\big(\{\varphi(f) : f\in C_c^\infty(\OO)\}\big)'',
\end{equation}
the double commutant in the GNS representation $\pi_\Omega$ of the Bunch--Davies (= conformal) vacuum state $\Omega_\FRW$. Standard Haag--Kastler axioms hold in this representation; existence of the Bunch--Davies vacuum for the conformally coupled massless scalar in conformally flat FRW is classical~\cite[\S4.6]{Wald1994}.

The conformal vacuum is, by definition, the state whose two-point function pulls back to the Minkowski Wightman function under the rescaling \eqref{eq:rescale}: $\langle \varphi(x)\varphi(y)\rangle_{\Omega_\FRW} = a(\eta_x)^{-1} a(\eta_y)^{-1} \langle \tilde\varphi(x)\tilde\varphi(y)\rangle_{\Omega_\Mink}$. This identifies $\Omega_\FRW$ with the Minkowski vacuum after the field redefinition \eqref{eq:rescale}. Equivalently, the GNS Hilbert spaces are canonically identified by a unitary $U:\HH_\FRW\to\HH_\Mink$ implementing $\varphi(f)\mapsto \tilde\varphi(a\cdot f)$, with $U\Omega_\FRW=\Omega_\Mink$ (see \S\ref{sec:U-isom}).

\subsection{Diamond algebras (Hislop--Longo) and Bisognano--Wichmann background}\label{sec:BW}

For a doubly-bounded causal diamond $D \subset \R^{1,3}$ (the causal completion of an open ball on a Cauchy slice), the modular structure is given by the Hislop--Longo theorem~\cite{HislopLongo1982}: for the free massless scalar field, the modular automorphism group of $(\AA(D), \Omega_\Mink)$ is implemented by a one-parameter subgroup of conformal transformations of Minkowski space (a M\"obius flow on the diamond fixing its tips), and $\AA(D)$ is a type III$_1$ factor with full Connes spectrum $\Gamma(\sigma)=\R$. This is the regime in which the present note operates.

For background and contrast we recall Bisognano--Wichmann~\cite{BisognanoWichmann1975, BisognanoWichmann1976}. In Minkowski space, fix a Rindler wedge $W_R = \{x \in \R^{1,3} : x^1 > |x^0|\}$. The Tomita--Takesaki modular operator of the pair $(\AA(W_R)_\Mink, \Omega_\Mink)$ has modular automorphism group given by the boost flow,
\begin{equation}
\label{eq:BW}
\sigma_s^\BW(\varphi(x)) = \varphi(\Lambda_{2\pi s}\,x),
\end{equation}
where $\Lambda_t$ is the Lorentz boost in the $(x^0,x^1)$-plane with rapidity $t$, and $\Delta_\BW = \exp(-2\pi K_\BW)$ with $K_\BW$ the boost generator restricted to $W_R$.

\smallskip

\noindent\emph{II$_\infty$ vs II$_1$ via the modular Hamiltonian spectrum.} The CLPW~\cite{CLPW2023} type-II classification distinguishes type II$_1$ from type II$_\infty$ by whether the modular Hamiltonian $K=-\log\Delta$ is bounded below (yielding II$_1$, as in the de Sitter static patch) or unbounded both ways (yielding II$_\infty$). For the Minkowski Rindler wedge in the vacuum, the modular Hamiltonian is the boost generator $K_\BW$, whose self-adjoint spectrum is all of $\R$ (the boost generator on Minkowski space has continuous unbounded spectrum on both ends). Hence the Connes spectrum $\Gamma(\sigma^\BW)=\R$ and the Connes--Takesaki crossed product $\AA(W_R)\rtimes_{\sigma^\BW}\R$ is a type II$_\infty$ factor. The same holds for the Hislop--Longo modular flow on a Minkowski diamond, since the M\"obius-flow generator (a conformal Killing field) is conjugate to a boost generator in the $\mathrm{SO}(2,4)$ conformal group of $\R^{1,3}$ and shares the same spectral type. Throughout the present note, we work in the II$_\infty$ regime; II$_1$ would only arise if the modular Hamiltonian on the FRW diamond were bounded below, which is not the case once it is conjugated to the M\"obius generator.

\subsection{The crossed product}\label{sec:crossed}

Given a von Neumann algebra $\MM$ and a one-parameter group of automorphisms $\sigma:\R\to\Aut(\MM)$, the crossed product $\MM\rtimes_\sigma \R$ is the von Neumann algebra on $L^2(\R, \HH)$ generated by $\pi(m)\xi(s) = \sigma_{-s}(m)\xi(s)$ and translation $u_t \xi(s) = \xi(s-t)$. When $\MM$ is type III$_1$ and $\sigma$ is a modular automorphism with full Connes spectrum, $\MM\rtimes_\sigma\R$ is a type II$_\infty$ factor (Connes--Takesaki). This is the construction underlying CLPW~\cite{CLPW2023, Witten2022}.

\section{Main theorem}\label{sec:main}

\subsection{The symplectic form on FRW (sympy-verified)}\label{sec:symplectic}

We verify off-shell and explicitly the conformal-covariance identity \eqref{eq:cc-identity} and its consequence that the field rescaling $\varphi=\tilde\varphi/a$ matches the FRW symplectic form to the Minkowski symplectic form, with no leftover $\partial_\eta a$ term. Both calculations are double-checked symbolically in sympy~1.12; the scripts (\texttt{symplectic\_check.py}, \texttt{conformal\_box\_check.py}) are bundled with the source.

\smallskip

\noindent\textbf{(a) Conformal covariance of the wave operator.} For the metric $g_{\mu\nu} = a(\eta)^2\,\eta_{\mu\nu}$, direct computation of the Christoffels gives
\begin{equation}\label{eq:Box-FRW-explicit}
\Box_g\,\varphi \;=\; \frac{1}{a^4}\,\partial_\mu\big(a^4\,g^{\mu\nu}\,\partial_\nu \varphi\big)
\;=\; \frac{1}{a^2}\,\Box_{\tilde g}\,\varphi \;-\; \frac{2\,(\partial_\eta a)}{a^3}\,\partial_\eta \varphi,
\end{equation}
and the Ricci scalar of $g$ in conformal coordinates is $R = 6\,a^{-3}\,(\partial_\eta^2 a)$ (\cite[\S5.1]{Wald1994}; we re-derived it from the Christoffels in our sympy script). Combining \eqref{eq:Box-FRW-explicit} with $\Box_{\tilde g}(a\,\varphi) = a\,\Box_{\tilde g}\,\varphi - 2(\partial_\eta a)\partial_\eta\varphi - (\partial_\eta^2 a)\,\varphi$ gives, off-shell and exactly,
\begin{equation}\label{eq:cc-cov-explicit}
\Box_{\tilde g}\big(a\,\varphi\big) \;=\; a^3\,\Big(\Box_g\,\varphi \;-\; \tfrac{1}{6}\,R\,\varphi\Big),
\end{equation}
which is \eqref{eq:cc-identity}. \emph{Sympy verification.} The script \texttt{conformal\_box\_check.py} computes both sides of \eqref{eq:cc-cov-explicit} in $(\eta,\mathbf{x})$ components for arbitrary smooth $a(\eta)$ and arbitrary $\varphi(\eta,\mathbf{x})$, and reports a difference of zero after \texttt{simplify}. The sign of the curvature term in the action and EOM \eqref{eq:S-cc} matches Wald's convention~\cite[\S4.6]{Wald1994}.

\smallskip

\noindent\textbf{(b) Symplectic form.} On a constant-$\eta$ Cauchy slice $\Sigma_\eta\subset D_\OO$, the canonical symplectic form for the conformally coupled scalar (Klein--Gordon current $j_\mu = \varphi_1\,\partial_\mu \varphi_2 - \varphi_2\,\partial_\mu \varphi_1$, contracted with the unit normal and integrated against the induced volume) reads
\begin{equation}\label{eq:sigma-FRW-def}
\sigma_\FRW(\varphi_1,\varphi_2) \;=\; \int_{\Sigma_\eta}\big(\varphi_1\,\partial_\eta \varphi_2 - \varphi_2\,\partial_\eta \varphi_1\big)\,a^2\,\mathrm{d}^3\mathbf{x},
\end{equation}
where the weight $a^2$ comes from $\sqrt{h}\,n^\eta = a^3\cdot a^{-1} = a^2$, with $\sqrt{h}=a^3$ the induced 3-volume and $n^\mu = (1/a,\mathbf{0})$ the future-directed unit timelike normal in $g_\FRW$. (The curvature term $-(R/6)\varphi^2$ in the action is symmetric in any product $\varphi_1\varphi_2$ and contributes nothing to the antisymmetrised current; the symplectic form for a conformally coupled scalar coincides with the standard Klein--Gordon form on a Cauchy slice.) Substituting $\varphi_i = \tilde\varphi_i/a$:
\begin{align*}
\sigma_\FRW(\varphi_1,\varphi_2) &= \int_{\Sigma_\eta} a^2\Big[\tfrac{\tilde\varphi_1}{a}\partial_\eta\!\big(\tfrac{\tilde\varphi_2}{a}\big) - \tfrac{\tilde\varphi_2}{a}\partial_\eta\!\big(\tfrac{\tilde\varphi_1}{a}\big)\Big]\mathrm{d}^3\mathbf{x}\\
&= \int_{\Sigma_\eta}\Big[\tilde\varphi_1\Big(\tfrac{\partial_\eta\tilde\varphi_2}{a} - \tfrac{(\partial_\eta a)\tilde\varphi_2}{a^2}\Big)\cdot a - \tilde\varphi_2\Big(\tfrac{\partial_\eta\tilde\varphi_1}{a} - \tfrac{(\partial_\eta a)\tilde\varphi_1}{a^2}\Big)\cdot a\Big]\mathrm{d}^3\mathbf{x}\\
&= \int_{\Sigma_\eta}\big(\tilde\varphi_1\partial_\eta\tilde\varphi_2 - \tilde\varphi_2\partial_\eta\tilde\varphi_1\big)\mathrm{d}^3\mathbf{x}
\;-\; \int_{\Sigma_\eta}\frac{\partial_\eta a}{a}\big(\tilde\varphi_1\tilde\varphi_2 - \tilde\varphi_2\tilde\varphi_1\big)\,\mathrm{d}^3\mathbf{x}\\
&= \sigma_\Mink(\tilde\varphi_1,\tilde\varphi_2)\;-\;0.
\end{align*}
The two $(\partial_\eta a)/a$ terms cancel \emph{by antisymmetry of the symplectic-current bracket}, since $\tilde\varphi_1\tilde\varphi_2 = \tilde\varphi_2\tilde\varphi_1$ pointwise on the Cauchy slice. \emph{Sympy verification.} The script \texttt{symplectic\_check.py} computes the integrand
$
a^2\big(\varphi_1\partial_\eta\varphi_2 - \varphi_2\partial_\eta\varphi_1\big) - \big(\tilde\varphi_1\partial_\eta\tilde\varphi_2 - \tilde\varphi_2\partial_\eta\tilde\varphi_1\big)
$
with $\varphi_i = \tilde\varphi_i/a$ for arbitrary smooth $a(\eta)$ and arbitrary smooth $\tilde\varphi_1,\tilde\varphi_2$, and reports a difference of zero after \texttt{simplify}.

\smallskip

\noindent At the level of test-function smearings we use the convention $\varphi(f) := \int \varphi(x)\, f(x)\, \sqrt{-g}\,\mathrm{d}^4 x$ on $g_\FRW$ and $\tilde\varphi(h) := \int \tilde\varphi(x)\, h(x)\,\mathrm{d}^4 x$ on Minkowski. Then $\varphi(f) = \tilde\varphi(a^3 f)$, so the natural test-function bijection $C_c^\infty(D_\OO)\to C_c^\infty(M_\OO)$ is $f\mapsto a^3 f$. Smoothness and strict positivity of $a$ on $\overline{D_\OO}$ make this a $C_c^\infty$-bijection (its inverse is $h\mapsto a^{-3} h$, which is also $C_c^\infty$). The symplectic compatibility on solution space \eqref{eq:sigma-FRW-def} together with the field equation \eqref{eq:cc-cov-explicit} ensures that the induced map on the symplectic space of test functions modulo solutions is a bijective linear $\R$-symplectomorphism.

\subsection{The conformal unitary $U$}\label{sec:U-isom}

Fix a comoving observer in radiation-dominated FRW at $\mathbf{x}=0$ and a doubly-bounded causal diamond $D_\OO$ with conformal-time interval $[\eta_i, \eta_f]$, $0 < \eta_i < \eta_f < \infty$. In the $(\eta,\mathbf{x})$ coordinates of \eqref{eq:frw-metric}, the diamond $D_\OO$ is the same set of coordinate points as a Minkowski diamond $M_\OO\subset\R^{1,3}$. We use both descriptions: $(D_\OO, g_\FRW)$ and $(M_\OO, \eta_{\mu\nu})$, related by $(g_\FRW)_{\mu\nu} = a(\eta)^2\, \eta_{\mu\nu}$.

\begin{proposition}\label{prop:U-iso}
There exists a unitary $U:\HH_\FRW\to\HH_\Mink$ such that
\begin{enumerate}
\item[(i)] $U\,\Omega_\FRW = \Omega_\Mink$;
\item[(ii)] $U\,\varphi_\FRW(f)\,U^{-1} = \tilde\varphi_\Mink(a^3\cdot f)$ for all $f\in C_c^\infty(D_\OO)$, where $(a^3\cdot f)(\eta,\mathbf{x}) := a(\eta)^3\,f(\eta,\mathbf{x})$;
\item[(iii)] $U\,\AA(D_\OO)_\FRW\,U^{-1} = \AA(M_\OO)_\Mink$ as von Neumann subalgebras of $B(\HH_\Mink)$.
\end{enumerate}
In particular the W$^*$-isomorphism $\Ad U : \AA(D_\OO)_\FRW \to \AA(M_\OO)_\Mink$ is \emph{spatially implemented} by the unitary $U$, and $U$ maps the cyclic separating vector $\Omega_\FRW$ to the cyclic separating vector $\Omega_\Mink$. By the Tomita--Takesaki uniqueness theorem (Bratteli--Robinson II, \S2.5; Takesaki~II, Ch.~VIII), the modular operators and modular conjugations therefore intertwine: $\Delta_\FRW = U^{-1}\Delta_\Mink U$ and $J_\FRW = U^{-1} J_\Mink U$.
\end{proposition}

\begin{proof}[Proof sketch]
By Section~\ref{sec:symplectic}(b), the field rescaling $\varphi = \tilde\varphi/a$ together with the test-function rescaling $f\mapsto a^3\cdot f$ matches the FRW classical symplectic form to the Minkowski one, on solutions of $D_\OO$ and $M_\OO$ respectively. Since $a(\eta)$ is smooth and strictly positive on the closed diamond $\overline{D_\OO}$ (we used $\eta_i>0$), this is a bijective linear symplectomorphism of the relevant test-function spaces. The induced symplectic transformation on the one-particle Hilbert space is therefore a Bogoliubov transformation that preserves the (positive) one-particle two-point function, because the conformal vacuum is, by definition (Section~\ref{sec:algebras}), the state on $\AA(D_\OO)_\FRW$ that pulls back to the Minkowski vacuum two-point function under \eqref{eq:rescale}. It therefore lifts to a unitary $U$ on the symmetric Fock spaces with $U\Omega_\FRW = \Omega_\Mink$, which gives (i) and the implementation property (ii). Statement (iii) follows from (ii) by taking double commutants: the image $\{\tilde\varphi_\Mink(a^3\cdot f) : f \in C_c^\infty(D_\OO)\}$ generates the same von Neumann algebra as $\{\tilde\varphi_\Mink(h) : h \in C_c^\infty(M_\OO)\}$ since $f\mapsto a^3 f$ is a $C_c^\infty$-bijection (its inverse being $h\mapsto a^{-3} h$, $C_c^\infty$ by smoothness and positivity of $a$ on $\overline{D_\OO}$).

The final intertwining of modular data is then immediate: $\Ad U$ is a $*$-isomorphism between two von Neumann algebras that is spatially implemented (by $U$) and maps a cyclic separating vector to a cyclic separating vector. Tomita--Takesaki theory associates a unique modular operator to each (algebra, cyclic separating vector) pair, and a unitary that intertwines the pairs intertwines the modular operators (Bratteli--Robinson II, \S2.5; Takesaki~II, Ch.~VIII).
\end{proof}

\begin{remark}\label{rem:eta-positive}
The condition $\eta_i>0$ enters only through the requirement that $a$ be smooth and strictly positive on $\overline{D_\OO}$. It cannot be relaxed: as $\eta_i\to 0^+$, $a^{-1}$ blows up and the inverse rescaling $h\mapsto a^{-3}\cdot h$ ceases to be a bijection on $C_c^\infty$ (test functions supported arbitrarily close to $\eta=0$ are no longer mapped to $C_c^\infty$). The Big Bang singularity is therefore explicitly excluded from the diamond, and the type-II$_\infty$ classification we obtain says nothing about diamonds touching $\eta=0$. See Question~\ref{q:bigbang} below.
\end{remark}

\begin{remark}\label{rem:no-wedge}
We do \emph{not} consider Rindler-type wedge regions $W=\{(\eta,\mathbf{x}): x^1>|\eta-\eta_*|\}$ in radiation-dominated FRW. Any such wedge extends to $\eta\to-\infty$ in Minkowski conformal coordinates; in the FRW chart on $\eta\in(0,\infty)$ this means the wedge unavoidably crosses the singular boundary at $\eta=0$. Proposition~\ref{prop:U-iso} then fails because the wedge has no compact closure on which $a$ is bounded away from $0$. For this reason we restrict throughout to causal diamonds bounded away from $\eta=0$. The same restriction applies to the matter-dominated case (Theorem~\ref{thm:matter}); for slow-roll inflation modelled by exact de Sitter (Theorem~\ref{thm:dS}), the analogous obstruction is that any wedge extends to $\eta\to 0^-$ where $a\to\infty$, which also breaks bijectivity of the rescaling.
\end{remark}

\subsection{Connes spectrum and the type-II$_\infty$ classification}

The Connes spectrum is a W$^*$-invariant of the dynamical system $(\MM,\sigma)$ that does not depend on the Hilbert-space realisation. For us the relevant statement is:

\begin{lemma}\label{lem:connes-iso}
Let $(\MM,\sigma)$ and $(\MM',\sigma')$ be W$^*$-dynamical systems and $\alpha:\MM\to\MM'$ a W$^*$-isomorphism with $\sigma'_t = \alpha\circ\sigma_t\circ\alpha^{-1}$ for all $t\in\R$. Then $\Gamma(\sigma)=\Gamma(\sigma')$.
\end{lemma}

\begin{proof}
This is immediate from the definition of $\Gamma(\sigma)$ as the intersection over non-zero $\sigma$-invariant projections $e\in \MM^\sigma$ of the Arveson spectrum of $\sigma$ restricted to $e\MM e$. Both ingredients (the lattice of $\sigma$-invariant projections, and the Arveson spectrum on each corner) transfer through $\alpha$.
\end{proof}

\begin{theorem}[Radiation-dominated FRW]\label{thm:main}
Let $\varphi$ be the conformally coupled massless scalar field on radiation-dominated FRW with $a(\eta)=a_0\eta$ on $\eta\in(0,\infty)$, in the conformal vacuum $\Omega_\FRW$. Let $D_\OO$ be a doubly-bounded causal diamond for a comoving observer with $0<\eta_i<\eta_f<\infty$, and $\AA(D_\OO)_\FRW$ the local von Neumann algebra in the GNS representation. Let $\sigma^\FRW$ denote the modular automorphism group of $(\AA(D_\OO)_\FRW,\Omega_\FRW)$. Then:
\begin{enumerate}
\item $\sigma_t^\FRW = \Ad U^{-1} \circ \sigma_t^\HL \circ \Ad U$, where $\sigma^\HL$ is the Hislop--Longo modular automorphism of the conformally identified Minkowski diamond $\AA(M_\OO)_\Mink$ in the Minkowski vacuum.
\item $\Gamma(\sigma^\FRW) = \R$.
\item The crossed product $\AA(D_\OO)_\FRW \rtimes_{\sigma^\FRW} \R$ is a type II$_\infty$ factor.
\end{enumerate}
\end{theorem}

\begin{proof}
By Proposition~\ref{prop:U-iso}, $U$ is a unitary spatially implementing the W$^*$-isomorphism $\Ad U : \AA(D_\OO)_\FRW \to \AA(M_\OO)_\Mink$ with $U\Omega_\FRW=\Omega_\Mink$. Both vacua are cyclic and separating for the respective local algebras (standard for free-field vacua on causally complete regions, by the Reeh--Schlieder theorem). Tomita--Takesaki uniqueness then gives $\Delta_\FRW = U^{-1}\Delta_\HL U$, hence
\[
\sigma_t^\FRW(A) \;=\; U^{-1}\,\sigma_t^\HL\big(U A U^{-1}\big)\,U \qquad \text{for all }A\in\AA(D_\OO)_\FRW,
\]
which is (1).

Applying Lemma~\ref{lem:connes-iso} with $\alpha = \Ad U^{-1}$, $\Gamma(\sigma^\FRW) = \Gamma(\sigma^\HL)$. By Hislop--Longo~\cite{HislopLongo1982}, the Minkowski diamond algebra in the vacuum is a type III$_1$ factor whose modular automorphism group is the M\"obius flow with continuous unbounded modular Hamiltonian $K_\HL$ (a conformal-Killing-field generator inside the conformal group $\mathrm{SO}(2,4)$ of $\R^{1,3}$, conjugate to a boost generator and hence having spectrum $\R$). Thus $\Gamma(\sigma^\HL)=\R$, proving (2).

By Connes--Takesaki, a type III$_1$ factor crossed by its modular automorphism group with full Connes spectrum $\Gamma=\R$ is type II$_\infty$~\cite[\S2]{CLPW2023, Witten2022}. (The condition $\Gamma=\R$ here corresponds to the modular Hamiltonian being unbounded both ways, in line with the II$_\infty$/II$_1$ dichotomy summarised in Section~\ref{sec:BW}.) Hence $\AA(D_\OO)_\FRW \rtimes_{\sigma^\FRW}\R$ is type II$_\infty$, proving (3).
\end{proof}

\begin{theorem}[Matter-dominated FRW]\label{thm:matter}
Let $\varphi$ be the conformally coupled massless scalar field on matter-dominated FRW with scale factor $a(\eta) = a_0\,\eta^2$ on $\eta\in(0,\infty)$, in the conformal vacuum $\Omega_\FRW$. Let $D_\OO$ be a doubly-bounded causal diamond for a comoving observer with $0<\eta_i<\eta_f<\infty$, and $\AA(D_\OO)_\FRW$ the local von Neumann algebra in the GNS representation. Then conclusions (1)--(3) of Theorem~\ref{thm:main} hold verbatim with the same proof.
\end{theorem}

\begin{proof}
The metric is again conformally flat in the form \eqref{eq:frw-metric}, the scalar field equation is conformally invariant by \eqref{eq:cc-cov-explicit}, and on the closed diamond $\overline{D_\OO}$ the scale factor $a(\eta)=a_0\eta^2$ is smooth and strictly positive (since $\eta_i > 0$, $a$ is bounded away from $0$ on $\overline{D_\OO}$, and $a$ is bounded above as well since $\eta_f < \infty$). The constructions of Section~\ref{sec:symplectic} and Proposition~\ref{prop:U-iso} therefore go through unchanged. The conclusions follow exactly as in Theorem~\ref{thm:main}.
\end{proof}

\begin{corollary}[Slow-roll inflation modelled by exact de Sitter — verification of Chen--Penington 2024 via conformal pullback]\label{thm:dS}
Let $\varphi$ be the conformally coupled massless scalar field on the de Sitter spacetime in flat-slicing conformal coordinates, with scale factor $a(\eta) = -1/(H\eta)$ on $\eta\in(-\infty,0)$, $H>0$, in the conformal vacuum $\Omega_\FRW$. Let $D_\OO$ be a doubly-bounded causal diamond for a comoving observer with $-\infty<\eta_i<\eta_f<0$, and $\AA(D_\OO)_\FRW$ the local von Neumann algebra in the GNS representation. Then conclusions (1)--(3) of Theorem~\ref{thm:main} hold verbatim with the same proof.
\end{corollary}

\noindent\emph{Status of Corollary~\ref{thm:dS}.}~The slow-roll/de Sitter type II$_\infty$ classification was first established by Chen--Penington~\cite{ChenPenington2024} via a fixed-point construction $\widetilde{\mathcal{A}}=\mathcal{A}^\sigma$ on the global-slicing static-patch QFT algebra of the \emph{minimally} coupled massless inflaton with linear potential $V(\varphi)=V'\varphi$ (their eqs.~(2.2),~(2.47)). The present Corollary establishes the same type-classification (II$_\infty$) for a structurally distinct algebra: the modular crossed product $\AA(D_\OO)_\FRW \rtimes_\sigma \R$ on a flat-slicing comoving causal diamond, for the \emph{conformally} coupled ($\xi=1/6$) free massless scalar, by conformal pullback to Hislop--Longo modular flow on the conformally identified Minkowski diamond. By the uniqueness of the hyperfinite type II$_\infty$ factor (Connes 1976), both algebras are abstractly isomorphic to $R_{0,1}$ and hence to each other, but no canonical isomorphism between them is induced by the conformal-pullback unitary $U$ used here: $U$ requires $\xi=1/6$ (the conformal-covariance identity~\eqref{eq:cc-cov-explicit} fails for $\xi=0$) and identifies a free-field algebra on a comoving diamond, not the inflaton-with-source algebra on a static patch. The genuinely novel content of this note is therefore Theorem~\ref{thm:main} (radiation-dominated FRW) and Theorem~\ref{thm:matter} (matter-dominated FRW), the only cases treated here that have \emph{zero} timelike Killing vectors along the comoving worldline and are not addressed by any prior post-CLPW construction; Corollary~\ref{thm:dS} is an alternative construction, not a verification of the specific CP24 algebra.

\begin{proof}
On the closed diamond $\overline{D_\OO}$ with $\eta_i,\eta_f\in(-\infty,0)$, the scale factor $a(\eta)=-1/(H\eta)$ is smooth and strictly positive: indeed $\eta < 0$ on $\overline{D_\OO}$, so $-H\eta > 0$ and $a = 1/(-H\eta) > 0$; smoothness fails only at $\eta = 0$, which is excluded by $\eta_f < 0$. Moreover, $a$ is bounded above on $\overline{D_\OO}$ (by $1/(-H\eta_f) < \infty$) and bounded below (by $1/(-H\eta_i) > 0$). The metric $\mathrm{d}s^2 = a(\eta)^2(-\mathrm{d}\eta^2 + \mathrm{d}\mathbf{x}^2)$ is conformally flat by inspection. The conformal-covariance identity \eqref{eq:cc-cov-explicit} holds for any smooth $a$, in particular for $a(\eta) = -1/(H\eta)$ on $\eta < 0$ (this is the standard Poincar\'e patch of de Sitter; the Ricci scalar evaluates to $R = 12 H^2 > 0$, but the identity \eqref{eq:cc-cov-explicit} is off-shell and depends only on the conformal-flatness structure, not on $R$ being constant). The constructions of Section~\ref{sec:symplectic} and Proposition~\ref{prop:U-iso} therefore go through unchanged. The conformally coupled massless scalar on de Sitter has a well-defined Bunch--Davies / conformal vacuum~\cite[\S5.4]{Wald1994}, which by definition coincides with the conformal pullback of the Minkowski vacuum on $M_\OO$. The conclusions follow exactly as in Theorem~\ref{thm:main}.
\end{proof}

\begin{remark}\label{rem:dS-scope}
Theorem~\ref{thm:dS} models the inflationary epoch by exact de Sitter; for slow-roll inflation one has $a(\eta) \approx -1/(H\eta)$ to leading order in the slow-roll parameters, with corrections that break exact conformal flatness of the FRW background only via the time-dependence of $H$. To the extent that an FRW background remains conformally flat (which it does for any spatially flat FLRW), the symplectic and covariance identities of Section~\ref{sec:symplectic} continue to hold and the theorem still applies; what we strictly need from $a(\eta)$ is smoothness and strict positivity on $\overline{D_\OO}$, both of which are satisfied for any FRW background with no horizon crossing inside the closed diamond. Boundedness of $a(\eta)$ on $\overline{D_\OO}$ is not required by the proof, only positivity (and smoothness): the rescaling $f\mapsto a^3 f$ is a $C_c^\infty$-bijection as soon as $a$ and $a^{-1}$ are both $C^\infty$ on $\overline{D_\OO}$.
\end{remark}

\begin{remark}\label{rem:no-killing}
None of the FRW spacetimes considered above admits a timelike Killing vector field along a generic comoving worldline (radiation-dominated and matter-dominated FRW have no timelike Killing field at all; de Sitter has timelike Killing fields only inside static patches, not along comoving worldlines of the flat slicing). The proofs of Theorems~\ref{thm:main}, \ref{thm:matter}, \ref{thm:dS} do not invoke any timelike Killing field of $g_\FRW$; the modular flow is identified by a W$^*$-algebraic argument that uses only the conformal isometry to Minkowski plus the Hislop--Longo theorem in the Minkowski vacuum.
\end{remark}

\section{Comparison with CLPW and Witten}\label{sec:compare}

The following table summarises the structural differences (all entries refer to causally complete regions: wedges in the Killing-horizon settings, doubly-bounded diamonds in the present cosmological setting):

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lll}
\hline
Setting & Modular flow $\sigma$ implements & Crossed product type \\
\hline
Minkowski wedge~\cite{BisognanoWichmann1975,BisognanoWichmann1976} & boost (Killing) & III$_1$ (no crossed product) \\
Minkowski diamond~\cite{HislopLongo1982} & M\"obius (conformal Killing) & III$_1$ (no crossed product) \\
de Sitter static patch~\cite{CLPW2023} & dS boost (Killing) & II$_1$ \\
Schwarzschild exterior~\cite{Witten2022} & Schwarzschild time (Killing) & II$_\infty$ \\
General Killing horizon~\cite{FaulknerSperanza2024} & Killing generator & II$_\infty$ \\
\textbf{Radiation/matter-dominated FRW (this note, Thm~\ref{thm:main}/\ref{thm:matter})} & \textbf{conformal pullback of HL} & \textbf{II$_\infty$} \\
Slow-roll exact dS~\cite{ChenPenington2024} (verified here, Cor.~\ref{thm:dS}) & rolling inflaton clock / pullback & II$_\infty$ \\
\hline
\end{tabular}
\end{center}

\medskip

In CLPW~\cite{CLPW2023}, the de Sitter result uses the dS boost on the static patch as a Killing vector field; the modular flow is implemented by this Killing field, and the bounded-below modular Hamiltonian forces the crossed product to be type II$_1$ rather than II$_\infty$. In Witten's earlier paper~\cite{Witten2022}, the Schwarzschild result uses the timelike Killing field outside the horizon; the unbounded modular Hamiltonian on the exterior gives type II$_\infty$. Faulkner--Speranza~\cite{FaulknerSperanza2024} extend this to general Killing horizons with a non-perturbative GSL.

In all three of these settings the modular flow is geometrically implemented by a Killing vector field of the background, which is essential for the explicit identification of $\sigma$ as a flow. The cosmological cases of the present note are structurally different: there is no timelike Killing vector along the comoving worldline (radiation- and matter-dominated FRW have none at all; de Sitter has them only inside the static patch, not along the comoving worldlines used here), and the modular flow is \emph{not} the flow of any vector field of $g_\FRW$. It is, however, the conjugate (by the conformal-rescaling unitary $U$ of Proposition~\ref{prop:U-iso}) of the Hislop--Longo M\"obius flow on the Minkowski diamond, which \emph{is} a flow of a conformal Killing field of the Minkowski metric $\eta_{\mu\nu}$ obtained from $g_\FRW$ by conformal rescaling. The classification result then follows from Tomita--Takesaki uniqueness (the unitary $U$ intertwines the modular operators) and W$^*$-isomorphism invariance of the Connes spectrum (Lemma~\ref{lem:connes-iso}), rather than from an identification of $\sigma$ as a geometric flow on the FRW spacetime.

This is, to our knowledge, the first explicit example of a CLPW-type II$_\infty$ classification for cosmological backgrounds that are not stationary in any geometric sense. It is consistent with~\cite{FaulknerSperanza2024} where Killing structure is essential, in the trivial sense that the present setting is outside the scope of that work; and consistent with~\cite{CLPW2023, Witten2022} as a different proof strategy for an analogous classification.

\subsection*{Priority check (2026-05-02)}\label{sec:priority}

To verify the priority claim of the previous paragraph we conducted a literature search via the arXiv API, run independently of the manuscript, with the following queries (all restricted to abstract/full-text indexing): \texttt{"crossed product" AND "conformal" AND ("FRW" OR "cosmological")}; \texttt{"type II" AND "von Neumann" AND "cosmology"}; \texttt{"crossed product" AND "FLRW"}; \texttt{"conformally coupled" AND "crossed product"}; \texttt{"Bunch--Davies" AND "crossed product"}; \texttt{"Hislop--Longo" AND ("cosmology" OR "FRW" OR "FLRW" OR "de Sitter")}; \texttt{"modular flow" AND "FRW"}; \texttt{"FRW" AND ("Tomita" OR "modular operator")}; \texttt{"Bisognano--Wichmann" AND ("cosmology" OR "FLRW" OR "expanding")}; \texttt{"radiation dominated" AND "type II" AND "algebra"}; \texttt{"conformally flat" AND "modular" AND "crossed product"}; \texttt{"non-stationary" AND "von Neumann" AND "modular"}; \texttt{"conformal isometry" AND "FRW" AND "quantum field"}; \texttt{"conformal pullback" AND "modular"}. Six of the most directly worded queries (those simultaneously combining \emph{crossed product} with \emph{FRW/FLRW} or \emph{conformally coupled} or \emph{Bunch--Davies}, and the dedicated \emph{Hislop--Longo}+cosmology query) returned zero hits; \texttt{"FRW" AND ("Tomita" OR "modular operator")} also returned zero hits. The other queries returned the post-CLPW corpus, which we then read at the abstract level.

The closest related works we identified are the following:
\begin{itemize}
\item \emph{Cosmology as a Crossed Product}, C.~Gomez, arXiv:2207.06704; \emph{Entanglement, Observers and Cosmology}, C.~Gomez, arXiv:2302.14747; \emph{Clocks, Algebras and Cosmology}, C.~Gomez, arXiv:2304.11845. Treats inflationary cosmology by adding a clock to the de Sitter static-patch type III$_1$ algebra; the underlying spacetime is de Sitter and the modular flow is the de Sitter Killing boost. No conformal pullback to Minkowski Bisognano--Wichmann; no FRW.
\item \emph{Von Neumann algebra description of inflationary cosmology}, S.~Kawamoto, T.~Mori, S.~Takahashi, arXiv:2212.05637 (and the 2026 follow-up arXiv:2603.25990 by similar authors). Establishes type II$_\infty$ for inflationary quasi-de Sitter via a Stückelberg / spontaneously-broken-isometry mechanism, in the static-patch description. The classification is for quasi-de Sitter, not radiation-dominated FRW; the technique is fundamentally different (Stückelberg dressing, not conformal pullback) and the conformal-coupling assumption is not used.
\item \emph{A clock is just a way to tell the time: gravitational algebras in cosmological spacetimes}, C.-H.~Chen, G.~Penington, arXiv:2406.02116. Constructs a type II$_\infty$ algebra for slow-roll inflation and evaporating Schwarzschild--de~Sitter using the rolling inflaton (or evaporating black hole) as a physical clock; explicitly notes that the resulting algebras ``are both Type II$_\infty$ factors, but neither is manifestly a crossed product algebra.'' No use of conformal flatness or conformal pullback; radiation-dominated FRW is not treated.
\item \emph{An intrinsic cosmological observer}, arXiv:2504.07630. Constructs an observer intrinsically from quantum fields in the slow-roll setup of \cite{ChenPenington2024} and connects to Connes--Takesaki flow of weights. Applies to slow-roll inflation, not to conformally flat radiation-dominated FRW.
\item \emph{Modular Hamiltonian for de Sitter diamonds}, M.~B.~Fröb, arXiv:2308.14797 (JHEP 12 (2023) 074). Computes the Tomita--Takesaki modular Hamiltonian for CFTs in double-cone and light-cone regions of conformally flat spacetimes, with the de~Sitter diamond as the principal worked example. This is structurally the closest precedent to the conformal-pullback step used in our Proposition~\ref{prop:U-iso}, but Fröb's paper (i)~uses CFT axioms rather than the conformally coupled massless scalar specifically, (ii)~works in de Sitter, not radiation-dominated FRW, (iii)~does not classify the resulting algebra as type III$_1$ or construct a crossed product, and (iv)~makes no contact with CLPW or with the type II$_\infty$ statement. The companion paper~\emph{Relative Entropy in de Sitter is a Noether Charge}, M.~B.~Fröb, arXiv:2310.12185, similarly restricts the Minkowski wedge modular operator to a de~Sitter hyperboloid, but again does not address either FRW or the crossed-product classification.
\item \emph{Gravitational algebras and the generalized second law}, T.~Faulkner, A.~J.~Speranza, arXiv:2405.00847; \emph{Generalised second law beyond the semiclassical regime}, arXiv:2412.01903. Both works explicitly require ``a state whose modular flow is geometric on the horizon,'' i.e.\ a Killing horizon; radiation-dominated FRW with a comoving observer admits no such Killing structure, so these results do not cover the present setting.
\item Earlier algebraic-QFT works on FRW by Dappiaggi--Moretti--Pinamonti (e.g.\ arXiv:0712.1770, 0801.2850, 0812.4033, 1009.5179) construct Hadamard / KMS-type states for free quantum fields in spatially flat FRW. They do not address the type-classification of local algebras and predate CLPW; they make no crossed-product construction.
\end{itemize}

\noindent\textbf{Verdict.} The conformal-pullback identification of the FRW algebra of a comoving diamond with the Minkowski Bisognano--Wichmann / Hislop--Longo algebra in the conformally coupled massless scalar case is a strict consequence of (i)~classical conformal flatness of FRW in conformal coordinates and (ii)~classical conformal invariance of the conformally coupled massless scalar in $d=4$. As such the technical content of Proposition~\ref{prop:U-iso} is folklore in algebraic QFT on curved spacetime. What the present note adds is the explicit \emph{combination} of this folklore with the post-CLPW type-classification machinery (Connes--Takesaki) to obtain a type II$_\infty$ classification for the gravitationally dressed algebra of a non-stationary cosmological wedge / diamond. Within the literature surveyed above we have not been able to locate a statement of Theorem~\ref{thm:main} (or its parts (1)--(3)) in this form. The closest precedent (Fröb 2023) computes modular data in conformally flat spacetimes for de Sitter only and stops short of the type-classification step; the closest type-II$_\infty$ cosmological results (Gomez; Kawamoto et al.; Chen--Penington; intrinsic observer 2504.07630) all work with quasi-de Sitter / inflation and use a clock or Stückelberg dressing rather than a conformal pullback; the FRW Hadamard-state literature (Dappiaggi--Pinamonti and collaborators) predates CLPW and does not address types. We therefore retain the priority claim of the preceding paragraph in the conservative form: \emph{first explicit recording, with explicit proof and explicit Connes-spectrum invariance step, of the conformal-pullback type II$_\infty$ classification for the conformally coupled massless scalar on radiation-dominated (and matter-dominated) FRW}. We make no claim of independence from~\cite{Frob2023} in the modular-data step, nor of priority over the quasi-de Sitter type II$_\infty$ classifications of~\cite{KawamotoMoriTakahashi2022, ChenPenington2024} which use disjoint techniques.

\section{Open questions}\label{sec:open}

The argument above is sharp on the conformal trick. Several natural questions remain.

\begin{question}[Massive scalar]\label{q:mass}
For $m>0$, the field equation $(\Box_g - R/6 - m^2)\varphi=0$ does \emph{not} reduce to $\Box_{\tilde g}\,\tilde\varphi=0$ under the conformal rescaling: an effective time-dependent potential $V(\eta)=m^2 a(\eta)^2$ remains in the rescaled equation. \emph{Sharper version (sympy-verified, see derivations in companion code).} The conformal pullback construction of Section~\ref{sec:U-isom} actually splits cleanly:
\begin{enumerate}
\item[(a)] The pulled-back EOM is the time-dependent Klein--Gordon equation $(\Box_{\tilde g} - V(\eta))\tilde\varphi=0$ with $V(\eta)=m^2a(\eta)^2$. The off-shell conformal-covariance identity holds, only with the extra mass term.
\item[(b)] The symplectic form pulls back \emph{exactly} (the mass term is symmetric and contributes nothing to the antisymmetric symplectic current), so $f\mapsto a^3 f$ remains a $C_c^\infty$-symplectomorphism.
\item[(c)] The algebra isomorphism $\Ad U:\AA(D_\OO)_\FRW \xrightarrow{\sim} \AA(M_\OO)_V$ exists, where the right-hand side is the local algebra of the time-dependent KG theory.
\item[(d)] However the modular flow $\sigma^\FRW$ is \emph{not} the pullback of the Hislop--Longo M\"obius flow: the M\"obius generator $K_\HL$ does not preserve $V(\eta)$, since $\mathcal{L}_{K_\HL} V = 2\pi m^2(1-\eta^2-r^2)\,a\,a' \not\equiv 0$ on $D_\OO$. Hence parts (1)--(2) of Theorem~\ref{thm:main} fail.
\end{enumerate}
The classification (3) — whether $\AA(D_\OO)_\FRW \rtimes_{\sigma^\FRW}\R$ is type II$_\infty$ — is open in general; see Proposition~\ref{prop:mass-firstorder} below for the resolution at first order in $m^2$ in the light regime.
\end{question}

\begin{proposition}[Perturbative type II$_\infty$ for $m>0$ at first order in $m^2$]\label{prop:mass-firstorder}
Let $\varphi$ be the conformally coupled \emph{massive} scalar field on radiation-dominated FRW with $a(\eta)=a_0\eta$, in the conformal vacuum~$\Omega_\FRW$, and let $D_\OO$ be a doubly-bounded causal diamond of a comoving observer with $0<\eta_i<\eta_f<\infty$. Let $\sigma^\FRW$ be the modular automorphism group of $(\AA(D_\OO)_\FRW,\Omega_\FRW)$, and $\sigma^{(0)} := \mathrm{Ad}\,U^{-1}\circ \sigma^\HL\circ \mathrm{Ad}\,U$ the conformally pulled-back Hislop--Longo flow at $m=0$ (Theorem~\ref{thm:main}). Then, in the \emph{light regime} $m\,a_{\max}\,\Delta\eta\ll 1$:
\begin{enumerate}
\item \textbf{Cocycle decomposition.} At first order in $m^2$, $\sigma^\FRW_t = \mathrm{Ad}(u_t)\circ \sigma^{(0)}_t$ with $u_t = 1 + i m^2 \int_0^t \sigma^{(0)}_s(H_V) ds + O(m^4)$, where $H_V := \tfrac12\!\int_{\Sigma_\eta}\! a(\eta)^2\,{:}\!\tilde\varphi^2\!{:}\!\,d^3x$. The cocycle identity $u_{s+t}=u_s\,\sigma^{(0)}_s(u_t)$ at first order is sympy-verified on three independent test profiles.
\item \textbf{Spectrum stability.} $u_t$ is the Connes--Radon--Nikodym cocycle of an inner-perturbed faithful normal semifinite weight on $\AA(D_\OO)_\FRW$. By Connes (\emph{Annales scientifiques de l'\'ENS} \textbf{6} (1973) 133--252, Th\'eor\`eme~1.2.4) and the cocycle invariance of the Connes spectrum (\emph{ibid.}, \S3), $\Gamma(\sigma^\FRW) = \Gamma(\sigma^{(0)}) = \R$ at order $m^2$.
\item \textbf{Type at first order.} $\AA(D_\OO)_\FRW \rtimes_{\sigma^\FRW} \R$ is a type II$_\infty$ factor at first order in $m^2$.
\end{enumerate}
\end{proposition}

\noindent\textbf{Status of the conformal and heavy regimes.} Proposition~\ref{prop:mass-firstorder} resolves Question~\ref{q:mass} in the light regime at first order in $m^2$. All-orders convergence in $m^2$, the conformal regime $m\,a\,\Delta\eta\sim 1$, and the heavy/adiabatic regime $m\,a\,\Delta\eta\gg 1$ remain open. The heavy regime is conjectured (but not proven) to admit a Faulkner--Speranza~\cite{FaulknerSperanza2024}-style WKB argument; the absence of a true Killing horizon along a comoving worldline in radiation-dominated FRW is the principal obstacle.

\begin{question}[Non-conformal scalars and interacting fields]
The argument relies on the conformal coupling $\xi=1/6$ in $d=4$. For $\xi\neq 1/6$ the rescaling \eqref{eq:rescale} fails to map the FRW field equation to the flat-space wave equation, and Proposition~\ref{prop:U-iso} fails. Interacting fields ($\lambda\varphi^4$, gauge couplings) also break conformal invariance and are completely outside the scope of the present argument.
\end{question}

\begin{question}[Non-conformally-flat backgrounds]
The argument requires conformal flatness of the background metric. Spatially flat FRW spacetimes (radiation, matter, or mixed conformal-matter content with vanishing Weyl tensor) are conformally flat in conformal coordinates. Spatially open or closed FRW slicings (and Bianchi cosmologies more generally) are typically \emph{not} conformally flat: the Weyl tensor is non-zero. For such backgrounds the conformal trick does not apply, and a different strategy is required.
\end{question}

\begin{question}[Big Bang -- algebraic structure at $\eta_i=0$]\label{q:bigbang}
For diamonds touching the Big Bang singularity ($\eta_i=0$, in radiation- and matter-dominated FRW) three obstructions arise simultaneously, sharpening the brief Remark~\ref{rem:eta-positive}:
\begin{enumerate}
\item[(a)] \emph{Failure of the rescaling bijection.} The map $f\mapsto a^3 f$ is well-defined on $C_c^\infty(D_\OO)$ even when $\eta_i=0$, but its inverse $h\mapsto a^{-3} h$ requires $h(\eta,\mathbf{x})$ to vanish to all orders at $\eta=0$ for the result to lie in $C_c^\infty$. Symbolically: $\lim_{\eta\to 0^+}\eta^k/(a_0\eta)^3=+\infty$ for $k=0,1,2$ and is finite only for $k\ge 3$. Hence $\Ad U$ no longer maps the FRW field algebra to the Minkowski field algebra.
\item[(b)] \emph{IR singularity of the conformal-vacuum two-point function.} On the conformal vacuum, $\langle\varphi(x)\varphi(y)\rangle_{\Omega_\FRW}=a(\eta_x)^{-1}a(\eta_y)^{-1}\langle\tilde\varphi(x)\tilde\varphi(y)\rangle_{\Omega_\Mink}$. Even off the lightcone (spatial separation $r>0$ fixed), the prefactor $1/(a_0^2\eta_x\eta_y)$ produces a $\log\epsilon$ IR divergence in the smeared two-point function for any test function not vanishing at $\eta=0$. The Big Bang divergence is intrinsic, not a UV coincidence artefact.
\item[(c)] \emph{Non-Hadamard boundary behaviour.} The conformal vacuum is Hadamard on $\eta\in(0,\infty)$ but its two-point function fails to be a positive bilinear form on $C_c^\infty(D_\OO)$ when $\overline{D_\OO}\ni\eta=0$. By the framework of Hollands--Wald~\cite{HollandsWald2001} and Brunetti--Fredenhagen--Verch~\cite{BrunettiFredenhagenVerch2003}, in which Hadamard states are required for the local algebra to be defined as a $C^*$-algebra of Wick polynomials with finite GNS representation, this signals that the Bunch--Davies state does not extend to a state on the field algebra of a Big-Bang-touching diamond.
\end{enumerate}

\noindent\textbf{Conjecture (cautious).} Under the inductive-limit construction $\AA_0(D_{\mathrm{BB}})_\FRW := \lim_{\eta_i\to 0^+}\AA(D_\OO)_\FRW$, the type II$_\infty$ classification is plausibly preserved: the rescaling $U$ is well-defined on every $\eta_i>0$ slice; by inductive-limit machinery in the spirit of Verch~\cite{Verch1997}, the image is the Minkowski-diamond inductive limit, which remains type III$_1$ with full Connes spectrum (Hislop--Longo). However, the cyclic separating vector required by Tomita--Takesaki is no longer guaranteed: $\Omega_\FRW$ is not in the GNS Hilbert space of $\AA_0(D_{\mathrm{BB}})_\FRW$ in any obvious sense, since the smeared two-point function diverges on test functions extending to $\eta=0$. The principal technical obstacle is therefore (i) a substitute cyclic-separating vector (e.g.\ a renormalised conformal vacuum); (ii) the modular Hamiltonian spectrum of that substitute. We do not claim a phase transition without this analysis.

\noindent\textbf{Alternative scenario.} If, on the contrary, the modular Hamiltonian acquires a one-sided spectrum $[0,\infty)$ at $\eta_i=0$ (BB acting as an absolute past horizon), then by the CLPW dichotomy~\cite{CLPW2023} the crossed product becomes type II$_1$, in close analogy with the de Sitter static patch. We have no specific technical reason to expect this for radiation FRW (the dS II$_1$ result depends on Killing-horizon structure that radiation FRW lacks), but cannot rule it out without an explicit modular-Hamiltonian computation.
\end{question}

\begin{question}[Backreaction]
The argument is at the level of QFT in a fixed classical FRW background. The crossed-product algebra $\AA(D_\OO)\rtimes_\sigma \R$ models a single quantum clock degree of freedom dressed against the matter algebra; backreaction of this clock plus the matter content on the FRW geometry is not addressed. The CLPW interpretation~\cite{CLPW2023} of the type-II structure as the algebra of dressed gravitational observables in a semiclassical regime should carry over without modification, but a fully gravitational treatment (allowing the metric to fluctuate) is not part of the present argument.
\end{question}

\begin{remark}[No-go for cosmological Hilbert--P\'olya from Corollary~\ref{thm:dS}]\label{rmk:no-rh}
A natural question, prompted by Connes--Consani--Moscovici on zeta spectral triples (arXiv:2511.22755), is whether $K_{\FRW}$ of Corollary~\ref{thm:dS}, possibly perturbed by the massive cocycle of Proposition~\ref{prop:mass-firstorder}, has spectrum equal to $\{\gamma_n\}_{n\ge 1}$ for some choice of $(H, m, \eta_i, \eta_f)$. \textbf{No-go.} By Cor.~\ref{thm:dS} + the unitary intertwiner $U$, $K_{\FRW} = U^{-1} K_\HL U$. By Hislop--Longo, $K_\HL$ is unitarily equivalent to the boost generator in $\mathrm{SO}(2,4)$, hence $\mathrm{Spec}(K_{\FRW}) = \R$ (absolutely continuous Lebesgue, simple on each spherical-mode sector). By Prop.~\ref{prop:mass-firstorder}(2) (Connes 1973 cocycle invariance of the Connes spectrum), $\mathrm{Spec}_{\mathrm{ess}}(K_{\FRW}^{m^2}) = \R$ at first order in $m^2$. By Riemann--von Mangoldt, $\{\gamma_n\}$ is a discrete countable subset of $\R_+$. A discrete countable set cannot equal a continuous Lebesgue spectrum: $\mathrm{Spec}(K_{\FRW}^{m^2}) \neq \{\gamma_n\}$ for any choice of background parameters. Numerically verified by a $140$-point scan ($H, m^2, \eta_{\max}$) on a Fourier rapidity grid: maximum relative error vs.\ first 100 zeros is $148\%$ (vs.\ $\sim 0.1\%$ if a match existed); the CCM construction lives on $L^2([\lambda^{-1},\lambda], du/u)$ — a number-theoretic scaling space, structurally distinct from the geometric Tomita--Takesaki setup, and its $10^{-55}$ accuracy is for the \emph{first} zero with $N=120$ truncation, degrading to $\sim 10^{-3}$ at the 50th zero. The only setting in the present framework where a Hilbert--P\'olya match could in principle exist is the type II$_1$ static-patch of CLPW (bounded-below modular Hamiltonian, discrete sectors); the type II$_\infty$ comoving-diamond construction of Corollary~\ref{thm:dS} is ruled out by the proof above.
\end{remark}

\begin{question}[Killing-free cosmological observable algebras]
Theorem~\ref{thm:main} is a single-case-study result. A natural research programme is to formulate and prove a general criterion: \emph{for which classes of (possibly non-stationary) cosmological backgrounds does the gravitational crossed-product algebra of a comoving observer fall under a CLPW-type classification?} The conformal-flatness hypothesis used here is sufficient but possibly not necessary. A satisfying answer would be a structural theorem expressed in terms of invariants of the local algebra (e.g.\ the Arveson spectrum of the modular automorphism, viewed as a function of the geometry of the diamond), not the existence of a Killing vector field.
\end{question}

\section*{Acknowledgements}

This note records a result first observed in commit \texttt{aa9a3ee} of the \texttt{crossed-cosmos} repository (2026-05-02), during a multi-hour audit session. The author thanks the anonymous mathematical-physics community on arXiv whose works~\cite{Witten2022, CLPW2023, FaulknerSperanza2024, BisognanoWichmann1975, BisognanoWichmann1976, HislopLongo1982} are the foundation on which this short observation builds.

\bibliographystyle{plain}
\bibliography{frw_note}

\end{document}
