% note.tex --- F3 wedge reconstruction draft note
% Wedge reconstruction on FRW past-light-cone causal diamonds:
% modular flow, obstructions, and a conditional theorem.
%
% Author: [ECI programme, generated 2026-05-03]
% Compile: pdflatex note ; bibtex note ; pdflatex note ; pdflatex note
%
% CITATION AUDIT NOTES:
%   * Faulkner-Lewkowycz = arXiv:1704.05464 (NOT 1704.05732, which is unrelated
%     random-hypergraph combinatorics)
%   * Almheiri-Dong-Harlow = arXiv:1411.7041 (JHEP 1504:163,2015) -- VERIFIED
%   * Dong-Harlow-Wall = arXiv:1601.05416 (PRL 117:021601,2016) -- VERIFIED
%   * Witten 2022 = arXiv:2112.12828 -- VERIFIED
%   * CLPW 2022 = arXiv:2206.10780 -- VERIFIED
%   * Faulkner-Speranza 2024 = arXiv:2405.00847 -- VERIFIED
%   * Chandrasekaran-Speranza 2020 = arXiv:2009.10739 -- VERIFIED
%   * Frob 2023 = arXiv:2308.14797 -- VERIFIED
%   * Casini-Huerta-Myers 2011 = arXiv:1102.0440 -- VERIFIED
%   * Mukohyama-Speranza arXiv:2402.10362 is NOT a horizons paper;
%     that identifier belongs to a quantum simulation paper (Hejazi et al.).
%     The Mukohyama-Speranza paper (stretched horizons) was NOT found on arXiv
%     under that ID -- treated as unverified and EXCLUDED from references.

\documentclass[11pt,a4paper]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[numbers,sort&compress]{natbib}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Open Question}

% Macros
\newcommand{\FRW}{{\mathrm{FRW}}}
\newcommand{\HL}{{\mathrm{HL}}}
\newcommand{\Mink}{{\mathrm{M}}}
\newcommand{\dS}{{\mathrm{dS}}}
\newcommand{\EW}{{\mathrm{EW}}}
\newcommand{\RT}{{\mathrm{RT}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\Sgen}{S_{\mathrm{gen}}}
\newcommand{\KFRW}{K_{\FRW}}
\newcommand{\aeta}{a(\eta)}
\newcommand{\aetac}{a(\eta_c)}
\newcommand{\DR}{D_R(\eta_c,x_c)}
\newcommand{\sigmat}{\sigma_t^{\FRW}}
\newcommand{\sigmats}{\sigma_s^{\FRW}}

\title{Wedge reconstruction on FRW past-light-cone causal diamonds:\\
       modular flow, obstructions, and a conditional theorem}

\author{[ECI programme --- draft note, 2026-05-03]}
\date{2026-05-03}

\begin{document}
\maketitle

\begin{abstract}
We formulate the analog of AdS entanglement-wedge reconstruction for
FRW comoving causal diamonds $\DR \subset \mathcal{M}_{\FRW}$.
The programme has three natural inputs: the closed-form modular Hamiltonian
$\KFRW$ of the companion carve-out note (Hislop--Longo pullback via the
Frob~\cite{Frob2023} conformal intertwiner), the Faulkner--Lewkowycz (FL)
modular-flow prescription for bulk reconstruction
\cite{FaulknerLewkowycz2017}, and the Witten~\cite{Witten2022}
type~II$_\infty$ crossed-product algebra on which $\Sgen[\DR]$ is
well-defined.
We establish the FRW modular flow explicitly:
$\sigmats[\phi_{\FRW}](\eta,x) = [a(\eta_s)/a(\eta)]\,\phi_{\FRW}(\eta_s,x_s)$,
where $(\eta_s,x_s)$ is the Minkowski M\"obius image and $a(\eta_s)/a(\eta)$
is the FRW conformal-time-dependent rescaling (Sympy-verified).
We then identify four obstructions to a literal wedge-reconstruction
theorem: (O1)~absence of a conformal boundary, (O2)~type-II$_\infty$ vs.\
type-I tensor-factorisation mismatch, (O3)~Hadamard-state conformal anomaly,
(O4)~absence of a QFT Markov property on a non-Killing boundary.
Under the Hadamard hypothesis (H1) and a QFT Markov-property hypothesis (H2)
for the FRW past-light-cone, we state a conditional theorem in which bulk
fields in the entanglement wedge of $\DR$ are reconstructible from modular
flow data on $\partial\DR$ via an FL-type kernel that includes the
$a(\eta_s)/a(\eta)$ conformal rescaling.
We show that (O1) and (O4) are the binding obstructions for a literal
(unconditional) reconstruction theorem, while (O2) and (O3) are
state-dependent and potentially fixable.  The note closes with a gap
analysis identifying what is already in the literature vs.\ what requires
new work.
\end{abstract}

\tableofcontents

% ---------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
% ---------------------------------------------------------------------------

\paragraph{AdS/CFT wedge reconstruction.}
Almheiri, Dong, and Harlow (ADH)~\cite{ADH2015} showed that in AdS/CFT, bulk
operators lying in the \emph{entanglement wedge} $\EW(A)$ of a boundary
subregion~$A$ can be reconstructed as operators in the CFT subalgebra $\A(A)$.
The entanglement wedge is the bulk domain of dependence of a Cauchy slice
whose boundary is the union of $A$ and the Ryu--Takayanagi (RT)
surface~$\RT(A)$.  Dong, Harlow, and Wall~\cite{DHW2016} gave an explicit
proof using quantum error correction.  Faulkner and Lewkowycz~\cite{FaulknerLewkowycz2017}
later showed that the modular flow of the boundary state is the key
reconstruction tool: bulk operators can be recovered from the modular-evolved
boundary operators via an explicit kernel.

\paragraph{Cosmological setting.}
In cosmological spacetimes, no conformal boundary exists and the AdS/CFT
dictionary does not directly apply.  However, two complementary developments
make a wedge-reconstruction question natural in FRW:

\begin{itemize}
\item Witten~\cite{Witten2022} and
  Chandrasekaran--Longo--Penington--Witten~\cite{CLPW2022} (CLPW) showed
  that the local algebra of an observer in quantum gravity is a type~II
  factor, on which a well-defined entropy and modular flow exist.
  The ECI companion frw\_typeII\_note establishes the type~II$_\infty$
  classification for the FRW comoving-diamond algebra without any Killing
  vector~\cite{ECI_frw_note}.

\item Frob~\cite{Frob2023} proved that the modular Hamiltonian for the
  conformally-coupled massless scalar on de~Sitter and FRW diamonds is
  the conformal pullback of the Hislop--Longo (HL)~\cite{HislopLongo1982}
  Minkowski-diamond modular Hamiltonian, and that the modular flow is
  geometric (a conformal transformation) in the conformal frame.
  The ECI carve-out note~\cite{ECI_kfrw_note} exploits this to give an
  explicit closed-form $\KFRW$.
\end{itemize}

\paragraph{The question we address.}
Given a comoving causal diamond $\DR \subset \mathcal{M}_{\FRW}$ and the
localised algebra $\A_{\FRW}(\DR)$, can we reconstruct FRW bulk fields in
the entanglement wedge of $\DR$ from modular flow data on the boundary
$\partial\DR$?  This is the natural AdS/CFT analog where the role of
``boundary CFT'' is played by data on $\partial\DR$ (the past-light-cone
boundary of the diamond), and the ``bulk'' is the interior of $\DR$.

\paragraph{Scope and disclaimer.}
The ECI v6.0.25 piste ``AdS/CFT in FRW via II$_\infty$ frame'' was falsified
(R6 carve-out): $\mathcal{I}^+_{\FRW}$ is spacelike (not null), the
intertwiner $U$ maps interior to interior, and the Bunch--Davies two-point
function vanishes pointwise as $\eta \to \infty$.  The present note does NOT
attempt to resurrect that piste.  We work strictly within the interior of the
comoving diamond $\DR$, treating ``boundary'' as $\partial\DR$ (not spatial
infinity).

\paragraph{Main deliverable (conditional).}
We state a conditional Theorem~\ref{thm:frw_wedge} (FRW wedge reconstruction),
and identify which hypotheses are verifiable and which constitute genuine
obstructions.  The honest answer (see Section~\ref{sec:obstructions}) is that
obstructions O1 and O4 are currently binding.

% ---------------------------------------------------------------------------
\section{ECI setup: FRW modular Hamiltonian and type~II$_\infty$ algebra}
\label{sec:setup}
% ---------------------------------------------------------------------------

\subsection{FRW geometry and comoving diamond}

Throughout, $\mathcal{M}_{\FRW}$ is spatially flat with metric
$g_{\FRW} = a(\eta)^2(-d\eta^2 + \delta_{ij}dx^i dx^j)$, $\eta\in(0,\infty)$.
The comoving causal diamond of size $R$ centered at $(\eta_c, x_c)$ is
\begin{equation}
  \DR = \bigl\{(\eta,x)\in\mathcal{M}_{\FRW} :
    |\eta-\eta_c| + |x-x_c| \le R,\; \eta \ge \eta_c - R\bigr\}.
\label{eq:diamond}
\end{equation}
We restrict to the conformally-coupled massless scalar $\phi_{\FRW}$ with
$\xi=1/6$ in $d=4$, and work in the conformal (Bunch--Davies) vacuum
$\Omega_{\FRW}$.

\subsection{Type II$_\infty$ algebra (ECI Theorem 3.5/3.6)}

By Frob~\cite{Frob2023} and ECI frw\_typeII\_note~\cite{ECI_frw_note}
(Theorems~3.5/3.6), the Weyl-algebra isomorphism
$\A_{\FRW}(\DR) \cong \A_{\Mink}(M_\DR)$ (where $M_\DR$ is the
conformally identified Minkowski diamond) holds as a spatial isomorphism
of von~Neumann algebras.  Consequently:
\begin{itemize}
\item $\A_{\FRW}(\DR)$ is a type~III$_1$ factor (Hislop--Longo via intertwiner $U$).
\item The crossed product $\N_{\FRW}(\DR) = \A_{\FRW}(\DR)\rtimes_\sigma \R$ is a
  type~II$_\infty$ factor, on which the Witten~\cite{Witten2022} trace
  $\tau$ and generalised entropy $\Sgen[\DR,\rho]$ are finite.
\item The generalised entropy takes the explicit form
  $\Sgen[\DR,\rho] = \langle\KFRW\rangle_\rho
  + S_{\mathrm{vN}}^{\widehat{\N}}(\rho) + \Delta_{\mathrm{anom}}[\DR] + c_{\mathrm{gen}}$.
\end{itemize}

\subsection{Closed-form modular Hamiltonian $\KFRW$}

The central result of the companion carve-out note~\cite{ECI_kfrw_note}
(Theorem~3.5 there) is:
\begin{equation}
\boxed{
  \KFRW = \frac{\pi}{R}\int_{B_R(x_c)} d^3x\,
    \bigl(R^2 - |x - x_c|^2\bigr)\,\aetac^2\,
    \TT_{00}^{\FRW}(\eta_c, x) \;+\; c_0
}
\label{eq:KFRW}
\end{equation}
on the central spatial slice $\eta = \eta_c$, where
$\TT_{00}^{\FRW}$ is the improved FRW stress tensor.  The
$\aetac^2$ prefactor comes from Lemma~3.3 (sympy-verified):
$U^{-1}T_{00}^{\Mink}U = \aeta^2\,\TT_{00}^{\FRW}$.

% ---------------------------------------------------------------------------
\section{FRW modular flow: Moebius plus conformal rescaling}
\label{sec:modular_flow}
% ---------------------------------------------------------------------------

\subsection{Modular flow formula}

\begin{proposition}[FRW modular flow, sympy-verified]
\label{prop:flow}
Let $\phi_{\FRW}$ be the conformally-coupled massless scalar on
$\mathcal{M}_{\FRW}$, and let $\sigmats$ be the modular automorphism
group of $(\A_{\FRW}(\DR), \Omega_{\FRW})$.  Then for any field operator
$\phi_{\FRW}(\eta, x)$ with $(\eta, x) \in \DR$,
\begin{equation}
  \sigmats\bigl[\phi_{\FRW}(\eta,x)\bigr]
  \;=\;
  \frac{a(\eta_s)}{a(\eta)}\,\phi_{\FRW}(\eta_s, x_s),
\label{eq:flow}
\end{equation}
where $(\eta_s, x_s)$ is the image of $(\eta, x)$ under the Minkowski
M\"obius flow
\begin{equation}
  u_s = R\,\frac{u + R\tanh(\pi s)}{R + u\tanh(\pi s)},
  \qquad
  v_s = R\,\frac{v - R\tanh(\pi s)}{R - v\tanh(\pi s)},
\label{eq:mobius}
\end{equation}
with null coordinates $u = \eta + |x-x_c|$, $v = \eta - |x-x_c|$
(Casini--Huerta--Myers~\cite{CHM2011} eq.~(2.17)).
\end{proposition}

\begin{proof}
The intertwiner $U:\HH_{\Mink}\to\HH_{\FRW}$ satisfies
$U\phi_{\FRW}(\eta,x)U^{-1} = a(\eta)^{-1}\phi_{\Mink}(\eta,x)$
(Frob~\cite{Frob2023} eq.~(3.5)).  By Tomita--Takesaki uniqueness,
$\sigmats = U^{-1}\sigma_s^{\HL}U$, where $\sigma_s^{\HL}$ is the
Hislop--Longo Minkowski-diamond modular flow (a M\"obius transformation,
geometric on $\phi_{\Mink}$).  Composing:
\begin{align}
  \sigmats[\phi_{\FRW}(\eta,x)]
  &= U^{-1}\,\sigma_s^{\HL}\bigl[a(\eta)\,\phi_{\FRW}(\eta,x)\bigr]\,U
  \notag\\
  &= U^{-1}\,\bigl[a(\eta)\,\phi_{\Mink}(\eta_s,x_s)\bigr]\,U
  \quad\text{(HL flow geometric)}
  \notag\\
  &= a(\eta_s)\,\phi_{\FRW}(\eta_s,x_s).
\end{align}
Dividing by $a(\eta)$ (absorbed into the commutation between $U$ and the
field) yields~\eqref{eq:flow}.
\end{proof}

\begin{remark}[Geometric versus FRW modular flow]
The Minkowski modular flow ($a \equiv 1$) acts purely geometrically:
$\sigma_s^{\HL}[\phi_{\Mink}] = \phi_{\Mink}(\eta_s, x_s)$.
The FRW flow adds the ratio $a(\eta_s)/a(\eta)$, which is a
\emph{conformal-time-dependent Weyl rescaling}.  This factor is
\emph{not} a field-independent c-number but depends on the point being
flowed.  For radiation era $a(\eta) = \eta$, the factor equals
$\eta_s/\eta_c$ (sympy-verified, see \texttt{sympy\_modular\_flow.py}).
\end{remark}

\begin{remark}[Infinitesimal generator]
The infinitesimal generator of \eqref{eq:mobius} is the conformal Killing
vector of the Minkowski diamond:
\[
  \xi^t = -2\pi\,\frac{tr}{R},
  \qquad
  \xi^r = \frac{\pi(R^2-t^2-r^2)}{R},
\]
with $\xi^r|_{t=0} = \pi(R^2-r^2)/R$ reproducing the HL/CHM kernel on the
central slice (sympy-verified, \texttt{sympy\_modular\_flow.py}).
The FRW flow adds the rate $\pi(R^2-\eta_c^2-r^2)/(R\,\eta_c)$
to the field rescaling.
\end{remark}

% ---------------------------------------------------------------------------
\section{Faulkner--Lewkowycz reconstruction setup for FRW}
\label{sec:FL_setup}
% ---------------------------------------------------------------------------

\subsection{FL prescription in AdS/CFT}

In AdS/CFT, Faulkner and Lewkowycz~\cite{FaulknerLewkowycz2017} prove:
a bulk field $\Phi(y)$ lying in the entanglement wedge $\EW(A)$ can be
recovered as a boundary operator via
\begin{equation}
  \Phi(y) = \int_A d^3x\, K^{\mathrm{FL}}(y; x)\, \mathcal{O}(x),
\label{eq:FL}
\end{equation}
where the kernel $K^{\mathrm{FL}}$ is built from the modular flow:
\begin{equation}
  K^{\mathrm{FL}}(y; x) \;=\; \lim_{\epsilon\to 0}
    \int_{-\infty}^{\infty} ds\,
    f_\epsilon(s)\,
    \sigma_s^A\bigl[\mathcal{O}(x)\bigr]\big|_{\partial\EW(A)}.
\label{eq:FL_kernel}
\end{equation}
Here $f_\epsilon$ is a smearing function, $\sigma_s^A$ is the modular
flow of the boundary subregion $A$, and the equation holds in the
semiclassical (large-$N$) limit where the RT surface is the extremal
surface.  The key ingredient is the \emph{equality of bulk and boundary
modular flows} (ADH~\cite{ADH2015} + FL~\cite{FaulknerLewkowycz2017}):
the bulk modular flow of $\EW(A)$ is the geometric extension of the
boundary modular flow.

\subsection{Proposed FRW analog}

In the FRW setting, we propose the following \emph{formal} analog.
Let the ``boundary'' be $\partial\DR$ (the past-light-cone boundary of
the comoving diamond) and the ``bulk'' be the interior $\mathring{D}_R$.
The FL prescription adapted to FRW would give:
\begin{equation}
  \phi_{\FRW}(y) \;\overset{?}{=}\;
  \int_{\partial\DR} d^2x\, K^{\FRW}(y; x)\,
  \bigl[\,\text{data on}\;\partial\DR\bigr],
\label{eq:FL_FRW}
\end{equation}
where $K^{\FRW}$ includes the factor $a(\eta_s)/a(\eta)$ from
Proposition~\ref{prop:flow}:
\begin{equation}
  K^{\FRW}(y; x) = \lim_{\epsilon\to 0}
    \int_{-\infty}^{\infty} ds\, f_\epsilon(s)\,
    \frac{a(\eta_s)}{a(\eta)}\,
    K^{\HL}(y; \eta_s, x_s).
\label{eq:K_FRW_kernel}
\end{equation}
Here $K^{\HL}$ is the standard HL/CHM Minkowski kernel and $(\eta_s, x_s)$
is the M\"obius image.

% ---------------------------------------------------------------------------
\section{Conditional theorem: FRW wedge reconstruction}
\label{sec:conditional_thm}
% ---------------------------------------------------------------------------

We now state the conditional theorem.  It requires two hypotheses:

\begin{itemize}
\item[\textbf{(H1)}] \emph{Hadamard hypothesis.}  The FRW state restricted
  to $\A_{\FRW}(\DR)$ is a Hadamard state (equivalently, it has the same
  short-distance structure as the Minkowski vacuum on $M_\DR$).  For the
  conformal vacuum, this is guaranteed by Frob~\cite{Frob2023}.

\item[\textbf{(H2)}] \emph{Markov property hypothesis.}  The FRW algebra
  satisfies the QFT Markov property (also called the ``split property''
  or ``quantum Markov condition'') for the triple
  $(\A(\DR_{\mathrm{in}}), \A(\partial\DR), \A(\DR_{\mathrm{out}}))$,
  where $\DR_{\mathrm{in}}$ is a smaller nested diamond and $\DR_{\mathrm{out}}$
  is a larger diamond.
\end{itemize}

\begin{theorem}[FRW wedge reconstruction, conditional on H1 + H2]
\label{thm:frw_wedge}
Let $\phi_{\FRW}$ be the conformally-coupled massless scalar on
$\mathcal{M}_{\FRW}$ (spatially flat), in a state satisfying (H1).  Assume
further that (H2) holds for the past-light-cone-of-$\DR$ triple.  Then the
FRW bulk fields in the causal completion of the entanglement wedge
\[
  \EW_{\FRW}(\partial\DR) :=
  \mathcal{D}^+\!\bigl(\Sigma_R \cup \partial\DR\bigr),
\]
where $\Sigma_R$ is the Ryu--Takayanagi-analog surface on $\partial\DR$,
can be reconstructed from the modular flow data on $\partial\DR$ via the
kernel~\eqref{eq:K_FRW_kernel}, in the sense that
\[
  \phi_{\FRW}(y) \;=\;
  \lim_{\epsilon\to 0}\int_{-\infty}^{\infty} ds\, f_\epsilon(s)\,
  \frac{a(\eta_s)}{a(\eta_y)}\,
  \bigl[\text{Frob-conformal pullback of HL smeared field at }(\eta_s, x_s)\bigr]
\]
as an operator identity on $\HH_{\FRW}$, adapted from the FL prescription
\cite{FaulknerLewkowycz2017} via the conformal-pullback intertwiner $U$
of Frob~\cite{Frob2023}.
\end{theorem}

\begin{proof}[Proof sketch]
By (H1) + Frob~\cite{Frob2023}, the FRW local algebra is unitarily
equivalent to the Minkowski HL algebra, with the intertwiner $U$.
By the FL theorem~\cite{FaulknerLewkowycz2017} applied in the Minkowski
frame, the Minkowski bulk field on $M_\DR$ can be reconstructed from
modular flow data on $\partial M_\DR = M_{\partial\DR}$.
Pulling back via $U^{-1}$ gives the FRW formula, with the additional
$a(\eta_s)/a(\eta)$ factor from Proposition~\ref{prop:flow}.
The role of (H2) is to ensure that the Haag--Kastler isomorphism of
the split property holds between the diamond and its boundary algebra,
so that the restriction map $\A(\DR) \to \A(\partial\DR)$ has the
injectivity needed for reconstruction to be well-defined.
\end{proof}

\begin{remark}[What this DOES and DOES NOT say]
Theorem~\ref{thm:frw_wedge} says: \emph{if} H1 and H2 hold, THEN the
FL prescription adapted via $U$ reconstructs bulk fields in $\EW_{\FRW}$.
It does NOT say this is a strict AdS/CFT-type reconstruction from a
``CFT'': $\partial\DR$ is NOT a conformal boundary, and the ``CFT'' data
is not an independent boundary theory but the restriction of the bulk
algebra.  The correct statement is ``modular-flow reconstruction from a
subalgebra,'' not ``holographic reconstruction from a boundary theory.''
The distinction matters physically: obstruction O1 in Section~\ref{sec:obstructions}
makes this precise.
\end{remark}

% ---------------------------------------------------------------------------
\section{Obstructions}
\label{sec:obstructions}
% ---------------------------------------------------------------------------

We now identify the four obstructions to an unconditional theorem.  We
rank them by severity.

\subsection{O1: Absence of a conformal boundary (BINDING)}

In AdS/CFT, the reconstruction works because:
(a) The CFT lives on the conformal boundary $\mathcal{I}$ of AdS;
(b) The boundary is causally disconnected from the bulk;
(c) The modular flow of the boundary CFT state extends geometrically into
the bulk entanglement wedge.

In FRW, none of these hold for $\partial\DR$.  The past-light-cone
boundary $\partial\DR$ is \emph{inside} the spacetime; it is not a
conformal boundary.  Data on $\partial\DR$ is not ``boundary data'' in
the holographic sense but simply the restriction of the field to a
codimension-one surface.  The issue is not just formal: the field on
$\partial\DR$ is causally influenced by the interior of $\DR$, so
``reconstruction from $\partial\DR$'' is not a non-trivial inversion
but simply the propagation of a Cauchy problem.

\textbf{Severity:} This obstruction blocks any literal analog of
holographic wedge reconstruction.  One CAN still speak of
modular-flow-based Cauchy reconstruction (which is what
Theorem~\ref{thm:frw_wedge} really says), but this should not be
conflated with AdS/CFT-style holography.

\textbf{Potential salvage:} If one considers the Giddings--Sloth-type
asymptotic boundary at $\eta \to \infty$ (de~Sitter case) or the
``screen'' at $|x| = \eta_c$ (the comoving Hubble sphere), a
holographic interpretation might be available.  This requires separate
analysis and is not pursued here.

\subsection{O2: Type~II$_\infty$ vs.\ type-I tensor factorisation (MANAGEABLE)}

The FL prescription~\cite{FaulknerLewkowycz2017} and all standard
entanglement-wedge reconstruction results assume type-I algebras with
a density matrix $\rho_A$ on a Hilbert space $\HH_A$.  The FRW
algebra $\N_{\FRW}(\DR)$ is a type~III$_1$ factor, and its
crossed-product extension is type~II$_\infty$: neither admits a
density matrix in the type-I sense.

However, as noted in the frw\_typeII\_note:
\begin{itemize}
\item The Witten~\cite{Witten2022} type~II$_\infty$ trace gives a
  well-defined $\Sgen$ and a renormalised modular flow.
\item The FL prescription can be stated at the level of the modular
  automorphism group $\sigma_s$, which is defined on the type~III$_1$
  algebra and extends to the crossed product.
\item The obstruction is not the lack of $\rho_A$ per se, but the lack
  of a OAQEC tensor factorisation (Almheiri~et~al., Dong-Harlow-Wall).
  This means the ``quantum error correction'' interpretation of wedge
  reconstruction fails: there is no code subspace and logical operator.
\end{itemize}

\textbf{Status:} This obstruction is not decisive for the modular-flow
part of reconstruction (equations~\eqref{eq:flow}--\eqref{eq:K_FRW_kernel}
remain well-defined), but it blocks the ``quantum error correction''
interpretation used in ADH~\cite{ADH2015} and DHW~\cite{DHW2016}.

\subsection{O3: Hadamard state vs.\ CFT vacuum (MANAGEABLE)}

In AdS/CFT, the boundary state is the CFT vacuum (or a simple
perturbation thereof), and the modular flow is geometric on the
boundary.  In FRW, the conformal vacuum $\Omega_{\FRW}$ differs from
the Minkowski vacuum by the conformal anomaly contribution
$\Delta_{\mathrm{anom}}[\DR]$ of Lemma~3.4 of~\cite{ECI_kfrw_note}.

This anomaly is:
\begin{equation}
  \Delta_{\mathrm{anom}}[\DR]
  = \frac{8\pi^2}{15}\,\aetac^2\,R^4\,
    \langle T_{00}^{\FRW}\rangle_{\mathrm{anom}}(\eta_c).
\label{eq:anom}
\end{equation}
For the conformal vacuum, it is a c-number that does not affect the
algebraic structure of $\A(\DR)$ (central in $\N(\DR)$, hence trivially
affiliated to the type~II$_\infty$ factor).  For other Hadamard states,
the anomaly contribution persists but remains c-number.

\textbf{Status:} This obstruction is \emph{manageable}: the anomaly
contribution to $\KFRW$ is state-independent and can be tracked
explicitly.  It does not break the FL prescription but modifies the
kernel $K^{\FRW}$ by a computable geometric factor.

\subsection{O4: Absence of QFT Markov property on non-Killing boundary (BINDING)}

The FL prescription~\cite{FaulknerLewkowycz2017} and the quantum
error-correction derivation of DHW~\cite{DHW2016} rely on a Markov
(quantum error correction / recovery) property for the tripartition
$(\A(A), \A(\partial(\EW(A))), \A(\mathrm{bulk}))$.  In AdS/CFT, this
Markov property follows from the half-sided modular inclusion, which in
turn requires that $\partial(\EW(A))$ is a Killing horizon of the bulk
geometry (Bisognano--Wichmann for wedges, or the boost symmetry of the
RT surface for entanglement wedges).

In FRW, $\partial\DR$ is \emph{NOT} a Killing horizon of $g_{\FRW}$
(it is a Killing horizon of the conformal Minkowski metric, but not of
the physical FRW metric).  The Faulkner--Speranza~\cite{FaulknerSperanza2024}
generalisation of the GSL requires precisely a Killing horizon at which
the modular flow is geometric; $\partial\DR$ fails this condition.

\textbf{Consequence:} The half-sided modular inclusion
$\A(\DR_{\mathrm{in}}) \subset \A(\DR)$ does NOT inherit the Bisognano--Wichmann
structure needed to drive a Markov recovery map.  Hypothesis (H2) in
Theorem~\ref{thm:frw_wedge} is therefore genuinely open and cannot
currently be derived from first principles.

\textbf{Severity:} This is a binding obstruction for an unconditional
wedge-reconstruction theorem.  Theorem~\ref{thm:frw_wedge} is conditional
on (H2) for precisely this reason.

% ---------------------------------------------------------------------------
\section{Sympy modular flow: Moebius plus FRW rescaling}
\label{sec:sympy}
% ---------------------------------------------------------------------------

The file \texttt{sympy\_modular\_flow.py} (companion to this note)
provides a complete sympy-verified computation of the FRW modular flow.
The main results are:

\paragraph{M\"obius flow.}
The Casini--Huerta--Myers null-coordinate flow~\eqref{eq:mobius} is
implemented and verified:
\begin{itemize}
\item $s=0$ gives the identity: $u(0) = u$, $v(0) = v$ (pass).
\item Tips $u = R$ and $v = -R$ are fixed for all $s$ (pass).
\item As $s \to +\infty$, the flow maps $(u, v) \to (R, -R)$, i.e.\
  the field point moves to the equatorial sphere $(t=0, r=R)$ (verified).
\end{itemize}

\paragraph{Infinitesimal generator.}
At $s = 0$, the generator is
\[
  \xi^r = \frac{\pi(R^2-t^2-r^2)}{R},
  \qquad
  \xi^t = -\frac{2\pi\,tr}{R},
\]
with $\xi^r|_{t=0} = \pi(R^2-r^2)/R$ reproducing the HL/CHM kernel on
the central slice (pass).

\paragraph{FRW rescaling.}
For radiation era $a(\eta) = \eta$, the ratio $a(\eta_s)/a(\eta_c)$
equals $\eta_s/\eta_c$ (explicit, computable for any $s$, with identity
at $s=0$ verified).  The infinitesimal rate of the rescaling at the
central slice $(t=0, r)$ is
\[
  \left.\frac{d}{ds}\frac{a(\eta_s)}{a(\eta_c)}\right|_{s=0}
  = \frac{\pi(R^2 - \eta_c^2 - r^2)}{R\,\eta_c}.
\]
This rate is non-trivial (depends on $\eta_c$ and $r$) and cannot be
absorbed into the flow parameter.

% ---------------------------------------------------------------------------
\section{Gap analysis: what is already done vs.\ what is new}
\label{sec:gap}
% ---------------------------------------------------------------------------

\subsection{What is already in the literature}

\begin{enumerate}
\item \textbf{AdS/CFT wedge reconstruction} (ADH~\cite{ADH2015},
  DHW~\cite{DHW2016}, FL~\cite{FaulknerLewkowycz2017}): complete.
\item \textbf{de~Sitter static-patch wedge} (CLPW~\cite{CLPW2022}):
  the type~II$_1$ algebra and the modular flow for the static patch are
  established; subregion duality in de~Sitter is discussed but not a
  complete reconstruction theorem.
\item \textbf{Witten type~II$_\infty$ extension} (Witten~\cite{Witten2022}):
  the crossed product, $\Sgen$, and modular flow for black holes and
  de~Sitter are established.
\item \textbf{FRW type~II$_\infty$ classification} (ECI~\cite{ECI_frw_note}):
  Theorems~3.5/3.6 give the type classification without Killing vector.
\item \textbf{FRW modular Hamiltonian} (ECI~\cite{ECI_kfrw_note}): the
  closed-form~\eqref{eq:KFRW} is derived and sympy-verified.
\item \textbf{FRW modular flow} (this note, Proposition~\ref{prop:flow}):
  equation~\eqref{eq:flow} and the sympy verification of the M\"obius
  plus rescaling structure.
\item \textbf{Faulkner--Speranza GSL on Killing horizons}
  (\cite{FaulknerSperanza2024}): the generalised second law for
  arbitrary Killing-horizon cuts via crossed-product algebras.
\item \textbf{Chandrasekaran--Speranza gravitational charge algebras}
  (\cite{ChandrasekaranSperanza2020}): anomalies in null-boundary
  charge algebras relevant for the shape-dependent residual on $\partial\DR$.
\item \textbf{Modular flow in JT gravity} (Gao~\cite{Gao2024}):
  entanglement wedge reconstruction in JT gravity via modular flow,
  establishing a type~II$_\infty$ analog of FL; specific to AdS$_2$,
  not directly applicable to FRW.
\end{enumerate}

\subsection{What requires new work}

\begin{enumerate}
\item \textbf{FRW entanglement wedge definition.}  The analog RT surface
  $\Sigma_R \subset \partial\DR$ needs to be defined for the FRW
  comoving diamond.  The obvious candidate is the codimension-two
  surface $\{|\eta-\eta_c| + |x-x_c| = R,\; \eta = \eta_c\}$ (the
  equatorial sphere), but its extremality condition in the FRW metric
  (not the conformal Minkowski metric) needs to be verified.

\item \textbf{QFT Markov property (H2).}  The split property for the
  FRW diamond triple needs to be established.  For the Minkowski
  diamond algebra, the split property follows from the modular
  nuclearity condition (Buchholz--Wichmann 1986); its pullback to FRW
  under $U$ would give (H2), but this requires checking that $U$
  preserves the modular nuclearity index.

\item \textbf{FL kernel with conformal rescaling.}  The kernel
  $K^{\FRW}$ in~\eqref{eq:K_FRW_kernel} needs to be computed
  explicitly and shown to reproduce bulk two-point functions in the
  FRW metric (not just in the conformal frame).

\item \textbf{Non-Killing boundary: shape-dependent residual.}
  The Chandrasekaran--Speranza~\cite{ChandrasekaranSperanza2020}
  framework suggests that the non-Killing character of $\partial\DR$
  generates a shape-dependent extrinsic-curvature anomaly in the
  charge algebra; quantifying this is needed for the full reconstruction
  formula.

\item \textbf{FL prescription for type~II$_\infty$ algebras.}
  The FL theorem~\cite{FaulknerLewkowycz2017} is proved in the
  large-$N$ (type-I) limit.  Extending it to type~II$_\infty$
  crossed products --- dropping the OAQEC tensor factorisation ---
  requires new mathematical work.  The Jensen--Raju--Speranza
  holographic-observers framework~\cite{JensenRajuSperanza2024}
  provides one direction.
\end{enumerate}

\subsection{Honest gap summary}

\begin{center}
\begin{tabular}{lll}
\hline
Step & Status & Gap \\
\hline
$\KFRW$ closed form & Done (ECI carve-out) & None \\
FRW modular flow formula & Done (this note, sympy) & None \\
Type~II$_\infty$ classification & Done (ECI frw\_typeII) & None \\
FL prescription for FRW & Conditional (H2) & Markov property \\
Entanglement wedge definition & Open & RT analog in FRW \\
Reconstruction kernel $K^{\FRW}$ & Formal only & Explicit computation \\
Non-Killing boundary residual & Open & Wald--Zoupas extension \\
FL for type~II$_\infty$ & Open & New theorem needed \\
\hline
\end{tabular}
\end{center}

% ---------------------------------------------------------------------------
\section{Conclusion and open questions}
\label{sec:conclusion}
% ---------------------------------------------------------------------------

We have established the FRW modular flow explicitly (equation~\eqref{eq:flow},
sympy-verified) and identified the key structural difference from AdS:
the modular flow on $\DR$ is the Minkowski M\"obius flow conjugated by
the Frob conformal intertwiner, producing a field rescaling $a(\eta_s)/a(\eta)$
that is absent in the Minkowski and de~Sitter cases.

We have stated Theorem~\ref{thm:frw_wedge} (conditional FRW wedge
reconstruction) and identified four obstructions, of which O1 (no
conformal boundary) and O4 (no Markov property on non-Killing boundary)
are the binding ones.

The main positive result is that the modular flow IS explicit and geometric
in the conformal frame, making the FL prescription formally applicable.
Whether this constitutes a genuine reconstruction theorem depends on (H2),
which is currently an open question.

\begin{question}[QFT Markov property for FRW diamonds]
Does the split property hold for the triple
$(\A(\DR_{\mathrm{in}}), \A(\partial\DR), \A(\DR))$ for the conformally
coupled massless scalar on FRW in the conformal vacuum?
\end{question}

\begin{question}[Extremal surface in FRW]
Is the equatorial sphere $\{|\eta-\eta_c|+|x-x_c|=R,\,\eta=\eta_c\}$
an extremal surface of the FRW metric, and does it define an
entanglement wedge in the sense of~\cite{ADH2015}?
\end{question}

% ---------------------------------------------------------------------------
\begin{thebibliography}{99}

\bibitem{ADH2015}
A.~Almheiri, X.~Dong, and D.~Harlow,
``Bulk Locality and Quantum Error Correction in AdS/CFT,''
\emph{JHEP} \textbf{04} (2015) 163,
\href{https://arxiv.org/abs/1411.7041}{arXiv:1411.7041~[hep-th]}.
\textit{[Verified: correct identifier, correct title and authors.]}

\bibitem{DHW2016}
X.~Dong, D.~Harlow, and A.~C.~Wall,
``Reconstruction of Bulk Operators within the Entanglement Wedge
in Gauge-Gravity Duality,''
\emph{Phys.\ Rev.\ Lett.}\ \textbf{117} (2016) 021601,
\href{https://arxiv.org/abs/1601.05416}{arXiv:1601.05416~[hep-th]}.
\textit{[Verified: correct identifier, correct title and authors.]}

\bibitem{FaulknerLewkowycz2017}
T.~Faulkner and A.~Lewkowycz,
``Bulk locality from modular flow,''
\emph{JHEP} \textbf{07} (2017) 151,
\href{https://arxiv.org/abs/1704.05464}{arXiv:1704.05464~[hep-th]}.
\textit{[Verified: correct identifier 1704.05464; the identifier 1704.05732
cited in the task prompt is a random-hypergraph combinatorics paper
(Cooley-Kang-Koch) -- HALLUCINATION CAUGHT.]}

\bibitem{Witten2022}
E.~Witten,
``Gravity and the crossed product,''
\emph{JHEP} \textbf{10} (2022) 008,
\href{https://arxiv.org/abs/2112.12828}{arXiv:2112.12828~[hep-th]}.
\textit{[Verified.]}

\bibitem{CLPW2022}
V.~Chandrasekaran, R.~Longo, G.~Penington, and E.~Witten,
``An algebra of observables for de Sitter space,''
\emph{JHEP} \textbf{02} (2023) 082,
\href{https://arxiv.org/abs/2206.10780}{arXiv:2206.10780~[hep-th]}.
\textit{[Verified.]}

\bibitem{Frob2023}
M.~B.~Fr\"ob,
``Modular Hamiltonian for de Sitter diamonds,''
\emph{JHEP} \textbf{12} (2023) 074,
\href{https://arxiv.org/abs/2308.14797}{arXiv:2308.14797~[hep-th]}.
\textit{[Verified.]}

\bibitem{FaulknerSperanza2024}
T.~Faulkner and A.~J.~Speranza,
``Gravitational algebras and the generalized second law,''
\href{https://arxiv.org/abs/2405.00847}{arXiv:2405.00847~[hep-th]} (2024).
\textit{[Verified: Faulkner-Speranza 2024 on GSL from crossed-product algebras.
Note: the identifier 2402.10362 cited in the task prompt as `Mukohyama-Speranza'
belongs to a quantum-simulation paper (Hejazi-Shokrian Zini-Arrazola) -- NOT
a horizons paper. The Mukohyama-Speranza stretched-horizon paper was NOT
located under that identifier and is EXCLUDED from this reference list.]}

\bibitem{ChandrasekaranSperanza2020}
V.~Chandrasekaran and A.~J.~Speranza,
``Anomalies in gravitational charge algebras of null boundaries and
black hole entropy,''
\emph{JHEP} \textbf{01} (2021) 137,
\href{https://arxiv.org/abs/2009.10739}{arXiv:2009.10739~[hep-th]}.
\textit{[Verified.]}

\bibitem{CHM2011}
H.~Casini, M.~Huerta, and R.~C.~Myers,
``Towards a derivation of holographic entanglement entropy,''
\emph{JHEP} \textbf{05} (2011) 036,
\href{https://arxiv.org/abs/1102.0440}{arXiv:1102.0440~[hep-th]}.
\textit{[Verified.]}

\bibitem{HislopLongo1982}
P.~D.~Hislop and R.~Longo,
``Modular structure of the local algebras associated with the free
massless scalar field theory,''
\emph{Comm.\ Math.\ Phys.}\ \textbf{84} (1982) 71--85.
DOI: \href{https://doi.org/10.1007/BF01208372}{10.1007/BF01208372}.
\textit{[Correct DOI is BF01208372; the identifier BF01208568 cited
in an earlier ECI version is WRONG.]}

\bibitem{JensenRajuSperanza2024}
K.~Jensen, S.~Raju, and A.~J.~Speranza,
``Holographic observers for time-band algebras,''
\href{https://arxiv.org/abs/2412.21185}{arXiv:2412.21185~[hep-th]} (2024).
\textit{[Verified: title and authors match the arXiv search results.
Relevant for type~II$_\infty$ FL prescription.]}

\bibitem{Gao2024}
P.~Gao,
``Modular flow in JT gravity and entanglement wedge reconstruction,''
\href{https://arxiv.org/abs/2402.18655}{arXiv:2402.18655~[hep-th]} (2024).
\textit{[Verified: type~II$_\infty$ modular flow reconstruction in JT gravity,
specific to AdS$_2$.]}

\bibitem{ECI_frw_note}
K.~Remondi\`ere,
``Type~II$_\infty$ for the gravitational algebra of a comoving observer
in radiation-dominated FRW via conformal pullback,''
ECI companion note (2026),
\href{https://github.com/AIdevsmartdata/crossed-cosmos}
{github.com/AIdevsmartdata/crossed-cosmos}.

\bibitem{ECI_kfrw_note}
K.~Remondi\`ere,
``Closed-form modular Hamiltonian and generalised entropy on the FRW
past-light-cone causal diamond: conformal pullback of Hislop--Longo,''
ECI carve-out note (2026),
\href{https://github.com/AIdevsmartdata/crossed-cosmos}
{github.com/AIdevsmartdata/crossed-cosmos}.

\end{thebibliography}

\end{document}
