%% modular_shadow_LMP_v2.tex
%% Standalone note for Letters in Mathematical Physics
%% "A Modular Lyapunov Bound for Finite-Rank Type-II_infty Crossed-Product Algebras"
%% v2 (2026-05-05): rework of v1 (553 lines, "saturation conjecture" framing) per A11/A19 audit:
%%   - Saturation downgraded to conjecture (free-QFT counter-examples preclude converse).
%%   - Bound theorem (lambda_K^mod <= 2pi/beta) presented as PROVABLE today via
%%     Parker et al. 2019 + Bisognano-Wichmann.
%%   - Vardian arXiv:2602.02675 cited as concurrent independent work (AdS/CFT scope).
%%   - FS24 over-attribution corrected (Faulkner-Speranza derive crossed-product S_gen,
%%     do NOT explicitly identify S_gen with C_K^mod; ECI conjecture).
%%   - A19 BEC bound 288 +/- 82 s^-1 added as concrete falsifier on Kolobov 2021 data.
%% All arXiv IDs live-verified 2026-05-05 (WebFetch, no Mistral).

\documentclass[12pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{geometry}
\geometry{margin=2.5cm}

%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}{Conjecture}
\newtheorem{remark}{Remark}
\newtheorem{corollary}[theorem]{Corollary}

\title{\textbf{A modular Lyapunov bound for finite-rank\\
       type-II$_\infty$ crossed-product algebras}\\[6pt]
       \large with a saturation conjecture and an analog-Hawking falsifier}

\author{%
  K\'evin Remondi\`ere\\[2pt]
  \small Independent researcher, Tarbes, France\\[2pt]
  \small \textit{Submitted to Letters in Mathematical Physics}
}

\date{Preprint: \today\quad(v2)}

\begin{document}

\maketitle

\begin{abstract}
Let $\mathcal{M}_R \otimes B(\mathcal{H}_n)$ be a finite-rank truncation of
the type-II$_\infty$ crossed-product von~Neumann algebra associated with an
observer's causal wedge at Bisognano--Wichmann KMS inverse temperature
$\beta = 2\pi$, and let $K_R$ be its modular Hamiltonian.
We prove that the modular Krylov (Lanczos) complexity
$\mathcal{C}_K^{\mathrm{mod}}(s)$ of any operator $\mathcal{O}$ under the
modular flow $\sigma_s = e^{i K_R s}\cdot e^{-i K_R s}$ satisfies the bound
\[
  \lambda_K^{\mathrm{mod}}
  \;:=\; \limsup_{s\to\infty}\frac{1}{s}\,\log \mathcal{C}_K^{\mathrm{mod}}(s)
  \;\le\; \frac{2\pi}{\beta}\;=\;1,
\]
matching the Maldacena--Shenker--Stanford chaos bound \emph{kinematically}
on the truncation.
The proof combines the universal operator-growth inequality of Parker,
Cao, Avdoshkin, Scaffidi, and Altman~\cite{Parker2019} with the
Bisognano--Wichmann theorem~\cite{BisognanoWichmann1976} and uses
no large-$N$, holography, or chaos hypothesis.
The converse statement -- that saturation $\lambda_K^{\mathrm{mod}} =
2\pi/\beta$ characterises type-II$_\infty$ chaos -- is shown to be
\emph{false in general}: free-QFT and integrable counter-examples
(Avdoshkin--Dymarsky~\cite{AvdoshkinDymarsky2019},
Camargo--Jahnke--Kim--Nishida~\cite{Camargo2023},
Sreeram--Kannan--Modak--Aravinda~\cite{Sreeram2025})
attain the same kinematic ceiling without underlying chaos.
We therefore retain saturation only as a conjectural \emph{sufficient
condition} (Conjecture~\ref{conj:msc-saturation}).
A concrete one-sided falsifier is then derived: for the
Kolobov--Golubkov--Mu\~{n}oz~de~Nova--Steinhauer~\cite{Kolobov2021}
analog black-hole BEC at $T_H = 0.35 \pm 0.10$~nK
(Mu\~{n}oz~de~Nova et al.~\cite{Steinhauer2019}), the bound predicts
$\Gamma_{\mathrm{meas}} \le 2\pi k_B T_H/\hbar = (288 \pm 82)\,\mathrm{s}^{-1}$
for any exterior-phonon temporal correlation rate.
We compare with the concurrent boundary-area-operator construction of
Vardian~\cite{Vardian2026} (AdS/CFT) and note that our scope (de~Sitter
static patch / generic type-II$_\infty$ truncation) is complementary.
\end{abstract}

% =====================================================================
\section{Introduction}
\label{sec:intro}
% =====================================================================

The Maldacena--Shenker--Stanford (MSS) bound on chaos~\cite{MSS2016}
\begin{equation}
  \lambda_L \;\le\; \frac{2\pi k_B T}{\hbar} \;=\; \frac{2\pi}{\beta}
  \label{eq:mss}
\end{equation}
limits the OTOC Lyapunov exponent of any thermal large-$N$ quantum system,
with maximally chaotic systems (SYK, near-extremal black holes)
saturating it.
Parker, Cao, Avdoshkin, Scaffidi, and Altman~\cite{Parker2019} subsequently
introduced the universal operator-growth hypothesis: in chaotic systems
the Lanczos coefficients $b_n$ of any few-body operator grow linearly,
$b_n \sim \alpha n$, and the Krylov (Lanczos) complexity satisfies
the rigorous inequality
\begin{equation}
  \lambda_K \;\le\; 2\alpha,\qquad
  \mathcal{C}_K(t) \;\le\; A \cosh(2\alpha t),
  \label{eq:parker}
\end{equation}
sharper than~\eqref{eq:mss} and free of large-$N$ assumptions.
For thermal correlators in $d$-dimensional CFT one has the universal
slope $\alpha = \pi/\beta$, recovering the MSS rate as a kinematic
upper envelope.

In parallel, Witten~\cite{Witten2022} and Chandrasekaran--Longo--Penington--Witten~\cite{CLPW2023}
showed that gravitational observable algebras with quantum reference
frames (QRFs) are type-II$_\infty$ crossed products
$\mathcal{A}_{\mathrm{II}_\infty} = \mathcal{A}_{\mathrm{III}_1}
\rtimes_{\sigma^\phi} \mathbb{R}$, restoring a semifinite trace and
finite generalised entropy.
De~Vuyst, Eccles, H\"{o}hn, and Kirklin~\cite{DEHK2025a,DEHK2025b}
established observer-dependent gravitational entropy in this framework,
and Faulkner--Speranza~\cite{FaulknerSperanza2024} derived the
generalised second law in the crossed-product setting.

\paragraph{Aim of this paper.}
We prove a clean kinematic statement:
on \emph{any} finite-rank truncation of a type-II$_\infty$ crossed
product at the canonical Bisognano--Wichmann temperature $\beta = 2\pi$,
the modular Krylov complexity is bounded by the MSS rate
(Theorem~\ref{thm:bound}, \S\ref{sec:bound}).
The proof combines the Parker--Cao--Avdoshkin--Scaffidi--Altman
inequality with the Bisognano--Wichmann KMS condition and is independent
of any large-$N$ or chaos assumption.

\paragraph{What this paper does \emph{not} prove.}
We do \emph{not} prove the converse: saturation of the bound does
not characterise type-II$_\infty$ chaos.
Free-QFT lattice models~\cite{AvdoshkinDymarsky2019},
free and interacting bounded-power-spectrum scalar
fields~\cite{Camargo2023}, and operator/initial-state-dependent
counter-examples~\cite{Sreeram2025} all attain the same kinematic
ceiling without exhibiting MSS-saturating chaos.
The earlier ECI ``modular shadow'' draft (v1, 2026-05-04) presented
saturation as the central object; this revision (v2) downgrades
saturation to a conjecture (Conjecture~\ref{conj:msc-saturation},
\S\ref{sec:saturation}) that requires additional dynamical input
beyond the type-II$_\infty$ structure.

\paragraph{Concurrent work.}
While preparing this manuscript we became aware of
Vardian~\cite{Vardian2026} (arXiv:2602.02675, February 2026),
``Modular Krylov Complexity as a Boundary Probe of Area Operator and
Entanglement Islands'', which uses modular Lanczos coefficients to
reconstruct the QES area operator from boundary dynamics in AdS/CFT.
Vardian's analysis is complementary to ours: it operates in the
holographic AdS/CFT setting and targets the area-operator spectrum,
whereas we work intrinsically in the finite-rank type-II$_\infty$
algebra at any KMS temperature, and aim at a generic Lyapunov bound
applicable to de~Sitter and BEC analog systems.
We compare scopes in \S\ref{sec:discussion}.

\paragraph{Organization.}
\S\ref{sec:setup} fixes the algebraic setup (type-II$_\infty$ crossed
product, finite-rank truncation, modular Krylov complexity).
\S\ref{sec:bound} states and proves the bound theorem.
\S\ref{sec:saturation} downgrades the saturation step to a conjecture
and reviews the free-QFT counter-examples.
\S\ref{sec:bec} applies the bound to the Kolobov--Steinhauer 2021
BEC analog black hole, deriving the falsifier
$\Gamma_{\mathrm{meas}} \le (288 \pm 82)\,\mathrm{s}^{-1}$.
\S\ref{sec:discussion} compares with Vardian and discusses scope.
\S\ref{sec:outlook} lists open problems.

% =====================================================================
\section{Setup: type-II$_\infty$ crossed product, finite-rank truncation,
         and modular Krylov complexity}
\label{sec:setup}
% =====================================================================

\subsection{The crossed-product algebra}

Let $\mathcal{A}_{\mathrm{III}_1}$ be the type-III$_1$ von~Neumann
algebra of observables in a relativistic QFT restricted to a Rindler
wedge $\mathcal{W}$, and let $\phi$ be the vacuum KMS state at the
Bisognano--Wichmann inverse temperature $\beta_{\mathrm{BW}} = 2\pi$.
Following Witten~\cite{Witten2022} and CLPW~\cite{CLPW2023}, the
crossed product with the modular automorphism group
\begin{equation}
  \mathcal{A}_{\mathrm{II}_\infty}
  \;=\;
  \mathcal{A}_{\mathrm{III}_1} \rtimes_{\sigma^\phi} \mathbb{R}
  \label{eq:crossed}
\end{equation}
is type-II$_\infty$, admits a faithful semifinite trace $\tau$, and
hosts a well-defined dressed modular Hamiltonian $K_\phi$ generating
the modular flow.
The Bisognano--Wichmann theorem~\cite{BisognanoWichmann1976} guarantees
$K_\phi = 2\pi K_{\mathrm{boost}}$ where $K_{\mathrm{boost}}$ is the
boost generator of the wedge.

\subsection{Finite-rank truncation}

Working with the full type-II$_\infty$ algebra requires care: $K_\phi$
is unbounded and the modular flow is generally non-trace-preserving.
For our bound theorem we work on the \emph{finite-rank truncation}
\begin{equation}
  \mathcal{M}_R \otimes B(\mathcal{H}_n)
  \;:=\;
  P_n \,\mathcal{A}_{\mathrm{II}_\infty}\, P_n
  \label{eq:truncation}
\end{equation}
where $P_n$ is the spectral projector of $K_\phi$ onto the modular
energy band $|K_\phi| \le \Lambda_n$ with $\dim(\mathrm{ran}\,P_n) = n$.
This truncation preserves the $\beta = 2\pi$ KMS condition (it is the
restriction of the GNS representation to a finite-dimensional subspace
on which $K_\phi$ acts as a bounded self-adjoint operator) and inherits
the semifinite trace from the type-II$_\infty$ parent.

\begin{remark}\label{rem:bw-truncation}
The Bisognano--Wichmann property survives the truncation in the
following sense: although $\mathcal{M}_R \otimes B(\mathcal{H}_n)$ is
not itself a wedge algebra, its modular flow restricted to operators
of energy $\le \Lambda_n$ agrees with the boost flow on those modes,
because the Bisognano--Wichmann modular operator commutes with the
spectral projection $P_n$.
This is the only structural fact about the truncation that we will
need.
\end{remark}

\subsection{Modular Krylov complexity}

For an operator $\mathcal{O} \in \mathcal{M}_R \otimes B(\mathcal{H}_n)$,
its modular evolution is
\begin{equation}
  \mathcal{O}(s) \;=\; e^{i K_\phi s}\,\mathcal{O}\,e^{-i K_\phi s},
  \qquad s \in \mathbb{R}.
  \label{eq:modular-flow}
\end{equation}
Equip the truncation with the KMS Gel'fand--Naimark--Segal inner product
$(A,B)_\phi := \tau(A^* B)$ (well-defined because $\tau$ is finite on
the truncation).
The Lanczos algorithm applied to the modular Liouvillian
$\mathcal{L} := [K_\phi,\,\cdot\,]$ produces an orthonormal Krylov
basis $\{|\mathcal{O}_n\rangle\}$ with tridiagonal coefficients
$\{b_n\}_{n\ge 1}$, and the modular Krylov complexity is
\begin{equation}
  \mathcal{C}_K^{\mathrm{mod}}(s)
  \;:=\;
  \sum_{n\ge 0} n\,|\varphi_n(s)|^2,
  \qquad
  \mathcal{O}(s) \;=\; \sum_{n\ge 0}\varphi_n(s)\,\mathcal{O}_n,
\end{equation}
following the modular construction of Caputa--Mag\'{a}n--Patramanis--Tonni~\cite{Caputa2024}
who showed $\lambda^{\mathrm{mod}}_L = 2\pi$ in dimensionless modular
time for 2D CFTs.

\subsection{Modular Lyapunov exponent}

The \emph{modular Lyapunov exponent} of $\mathcal{O}$ is
\begin{equation}
  \lambda_K^{\mathrm{mod}}(\mathcal{O})
  \;:=\;
  \limsup_{s\to\infty}\;\frac{1}{s}\,
  \log \mathcal{C}_K^{\mathrm{mod}}(s),
\end{equation}
when this limit is finite.
Our central object is the supremum of $\lambda_K^{\mathrm{mod}}(\mathcal{O})$
over operators $\mathcal{O}$ in a dense $\ast$-subalgebra of the
truncation; the bound theorem (next section) controls this supremum.

% =====================================================================
\section{The modular Lyapunov bound: a provable theorem}
\label{sec:bound}
% =====================================================================

\begin{theorem}[Modular Lyapunov bound on finite-rank type-II$_\infty$
truncations]\label{thm:bound}
Let $\mathcal{M}_R \otimes B(\mathcal{H}_n)$ be the finite-rank
truncation~\eqref{eq:truncation} of the type-II$_\infty$ crossed
product~\eqref{eq:crossed} at Bisognano--Wichmann KMS inverse
temperature $\beta = 2\pi$, equipped with the KMS--GNS inner product
$(A,B)_\phi = \tau(A^* B)$ and the modular flow $\sigma_s$
of~\eqref{eq:modular-flow}.
For every $\mathcal{O} \in \mathcal{M}_R \otimes B(\mathcal{H}_n)$
with finite modular two-point function, the Lanczos coefficients
$b_n(\mathcal{O})$ produced by the Lanczos algorithm with respect
to $(\,\cdot\,,\,\cdot\,)_\phi$ and Liouvillian
$\mathcal{L} = [K_\phi,\,\cdot\,]$ satisfy
\begin{equation}
  b_n(\mathcal{O}) \;\le\; n\cdot\frac{\pi}{\beta} + o(n),
  \qquad n\to\infty,
  \label{eq:bn-bound}
\end{equation}
and consequently the modular Krylov complexity satisfies
\begin{equation}
  \mathcal{C}_K^{\mathrm{mod}}(s)
  \;\le\;
  A\,\cosh\!\bigl(2 b_1(\mathcal{O}) s\bigr),
  \qquad
  \lambda_K^{\mathrm{mod}}(\mathcal{O})
  \;\le\; \frac{2\pi}{\beta} \;=\; 1,
  \label{eq:bound-main}
\end{equation}
where $A>0$ depends on $\|\mathcal{O}\|_\phi$ but not on $s$.
\end{theorem}

\begin{proof}[Proof sketch (full proof in Appendix~A of an extended version)]
The argument has three steps.

\medskip
\noindent\textbf{Step 1 (Lanczos--Liouville inequality on KMS--GNS).}
The Lanczos algorithm on the KMS--GNS inner product produces a
tridiagonal representation of $\mathcal{L}$:
\[
  \mathcal{L}\,\mathcal{O}_n
  \;=\; b_{n+1}\,\mathcal{O}_{n+1} + b_n\,\mathcal{O}_{n-1}
\]
(with $a_n = 0$ for hermitian $\mathcal{L}$ on the KMS state, by
KMS reflection symmetry).
Parker et al.~\cite[\S II.B]{Parker2019} prove that for any such
tridiagonal evolution
\[
  \mathcal{C}_K(s)
  \;\le\;
  A\,\cosh(2\,b_{\max}\,s),\qquad
  b_{\max} := \sup_n \frac{b_n}{n}.
\]
This is a purely algebraic inequality from operator norms on the
tridiagonal matrix; nothing about chaos enters at this step.

\medskip
\noindent\textbf{Step 2 (Bisognano--Wichmann moment bound).}
The Lanczos coefficients $b_n$ are determined by the moments
$M_{2k} = (\mathcal{O},\mathcal{L}^{2k}\mathcal{O})_\phi$ of the
modular two-point function via the Hankel determinant formula
\[
  b_n^2
  \;=\;
  \frac{\det H_n \cdot \det H_{n-2}}{(\det H_{n-1})^2},
  \qquad
  (H_k)_{ij} = M_{2(i+j)}.
\]
On the truncation, by Remark~\ref{rem:bw-truncation} the modular
flow agrees with the boost flow on $\le \Lambda_n$ modes, and the
boost two-point function in the Bisognano--Wichmann KMS state has
the universal hyperbolic decay
\[
  C(s)
  \;=\;
  (2\sinh(\pi s/\beta))^{-\Delta}\cdot(\text{spatial factors})
\]
with $\beta = 2\pi$.
The moments of this $\sinh^{-\Delta}$ kernel have the universal
asymptotic form $M_{2k} \sim C_\Delta \,(2\pi k/\beta)^{2k}$
(standard hypergeometric expansion), which by the Hankel--Lanczos
inverse map yields exactly $b_n \le n\cdot\pi/\beta + o(n)$.

\medskip
\noindent\textbf{Step 3 (combine).}
Substituting Step~2 into Step~1 gives~\eqref{eq:bound-main}.
The Lyapunov exponent reads off the cosh growth as
$\lambda_K^{\mathrm{mod}} \le 2 \cdot (\pi/\beta) = 2\pi/\beta = 1$
in $\beta = 2\pi$ units.
\end{proof}

\begin{remark}
Theorem~\ref{thm:bound} is purely kinematic.
It uses (i) the Bisognano--Wichmann KMS structure (algebraic input)
and (ii) the Parker--Cao--Avdoshkin--Scaffidi--Altman tridiagonal
inequality (operator-norm input).
No chaos hypothesis, no holography, no large-$N$ assumption is invoked.
\end{remark}

\begin{remark}\label{rem:caputa-bridge}
For 2D CFT in the Bisognano--Wichmann KMS state, Caputa--Mag\'{a}n--Patramanis--Tonni~\cite{Caputa2024}
compute $b_n^{\mathrm{CFT}} = n\cdot\pi/\beta$ exactly,
saturating~\eqref{eq:bn-bound} as an equality.
Thus the bound is tight on the CFT class.
The full type-II$_\infty$ saturation question (do all dressed
type-II$_\infty$ algebras saturate?) is open --
see~\S\ref{sec:saturation}.
\end{remark}

\begin{corollary}[MSS bound recovery]
For the truncation $\mathcal{M}_R \otimes B(\mathcal{H}_n)$ at any
KMS inverse temperature $\beta$, the modular Lyapunov exponent
satisfies $\lambda_K^{\mathrm{mod}} \le 2\pi/\beta$,
which is the MSS rate~\eqref{eq:mss}.
\end{corollary}

\begin{proof}
Replace $\beta = 2\pi$ by general $\beta$ in the proof of
Theorem~\ref{thm:bound}; the hyperbolic moment kernel becomes
$(2\sinh(\pi s/\beta))^{-\Delta}$ and the universal Lanczos slope
becomes $\pi/\beta$, giving the MSS rate.
\end{proof}

% =====================================================================
\section{Saturation as a conjecture: free-QFT counter-examples to the
         converse}
\label{sec:saturation}
% =====================================================================

The original ECI ``modular shadow'' draft~\cite{ECI_v6} proposed the
stronger statement that saturation $\lambda_K^{\mathrm{mod}} = 2\pi/\beta$
is a structural fingerprint of type-II$_\infty$.
On audit (A11, 2026-05-05), this converse fails: there exist
non-chaotic systems that saturate the kinematic ceiling.
We collect three counter-examples and then state what survives as a
conjecture.

\paragraph{Counter-example 1 (free-scalar lattice; Avdoshkin--Dymarsky 2019).}
Avdoshkin and Dymarsky~\cite{AvdoshkinDymarsky2019} show that
operator growth under imaginary-time evolution in 1D lattice models
can attain exponential Lanczos-coefficient growth even in integrable
regimes, with a 1D superexponential frequency suppression ensuring
that ``one-dimensional systems are special''.
This means linear $b_n$ and exponential $\mathcal{C}_K$ are
\emph{not} signatures of chaos; they appear in integrable lattice
systems as well.

\paragraph{Counter-example 2 (free \& interacting massive scalar QFT;
Camargo et al.\ 2023).}
Camargo, Jahnke, Kim, and Nishida~\cite{Camargo2023} compute
Krylov complexity in free and weakly interacting massive scalar
field theories and find non-chaotic regimes where the Lanczos
coefficients exhibit the same linear envelope as in maximally
chaotic systems, with mass and UV cutoff modifying the prefactor
but preserving the kinematic form.
Their work refutes the reading that linear $b_n \to 2\pi/\beta$
implies type-II$_\infty$ MSS-saturating chaos.

\paragraph{Counter-example 3 (initial-operator dependence;
Sreeram et al.\ 2025).}
Sreeram, Kannan, Modak, and Aravinda~\cite{Sreeram2025} (PRE~112
L032203) prove that Krylov complexity depends monotonically on the
inverse participation ratio (IPR) of the initial operator in the
Hamiltonian eigenbasis, and that this dependence persists in fully
chaotic regimes.
Saturation of $\lambda_K^{\mathrm{mod}}$ is therefore not a state-
or operator-independent invariant of the algebra: it varies even
within a fixed maximally-chaotic dynamics.

\paragraph{What survives as a conjecture.}
The strong claim of v1, ``saturation $\Leftrightarrow$ type-II$_\infty$'',
is refuted by counter-examples 1--3.
What remains plausible is the weaker direction: type-II$_\infty$ at
$\beta = 2\pi$ is \emph{sufficient} for the equality
$b_n = n\pi/\beta$, even if not necessary.

\begin{conjecture}[Modular saturation for dressed type-II$_\infty$
algebras]\label{conj:msc-saturation}
\emph{[CONJECTURE; v2 downgraded.]}
Let $\mathcal{A}_{\mathrm{II}_\infty}$ be a type-II$_\infty$ crossed
product~\eqref{eq:crossed} arising from a QRF-dressing of a wedge or
static-patch algebra, with the canonical $\beta = 2\pi$ KMS state.
Then on a sequence of finite-rank truncations
$\mathcal{M}_R \otimes B(\mathcal{H}_n)$, $n \to \infty$, the
Lanczos coefficients of any few-mode operator $\mathcal{O}$ in a
dense subalgebra satisfy
\[
  b_n(\mathcal{O})
  \;\xrightarrow[n\to\infty]{}\;
  n\cdot\frac{\pi}{\beta},
\]
saturating the bound~\eqref{eq:bn-bound} and yielding
$\lambda_K^{\mathrm{mod}}(\mathcal{O}) = 2\pi/\beta$ in the
large-truncation limit.
\end{conjecture}

\begin{remark}
Conjecture~\ref{conj:msc-saturation} is consistent with the
2D-CFT computation of Caputa et al.~\cite{Caputa2024}
(Remark~\ref{rem:caputa-bridge}), with the SYK Krylov saturation
of Parker et al.~\cite{Parker2019}, and with the DSSYK / sine-dilaton
holographic checks of Heller, Papalini, and
Schuhmann~\cite{HellerPapalini2024}.
We do not propose it as a theorem; the counter-examples above
show that the converse fails, and even a one-sided proof of the
sufficient direction is beyond current methods.
\end{remark}

\paragraph{Erratum: Faulkner--Speranza identification.}
The v1 draft of this work attributed to Faulkner and
Speranza~\cite{FaulknerSperanza2024} the identification
$S_{\mathrm{gen}} \leftrightarrow \mathcal{C}_K^{\mathrm{mod}}$.
On rereading, FS24 derive the crossed-product generalised entropy
and prove the GSL via Wall monotonicity; they do \emph{not}
explicitly identify $S_{\mathrm{gen}}$ with the modular Krylov
complexity.
The bridge $S_{\mathrm{gen}} \leftrightarrow \mathcal{C}_K^{\mathrm{mod}}$
is a working conjecture of the ECI programme, not a theorem of FS24.
We thank the A11/A19 audit for catching this over-attribution.

% =====================================================================
\section{Application: a falsifier on Kolobov 2021 BEC analog Hawking
         radiation}
\label{sec:bec}
% =====================================================================

We now derive a concrete observable consequence of
Theorem~\ref{thm:bound} for the analog black-hole BEC platform of
Kolobov, Golubkov, Mu\~{n}oz~de~Nova, and Steinhauer~\cite{Kolobov2021}
(Nat.~Phys.~17, 362, 2021).

\subsection{Hawking temperature and the predicted bound}

Mu\~{n}oz~de~Nova et al.~\cite{Steinhauer2019} (Nature 569, 688, 2019)
measured the analog Hawking temperature in the BEC sonic-horizon
platform as
\begin{equation}
  T_H \;=\; 0.35 \pm 0.10\,\mathrm{nK},
\end{equation}
implying the Bisognano--Wichmann KMS rate
\begin{equation}
  \boxed{\;
  \Gamma_{\mathrm{bound}}
  \;:=\;
  \frac{2\pi k_B T_H}{\hbar}
  \;=\;
  (288 \pm 82)\,\mathrm{s}^{-1}
  \;}
  \label{eq:gamma-bound}
\end{equation}
($k_B/\hbar = 1.31 \times 10^{11}\,\mathrm{rad}\,\mathrm{s}^{-1}\,\mathrm{K}^{-1}$).

By Theorem~\ref{thm:bound} applied to the (assumed) finite-rank
type-II$_\infty$ truncation of the exterior phonon algebra at the
analog horizon, any temporal correlator of exterior phonons must
have its envelope decaying with rate $\le \Gamma_{\mathrm{bound}}$:
\begin{equation}
  \Gamma_{\mathrm{meas}}
  \;\le\;
  \Gamma_{\mathrm{bound}}
  \;=\;
  (288 \pm 82)\,\mathrm{s}^{-1},
  \label{eq:gamma-pred}
\end{equation}
where $\Gamma_{\mathrm{meas}}$ is the exponential-decay rate of any
of: (i) the second-order coherence $g^{(2)}(\tau)$; (ii) the
two-point density correlator $G^{(2)}(z_1,z_2,t)$ envelope;
(iii) the Hawking-flux ramp-up rate.

\subsection{Why this is a clean falsifier}

The Kolobov~2021 dataset~\cite{Kolobov2021} contains time-resolved
density correlations at six temporal points across a 124-day,
97{,}000-repetition acquisition: the temporal envelope of
$G^{(2)}(z_1,z_2,t)$ can be extracted directly from the existing
data without new experimental runs.

\begin{proposition}[BEC bound falsifier]
If a re-analysis of Kolobov~2021 data yields
$\Gamma_{\mathrm{meas}} > 1.5 \times \Gamma_{\mathrm{bound}}
= 432\,\mathrm{s}^{-1}$ at $5\sigma$, then the type-II$_\infty$
finite-rank scaffolding underlying Theorem~\ref{thm:bound} is
falsified for the analog-BEC exterior phonon algebra.
\end{proposition}

\paragraph{Robustness to integrability counter-examples.}
Crucially, the bound~\eqref{eq:gamma-pred} is the BOUND form,
not the saturation form: it is therefore robust to the free-QFT /
integrability counter-examples
of~\S\ref{sec:saturation}~\cite{AvdoshkinDymarsky2019,Camargo2023,Sreeram2025}.
Even if the BEC phonon system is effectively integrable on the
relevant timescales, the bound still holds; only an observed
\emph{violation} would refute the type-II$_\infty$ kinematic
prefactor.

\paragraph{Cross-platform redundancy.}
Independent platforms can cross-check~\eqref{eq:gamma-pred}:
polariton-condensate analog black-hole mergers
(Solnyshkov--Paquelier--Balmisse--Malpuech 2026,
arXiv:2603.01664~\cite{Solnyshkov2026}) and water-tank vortex QNM
decay rates~\cite{TorresPatrick2017} provide independent substrates.
A UV-finite volume-law entanglement negativity
(Chandran--Fischer 2026, arXiv:2604.02075~\cite{ChandranFischer2026})
likewise yields an entanglement-growth rate bounded
by~\eqref{eq:gamma-pred} via Theorem~\ref{thm:bound}.

% =====================================================================
\section{Discussion: comparison with Vardian (2026) and scope}
\label{sec:discussion}
% =====================================================================

\paragraph{Vardian's concurrent boundary-area-operator construction.}
Vardian~\cite{Vardian2026} (arXiv:2602.02675, February 2026,
single-author hep-th) reconstructs the QES area operator from
boundary modular Lanczos coefficients in AdS/CFT, using
operator-algebra quantum error correction.
Specifically: from boundary modular dynamics,
the spectrum of the boundary modular Hamiltonian is extracted via
the Lanczos algorithm; its central component is identified with the
area operator of the dual quantum extremal surface.
Applications discussed include diagnosing island formation
and the Page transition for evaporating black holes.

\paragraph{Comparison: complementary scopes.}
Vardian's work and ours overlap in their use of modular Lanczos
coefficients on a type-II$_\infty$-like algebra, but the scopes
differ along three axes:
\begin{itemize}
\item \textbf{Holographic context.}
Vardian works in AdS/CFT and uses holographic boundary--bulk duality
explicitly; we work intrinsically in the finite-rank truncation
without invoking AdS/CFT and our results apply to any wedge or
static-patch algebra (de~Sitter, Rindler, BEC analog horizon).
\item \textbf{Object of study.}
Vardian targets the area operator (a specific element of the
modular Hamiltonian's spectrum); we target the kinematic Lyapunov
ceiling on Krylov growth (an envelope statement on all operators).
\item \textbf{Claim type.}
Vardian provides a constructive boundary--bulk dictionary;
we provide a one-sided inequality with concrete falsifier on a
laboratory analog system.
\end{itemize}
The two works together suggest that modular Krylov techniques are
becoming a standard tool for type-II$_\infty$ algebras, with
Vardian probing the holographic side and our bound theorem
covering the non-holographic side (de~Sitter, BEC, and generic
truncations).

\paragraph{Scope of Theorem~\ref{thm:bound}.}
Our theorem applies to any finite-rank truncation of a type-II$_\infty$
crossed product at canonical KMS temperature.
This covers:
\begin{enumerate}
\item Rindler wedge of a free QFT (where the bound is tight by
Caputa et al.~\cite{Caputa2024});
\item BTZ exterior in $\mathrm{AdS}_3$ (the holographic paradigm);
\item de~Sitter static patch~\cite{CLPW2023}, with $\beta = 2\pi/H$;
\item BEC and other analog-Hawking systems with a sonic horizon
(Kolobov 2021, Solnyshkov 2026);
\item the SYK / DSSYK / sine-dilaton family at large
$N$~\cite{Parker2019,HellerPapalini2024}.
\end{enumerate}
It does \emph{not} apply directly to:
\begin{enumerate}
\item Type-III$_1$ algebras at finite $N$ without a QRF dressing
(no semifinite trace, modular flow ergodic);
\item Cosmological algebras (Bianchi~IX, FRW) where the type
classification is open;
\item Type-I or II$_1$ models without boost geometry (no
Bisognano--Wichmann input).
\end{enumerate}

\paragraph{Modular vs physical Lyapunov exponents.}
The bound~\eqref{eq:bound-main} is on the modular Lyapunov exponent
$\lambda_K^{\mathrm{mod}}$, defined via modular flow.
Translating to the physical Lyapunov exponent
$\lambda_L^{\mathrm{OTOC}}$ requires the physical Hamiltonian to
align with the modular Hamiltonian under the
Connes--Rovelli~\cite{ConnesRovelli1994} thermal-time hypothesis.
For boost-symmetric wedges this alignment is automatic
(Bisognano--Wichmann); for general dynamical settings it is an
additional input.

% =====================================================================
\section{Outlook}
\label{sec:outlook}
% =====================================================================

\begin{enumerate}
\item \textbf{Lifting the truncation.}
Theorem~\ref{thm:bound} is stated on the finite-rank truncation
$\mathcal{M}_R \otimes B(\mathcal{H}_n)$.
A genuine type-II$_\infty$ statement requires a uniform-in-$n$
control of $b_n$ and a passage to the inductive-limit algebra.
The Hankel-determinant moment estimate of Step~2 in our proof
sketch is uniform in $n$ on the BW class, but the operator
$\mathcal{O}$ must lie in a dense $\ast$-subalgebra of the
type-II$_\infty$ parent.
We expect the lifted statement to hold; a complete proof is
left for future work.

\item \textbf{de~Sitter static patch.}
CLPW~\cite{CLPW2023} establish the type-II$_\infty$ structure
of the dS static patch with $\beta = 2\pi/H$ (Gibbons--Hawking
temperature).
By Theorem~\ref{thm:bound}, the modular Lyapunov exponent of
any few-mode operator on the dS static patch is bounded by
$H$ (the Hubble rate).
A concrete observable (e.g.\ a primordial-tensor-mode decoherence
rate) testing this bound would be of cosmological interest.

\item \textbf{Bianchi~IX and homogeneous chaotic cosmologies.}
The mixmaster Bianchi~IX universe exhibits chaotic dynamics with
a known modular structure on the kasner-segment wedge algebra.
Whether this algebra is type-II$_\infty$ and whether its modular
Krylov exponent saturates~\eqref{eq:bound-main} is open.

\item \textbf{Connes--St\o{}rmer entropy.}
The Connes--St\o{}rmer entropy on type-II$_1$ reductions of our
truncation provides an alternative diagnostic
(see Connes--Rovelli~\cite{ConnesRovelli1994}); whether it
inherits the same kinematic ceiling is an open question.

\item \textbf{Bridge $S_{\mathrm{gen}} \leftrightarrow
\mathcal{C}_K^{\mathrm{mod}}$.}
The bridge between the FS24 generalised entropy and modular
Krylov complexity remains a working conjecture of the ECI
programme.
A rigorous derivation would close the loop between
Theorem~\ref{thm:bound} and the Faulkner--Speranza GSL.

\item \textbf{Relation to Vardian's area-operator reconstruction.}
The boundary area operator of Vardian~\cite{Vardian2026} sits in
the central part of the modular Hamiltonian.
Its Krylov-Lanczos signature should sit at a specific level of
our truncation; making this precise would unify the two approaches.
\end{enumerate}

% =====================================================================
\section*{Acknowledgments}
% =====================================================================

This work was extracted and reworked from the ECI framework
(v6.0.53, Zenodo record \texttt{10.5281/zenodo.20030684}).
The split between the (provable) bound and the (conjectural)
saturation, and the identification of the Avdoshkin--Dymarsky /
Camargo et al.\ / Sreeram et al.\ counter-examples, owe to the A11
audit (2026-05-05).
The discovery of Vardian~\cite{Vardian2026} as concurrent
independent work was made via the same audit, with arXiv ID
2602.02675 live-verified.
Erratum on the Faulkner--Speranza over-attribution (\S\ref{sec:saturation})
follows the A11 cross-check.
We thank [collaborators] for discussions.

% =====================================================================
\begin{thebibliography}{99}
% =====================================================================

\bibitem{MSS2016}
J.~Maldacena, S.~H.~Shenker, and D.~Stanford,
``A bound on chaos,''
\textit{JHEP} \textbf{08} (2016) 106,
\href{https://arxiv.org/abs/1503.01409}{arXiv:1503.01409 [hep-th]}.

\bibitem{Parker2019}
D.~E.~Parker, X.~Cao, A.~Avdoshkin, T.~Scaffidi, and E.~Altman,
``A Universal Operator Growth Hypothesis,''
\textit{Phys.\ Rev.\ X} \textbf{9} (2019) 041017,
\href{https://arxiv.org/abs/1812.08657}{arXiv:1812.08657 [cond-mat.str-el]}.

\bibitem{Caputa2024}
P.~Caputa, J.~M.~Mag\'{a}n, D.~Patramanis, and E.~Tonni,
``Krylov complexity of modular Hamiltonian evolution,''
\textit{Phys.\ Rev.\ D} \textbf{109} (2024) 086004,
\href{https://arxiv.org/abs/2306.14732}{arXiv:2306.14732 [hep-th]}.

\bibitem{Witten2022}
E.~Witten,
``Gravity and the crossed product,''
\textit{JHEP} \textbf{10} (2022) 008,
\href{https://arxiv.org/abs/2112.12828}{arXiv:2112.12828 [hep-th]}.

\bibitem{CLPW2023}
V.~Chandrasekaran, R.~Longo, G.~Penington, and E.~Witten,
``An algebra of observables for de~Sitter space,''
\textit{JHEP} \textbf{02} (2023) 082,
\href{https://arxiv.org/abs/2206.10780}{arXiv:2206.10780 [hep-th]}.

\bibitem{FaulknerSperanza2024}
T.~Faulkner and A.~J.~Speranza,
``Gravitational algebras and the generalized second law,''
\textit{JHEP} \textbf{11} (2024) 099,
\href{https://arxiv.org/abs/2405.00847}{arXiv:2405.00847 [hep-th]}.
%% v2 erratum: FS24 derives crossed-product S_gen and the GSL via Wall monotonicity;
%% it does NOT explicitly identify S_gen with modular Krylov complexity.
%% The bridge S_gen ↔ C_K^mod is an ECI working conjecture, not a FS24 theorem.

\bibitem{DEHK2025a}
J.~De~Vuyst, S.~Eccles, P.~A.~H\"{o}hn, and J.~Kirklin,
``Crossed products and quantum reference frames: on the observer-dependence
of gravitational entropy,''
\textit{JHEP} \textbf{07} (2025) 063,
\href{https://arxiv.org/abs/2412.15502}{arXiv:2412.15502 [hep-th]}.

\bibitem{DEHK2025b}
J.~De~Vuyst, S.~Eccles, P.~A.~H\"{o}hn, and J.~Kirklin,
``Gravitational entropy is observer-dependent,''
\textit{JHEP} \textbf{07} (2025) 146,
\href{https://arxiv.org/abs/2405.00114}{arXiv:2405.00114 [hep-th]}.

\bibitem{HellerPapalini2024}
M.~P.~Heller, A.~Papalini, and T.~Schuhmann,
``Krylov spread complexity as holographic complexity beyond JT gravity,''
\textit{Phys.\ Rev.\ Lett.}\ \textbf{135} (2025) 151602,
\href{https://arxiv.org/abs/2412.17785}{arXiv:2412.17785 [hep-th]}.

\bibitem{BisognanoWichmann1976}
J.~J.~Bisognano and E.~H.~Wichmann,
``On the duality condition for a Hermitian scalar field,''
\textit{J.\ Math.\ Phys.}\ \textbf{17} (1976) 303.

\bibitem{ConnesRovelli1994}
A.~Connes and C.~Rovelli,
``Von Neumann algebra automorphisms and time-thermodynamics relation in
generally covariant quantum theories,''
\textit{Class.\ Quantum Grav.}\ \textbf{11} (1994) 2899--2917.

%% --- Vardian (concurrent independent work, must cite) ---
\bibitem{Vardian2026}
N.~Vardian,
``Modular Krylov Complexity as a Boundary Probe of Area Operator and
Entanglement Islands,''
\href{https://arxiv.org/abs/2602.02675}{arXiv:2602.02675 [hep-th]}
(February 2026).
%% Single-author hep-th. Live-verified 2026-05-05.

%% --- Free-QFT counter-examples to the saturation converse ---
\bibitem{AvdoshkinDymarsky2019}
A.~Avdoshkin and A.~Dymarsky,
``Euclidean operator growth and quantum chaos,''
\textit{Phys.\ Rev.\ Research} \textbf{2} (2020) 043234,
\href{https://arxiv.org/abs/1911.09672}{arXiv:1911.09672 [cond-mat.stat-mech]}.

\bibitem{Camargo2023}
H.~A.~Camargo, V.~Jahnke, K.-Y.~Kim, and M.~Nishida,
``Krylov Complexity in Free and Interacting Scalar Field Theories with
Bounded Power Spectrum,''
\textit{JHEP} \textbf{05} (2023) 226,
\href{https://arxiv.org/abs/2212.14702}{arXiv:2212.14702 [hep-th]}.

\bibitem{Sreeram2025}
Sreeram~PG, J.~Bharathi~Kannan, R.~Modak, and S.~Aravinda,
``Dependence of Krylov complexity on the initial operator and state,''
\textit{Phys.\ Rev.\ E} \textbf{112} (2025) L032203,
\href{https://arxiv.org/abs/2503.03400}{arXiv:2503.03400 [quant-ph]}.

%% --- BEC analog Hawking platforms (A19 BOUND falsifier) ---
\bibitem{Kolobov2021}
V.~I.~Kolobov, K.~Golubkov, J.~R.~Mu\~{n}oz~de~Nova, and J.~Steinhauer,
``Observation of stationary spontaneous Hawking radiation and the time
evolution of an analogue black hole,''
\textit{Nat.\ Phys.}\ \textbf{17} (2021) 362--367,
\href{https://arxiv.org/abs/1910.09363}{arXiv:1910.09363 [gr-qc]}.

\bibitem{Steinhauer2019}
J.~R.~Mu\~{n}oz~de~Nova, K.~Golubkov, V.~I.~Kolobov, and J.~Steinhauer,
``Observation of thermal Hawking radiation at the Hawking temperature in
an analogue black hole,''
\textit{Nature} \textbf{569} (2019) 688--691,
\href{https://arxiv.org/abs/1809.00913}{arXiv:1809.00913 [gr-qc]}.

\bibitem{Solnyshkov2026}
D.~D.~Solnyshkov, I.~Paquelier, A.~Balmisse, and G.~Malpuech,
``Polariton-condensate analog black-hole merger'' (working title),
\href{https://arxiv.org/abs/2603.01664}{arXiv:2603.01664}
(2026).

\bibitem{ChandranFischer2026}
A.~Chandran and U.~R.~Fischer,
``UV-finite volume-law entanglement negativity'' (working title),
\href{https://arxiv.org/abs/2604.02075}{arXiv:2604.02075}
(2026).

\bibitem{TorresPatrick2017}
T.~Torres, S.~Patrick, A.~Coutant, M.~Richartz, E.~W.~Tedford, and S.~Weinfurtner,
``Rotational superradiant scattering in a vortex flow,''
\textit{Nat.\ Phys.}\ \textbf{13} (2017) 833--836.

%% --- ECI internal references ---
\bibitem{ECI_v6}
[ECI author(s)],
``Entanglement--Complexity--Information: a phenomenological architecture
for quantum gravity phenomenology (v6.0.53),''
Zenodo record \texttt{10.5281/zenodo.20030684} (2026).

\end{thebibliography}

\end{document}
