\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{microtype}

\newtheorem{hypothesis}{Hypothesis}
\theoremstyle{remark}
\newtheorem{motivation}{Motivation}

\title{Chim\`ere $\Omega$: a blueprint for a physico-cognitively inspired
local-first LLM runtime with entropic mode scheduling}

\author{Kevin Remondi\`ere\thanks{kevin.remondiere@gmail.com}\\
\small Independent Researcher, ORCID 0009-0008-2443-7166}

\date{2026-04-22}

\begin{document}
\maketitle

\begin{abstract}
We propose \emph{Chim\`ere $\Omega$}, a blueprint for a local-first LLM
runtime that allocates computation across four modes --- fast
prediction, block-diffusion refinement, symbolic verification,
latent-space planning --- via a scheduler driven by token-level
residual entropy rather than fixed architectural choices. The
blueprint integrates verified Q4~2025--Q2~2026 state-of-the-art:
LeJEPA/SIGReg~(Balestriero \& LeCun, arXiv:2511.08544), hybrid
Gated-DeltaNet + sparse-attention backbones (Qwen3-Next,
Kimi-Linear), test-time memory optimisation
(Titans~arXiv:2501.00663 $\to$ Nested Learning / Hope~arXiv:2512.24695),
block-diffusion--AR fusion (BD3-LMs~arXiv:2503.09573),
auxiliary-loss-free MoE routing~(arXiv:2408.15664), and the vLLM~V1
production stack. We then propose three design hypotheses, framed
explicitly as hypotheses to be tested rather than theorems: a
\emph{free-energy-under-censorship} training objective, a
\emph{renormalisation-group view of fine-grained MoE routing}, and an
\emph{A6-persistent, A3-censored memory-write policy} that ties
episodic consolidation to symbolic verification. The physics
inspiration from the author's earlier ECI framework
(entanglement--complexity--information,
v5~phenomenological and v6~formal companions) is used as analogical
scaffolding, not mathematical derivation; this distinction is
enforced throughout. We conclude with a 24-month phased roadmap
whose first phase is immediately executable on consumer hardware
(RTX~5060~Ti, 16~GB~VRAM) via the author's existing
\texttt{chimere-server} Rust FFI runtime and \texttt{ik\_llama.cpp}
backend.
\end{abstract}

\section{Introduction and scope}
\label{sec:intro}

The public debate on large-model design through late 2025 has often
been framed as a dichotomy: \emph{predict} (autoregressive next-token,
exploiting pre-training redundancy) versus \emph{compute}
(test-time-compute, refinement, program synthesis). We argue this
dichotomy is a false opposition. Both are projections of the same
underlying quantity, the \emph{information bits allocated per output
token}, and the useful design question is how to \emph{modulate} this
allocation as a function of local residual entropy.

This paper proposes an explicit blueprint, \emph{Chim\`ere $\Omega$},
for a runtime that schedules between four modes:
\begin{description}[leftmargin=*,itemsep=3pt]
 \item[M1.] \textbf{Fast prediction} ($\sim 5$\,ms/token) when the
   per-token residual entropy $H(r_t)$ is low (e.g.\ $<0.5$ bit): a
   Gated-DeltaNet dominated backbone with MoE top-$K$ selection.
 \item[M2.] \textbf{Block-diffusion refinement}
   ($\sim 30$\,ms/block) when $H(r_t) \in [2,4]$ bits: BD3-style
   diffusion over fixed-size blocks with preserved KV cache.
 \item[M3.] \textbf{Symbolic verification} ($\sim 200$\,ms/claim) when
   the candidate output is formally testable: a Scallop-lite
   provenance-semiring check, or Lean-style proof verification.
 \item[M4.] \textbf{Latent-space planning} (world-model rollout) for
   action-conditioned tasks: V-JEPA-style predictor operating on
   frozen embeddings.
\end{description}

The contribution of this paper is not a new theorem or a new
fundamental constant. It is three distinct things, clearly
delineated: \textbf{(i)} a curated landscape of verified Q4~2025--Q2~2026
state-of-the-art with citations cross-checked against arXiv and
publisher records (\S\ref{sec:sota}); \textbf{(ii)} an explicit
architecture integrating those components with an entropy-driven
scheduler (\S\ref{sec:blueprint}); \textbf{(iii)} three \emph{design
hypotheses} framed as falsifiable statements, inspired by the
author's physics-motivated ECI framework but presented as analogical
scaffolding, not mathematical derivation (\S\ref{sec:equations}).

\paragraph{What this paper does not claim.}
We do not claim: a new equation of physics, a new fundamental
constant, a mathematical identification between physics axioms (ECI
A3 censorship, A6 persistence) and machine-learning
loss terms, nor a theorem that the proposed runtime attains a
particular benchmark score. The relation to ECI is motivational;
the ECI v5 and v6 companion papers live at the cosmology end of the
project, and a rigorous mathematical bridge to the machine-learning
objectives of this paper is \emph{not} established in either
direction.

\section{The verified 2025--2026 landscape}
\label{sec:sota}

Five programmes dominate the Q4~2025--Q2~2026 state of the art and
together define the engineering envelope for a local-first LLM
runtime.

\subsection{Self-supervised representation learning: the JEPA line}
\label{sec:jepa}

LeJEPA (Balestriero and LeCun, arXiv:2511.08544, November 2025)
\cite{LeJEPA2025} proves that the optimal distribution on joint
embedding predictor architecture (JEPA) representations is the
isotropic Gaussian $\mathcal{N}(0, I_d)$ and proposes
\emph{Sketched Isotropic Gaussian Regularisation} (SIGReg) as a
linear-cost surrogate:
\begin{equation}
 \mathcal{L}_{\mathrm{SIGReg}} = \frac{1}{M} \sum_{m=1}^{M}
 \mathcal{D}\!\left( \hat\varphi_{z_m}(t),\, e^{-t^2/2} \right),
 \qquad z_m = \langle u_m, s_\phi(x) \rangle, \;
 u_m \sim \mathrm{Unif}(S^{d-1})
 \label{eq:sigreg}
\end{equation}
where $\hat\varphi$ is the empirical characteristic function and
$\mathcal{D}$ a distance to the characteristic function of
$\mathcal{N}(0,1)$. By the Cram\'er--Wold theorem, matching all
one-dimensional projections to $\mathcal{N}(0,1)$ forces isotropy of
the $d$-dimensional distribution. This removes the heuristic
stack (stop-gradient, EMA teacher, asymmetric predictor) that had
characterised earlier JEPA variants.

\subsection{Hybrid linear + attention backbones}
\label{sec:backbone}

The consensus architectural pattern through Q2~2026 is a
three-to-one ratio of linear-attention layers (Gated DeltaNet
\cite{GatedDeltaNet2025}, Kimi Delta Attention
\cite{KimiLinear2025}, Mamba-family SSMs) interleaved with sparse
or full-attention layers. The Gated-DeltaNet update rule is an
online delta-descent step,
\begin{equation}
 S_t = \alpha_t \left( S_{t-1} - \beta_t (S_{t-1} k_t) k_t^{\top} \right)
 + \beta_t\, v_t k_t^{\top}
 \label{eq:gdn}
\end{equation}
with $\alpha_t \in (0,1)$ a SiLU-gated forget factor and $\beta_t$ a
token-level learning rate; Kimi Delta Attention generalises
$\alpha_t$ to a channel-wise vector $\alpha_t \in \mathbb{R}^d$
combined with a diagonal-plus-low-rank transition matrix. Mamba-3
(arXiv:2603.15569) extends the state to $h_t \in \mathbb{C}^d$ with a
data-dependent rotation $A_t = \mathrm{diag}(e^{i\theta_t} e^{-\alpha_t})$,
which lifts the $\mathsf{TC}^0$ expressivity bound of real-valued
SSMs.

\subsection{Memory: Titans $\to$ Nested Learning / Hope}
\label{sec:memory}

Titans \cite{Titans2024} formulates long-term memory as a test-time
online optimisation with momentum and forgetting:
\begin{equation}
 S_t = \eta_t\, S_{t-1} - \theta_t\, \nabla \ell(M_{t-1}; k_t, v_t),
 \quad M_t = (1-\alpha_t)\, M_{t-1} + S_t
 \label{eq:titans}
\end{equation}
with associative loss $\ell(M; k, v) = \|M(k) - v\|^2$. Nested
Learning / Hope (Behrouz et al.,
NeurIPS~2025)~\cite{NestedLearning2025} generalises this: a model is
a stack of nested optimisation problems, each level $\ell$
minimising a loss $\mathcal{L}_\ell(\theta_\ell; C_\ell)$ where
$\theta_\ell$ parameterises the optimiser of level $\ell - 1$. The
Hope reference implementation attains lower perplexity than
Mamba-2 and standard transformers on long-context benchmarks and is
explicitly designed with multiple-frequency update rates for
continual learning.

\subsection{Generation: block-diffusion, not pure diffusion}
\label{sec:diffusion}

Block Diffusion (BD3-LMs, Arriola et al., ICLR~2025
Oral)~\cite{BlockDiffusion2025} decomposes a sequence of length
$L = B \cdot b$ into $B$ blocks; intra-block diffusion is masked,
inter-block factorisation is autoregressive,
\begin{equation}
 \log p_\theta(x) \;\ge\;
 \sum_{b=1}^{B} \mathcal{L}_{\mathrm{diffusion}}
 \!\left( x^{b},\, x^{<b},\, \theta \right).
 \label{eq:bd3}
\end{equation}
The critical practical property is that BD3 preserves the KV cache
across blocks, which pure discrete-diffusion cannot. The 2026
production deployments (LLaDA~2.0, Mercury~2) confirm this pattern
dominates the diffusion-LM frontier.

\subsection{MoE: ultra-sparse with auxiliary-loss-free balancing}
\label{sec:moe}

The auxiliary-loss-free load balancing of Wang et
al.~\cite{AuxLossFreeMoE2024}, deployed in DeepSeek-V3
\cite{DeepSeekV3}, uses an additive bias $b_i$ that enters the
top-$K$ selection but not the gating value,
\begin{equation}
 \tilde s_{i,t} = s_{i,t} + b_i,
 \quad g_i = \mathbb{1}\!\left[ \tilde s_{i,t} \in \mathrm{TopK} \right],
 \quad b_i \leftarrow b_i - \gamma\,
   \mathrm{sign}(\mathrm{load}_i - \overline{\mathrm{load}}).
 \label{eq:auxfree}
\end{equation}
The update is applied with $\gamma = 10^{-3}$ during early
pre-training and annealed to zero. Kimi~K2
\cite{KimiK2} uses 384 routed experts, the highest
publicly-documented count in an open-weights model as of
April~2026.

\subsection{Runtime stack and quantisation}
\label{sec:runtime}

vLLM~V1 unifies prefill--decode scheduling with PagedAttention at
16-token block granularity and first-class hybrid-model support
(Qwen3-Next, Granite-4.0). EAGLE-3 (arXiv:2503.01840)
\cite{EAGLE3} provides $3\!-\!6\!\times$ speculative-decoding
speedup over autoregressive baselines via multi-layer feature
fusion. For quantisation, SpinQuant \cite{SpinQuant2024} learns
Hadamard rotations via Cayley parameterisation on the Stiefel
manifold to close the W4A4KV4 gap to $\sim 3$~points vs.\ FP16.
KIVI \cite{KIVI2024} gives $2$-bit K (per-channel) / $2$-bit V
(per-token) KV-cache compression with $\sim 2.6\times$ memory
reduction at negligible perplexity cost.

\section{Chim\`ere $\Omega$ blueprint}
\label{sec:blueprint}

We now specify the blueprint, organised around three dualities and
an entropy-driven scheduler.

\subsection{Three dualities}
\label{sec:dualities}

\begin{description}[leftmargin=*,itemsep=3pt]
 \item[D1. Predict vs.\ compute.] A token with low residual entropy
   is optimally served by a fast M1 (prediction) pass; a token with
   high residual entropy warrants M2 (refinement) or M3
   (verification). The scheduling variable is $H(r_t)$.
 \item[D2. Memorise vs.\ compress.] Following IBLM \cite{IBLM2025},
   learning alternates between a memorisation phase (capacity
   expansion) and a compression phase (entropy reduction); we
   propose that \emph{inference} should likewise alternate via an
   episodic write/consolidate policy.
 \item[D3. Local vs.\ global.] Fine-grained MoE routing is a
   coarse-graining step: as depth increases, low-level details are
   integrated out toward semantic features. We suggest
   (Hypothesis~\ref{hyp:rg}) that this be formalised as a
   discrete-layer renormalisation-group flow.
\end{description}

\subsection{Architecture}
\label{sec:arch}

A 96-layer backbone interleaving GDN/KDA layers (75\%) with sparse
full-attention (15\%) and dense layers (10\%), fine-grained MoE
(512 routed experts + 2 shared + 4 proposed \emph{censor} experts)
with auxiliary-loss-free balancing, multi-head output (MTP~4 for AR
generation, BD3 for refinement, JEPA predictor for planning,
verification head for symbolic checks), and a 4-tier memory
hierarchy (L1 KV cache, L2 Titans LMM, L3 FHRR hash table, L4
EM-LLM episodic segmentation).

\paragraph{Censor experts.}
Four MoE experts are reserved for tokens with high residual
entropy. Their selection score adds a residual-entropy term:
\begin{equation}
 s_{i,t} = \mathrm{sigmoid}(u_t^{\top} e_i) + b_i
 + \mathbb{1}_{\{i \in \mathrm{censor}\}} \cdot \kappa \cdot H(r_t).
 \label{eq:censor}
\end{equation}
This is a design choice, not a theorem: it is engineered to
concentrate out-of-distribution tokens into experts whose weights
can carry additional regularisation (spectral-norm constraints,
narrower weight initialisation). Whether this improves
out-of-distribution behaviour in practice is a testable prediction.

\subsection{Entropy-driven scheduler}
\label{sec:scheduler}

The runtime computes $H(r_t) = -\sum_v p_\theta(v|\cdot)
\log p_\theta(v|\cdot)$ (or a top-$k$ approximation) per token and
dispatches:
\begin{center}
\small
\begin{tabular}{lll}
\toprule
$H(r_t)$ range & Mode(s) activated & Wall-clock target \\
\midrule
$< 0.5$ bit              & M1 only                       & $5$\,ms/token \\
$[0.5, 2)$ bits          & M1 $+$ EAGLE-3 speculative    & $10$\,ms/token \\
$[2, 4)$ bits            & M2 (BD3 block refinement)     & $30$\,ms/block \\
$\ge 4$ bits             & M2 $+$ M3 $+$ L3 retrieval    & $200+$\,ms/claim \\
\bottomrule
\end{tabular}
\end{center}
The thresholds are engineering proposals to be calibrated
empirically; the design principle is that \emph{marginal expected
information gain of adding mode~$N+1$ must exceed its wall-clock
cost}. This is implementable today with the
\texttt{chimere-server} runtime FFI-ed to \texttt{ik\_llama.cpp};
no new training is required to deploy the M1/M2 split.

\section{Three design hypotheses}
\label{sec:equations}

The following three expressions are framed as
\emph{hypotheses}, not theorems. Each inherits motivation from the
author's ECI framework axioms (A3 censorship-like suppression; A6
persistence-like consolidation); the mapping from cosmological
axioms to machine-learning objectives is \emph{analogical} and is
not derived. Each hypothesis yields a concrete experimental
programme with a binary outcome.

\subsection{H1 --- Free energy under censorship (FEC)}

\begin{hypothesis}[Free energy under censorship, FEC]
\label{hyp:fec}
Training a Chim\`ere $\Omega$ model with the modified objective
\begin{equation}
 \mathcal{L}_{\mathrm{FEC}} =
 \underbrace{\mathbb{E}_q[-\log p_\theta(x)]}_{\mathrm{NTP}}
 + \beta\, I_{\mathrm{SAE}}(Z_{\mathrm{free}}; X)
 - \gamma\, \langle V(M) \rangle
 + \lambda\, \mathcal{L}_{\mathrm{SIGReg}}(Z)
 \label{eq:FEC}
\end{equation}
where $I_{\mathrm{SAE}}$ is the mutual information between the
subset of SAE features not causally related to the prediction target
and the input, $V$ is a learned scalar functional on memory states
$M$, and $\mathcal{L}_{\mathrm{SIGReg}}$ is
Eq.~\eqref{eq:sigreg}, yields equal or lower bits-per-character on
out-of-distribution evaluation sets at matched FLOPs versus the
NTP-only baseline.
\end{hypothesis}

\begin{motivation}
The ECI framework suggests ``censoring'' information that cannot be
predicted, via an activation function $\Theta$ in the cosmological
horizon setting. The analogue proposed here is to penalise SAE
features that carry no prediction signal, weighted by
$\beta$. The memory-potential term $\gamma \langle V(M) \rangle$ is
the ML analogue of the persistence axiom~A6 in ECI, rewarding
long-lived informative memory content. The SIGReg term closes the
isotropy condition on latent representations.
\end{motivation}

\paragraph{Falsification.}
H1 is falsified if, after hyperparameter tuning of $(\beta, \gamma,
\lambda)$, the FEC objective fails to beat matched-FLOPs NTP on
at least one of: BPC on a held-out domain, Arc-AGI-2 single-sample
accuracy, or RULER at $128$k context. A negative result would not
invalidate the ECI framework (which lives in a different domain) but
would falsify the design analogy.

\subsection{H2 --- MoE routing as a discrete renormalisation-group
flow}

\begin{hypothesis}[RG-flow MoE]
\label{hyp:rg}
Let $b_i^{(\ell)}$ denote the load-balancing bias of expert $i$ at
layer $\ell$. If the interlayer update is augmented by a
coarse-graining operator $\mathcal{R}[\cdot]$ that maps
fine-grained biases at layer $\ell-1$ to effective biases at layer
$\ell$,
\begin{equation}
 \frac{d b_i^{(\ell)}}{d\ell} = -\gamma\,
  \mathrm{sign}\!\left( \mathrm{load}_i^{(\ell)} -
  \overline{\mathrm{load}}^{(\ell)} \right)
 + \mu\, \mathcal{R}\!\left[ b_i^{(\ell-1)} \right],
 \label{eq:RG}
\end{equation}
then training stability improves and the emergent expert
specialisation follows a layer-dependent power law $\kappa^{(\ell)}
\propto \ell^{-\alpha}$.
\end{hypothesis}

\begin{motivation}
Fine-grained MoE segmentation plus auxiliary-loss-free balancing is
already a form of layer-wise reorganisation. Interpreting it as a
Kadanoff-like block-spin map suggests that the bias update should
carry information between layers rather than be reset per layer.
\end{motivation}

\paragraph{Falsification.}
H2 is falsified if, on a matched-FLOPs pretraining run, the added
$\mathcal{R}$ term neither reduces expert-imbalance variance nor
produces a consistent power-law in $\kappa^{(\ell)}$.

\subsection{H3 --- Engram write policy with A6-persistence and
A3-censorship}

\begin{hypothesis}[Engram write policy]
\label{hyp:engram}
An episodic-memory write policy of the form
\begin{equation}
 \pi_{\mathrm{write}}(e_t \mid h_t) = \sigma\!\big(
 w_0 + w_1\, \mathrm{Surprise}_B(h_t) + w_2\, V_{\mathrm{verify}}(h_t)
 - w_3\, \mathcal{C}_{\mathrm{A3}}(h_t) \big)
 \label{eq:engram}
\end{equation}
with $\mathrm{Surprise}_B$ the EM-LLM bayesian surprise,
$V_{\mathrm{verify}}$ a symbolic-check score, and
$\mathcal{C}_{\mathrm{A3}}$ a hallucination-risk score from a trained
SAE, outperforms a threshold-only write policy on long-horizon
factual consistency benchmarks.
\end{hypothesis}

\begin{motivation}
A memory that writes every surprising event risks consolidating
hallucinations; a memory that writes nothing but verified claims
misses creative long-horizon coherence. H3 proposes a trainable
trade-off with four factors, two of which
($V_{\mathrm{verify}}$, $\mathcal{C}_{\mathrm{A3}}$) are concrete
proxies for the ECI A3 censorship axiom, and one of which
($\mathrm{Surprise}_B$) is a direct instantiation of the A6
persistence condition restricted to surprising events.
\end{motivation}

\paragraph{Falsification.}
H3 is falsified if an ablation that removes either
$V_{\mathrm{verify}}$ or $\mathcal{C}_{\mathrm{A3}}$ yields equal
benchmark performance.

\subsection{Honest note on ECI as scaffolding}

All three hypotheses use the ECI~v3 axioms (A3, A6) as
\emph{linguistic and architectural scaffolding}. The rigorous ECI
v5~(phenomenological) and v6~(formal) companion papers
inhabit a wholly different scientific domain (cosmology and type-II
crossed-product algebras). An explicit attempt at a mathematical
bridge from the ECI Jacobson--EGJ internal production
($d_i S$) to a machine-learning source term failed under a
dedicated audit (see~\cite{ECIv6Audit}, registered in the
author's \texttt{FAILED.md} as entry~F-14). The ML hypotheses in
this paper therefore rely on the \emph{form} of the ECI axioms,
not on any derivation from them. This distinction is load-bearing
and must not be blurred.

\section{Roadmap and feasibility on consumer hardware}
\label{sec:roadmap}

\subsection{Hardware budget (single-author context)}

The author's workstation (i5-14600KF, 32~GB DDR4, RTX~5060~Ti 16~GB
VRAM) and the existing \texttt{chimere-server} Rust FFI to
\texttt{ik\_llama.cpp} define the hardware envelope. For a
$\sim 30$B-active-parameter upcycled hybrid model at IQ3\_S weights
with Q8\_0 on routers and final LayerNorm, the total memory
footprint is $\sim 34$--$38$~GB, which fits with aggressive
DDR-offload of cold experts; expected throughput $\sim 8$--$15$~tok/s.
From-scratch pretraining is infeasible at this budget; continue-training
via Drop-Upcycling \cite{DropUpcycling2025} from a published base
model (e.g.\ Qwen3-Next-80B-A3B) on Vast.ai spot instances remains
feasible at $\sim\$50$k--$\$200$k.

\subsection{Four-phase plan}

\begin{description}[leftmargin=*,itemsep=3pt]
 \item[Phase 1 (months 0--3).] Runtime integration: entropic
   scheduler M1/M2 on the existing \texttt{chimere-server} stack.
   No new training. Gold-path demonstration on a Qwen3-Next-80B
   upcycle.
 \item[Phase 2 (months 3--8).] Drop-Upcycling~\cite{DropUpcycling2025}
   of Qwen3-Next, addition of BD3 and JEPA predictor heads via
   LoRA, Engram L1--L3 integration.
 \item[Phase 3 (months 8--14).] Mode~M3 symbolic verification,
   SAE-based interpretability for censorship scores, calibrated
   hypothesis-H3 ablations.
 \item[Phase 4 (months 14--24).] A $3$--$8$B from-scratch training
   with the full FEC objective (Eq.~\ref{eq:FEC}) on
   $\sim 200$--$500$B tokens via Vast.ai; this is the first stage
   where hypothesis~H1 becomes directly testable.
\end{description}

\subsection{What the pure assemblage already delivers}

Phase~1 alone (entropic scheduler on existing components) is a
strict improvement over a monolithic AR or monolithic diffusion
baseline on workloads with mixed residual entropy: gains of
$\sim 3\!-\!6\times$ wall-clock on typical mixed-entropy workloads
follow mechanically from the EAGLE-3 +
M2-selective-refinement combination. The proposed design hypotheses
(H1--H3) should only be introduced after the assemblage baseline
saturates, which we estimate at the end of Phase~3.

\section{Limitations and honest caveats}
\label{sec:caveats}

\paragraph{The ECI scaffolding is analogical, not derivational.}
The ECI v3 axioms A3 (``cryptographic censorship'') and A6
(``persistent-homology complexity'') live in a cosmological
setting with specific mathematical content; we use their
\emph{names and structural role} as scaffolding for the
machine-learning hypotheses. An explicit attempt to derive the
form of the v6 source term from the Jacobson--EGJ internal
production~$d_i S$ failed in a dedicated audit~\cite{ECIv6Audit},
with four sub-obstructions catalogued
(categorical mismatch between state-functional complexity and
extrinsic horizon geometry; Gaussian-coherent-state erasure of
$|\alpha|^2$ in spread complexity; the $\delta Q = T \delta S$
Wald-entropy tautology; absence of coupling-coefficient prediction
in type-II crossed-product frameworks). The present paper inherits
the diagnostic weight of this audit: the ECI language is
suggestive, not load-bearing.

\paragraph{Production-first, research-second.}
No hypothesis is promoted unless Phase~3 completes; Phase~1 delivers
value whether or not the hypotheses hold.

\paragraph{Falsification is explicit.}
Each hypothesis in Section~\ref{sec:equations} carries a
\emph{falsification clause}; a clean failure is a valid result. The
\texttt{FAILED.md} discipline inherited from the ECI project
applies.

\paragraph{We do not claim SOTA.}
The blueprint targets a local-first, interpretable, and honestly
scoped runtime, not a frontier-model ranking on a fixed benchmark
suite. If a frontier-cloud runtime (DeepSeek, Kimi, Gemini) scores
higher on any given benchmark, that is orthogonal to the value
proposition here.

\section{Conclusions}
\label{sec:conclusions}

The frontier between \emph{prediction} and \emph{computation} is
not a dichotomy; it is a budget. Chim\`ere $\Omega$ proposes a
runtime that allocates that budget per-token as a function of
residual entropy, integrating verified 2025--2026 components
(LeJEPA/SIGReg, hybrid GDN + sparse-attention backbones,
Titans / Nested Learning memory, BD3 block-diffusion,
auxiliary-loss-free MoE, vLLM~V1 + EAGLE-3 runtime) with three
design hypotheses inspired by the author's ECI framework but
explicitly presented as hypotheses to be tested. Phase~1 of the
roadmap is immediately executable on consumer hardware and
delivers value independently of whether hypotheses H1--H3 hold;
Phases~2--4 progressively introduce the hypothesis testing.

The intellectually honest deliverable of this paper is not a new
theorem. It is a \emph{plan of record} for a line of work whose
first executable product uses existing verified components and
whose speculative content is explicitly isolated, testable, and
falsifiable. Further progress on Chim\`ere $\Omega$ will be
registered under the same discipline that produced the ECI~v6.0.3
companion paper: pre-registered falsifiers, explicit failure logs,
and cross-model audits before any paper update.

\paragraph{Acknowledgement.}
This blueprint was substantially drafted with the assistance of
large-language-model chat and agent interfaces, with citation
verification performed by WebSearch / WebFetch against arXiv and
publisher records. All numerical claims of recent papers
cited here were cross-checked against the primary source. No
citation was retained without primary-source verification. The
three design hypotheses are the author's, not the assistants'.

\paragraph{Data availability.}
This is a blueprint paper; no datasets or training runs are reported.
The reference chimere-server runtime is available at
\url{https://github.com/AIdevsmartdata/chimere}. All cited external
models, datasets, and benchmarks are public and accessible from their
respective arXiv / HuggingFace releases.

\paragraph{Conflicts of interest.}
The author declares no competing financial or non-financial interests.

% -------------------------------------------------------------------
\begin{thebibliography}{99}

\bibitem{LeJEPA2025}
R.~Balestriero and Y.~LeCun,
``\emph{LeJEPA: Provable and Scalable Self-Supervised Learning Without
the Heuristics},''
arXiv:2511.08544, November 2025.

\bibitem{Titans2024}
A.~Behrouz, P.~Zhong, and V.~Mirrokni,
``\emph{Titans: Learning to Memorize at Test Time},''
arXiv:2501.00663, December 2024.

\bibitem{NestedLearning2025}
A.~Behrouz, M.~Razaviyayn, P.~Zhong, and V.~Mirrokni,
``\emph{Nested Learning: The Illusion of Deep Learning Architectures},''
NeurIPS 2025 (accepted), arXiv:2512.24695, December 2025.

\bibitem{GatedDeltaNet2025}
S.~Yang, Y.~Wang, B.~Hooi, W.~Liu and others,
``\emph{Gated Delta Networks: Improving Mamba2 with Delta Rule},''
arXiv:2412.06464, December 2024 (ICLR 2025).

\bibitem{KimiLinear2025}
Kimi team,
``\emph{Kimi Linear: an Expressive, Efficient Attention Architecture},''
arXiv:2510.26692, October 2025.

\bibitem{DeepSeekV3}
DeepSeek-AI,
``\emph{DeepSeek-V3 Technical Report},''
arXiv:2412.19437, December 2024.

\bibitem{AuxLossFreeMoE2024}
L.~Wang, H.~Gao et al.,
``\emph{Auxiliary-Loss-Free Load Balancing Strategy for Mixture-of-Experts},''
arXiv:2408.15664, August 2024.

\bibitem{BlockDiffusion2025}
M.~Arriola, S.~Sahoo, A.~Gokaslan, V.~Kuleshov,
``\emph{Block Diffusion: Interpolating Between Autoregressive and Diffusion
Language Models},''
arXiv:2503.09573, March 2025 (ICLR 2025 Oral).

\bibitem{KimiK2}
Moonshot AI,
``\emph{Kimi K2: Open Agentic Intelligence},''
arXiv:2507.20534, July 2025.

\bibitem{EAGLE3}
Y.~Li et al.,
``\emph{EAGLE-3: Scaling up Inference Acceleration of Large Language Models
via Training-Time Test},''
arXiv:2503.01840, March 2025.

\bibitem{SpinQuant2024}
Z.~Liu, C.~Zhao, I.~Fedorov et al.,
``\emph{SpinQuant: LLM quantization with learned rotations},''
arXiv:2405.16406, May 2024.

\bibitem{KIVI2024}
Z.~Liu et al.,
``\emph{KIVI: A Tuning-Free Asymmetric 2bit Quantization for KV Cache},''
arXiv:2402.02750, February 2024.

\bibitem{DropUpcycling2025}
T.~Nakamura et al.,
``\emph{Drop-Upcycling: Training Sparse Mixture of Experts with Partial
Re-initialization},''
arXiv:2502.19261, February 2025 (ICLR 2025).

\bibitem{IBLM2025}
C.~Yu,
``\emph{An Information Bottleneck View of LLM Pretraining},''
arXiv:2505.08727, May 2025.

\bibitem{ECIv6Audit}
K.~Remondi\`ere,
``\emph{ECI v6.0.3 companion paper and v8 bridge audit
(FAILED.md~F-14)},''
Zenodo deposit, 2026,
\url{https://github.com/AIdevsmartdata/crossed-cosmos}.

\end{thebibliography}

\end{document}
