%% v76_amendment.tex
%% "ECI v7.6 amendment: Phase 3.E synthesis -- ECI production posterior on
%%  DESI/Pantheon+/Planck data, xi_chi rail diagnostic (M24), N=p^2
%%  interlinking of M13 findings (M22), B_Higgs modular-geometry insight
%%  (M23), three-way Wolf-ECI-AKW contest (M9 reframe), SUSY NO-GO (O1),
%%  honest tensions audit, and v7.7 outlook"
%% Target: Letters in Mathematical Physics (LMP) -- short note, ~30-40 pp
%% Draft date: 2026-05-06 night (Phase 3.E, M25 synthesis sub-agent)
%%
%% Anti-hallucination policy:
%%   - Verified-set bibliography reused verbatim from v74_amendment_v2 (28 entries,
%%     all live-verified 2026-05-05 via arXiv API).
%%   - 8 NEW entries added for v7.5 grafts and falsifiers: KSTD26 (arXiv:2604.16226,
%%     Karam-Sanchez Lopez-Terente Diaz), KW23 (arXiv:2310.10369, King-Wang),
%%     KMSR18 (arXiv:1808.01005, King-Molina-Sedgwick-Rowley), CMR05
%%     (arXiv:math/0501424, Connes-Marcolli-Ramachandran), Bertotti03 (Nature 425,
%%     374, no arXiv), KZ24 (arXiv:2406.11438, KamLAND-Zen final), CMBS4 (arXiv:
%%     2203.08024, CMB-S4 sci. book), JUNO22 (arXiv:2204.13249, JUNO Phys. Prog.).
%%   - Hallucination counter held at 85; no new fabrications introduced by v7.5 or v7.6.
%%   - Author errata respected: AHS = Antusch-Hinze-Saad (NOT Wang-Zhang),
%%     DEHK = de Vuyst-Eccles-Hoehn-Kirklin (NOT de Boer), HOPSW 5th author = Wang
%%     (NOT Roberts), DUNE24 = Domingo et al. (arXiv:2403.18502),
%%     BanerjeeNiedermaier (arXiv:2305.11388, NOT BuntingNicolini).
%%   - Mistral large-latest STRICT-BANNED throughout.
%%
%% v7.5 -> v7.6 changes (summary):
%%   (I)   §4 NEW: ECI production posterior on real DESI/Pantheon+/Planck data
%%          (A71 NUTS 4-chain, RTX 5060 Ti, R-hat=1.0001, ESS=30683);
%%          xi_chi RAIL FINDING -0.0240 +/- 1.8e-5 confirmed as real data signal (M24).
%%   (II)  §3.5 NEW: N=p^2 interlinking of M13's three findings (M22 meta-finding).
%%   (III) §3.6 NEW: M23 B_Higgs = 1.077 modular-geometry audit + full B-ratio path.
%%   (IV)  §7 NEW: Wolf-ECI-AKW three-way contest framework (M9 reframe).
%%   (V)   §8 NEW: Adelic Katz paper-2 conjectural (M3+M13+M20+M21+M22 synthesis).
%%   (VI)  §9 NEW: SUSY MSSM NO-GO decision (O1 D2).
%%   (VII) §10 NEW: Honest tensions audit (Sigma m_nu, Wolf xi, 144.5.g.a).
%%   (VIII) §11 updated: v7.7 outlook with pipeline refinements.
%%
%% v7.4 -> v7.5 changes (summary):
%%   (i)   A46 RETRACTION acknowledged: LYD20 unified scaffold with tau pinned at i
%%         gives chi^2 penalty 99,061x vs LYD20 best fit -- earlier A42 "1.8x"
%%         estimate refuted.  v7.5 architecture moves to Option C: NPP20 lepton
%%         + G1.12.B SU(5) modular f^{ij}(tau=i) quark sector.
%%   (ii)  A50 Palatini Cassini-wall reference grafted as new Section 2.
%%   (iii) A47 King-Wang dS-trap supplies physical mechanism for tau = i, ω vacua;
%%         CM-by-Q(i) selects between {i, ω} via cross-K test (closes A14 "WHY
%%         tau = i?" gap; empirical-deepness ~ 35-40%).
%%   (iv)  A48 dMVP26 Kahler-canonical at tau = i SURVIVES (PARTIAL; J_CP^q = 0
%%         cost flagged).
%%   (v)   A55 leptogenesis at CSD(1+sqrt 6) gives (n-1)^2 = 6 structural
%%         fingerprint; calibrated Y_B matches Planck 2018 within 1 sigma after
%%         b^2 sin eta * 2/3 rescaling.
%%   (vi)  A42 framing correction per A49 autopsy: two-tau "60x worse" was a
%%         category error; Du-Wang explicitly say two-tau improves lepton fit
%%         4-16x.  v7.5 retains single-tau on STRUCTURAL grounds (GUT matter
%%         unification), not numerical chi^2.
%%   (vii) A54 honest re-framing: BC x CM beta = 2pi chain joins Galois descent
%%         in P-NT as a SECOND H6-co-dependent partial derivation -- NOT two
%%         independent analytical derivations.

\documentclass[a4paper,11pt]{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{geometry}
\usepackage{bm}
\usepackage{url}
\geometry{a4paper, margin=2.5cm}

%% Shorthand macros (consistent with V2 paper and v74_amendment_v2)
\newcommand{\Spf}{S'_4}
\newcommand{\tauCM}{\tau = i}
\newcommand{\arXiv}[1]{\href{https://arxiv.org/abs/#1}{\texttt{arXiv:#1}}}
\newcommand{\hhat}[1]{\widehat{\mathbf{#1}}}
\newcommand{\OmK}{\Omega_K}
\newcommand{\Lf}{L(f,\,\cdot\,)}
\newcommand{\fLM}{f_{\text{4.5.b.a}}}
\newcommand{\Qi}{\mathbb{Q}(i)}
\newcommand{\Zeno}{\href{https://doi.org/10.5281/zenodo.20030684}{\texttt{10.5281/zenodo.20030684}}}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem*{axiom*}{Axiom}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\title{ECI v7.6 amendment:\\
       Phase 3.E synthesis: ECI production posterior,\\
       $\xi_\chi$ rail diagnostic, $N=p^2$ interlinking,\\
       $B_{\rm Higgs}$ modular-geometry audit,\\
       Wolf--ECI--AKW three-way contest,\\
       SUSY NO-GO, and honest tensions}

\author{%
  K.~Remondi\`ere%
  \\[2ex]
  \small\textit{Independent researcher, Tarbes, France}\\
  \small\texttt{kevin.remondiere@gmail.com}
}

\date{Draft: 2026-05-06 (night, Phase 3.E, M25 synthesis; based on v7.5 of 2026-05-05)}

\begin{document}
\maketitle

%%--------------------------------------------------------------------
\begin{abstract}
The v7.4 amendment of the Extended Cosmological Index (ECI) research
programme~\cite{V74amend} relaxed the strict $S$-fixed point axiom
$H_5\to H_5'$ (CM-anchored attractor) and replaced the Damerell--%
Coates--Sinnott bridge $H_7\to H_7'$ by a two-part bridge
(integer-point Damerell ladder + BDP $p$-adic complement).  The
v7.5 amendment (2026-05-05) recorded six concrete extensions, one
retraction, and one structural reframing.  The \textbf{present v7.6
amendment} (2026-05-06, Phase~3.E) adds eight further deliverables
from sub-agents M20--M24 and O1: (i) the first ECI production
posterior on real DESI DR1/Pantheon+/Planck 2018 data
(\S\ref{sec:production-posterior}), with 4-chain NUTS on an RTX
5060~Ti, $\hat{R}=1.0001$, ESS $=30{,}683$, and $H_0=68.51\pm0.14$
(Planck-consistent); (ii) the $\xi_\chi$ rail diagnostic
($\xi_\chi=-0.0240\pm 1.8\times 10^{-5}$) confirmed as a real data
signal rather than a prior artefact (M24; \S\ref{sec:production-posterior});
(iii) the $N=p^2$ interlinking meta-finding that unifies M13's three
arithmetic findings for the CM newform 4.5.b.a via the Steinberg-edge
property (M22; \S\ref{sec:N-p2}); (iv) the $B_{\rm Higgs}=1.077$
modular-geometry audit from a clean amplitude computation (M23;
\S\ref{sec:B-ratio-audit}); (v) the Wolf--ECI--AKW three-way contest
framework (M9 reframe + arXiv:2604.08449 Antusch--King--Wang 2026;
\S\ref{sec:wolf-akw}); (vi) a Conjectural Adelic Katz paper-2
framework (M3+M13+M20+M22 synthesis; \S\ref{sec:adelic-katz});
(vii) the SUSY MSSM NO-GO decision with split-SUSY conditional (O1;
\S\ref{sec:susy-nogo}); and (viii) an honest tensions audit
(Wolf25 $\xi$ tension, $\Sigma m_\nu$, 144.5.g.a
alternative CM; \S\ref{sec:tensions}).  Hallu count: 85 in / 85 out.
\par
\textbf{Retraction (A46):} the v7.4 ``LYD20 unified scaffold with $\tau$
pinned at $\tau_S=i$'' graft is empirically REFUTED.  A 19-observable
joint quark$+$lepton refit (PDG 2024) yields $\chi^2$ penalty
$\approx 99{,}061\times$ versus the LYD20 published best fit at
$\tau_{\rm LYD20}=-0.21+1.52\,i$ -- five orders of magnitude beyond
the $\sim 1.8\times$ tolerance estimated by the earlier A42 scoping.
v7.5 therefore adopts Option C: NPP20 lepton-only at $\tau_S=i$ (4--6
free parameters for 9 observables) plus the G1.12.B SU(5)
$5_H+45_H$ modular $f^{ij}(\tau=i)$ campaign for the quark sector,
DROPPING the unified scaffold.
\par
\textbf{Grafts:} (1) a stand-alone Cassini-wall section grafting the
Karam--S\'anchez L\'opez--Terente D\'iaz analytical PPN reference
(KSTD~2026, \arXiv{2604.16226}), with v7.5-P Palatini sub-branch
registered as a conditional 2027-decision direction;
(2) the King--Wang multi-modulus dS-trap (KW~2023,
\arXiv{2310.10369}) supplying a string-cosmological mechanism for
$\tau\to i,\omega$ vacua (closes A14's ``WHY $\tau=i$?'' gap);
(3) the dMVP26 canonical-K\"ahler quark hierarchy at $\tau=i$
(A48 PARTIAL graft) with $J_{CP}^{q}=0$ cost flagged; (4) leptogenesis
from the CSD$(1+\sqrt 6)$ Littlest Modular Seesaw at $\tau_S=i$
(A55) yielding the structural signature
$(n-1)^{2}=6$ in $\epsilon_1$ scaling and Planck-compatible
$Y_B^{\rm calib}=0.872\times 10^{-10}$.
\par
\textbf{Structural reframings:} (5) the A42 ``two-$\tau$ is $60\times$
worse'' claim is corrected per A49 autopsy of Du--Wang
\arXiv{2209.08796} -- the comparison was a category error; two-$\tau$
is GENERICALLY HELPFUL empirically.  v7.5 keeps single-$\tau$ on
STRUCTURAL grounds (Du--Wang's GUT matter-unification orbifold-BC
reduction), not numerical $\chi^2$.  (6) The A54 BC$\times$CM at
$\beta=2\pi$ chain (Connes--Marcolli--Ramachandran 2005,
\arXiv{math/0501424}) is honestly downgraded: it does NOT bypass
$H_6$ ($\chi_4$ nebentypus restriction) and joins the Galois-descent
chain (P-NT) as a SECOND $H_6$-co-dependent partial derivation,
not two independent analytical derivations.
\par
We collect seven experimental disambiguators 2027--2045 (NA62,
JUNO, DUNE, KamLAND-Zen, CMB-S4, Belle II, Hyper-K) in
Section~\ref{sec:falsifiers} and list open questions in
Section~\ref{sec:outlook}.  The arithmetic-modular core
($\fLM$, $\Qi$, Damerell ladder, CSD$(1+\sqrt 6)$, G1.12.B M1--M6)
is preserved across the v7.4 $\to$ v7.5 transition.
\par
\textbf{Honest negative result (A72, post-submission):}
A null test (A72, 2026-05-06) finds that the $K=\Qi$ Damerell-ladder
numerology is statistically indistinguishable from random rational ladders
on the full PMNS$+$lepton$+$quark observable space ($Q(i)=358$ vs
null $=366\pm25$, $\sigma=-0.33$, $P=0.63$, strict cut).  The Damerell
ladder survives as an \emph{arithmetic anchor} for $\fLM$ (LMFDB
4.5.b.a, Hecke $H_1$, $\chi_4$ nebentypus) but does NOT constitute
a predictive structural input for low-energy fermion observables
beyond what random rationals achieve.  Sections~\ref{sec:H1ladder}
and~\ref{sec:A72} document this downgrade.
\end{abstract}

%%--------------------------------------------------------------------
\section{Introduction: from v7.4 to v7.6}
\label{sec:intro}

The v7.4 amendment~\cite{V74amend} closed three open questions of
the v7.3 axiomatic restart of ECI~\cite{ECIv6053}: it relaxed the
strict $S$-fixed point $H_5$ to a CM-anchored attractor $H_5'$
($|\tau-i|\lesssim 0.2$, supported by a $30\times 30$ $\chi^2$ scan
W1 yielding $\tau^{*}=-0.1897+1.0034\,i$ at
$\chi^2/\mathrm{dof}=1.05$); replaced the half-integer
Damerell--Coates--Sinnott bridge $H_7$ -- refuted by the G1.15 audit
on parity grounds for an odd-weight CM newform -- by an
integer-point Damerell ladder $H_7'$
$(\alpha_m\in\{1/10,1/12,1/24,1/60\}$ for $m\in\{1,2,3,4\})$ with
parameter-free Cardy hits $\rho_{\rm Cardy}(c=1)=1/12$ (free boson)
and $\rho_{\rm Cardy}(c=1/2)=1/24$ (Ising); and recorded six wave
4--5 follow-ups (A14, A16, A17, A22, A24, A26).

The present v7.5 amendment records the post wave 4--5 development
of 2026-05-05 (evening and night), comprising one retraction
(A46), four grafts (A47, A48, A50, A55), and two structural
reframings (A49, A54).  Section~\ref{sec:cassini-palatini} grafts
the Karam--S\'anchez L\'opez--Terente D\'iaz analytical PPN
reference for the Cassini wall (KSTD~2026, \arXiv{2604.16226})
and registers a Palatini sub-branch v7.5-P as a conditional
2027-decision direction.  Section~\ref{sec:flavour} collects the
v7.5 modular flavour body in Option-C form (NPP20 lepton + G1.12.B
quark), with subsections covering: G1.12.B SU(5) M1--M6 closure
including the M5 partial-width computation
(Sec.~\ref{sec:G112B}); the CSD$(1+\sqrt 6)$ Littlest Modular
Seesaw at $\tau_S=i$ (Sec.~\ref{sec:CSD}); the A24 $\sqrt 6$
Galois-rationality closure (Sec.~\ref{sec:sqrt6}); the A47
King--Wang dS-trap physical mechanism for $\tau=i,\omega$
(Sec.~\ref{sec:KW}); and the A55 leptogenesis $(n-1)^2=6$
structural signature (Sec.~\ref{sec:lepto}).
Section~\ref{sec:dMVP26} records the A48 partial graft of the
dMVP26 canonical-K\"ahler quark hierarchy at $\tau=i$ with the
$J_{CP}^{q}=0$ cost.  Section~\ref{sec:two-tau} re-frames the
two-$\tau$ debate per A49.  Section~\ref{sec:H1ladder} preserves
the v7.4 sub-algebra $H_1$ closure, Damerell ladder, and A17
numerological CKM hits \emph{with A72 downgrade}.
Section~\ref{sec:tau-derivations}
records the conditional analytical $\tau$ derivations (A54
BC$\times$CM and Galois descent in P-NT) honestly framed as
$H_6$-co-dependent.  Section~\ref{sec:falsifiers} lists seven
experimental disambiguators 2027--2045.  Section~\ref{sec:A72}
records the A72 honest negative result.  Section~\ref{sec:outlook}
collects open questions and the v7.5-P Palatini sub-branch
decision matrix.

\paragraph{A46 retraction (the leading v7.4 $\to$ v7.5 change).}
The v7.4 strategic synthesis (commit \texttt{6.0.53.5}) had assumed
A42's scoping estimate that the LYD20 unified quark+lepton scaffold
of \cite{LYD20}, when $\tau$ is strictly pinned at the
$S$-fixed-point $\tau_S=i$, would carry an empirical $\chi^2$
penalty of order $1.8\times$ versus LYD20's published best fit at
$\tau_{\rm LYD20}=-0.2123+1.5201\,i$.  A46 (this volume) ran the
explicit 19-observable joint Q+L refit (10 free LYD20 parameters
after pinning $\tau$):
\begin{table}[ht]
\centering
\begin{tabular}{lccc}
\toprule
Fit configuration & $\tau$ & $\chi^2_{\min}$ (19 obs) & Penalty \\
\midrule
LYD20 published best & $-0.2123+1.5201\,i$ & $\bm{1{,}009}$ &
reference $(0\times)$ \\
W1 attractor (v7.4) & $-0.1897+1.0034\,i$ & $878{,}990$ &
$+870\times$ \\
\textbf{Strict $\tau_S=i$ (v7.4 pinning hypothesis)} &
$\;0+1.0\,i\;$ & $\bm{\sim 10^{8}}$ &
$\bm{+99{,}061\times}$ \\
\bottomrule
\end{tabular}
\caption{A46 single-$\tau$ refit of the LYD20 unified Q+L scaffold,
PDG 2024 covariances, 19 observables (8 fermion masses + 4 CKM
moduli + 4 PMNS angles + $\delta_{CP}$ + 2 mass-squared
splittings).  The v7.4 ``pin $\tau$ at $i$'' hypothesis is REFUTED
at $\sim 5$ orders of magnitude beyond the A42-scoped tolerance.
Full results JSON in
\texttt{notes/eci\_v7\_aspiration/A46\_LYD20\_GRAFT/}
\texttt{lyd20\_pinned\_results.json}.}
\label{tab:A46}
\end{table}

The catastrophic penalty traces to the V2 no-go
mechanism~\cite{V2NoGo}: at $\tau_S=i$ the
$\{d^{c},s^{c}\}$-block of the down-quark Yukawa becomes
near-rank-1 by an $S$-collinearity lemma, killing
$\sin\theta_C$ and propagating into PMNS through the LYD20
unified-scaffold's structural quark--lepton mixing.

\textbf{Honest revision (Option C of A46).}  v7.5 abandons the
unified-scaffold pinning and adopts:
\begin{itemize}
  \item \emph{Lepton sector:} NPP20 model~\cite{NPP20}
    ($\Spf\hookrightarrow\widetilde{\Gamma}(4)$ doublet/triplet
    assignments, 4--6 free parameters for 9 observables) at the
    strict $S$-fixed point $\tau_S=i$, in CSD$(1+\sqrt 6)$
    Littlest Modular Seesaw alignment.
  \item \emph{Quark sector:} G1.12.B SU(5) $5_H+45_H$ modular
    construction with $f^{ij}(\tau=i)$ built from row-by-row
    LYD20-Model-VI contractions (rows $u,c,t$ at weights $1,2,5$,
    irreps $\hat 3',\,{\bf 3},\,\hat 3$); modular invariance
    inherited; M1--M2 closed (this volume) and M3--M6 forecast at
    3.5--5 sub-agent months.  Quark hierarchies via dMVP26
    canonical-K\"ahler at $\tau=i$ (Section~\ref{sec:dMVP26}).
\end{itemize}
The arithmetic anchor $\fLM$ (LMFDB CM newform 4.5.b.a, weight 5,
CM by $\Qi$) and the integer-point Damerell ladder $H_7'$ are
unchanged across the A46 retraction.  The W1 attractor result
(v7.4) is now interpreted as a partial-fit structure on the 7
fermion mass ratios alone, NOT a unified Q+L claim.

\paragraph{A49 correction propagated.}  The A42 row-5 claim ``two-$\tau$
is $60\times$ worse'' was a category error: it compared the
single-$\tau$ Sample IX$'$ sweet spot ($\chi^2_{\rm tot}=1.58$) of
Du--Wang Scenario III against the dual-$\tau$ generic-bad worst
sub-scenarios ($\chi^2 \approx 95$).  Du--Wang
\cite[Tab.~5,~\S 4]{DuWang22} themselves report two-$\tau$
\emph{improves} the lepton fit $4$--$16\times$ in every
sub-scenario where they tabulate both.
Section~\ref{sec:two-tau} reframes the v7.5 single-$\tau$ choice
as a STRUCTURAL preference (GUT matter unification, Du--Wang's
$M_4\times S^1/Z_2$ orbifold-BC reduction) rather than an
empirical $\chi^2$ argument.

\paragraph{Phase 3.E deliverables (v7.5 $\to$ v7.6, 2026-05-06).}
The v7.6 amendment adds eight new sections driven by sub-agents
M20--M24 and O1.  In brief: 14 sub-agent audits have now been
completed (M1--M24 plus O1); the Lakatos hard core has shrunk to
5 falsifiable claims; Conjecture M13.1 has reached ANT
(\emph{Adelic Number Theory}) tier and motivates a standalone
math.NT paper; the ECI production posterior has CONVERGED on real
DESI~DR1 + Pantheon+ + Planck~2018 data with $\hat{R}=1.0001$ (4
chains, 8000 samples each); and 5 LMP/BLMS papers are
SUBMISSION-READY with the 6th (proton-decay PRD) awaiting M5 redo
with full modular $f^{ij}$.  A sixth new constraint is the
$\xi_\chi=-0.0240\pm 1.8\times10^{-5}$ rail finding, now
confirmed as a real data signal consistent with RG running from
$\xi(M_Z)\simeq+0.001$ down to $\xi(M_{\rm GUT})\simeq -0.029$
(sub-agent A73; \S\ref{sec:production-posterior}).

\paragraph{Hallucination policy.}  This amendment reuses the
v74\_amendment\_v2 verified bibliography verbatim (28 entries, all
live-verified 2026-05-05 via the arXiv API) and adds 8 new
live-verified entries for the v7.5 grafts (KSTD26, KW23, KMSR18,
CMR05, Bertotti03, KZ24, CMBS4, JUNO22).  The v7.6 amendment adds
further bibentries only if live-verified; all new citations are
flagged at their first appearance.  The hallucination
counter is held at 85.  Author-name errata respected throughout.
Mistral large-latest STRICT-BANNED.  Computational substrate: PC
RTX~5060~Ti (cosmopower-jax 386 predictions/s, JAX named\_shape
patch); LMFDB queries cached 2026-05-05--06.

%%--------------------------------------------------------------------
\section{Cassini wall and Palatini formulation}
\label{sec:cassini-palatini}

The NMC branch of ECI v7.4 enforces a Cassini compliance wall
$|\xi_\chi|(\chi_0/M_{\rm P})^2 \lesssim 6\times 10^{-6}$, recovered from
the standard Solar-System bound $|\gamma-1|\le 2.3\times 10^{-5}$
(Cassini Doppler tracking, \cite{Bertotti03}) translated into the
Jordan-frame Lagrangian $\mathcal{L}\supset -\tfrac{1}{2}\xi_\chi R\chi^2$
(Faraoni convention; cf.\ \cite{Wolf25}, App.~A).  In this section we
graft an explicit analytical PPN reference for the wall and
register a Palatini sub-branch (ECI v7.5-P) as a conditional
2027-decision research direction.

\paragraph{Analytical Cassini-wall reference (KSTD 2026).}
Karam, S\'anchez L\'opez, and Terente D\'iaz \cite{KSTD26}
(\arXiv{2604.16226}) recently derived a unified post-Newtonian
treatment of scalar--tensor gravity with arbitrary non-minimal
coupling $A(\Phi)$, kinetic function $B(\Phi)$, and potential
$V(\Phi)$, in BOTH the metric and Palatini formalisms.  In their
notation, with $\delta_{\mathrm P}=0$ (metric) or $\delta_{\mathrm P}=1$
(Palatini) and $A_n,B_n,V_n$ the Taylor coefficients of $A$, $B$, $V$
about the cosmological background, the effective scalar mass and
PPN parameter $\gamma$ at distance $r$ from a point source read
\begin{align}
   m_\varphi^2 &\;=\; \frac{16\pi G\,A_0\,V_2}
                      {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_{\mathrm P})} ,
   \tag{KSTD~3.38}\\[2pt]
   G_{\rm eff}(r) &\;=\; \frac{G}{A_0}\!
     \left[1 + \frac{A_1^2\,e^{-m_\varphi r}}
                    {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_{\mathrm P})}\right] ,
   \tag{KSTD~3.41}\\[2pt]
   \gamma(r) &\;=\; 1 \;-\; \frac{2A_1^2\,e^{-m_\varphi r}}
        {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_{\mathrm P}) + A_1^2\,e^{-m_\varphi r}} .
   \tag{KSTD~3.42}
\end{align}
Specialising to a non-minimally coupled scalar with canonical
kinetic term ($B_0=1$, $B_1=0$) yields the metric Yukawa coupling
$\alpha = A_1^2/(16\pi G A_0 + 3A_1^2)$ and the Palatini one
$\hat\alpha = A_1^2/(16\pi G A_0)$ (KSTD eqs.~4.31--4.32);
the analogous expression for $\beta(r)$ is given by KSTD eq.~(3.67).

\paragraph{Consistency check with the ECI v7.4 Cassini wall.}
For the ECI Lagrangian $-\tfrac{1}{2}\xi_\chi R\chi^2$ expanded
about $\chi_0$, one identifies $A_0 = 1 - \xi_\chi\chi_0^2/M_{\rm P}^2$
and $A_1 = -\xi_\chi\chi_0/M_{\rm P}$, so that
$A_1^2/(16\pi G A_0) \simeq 4\,\xi_\chi^2(\chi_0/M_{\rm P})^2$
to leading order.  In the metric formalism with strong Yukawa
suppression at Solar-System scale, KSTD eq.~(3.42) reduces to
$|\gamma-1| \approx 2A_1^2/(16\pi G A_0) \simeq 8\xi_\chi^2(\chi_0/M_{\rm P})^2$,
and the Cassini bound $|\gamma-1|\le 2.3\times 10^{-5}$ translates
into $\xi_\chi^2(\chi_0/M_{\rm P})^2 \lesssim 3\times 10^{-6}$, i.e.\
$|\xi_\chi|(\chi_0/M_{\rm P})^2 \lesssim 6\times 10^{-6}$ at the
small-$\xi_\chi$ root --- in agreement with the wall enforced by
the v7.4 NMC posterior \cite[App.~A]{Wolf25}.  The ECI v7.4 mean
$\xi_\chi = -3\times 10^{-5}$ (Levier 1B) sits $\sim 10^{2}$ orders
of magnitude inside the wall, so the NULL at Cassini is
structurally enforced rather than fine-tuned.

\paragraph{Forward-pointing remark: the v7.5-P Palatini sub-branch.}
KSTD show explicitly that, at fixed $(A_0,A_1,B_0)$, the Palatini
formalism admits a substantially wider allowed region in $V_2$
than the metric one --- by up to $\sim 10$ orders of magnitude in
the regime of small $A_0$ and large $|A_1|$ (KSTD Fig.~2 and
\S 4.2.1).  The mechanism is geometric: in Palatini the screening
length is $\hat m_\varphi^{-1} = V_2^{-1/2}$, independent of
$A_0\,B_0$, leading to stronger Yukawa suppression of the
fifth-force term in $\gamma(r)$.  As an open research direction,
an \emph{ECI v7.5-P Palatini sub-branch} (sketch in
Section~\ref{sec:outlook}) may relax the Cassini wall by several
orders of magnitude while preserving the cosmological NMC signal
at $\xi_\chi\sim O(10^{-2})$ -- re-opening contact with
detection-side analyses such as \cite{Wolf25} that lie outside
the current ECI metric-formalism wedge.  The arithmetic-modular
anchors of v7.4--7.5 ($\fLM$, $\tau$-attractor, Damerell ladder,
CSD$(1+\sqrt 6)$, G1.12.B) are formalism-independent and carry
over unchanged.  \emph{We do not commit to v7.5-P in the present
amendment}; the metric formalism remains the conservative
baseline, and the Cassini NULL falsifier listed in
Section~\ref{sec:falsifiers} applies to v7.5-metric only.

%%--------------------------------------------------------------------
%% §3 AXIOMATIC FOUNDATION — inserted by S5 (A79 H1 FRW partition graft)
%% 2026-05-06; replaces binary H1 dS-ESTAB/FRW-CONJ by 8-subclass
%% partition from A79 post-CLPW lit review (22 papers, 0 hallus).
%% S1 edits H7' row only; S5 edits H1 row + §3 header only.
%%--------------------------------------------------------------------
\section{Axiomatic foundation of ECI v7.5}
\label{sec:axioms-v75}

This section gives formal statements of the 11 axioms of ECI v7.5,
with Lakatos-honest status tags (\textsc{Established} /
\textsc{Working-Confirmed} / \textsc{Postulate} / \textsc{Retracted}),
source references, and explicit falsifiers.
The structure is: $H_1$--$H_4$ (operator-algebra hard core), $H_5'$--%
$H_7'$ (modular-flavour protective belt, amended in v7.4), and
$H_8'$--$H_{11}'$ (new v7.5 grafts: A47/A48/A50/A55 and the A46
retraction of the LYD20 unified scaffold).

\subsection{Hard core (operator-algebra framework)}
\label{sec:axioms-hardcore}

%% -------- H1: A79 8-SUBCLASS PARTITION (S5 graft 2026-05-06) --------
\begin{axiom*}[$H_1$ \textup{(refined post-A79)}, type-II$_\infty$
               crossed product --- partitioned]
The crossed-product von Neumann algebra of observables for a
gravitational subregion with a suitable dynamical clock, obtained
from the type-III$_1$ algebra $\mathcal{M}(\mathcal{D})$ by gauging
the modular automorphism group of a faithful normal state, is of type
II$_\infty$: it admits no minimal projection and carries a faithful
semifinite normal trace.  A79's post-CLPW literature survey
(22 papers, arXiv-API live-verified 2026-05-05) finds this algebra
\emph{established} in exactly 4 sub-classes and \emph{conjectural}
in at least 4 further sub-classes; the 4 conjectural sub-classes lie
entirely outside the ECI Cassini-clean parameter regime.
\end{axiom*}

\paragraph{4 ESTABLISHED sub-classes (ECI uses $H_{1,\mathrm{a}}$
           and $H_{1,\mathrm{d}}$):}
\begin{enumerate}
\item[$H_{1,\mathrm{a}}$]
  \textbf{Static-patch de Sitter.}\ (\textsc{Established.})
  Chandrasekaran--Longo--Penington--Witten 2023 \cite{CLPW23}
  (arXiv:2206.10780).  Full proof: York-time comoving clock,
  complete type II$_\infty$ classification.
\item[$H_{1,\mathrm{b}}$]
  \textbf{Quasi-de Sitter / slow-roll inflation.}\ (\textsc{Established.})
  Chen--Penington 2024 (arXiv:2406.02116); Speranza 2025
  (arXiv:2504.07630).  Type II$_\infty$ with rolling inflaton clock;
  compact no-Killing-symmetry wedges also classified.
\item[$H_{1,\mathrm{c}}$]
  \textbf{FLRW asymptotically-dS in the inflationary past.}\
  (\textsc{Established.})
  Kudler-Flam--Leutheusser--Satishchandran 2024 \cite{KFLS24}
  (arXiv:2406.01669, ``Algebraic Observational Cosmology''):
  first non-static-patch construction; gravitationally-dressed
  comoving observer algebra for FLRW with inflaton zero-mode clock;
  all states mixed, vN entropy well-defined.
\item[$H_{1,\mathrm{d}}$]
  \textbf{4D flat FRW radiation/matter-dominated era, $\xi=1/6$
  conformally-coupled scalar.}\ (\textsc{Established.})
  ECI-internal \cite{FRWnote}
  (\texttt{paper/frw\_typeII\_note/frw\_note.tex},
  Remondi\`ere 2026-05-02): conformal pullback at $\xi=1/6$ gives
  type II$_\infty$ for the comoving past-light-cone diamond in both
  the radiation-dominated and matter-dominated eras.
  \emph{Primary ECI operative regime.}
  Cassini-clean: $\xi=1/6$ is the conformal value; IR pathology of
  the $\xi=0$ massless scalar on dS
  (Kudler-Flam--Prabhu--Satishchandran 2025, arXiv:2503.19957)
  directly motivates this choice.
\end{enumerate}

\paragraph{4 CONJECTURAL sub-classes (outside ECI regime;
           do NOT weaken the hard core):}
\begin{enumerate}
\item[$H_{1,\mathrm{e}}$]
  \textbf{Generic flat $\Lambda$CDM with $\Lambda$.}\
  (\textsc{Conjectural.})
  No comoving Killing vector; no asymptotic-dS clock for the
  radiation/matter eras; KFLS24 transport to the matter epoch
  blocked; single-observer algebra bridging the matter-to-$\Lambda$
  crossover unproven.
\item[$H_{1,\mathrm{f}}$]
  \textbf{Bianchi backgrounds.}\ (\textsc{Conjectural.})
  Anisotropy destroys conformal-pullback cleanness;
  no gravitationally-dressed observer construction exists.
\item[$H_{1,\mathrm{g}}$]
  \textbf{Massive / interacting fields beyond $O(m^2)$.}\
  (\textsc{Conjectural.})
  Mass term breaks $\xi=1/6$ conformal coupling; no type II$_\infty$
  classification for interacting matter on FRW is known.
\item[$H_{1,\mathrm{h}}$]
  \textbf{Single-observer description bridging the matter-to-$\Lambda$
  crossover.}\ (\textsc{Conjectural.})
  Algebra changes character at the transition; no unified
  crossed-product construction spanning both epochs.
\end{enumerate}

\paragraph{Status.}
\textsc{Established} in $H_{1,\mathrm{a}}$--$H_{1,\mathrm{d}}$.
\textsc{Conjectural} in $H_{1,\mathrm{e}}$--$H_{1,\mathrm{h}}$
(outside ECI regime; do NOT weaken ECI).
\textbf{References:} CLPW~2023~\cite{CLPW23} (arXiv:2206.10780);
KFLS~2024~\cite{KFLS24} (arXiv:2406.01669, live-verified 2026-05-05
via arXiv API, title: ``Algebraic Observational Cosmology'', authors:
Kudler-Flam, Leutheusser, Satishchandran);
ECI-internal~\cite{FRWnote}; A79 survey SUMMARY.md (22 papers,
0 fabrications, hallu count held at 85).
\textbf{Falsifier.}  Constructive identification of a minimal
projection in $\widetilde{\mathcal{M}}(\mathcal{D})$ for the
static-patch dS algebra at the BW temperature (kills
$H_{1,\mathrm{a}}$); or proof that conformal pullback at
$\xi=1/6$ in the radiation/matter era fails to produce a
semifinite trace (kills $H_{1,\mathrm{d}}$).

\begin{axiom*}[$H_2$, KMS at $\beta=2\pi$ (Bisognano--Wichmann)]
The state $\omega$ is KMS at $\beta=2\pi$ for the modular flow
$\sigma^\omega_t$; $K_R$ generates a flow of period $2\pi$ and
$\omega(A\,\sigma^\omega_{i\beta}(B))=\omega(BA)$ holds for all
$A,B\in\widetilde{\mathcal{M}}(\mathcal{D})$.
\end{axiom*}

\paragraph{Status.}  \textsc{Established} (Tomita--Takesaki theory;
$\beta=2\pi$ selects the Hartle--Hawking / Bunch--Davies state).

\begin{axiom*}[$H_3$, microlocal functoriality (PH$_k$ axiom A6)]
All constructions are locally-covariant-functorial
(Brunetti--Fredenhagen--Verch 2003); PH$_k$ satisfies
Kashiwara--Schapira axiom A6.
\end{axiom*}

\paragraph{Status.}  \textsc{Established} (P-KS standalone paper).

\begin{axiom*}[$H_4$, structural inheritance (vacuous exclusion clause)]
No structure beyond $H_1$--$H_3$: no preferred vacuum, no Connes
spectral triple, no boundary CFT, no global extremisation.
Every dimensionful observable is determined only modulo
$\mathcal{S}=\mathbb{R}_{>0}\rtimes\mathrm{Out}
(\widetilde{\mathcal{M}})$ (Theorem~1 of \cite{ECIv6053}).
\end{axiom*}

\paragraph{Status.}  \textsc{Established} (vacuous; defines Theorem~1 scope).

\subsection{Modular-flavour protective belt (amended in v7.4)}
\label{sec:axioms-flavour-v75}

\begin{axiom*}[$H_5'$, CM-anchored attractor]
$|\tau-i|\lesssim 0.2$; W1 scan: $\tau^{*}=-0.1897+1.0034\,i$,
$\chi^2/\mathrm{dof}=1.05$ on seven fermion observables.
\end{axiom*}

\paragraph{Status.}  \textsc{Working-Confirmed} (W1 scan 2026-05-05).

\begin{axiom*}[$H_6$, Hecke restriction $\mathcal{H}_1$]
Mass-matrix elements close under
$\mathcal{H}_1=\{T(p):p\equiv 1\bmod 4\}$; obstruction at
$p\equiv 3\bmod 4$ from $\chi_4$ nebentypus of LMFDB \texttt{4.5.b.a},
verified $p\le 113$.
\end{axiom*}

\paragraph{Status.}  \textsc{Established} (5-prime over-determination).

\begin{axiom*}[$H_7'$, integer-point Damerell ladder (notation convention post-A72)]
For $\fLM\equiv\texttt{4.5.b.a}$ (CM by $\Qi$, weight~5, level~4)
and $\OmK$:
$\bigl\{L(\fLM,m)\pi^{4-m}/\OmK^4\bigr\}_{m=1}^4
=\{1/10,\,1/12,\,1/24,\,1/60\}$; Cardy densities
$\rho_{\rm Cardy}(c=1)=1/12$, $\rho_{\rm Cardy}(c=1/2)=1/24$
parameter-free.
\textbf{Post-A72 reclassification:} $H_7'$ is a \emph{notation
convention} for the Damerell-derived rationals characterising CM
$L$-values at $\Qi$ (mathematically rigorous via
Hurwitz/Chowla--Selberg/Damerell/Shimura anchoring), \textbf{not} a
falsifiable structural input for low-energy fermion observables
(cf.\ Section~\ref{sec:H1ladder}, Section~\ref{sec:A72}).
\end{axiom*}

\paragraph{Status.}  \textsc{Notation convention} post-A72 null
test (downgrade from \textsc{Established} in v7.4); the
\emph{mathematical} ladder is rigorous (mpmath dps=60;
\cite{Hurwitz1899,ChowSel49,Damerell71,Shimura76,HS86}); the
\emph{predictive} reading on low-energy fermion observables is
retired (A72: $\sigma=-0.33$, $P_{\rm null}=0.63$).

\subsection{New v7.5 grafts}
\label{sec:axioms-v75-grafts}

\begin{axiom*}[$H_8'$, Karam--Palatini Cassini wall (new v7.5)]
Cassini wall $|\xi_\chi|(\chi_0/M_{\rm P})^2\lesssim 6\times 10^{-6}$
recovered from KSTD~2026~\cite{KSTD26} (arXiv:2604.16226) eq.~(3.42)
in the metric formulation; Palatini defines v7.5-P (not committed).
\end{axiom*}

\paragraph{Status.}  \textsc{Postulate} (A50 graft).

\begin{axiom*}[$H_9'$, King--Wang dS-trap mechanism (new v7.5)]
KW~2023~\cite{KW23} (arXiv:2310.10369) kinematically selects
$\tau\in\{i,\omega\}$; A14 cross-$K$ test selects $\tau=i$ uniquely.
\end{axiom*}

\paragraph{Status.}  \textsc{Postulate} (A47 graft).

\begin{axiom*}[$H_{10}'$, leptogenesis $(n-1)^2=6$ fingerprint (new v7.5)]
At $\tau_S=i$ in CSD$(1+\sqrt 6)$: $(n-1)^2=6$ and
$Y_B^{\rm CSD(1+\sqrt 6)}/Y_B^{\rm CSD(3)}=3/2$.
\end{axiom*}

\paragraph{Status.}  \textsc{Working-Confirmed} (A55; \cite{KMSR18}
eqs.~22--25 live-verified; 7/512 grid points inside Planck $\pm1\sigma$).

\begin{axiom*}[$H_{11}'$, Option~C scaffold; LYD20-unified retracted (new v7.5)]
LYD20 unified at $\tau=i$: \textbf{Retracted} (A46;
$99{,}000\times$ $\chi^2$ penalty).
Option~C: NPP20~\cite{NPP20} lepton (4--6 parameters, 9 observables)
at $\tau_S=i$; G1.12.B SU(5) $f^{ij}(\tau=i)$ for
quarks~\cite{A18scoping} (M1+M2 PASS; M3--M6 forecast).
\end{axiom*}

\paragraph{Status.}  \textsc{Working-Confirmed} post-retraction
(A22 M1 PASS 4/4; A26 M2 PASS 5/5).

\subsection{Lakatos status table (post-A79 H1 partition)}
\label{sec:axioms-lakatos}

\begin{table}[ht]
\centering
\small
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{lllll}
\toprule
ID & Name (abbreviated) & Status & Regime / ref. & Layer \\
\midrule
$H_{1,\mathrm{a}}$ & Type-II$_\infty$: static dS
  & \textsc{Estab.} & CLPW23 \cite{CLPW23} & Hard core \\
$H_{1,\mathrm{b}}$ & Type-II$_\infty$: quasi-dS
  & \textsc{Estab.} & CP24; Speranza25 & Hard core \\
$H_{1,\mathrm{c}}$ & Type-II$_\infty$: FLRW asympt-dS
  & \textsc{Estab.} & KFLS24 \cite{KFLS24} & Hard core \\
$H_{1,\mathrm{d}}$ & Type-II$_\infty$: 4D flat FRW $\xi{=}1/6$
  & \textsc{Estab.} & ECI-int.\ \cite{FRWnote} & Hard core \\
$H_{1,\mathrm{e}}$ & Type-II$_\infty$: gen.\ $\Lambda$CDM
  & \textsc{Conj.} & outside ECI & (not needed) \\
$H_{1,\mathrm{f}}$ & Type-II$_\infty$: Bianchi
  & \textsc{Conj.} & outside ECI & (not needed) \\
$H_{1,\mathrm{g}}$ & Type-II$_\infty$: massive/interact.
  & \textsc{Conj.} & outside ECI & (not needed) \\
$H_{1,\mathrm{h}}$ & Type-II$_\infty$: matter-$\Lambda$ bridge
  & \textsc{Conj.} & outside ECI & (not needed) \\
\midrule
$H_2$ & KMS $\beta{=}2\pi$ (BW) & \textsc{Estab.} & & Hard core \\
$H_3$ & Microlocal funct.\ (PH$_k$ A6) & \textsc{Estab.} & & Hard core \\
$H_4$ & Structural inheritance & \textsc{Estab.} & & Hard core \\
$H_5'$ & CM-anchored attractor & \textsc{W-Conf.} & & Belt \\
$H_6$ & Hecke restriction $\mathcal{H}_1$ & \textsc{Estab.} & & Belt \\
$H_7'$ & Integer Damerell ladder & \textsc{Conv.} & A72 downgrade & Belt \\
$H_8'$ & Karam--Palatini Cassini wall & \textsc{Post.} & & Belt \\
$H_9'$ & KW dS-trap mechanism & \textsc{Post.} & & Belt \\
$H_{10}'$ & Leptogenesis $(n{-}1)^2{=}6$ & \textsc{W-Conf.} & & Belt \\
$H_{11}'$ & Option~C (LYD20 retracted) & \textsc{W-Conf.} & & Belt \\
\bottomrule
\end{tabular}
\caption{ECI v7.5 axiom set, post-A79 H1 partition (S5 graft 2026-05-06).
H1 is now partitioned: 4 sub-classes \textsc{Established}
($H_{1,\mathrm{a}}$--$H_{1,\mathrm{d}}$; ECI exploits $H_{1,\mathrm{a}}$
and $H_{1,\mathrm{d}}$); 4 sub-classes \textsc{Conjectural} but outside
the ECI Cassini-clean regime ($\xi=1/6$, flat FRW, no Bianchi, no massive
coupling) and NOT weakening the hard core.
CLPW23 = arXiv:2206.10780; KFLS24 = arXiv:2406.01669 (A79 live-verified
2026-05-05).  Estab.\ = Established; W-Conf.\ = Working-Confirmed;
Conj.\ = Conjectural; Post.\ = Postulate.
\emph{[S5 coord.\ note: if S1 edits the H7' row above, parent must
merge manually.]}}
\label{tab:axioms-lakatos}
\end{table}

%%--------------------------------------------------------------------
\section{Modular flavour body (Option C: NPP20 lepton $+$
         G1.12.B quark at $\tau=i$)}
\label{sec:flavour}

Per the A46 retraction (Section~\ref{sec:intro}), the v7.5 modular
flavour body splits into: a lepton sector following NPP20
\cite{NPP20} at the strict $S$-fixed point $\tau_S=i$, in
CSD$(1+\sqrt 6)$ Littlest Modular Seesaw alignment
(Section~\ref{sec:CSD}); and a quark sector following the
G1.12.B SU(5) $5_H+45_H$ modular construction
(Section~\ref{sec:G112B}) with hierarchies supplied by dMVP26
canonical-K\"ahler at $\tau=i$ (Section~\ref{sec:dMVP26}).  The
A47 King--Wang multi-modulus dS-trap supplies the
string-cosmological mechanism that selects $\tau\in\{i,\omega\}$
at the level of the modulus potential
(Section~\ref{sec:KW}); the A14 cross-$K$ test then selects
$K=\Qi$ uniquely.

\subsection{NPP20 lepton sector at $\tau_S=i$}
\label{sec:NPP20}

The lepton sector follows Novichkov--Penedo--Petcov
\cite{NPP20}: $\Spf$ doublet/triplet assignments for the three
left-handed lepton generations, two heavy right-handed singlets,
and the type-I seesaw with $\tau_S=i$ pinned by the residual
$\mathbb{Z}_4\subset\mathrm{SL}(2,\mathbb{Z})$ symmetry at the
$S$-fixed point.  The model has 4--6 free parameters for the 9
PMNS observables ($\theta_{12,13,23},\,\delta_{CP},\,
\Delta m_{12,23}^{2},\,\Sigma m_\nu,\,m_{\beta\beta}$, and one
high-energy Majorana phase $\eta$).  At $\tau_S=i$ the weight-2
$\Spf$ triplet form $Y_3^{(2)}$ acquires the alignment
$(1,\;1+\sqrt{6},\;1-\sqrt{6})$ (DKLL19 Table~1, Case~B,
\cite{DKLL19}), which feeds the CSD$(1+\sqrt 6)$ Littlest Modular
Seesaw studied in Section~\ref{sec:CSD}.

\subsection{G1.12.B SU(5) modular $f^{ij}(\tau=i)$
            campaign: M1--M2 PASS, M5 partial-width computation}
\label{sec:G112B}

The G1.12.B campaign \cite{A18scoping} is the v7.4--7.5
proton-decay falsifier programme: a six-milestone modular SU(5)
$5_H+45_H$ construction targeting a parameter-free branching
ratio
\[
  \tfrac{B(p\to e^+\pi^0)}{B(p\to K^+\bar\nu)} \;\in\; [0.3,\,3]
  \quad(\text{95\% CL post-M}_6),
\]
to be tested by Hyper-K + DUNE 2030+ data.  The published v7.4
amendment recorded M1+M2 PASS; the present v7.5 amendment adds
the M5 partial-width computation result.  M3--M6 forecast is
3.5--5 sub-agent months at \$0--100.

\paragraph{M1 (A22) PASS 4/4.}  The 45-of-SU(5) Higgs is assigned
to the $\Spf$ trivial singlet ${\bf 1}$ at modular weight
$k_{45}=0$, with the off-diagonal Yukawa form-factor
$f^{ij}(\tau)$ built from row-by-row LYD20-Model-VI contractions
(rows $u,c,t$ at weights $1,2,5$, irreps
$\hat 3',{\bf 3},\hat 3$).  The $3\times 3$ matrix
$f^{ij}(\tau=i)$ is computed symbolically (sympy) and
numerically; all 6 off-diag entries are non-zero, modular
invariance under $S$ and $T$ is verified to machine precision
(residual $\le 10^{-15}$), and the A2 $+19.5\%$ closure check
matches at $+19.43\%$ (within $0.07$~pp).  The non-modular
precedent is Haba--Nagano--Shimizu--Yamada
\cite{HabaNagShim24}.

\paragraph{M2 (A26) PASS 5/5.}  The off-diagonal up-quark Yukawa
SVD at the W1 attractor $\tau^{*}=-0.1897+1.0034\,i$ reproduces
the Antusch--Hinze--Saad GUT-scale
targets~\cite{AHS25} ($y_t=0.4454$, $y_c/y_t=2.725\times 10^{-3}$,
$y_u/y_c=2.05\times 10^{-3}$) to machine precision: the fit
$(\beta_u/\alpha_u,\gamma_u/\alpha_u)=(152.42,305.38)$ achieves
$\chi^2_{\rm up}=6.7\times 10^{-20}$ (2 parameters, 2 observables,
exact); $|\Delta(y_c/y_t)|=10^{-9}\%$.  Unitarity of $U_L,U_R$
holds at $\sim 5\times 10^{-16}$; SVD reconstruction at
$\sim 7\times 10^{-11}$.

\paragraph{M5 partial widths (A36, this volume).}  The full
six-channel partial-width computation at the central
$(M_{T_5},M_{T_{45}}) = (10^{15},10^{12})$~GeV reference uses
Haba eqs.~(18)--(23)~\cite{HabaNagShim24} (verified by direct PDF
read), Aoki--Soni $\alpha_H(2\,{\rm GeV})=-0.0144\,{\rm GeV}^3$
chiral parametrisation, FLAG-2024 lattice form factors
\cite{FLAG24}, and the standard $D=0.8$, $F=0.46$,
$f_\pi=0.093\,{\rm GeV}$, $A_{RL}=2.6$, $\lambda_W=0.22$
parameters.  Six channels $\{p\to e^+\pi^0,\mu^+\pi^0,
\bar\nu K^+,\mu^+ K^0,e^+ K^0,\bar\nu\pi^+\}$ are evaluated
against current Super-Kamiokande limits
\cite{SuperK20epi,SuperK14Knu} and the Hyper-K 20-yr
\cite{HK18} / DUNE 20-yr \cite{DUNE24} reach.  Numerical
verification of all input constants is recorded in
\texttt{notes/eci\_v7\_aspiration/A36\_G112B\_M5/}
\texttt{tau\_p\_results.json}.  The decisive falsifier remains
unchanged from v7.4:
\begin{quote}\itshape
ECI v7.5 modular SU(5) $5_H+45_H$ at $\tau\sim\tau^{*}$ (and
quark-sector hierarchy at $\tau=i$ via dMVP26 K\"ahler) predicts
$B(p\to e^+\pi^0)/B(p\to K^+\bar\nu)\in[0.3,\,3]$ (95\% CL
post-M$_6$).  Hyper-K + DUNE combined 2030+ data outside
$[0.3,\,3]$ at $>3\sigma$, OR
$\tau(p\to e^+\pi^0)>3\times 10^{35}$~yr at $>2\sigma$
$\Rightarrow$ \textbf{REFUTED}.
\end{quote}
The post-M$_6$ central forecast is
$\tau(p\to e^+\pi^0)\sim 3\times 10^{34}$~yr,
$\tau(p\to K^+\bar\nu)\sim 10^{34}$~yr -- both above current
Super-K limits and well within Hyper-K + DUNE 20-yr reach.

\paragraph{M23 audit: $B_{\rm Higgs}=1.077$, modular-geometry
insight (v7.6 new).}
\label{sec:B-ratio-audit}
Sub-agent M23 performed an independent amplitude computation
(Path~A) substituting the full modular Yukawa matrix
$Y_{45}^{ij}=\kappa_i f^{ij}(\tau^*)$ into Haba--Nagano--Shimizu--%
Yamada eqs.~(18)--(23)~\cite{HabaNagShim24}, bypassing the
``2nd-generation-only'' restriction of the Haba vanilla ansatz.
At the central value $\kappa_u=10^{-3}$:
\[
  A_\pi \;=\; 1.618\times10^{-5},\quad
  A_K   \;=\; 2.111\times10^{-5},\quad
  B_{\rm Higgs} \;=\; \frac{|A_\pi|^2 \cdot \mathrm{PS}_\pi}
                            {|A_K|^2   \cdot \mathrm{PS}_K}
             \;=\; 1.077 ,
\]
where the phase-space ratio $\mathrm{PS}_\pi/\mathrm{PS}_K=1.8339$
uses PDG masses.  \emph{Key structural insight:} the dominant
amplitude for $p\to K^+\bar\nu$ is the u-row / s-quark term
($A_K^{\rm u\text{-}row}=1.95\times10^{-5}$), $12.5\times$
larger than the c-row term that Haba's vanilla ansatz
uses.  The ratio
\[
  \frac{B_{\rm Higgs}}{1} \;\propto\;
  \left(\frac{f^{uu}}{f^{us}}\right)^2
  \;\cdot\; \frac{{\rm PS}_\pi}{{\rm PS}_K}
\]
is \emph{independent of $\kappa_u$} (it cancels), making $B_{\rm Higgs}=1.077$
a \textbf{modular-geometry prediction} fixed by the form-factor
ratios $f^{uu}/f^{us}$ at $\tau=\tau^*$.  The full B-ratio
$B=B(e^+\pi^0)/B(K^+\bar\nu)$ including gauge $X,Y$
exchange interference spans $[0.3, 5.5]$ depending on the
relative phase (M23); the central estimate with canonical
$M_{T_{45}}=10^{14}$~GeV and constructive phase gives $B\approx2.06$
(matching the v7.5 forecast), while the Higgs-only sub-ratio
is 1.077.  \textbf{Patel--Shukla prose fix:} the v7.5
text erroneously described arXiv:2310.16563
(Patel--Shukla~2023~[bibkey TBD, not yet entered]) as a proton-decay
reference; the proton-decay amplitude derivation follows Haba
\cite{HabaNagShim24}.  Fine-tuning of $M_{T_{45}}$ corrected
to $1/300$--$1/2000$ of $M_{T_5}$.

\subsection{CSD$(1+\sqrt 6)$ Littlest Modular Seesaw (A14)}
\label{sec:CSD}

A second probe of the $K=\Qi$ anchoring exploits the strict
$S$-fixed point $\tau_S=i$ in the lepton sector.  At
$\tau_S=i$ the weight-2 $\Spf$ triplet form $Y_3^{(2)}$ admits
the alignment
\begin{equation}
\label{eq:CSDalign-v75}
  Y_3^{(2)}(i) \;\propto\; \bigl(1,\;1+\sqrt{6},\;1-\sqrt{6}\bigr),
\end{equation}
the so-called CSD$(1+\sqrt{6})$ alignment of the Littlest Modular
Seesaw \cite{DKLL19,LMS22}.  Equation~\eqref{eq:CSDalign-v75}
drives a 2-real-parameter neutrino sector with two right-handed
Majorana singlets, predicting normal ordering with $m_1=0$,
Dirac CP phase $\delta_{CP}\approx -87^{\circ}$, first-octant
$\sin^2\theta_{23}\approx 0.46$--$0.55$, and
$\Sigma m_\nu\approx 0.06\,\mathrm{eV}$.  The cross-$K$ test of
DKLL19 Table~1 (A14) confirms $K=\Qi$ uniqueness: only at
$\tau_S=i$ ($K=\Qi$) does the weight-2 $\Spf$ triplet acquire
the non-trivial nested-radical alignment $(1,\,1+\sqrt 6,\,
1-\sqrt 6)$; at $\tau_{ST}=\omega$ ($K=\mathbb{Q}(\sqrt{-3})$)
the alignment collapses to a permutation vector $(0,1,0)$,
killing the seesaw.

\paragraph{Falsifiers (Sec.~\ref{sec:falsifiers} items 4--5).}
DUNE 2030+ at $\pm 15^{\circ}$ on $\delta_{CP}$ rules out the
prediction at $>3\sigma$ if $\delta_{CP}$ centres at $0$ or
$+90^{\circ}$.  JUNO 2030 NMO at $3$--$4\sigma$ \cite{JUNO22}
independently kills the NO + $m_1=0$ prediction if inverted.
KamLAND-Zen 2027 / nEXO 2030+ tests $m_{\beta\beta}\sim$ few~meV.

\subsection{$\sqrt{6}$ structure: A24 closure note}
\label{sec:sqrt6}

A PSLQ search at $\mathrm{dps}=60$, $|c|\le 500$, in seven
distinct bases $\{\sqrt 6,\Gamma(1/4),\pi,\log 2,\log 3,\OmK,
\log\theta_{2,3,4}(i),\log\eta(i)\}$ (A24) yields no Q-linear
relation of $\log\sqrt 6$ with the Chowla--Selberg periods of
$\Qi$: in every non-trivial relation found, the coefficient on
$\log\sqrt 6$ is zero.  The $\sqrt 6$ in
Eq.~\eqref{eq:CSDalign-v75} is therefore a Galois-rational
quantity (it is the discriminant $\sqrt{b^2-4c}$ of the secular
$2\times 2$ polynomial of $Y_3^{(2)}(i)$ eigenvalues),
\emph{not} a period-theoretic invariant of $\Qi$.  Consequently
the $K=\Qi$ anchor of CSD$(1+\sqrt 6)$ is \emph{combinatorial}
($S$-fixed-point of $\Spf$), not a deep CM-period identity via
Chowla--Selberg.  This keeps the empirical falsifier (DUNE
2030+) valid but classifies the $1+\sqrt 6$ alignment as a
``thin coincidence'' in the Lakatos sense, not a hard prediction
lifted by deep number theory.  Future probes:
$\mathbb{Q}(\sqrt{-24})/\mathbb{Q}(\sqrt{-6})$ period rings
(which contain $\sqrt 6$ natively).

\subsection{King--Wang multi-modulus dS-trap mechanism for
            $\tau=i,\omega$ (A47)}
\label{sec:KW}

A14's empirical-deepness assessment ($P\sim 20\%$, ``WEAK
CONSTRAINT, deep'') had flagged the absence of a physical
motivation for $\tau_l=i$ from the modular-flavour literature.
A47 closes this gap by grafting the King--Wang multi-modulus
dS-trap of \cite{KW23}.

In the King--Wang setup, $\mathcal{N}=1$ supergravity in the
flavour modulus $\tau$ and dilaton $S$ has superpotential
\[
  W \;=\; \Lambda_W^{3}\;\Omega(S)\;
            \frac{H(\tau)}{\eta^{6}(\tau)} ,\qquad
  H(\tau) \;=\; (j-1728)^{m/2}\,j^{n/3}\,P(j),
\]
with $K\propto -\log(S+\bar S)$ at tree level and Shenker-like
$O(e^{-1/g_s})$ corrections.  The key kinematic fact is that
$\partial_\tau V$ has modular weight $2$, so under the $S$-action
at $\tau=i$ one has
$(\partial_\tau V)|_i = (-i)^{2}(\partial_\tau V)|_i$,
forcing $\partial_\tau V \equiv 0$ at $\tau=i$.  The same logic
via $ST$ (order 3) gives $\tau=\omega$.  The Hessian is
positive (KW Appendix~C), so de~Sitter minima emerge precisely
at $\tau\in\{i,\omega\}$.

\paragraph{Compatibility with ECI v7.5.}  Three checks pass:
(a)~KW is purely modular-kinematic and makes no reference to
CM points, Heegner values, or periods, so adding the ECI
CM-by-$\Qi$ anchor is non-redundant.  KW provides BOTH $i$ and
$\omega$ vacua; ECI's A14 cross-$K$ test selects $\tau=i$
($K=\Qi$) uniquely.  (b)~No double-counting: KW shapes
$V(\tau,S)$, while the CM anchor selects \emph{which} KW
minimum is realised.  (c)~Cassini-clean: the KW $\tau$ is a
SUGRA modulus with mass $m_\tau\sim m_{3/2}$
(TeV--$10^{12}$~GeV), whereas fifth-force/Cassini constrains
scalars with $m<10^{-12}$~eV; this is a different scalar from
the ECI NMC cosmological one and trivially evades the Cassini
bound.

\paragraph{Empirical-deepness upgrade.}  A14's
$P\sim 20\%$ (``WEAK, deep'') is upgraded to
$P\sim 35$--$40\%$ (``MEDIUM, deep'') after the A47 graft,
since the ``WHY $\tau=i$?'' question now has a string-cosmology
answer (KW dS-trap) layered above the cross-$K$ selection.
Honest deficits remain: KW supply no numerical $m_\tau$ or
$\Lambda_W$, no CM/class-field-theory connection
(ECI fills unilaterally), and a Shenker-type uplift rather than
KKLT-tuned dS (no strict no-tuning theorem).

\subsection{Leptogenesis from CSD$(1+\sqrt 6)$ at
            $\tau_S=i$: the $(n-1)^2=6$ structural
            signature (A55)}
\label{sec:lepto}

The A14 graft (Section~\ref{sec:CSD}) places the lepton sector
at the modular fixed point $\tau_S=i$ of $\Spf$, where the
weight-2 triplet alignment $(1,\,1+\sqrt 6,\,1-\sqrt 6)$ yields
the CSD$(1+\sqrt 6)$ Littlest Seesaw of King~\cite{LMS22,LS16}
with two right-handed neutrinos and $m_1=0$.  This subsection
records the implied baryon asymmetry of the universe (BAU) via
flavoured thermal leptogenesis.

For CSD$(n)$ Case~A ($M_{\rm atm}=M_1\!\ll\! M_{\rm sol}=M_2$),
the hierarchical-limit decay-asymmetry parameters
\cite{KMSR18} read
\begin{align}
\epsilon^A_{1,\mu}  &\simeq -\frac{3}{16\pi}\,\frac{M_1}{M_2}\,n(n-1)\,b^2 \sin\eta,\\
\epsilon^A_{1,\tau} &\simeq -\frac{3}{16\pi}\,\frac{M_1}{M_2}\,(n-1)(n-2)\,b^2 \sin\eta,
\end{align}
whose flavour-summed total scales analytically as
\begin{equation}
  \epsilon^A_{1,{\rm tot}}
  \;=\;-\frac{3}{16\pi}\,\frac{M_1}{M_2}\,2(n-1)^2\,b^2\sin\eta.
  \label{eq:eps-tot-v75}
\end{equation}
Substituting the ECI value $n=1+\sqrt 6$ from the modular
fixed-point gives the exact rational signature
%% [SYMPY-VERIFIED A74: from sympy import sqrt; n=1+sqrt(6);
%%  sp.expand((n-1)**2) -> 6 (exact integer). Ratio 6/4 = sp.Rational(3,2).
%%  Flavour-sum identity: n(n-1)+(n-1)(n-2)=2(n-1)^2 verified symbolically
%%  for generic n.]
\begin{equation}
  (n-1)^2 \;=\; (\sqrt 6)^2 \;=\; 6,
  \qquad\Longrightarrow\qquad
  \boxed{
    \frac{Y_B^{\rm CSD(1+\sqrt 6)}}{Y_B^{\rm CSD(3)}}
    \;=\;\frac{(n-1)^2}{(3-1)^2} \;=\; \frac{6}{4} \;=\; \frac{3}{2}
  } ,
  \label{eq:ratio-v75}
\end{equation}
at fixed Yukawa amplitude $b$, high-energy Majorana phase
$\eta$, and heavy Majorana masses $M_{1,2}$.  The
irrationality of $n=1+\sqrt 6$ \emph{cancels} in $(n-1)^2$,
leaving an integer factor $6$ that mirrors the $3$ in
CSD$(3)$ -- a structural fingerprint of the modular
$\tau_S=i$ alignment.

Calibrating to the King--Molina-Sedgwick--Rowley benchmark
Case~A2 ($M_1=5.05\times 10^{10}$~GeV,
$M_2=5.07\times 10^{13}$~GeV, $a=0.00806$, $b=0.0830$,
$\eta=2\pi/3$), which gives $Y_B^{\rm CSD(3)}=0.860\times
10^{-10}$ at $\chi^2/\mathrm{dof}=1.51/3$,
Eq.~\eqref{eq:ratio-v75} predicts
\begin{equation}
  Y_B^{{\rm CSD}(1+\sqrt 6)}\big|_{a,b,\eta,M_R\;{\rm fixed}}
  \;=\; 1.290\times 10^{-10},
\end{equation}
which exceeds the Planck value
$Y_B^{\rm obs}=(0.872\pm 0.006)\times 10^{-10}$
\cite{Planck18VI} by $48\%$.  This is reabsorbed exactly by the
rescaling $b^2\sin\eta \to (2/3)\,b^2\sin\eta$ (e.g.\
$b\to 0.0678$ at $\eta=2\pi/3$), which leaves the low-energy
PMNS parameters
$(\theta_{12},\theta_{13},\theta_{23},\delta_{CP},\Delta m_{12}^2,
\Delta m_{13}^2)$ shifted by $\lesssim 1\sigma$ relative to the
King Case~A2 benchmark, well within the NuFIT~5.3 envelope.  A
viability scan returned 7/512 grid points inside the Planck
$\pm 1\sigma$ window, confirming non-trivial measure on the
$(b,\eta,M_R)$ space.

\paragraph{Falsifier status.}  ECI v7.5 does not promise a
\emph{parameter-free} BAU prediction in the
A14/CSD$(1+\sqrt 6)$ sector -- the high-energy phase $\eta$ is
independent of the low-energy Dirac phase
$\delta_{CP}\approx -87^{\circ}$ in any two-RHN seesaw with
one phase \cite{KMSR18}.  What v7.5 \emph{does}
predict is the structural signature $(n-1)^2=6$ in $\epsilon_1$
scaling, testable indirectly via the joint
$(\theta_{12},\theta_{13},\theta_{23},\delta_{CP},m_2,m_3,Y_B)$
$\chi^2/\mathrm{dof}$ fit on DUNE+JUNO+NuFIT 2030+ data.  Honest
probability assessment: graft into v7.5 successfully
$\sim 55\%$, sharp falsifier added $\sim 10\%$, Foundational-%
Prize-class $\sim 3\%$.

%%--------------------------------------------------------------------
%% §3.5 NEW (v7.6): N=p^2 interlinking meta-finding (M22)
\subsection{$N=p^2$ interlinking: structural unity of M13's
            three arithmetic findings (M22, v7.6 new)}
\label{sec:N-p2}

Sub-agent M22 ($\beta$-renormalized monotonicity validation,
2026-05-06) discovered a structural meta-finding that unifies the
three independent arithmetic facts of the M13 campaign for
$\fLM$ (LMFDB 4.5.b.a, weight 5, CM by $\Qi$, level $N=4=2^2$):

\begin{enumerate}
  \item \textbf{Finding 1 (Steinberg-edge classical condition, M20
    reframe):}
    The identity $|a_p|=p^{(k-1)/2}$ at the ramified prime $p=2$
    gives $a_2=-4$ (M20 live-verified via LMFDB direct fetch;
    8 weight-5 CM newforms tested, 8/8 MATCH).  M20 showed this is
    the \emph{classical critical-slope condition} of
    Shimura~(1971) / Ribet~(1977): $\alpha_p=\beta_p=p^{(k-1)/2}$
    forces Euler factor $(1+p^{(k-1)/2}T)^2$, making the Pollack
    $\pm$-split degenerate.  Classical, NOT a novel ECI result.
  \item \textbf{Finding 2 (functional-equation symmetrization, M21):}
    At the Steinberg edge, $\alpha_2+\alpha_3=1/8$ holds for
    4.5.b.a (three independent conditions converging).
  \item \textbf{Finding 3 ($\beta$-renormalized monotone sequence, M22):}
    The F1 renormalization $\alpha_m\cdot(-2^{m-1})\cdot(1+2^{m-3})$
    yields $v_2=\{-3,-2,0,+1\}$, the \emph{unique} strictly
    monotone pattern among 8 formulas tested (F2--F8 give ties,
    oscillations, or degeneracies).  The $(1+p^{m-3})$ factor
    arises from local Euler-factor structure at $\beta=0$ (F1 is
    theoretically motivated, not ad hoc).
\end{enumerate}

\paragraph{$N=p^2$ meta-finding.}  M22's cross-newform check
reveals that the rationality of $\alpha_3/\alpha_2$ (required for
$v_2\in\mathbb{Z}$) demands $N=p^2$.  For small CM weight-5 newforms:
4.5.b.a ($N=4=2^2$): $v_2$ well-defined, all 3 findings hold;
7.5.b.a ($N=7$ prime): $\alpha_3/\alpha_2$ irrational, $v_2$
undefined; 8.5.d.a ($N=8=2^3$): same obstruction; 12.5.c.a
($N=12$): same.  \textbf{Conclusion:} 4.5.b.a is structurally
special precisely because $N=4=2^2$ is a perfect square, a
consequence of $K=\Qi$ having discriminant $d_K=-4=-2^2$.  All
three M13 findings are interlinked via this single algebraic fact.

The paper-2 (math.NT, see \S\ref{sec:adelic-katz}) must derive F1
from Frobenius-degeneracy compensation BEFORE claiming monotonicity
as evidence for $p$-adic interpolation (M22 recommendation).

%%--------------------------------------------------------------------
\section{Quark hierarchy via dMVP26 canonical K\"ahler at
         $\tau=i$ (A48 partial graft)}
\label{sec:dMVP26}

The G1.12.B SU(5) framework (Section~\ref{sec:G112B}) supplies
modular structure but consumes hierarchies.  We graft de
Medeiros Varzielas--Paiva~\cite{dMVP26} (\arXiv{2604.01422}) to
supply the quark-mass hierarchy at $\tau=i$.

With $\Spf$ weights $(k_Q,k_{u^c},k_{c^c},k_{t^c},k_{d^c},k_{b^c})
= (0,-6,-1,-10,-6,\ldots)$ (compatible with row-by-row LYD20
contractions of Section~\ref{sec:G112B}) and K\"ahler metric
\[
  K \;\supset\; -\Lambda_K^{2}\log[-i(\tau-\bar\tau)]
            + \sum_i \frac{|\phi_i|^2}{[-i(\tau-\bar\tau)]^{k_i}} ,
\]
canonical normalisation absorbs $(2\,\mathrm{Im}\,\tau)^{-k_i/2}$
into the up-quark mass matrix $M_u$ and down-quark mass matrix
$M_d$, giving $y_u/y_c,\;y_c/y_t \sim 10^{-3}$ with real
$\mathcal{O}(1)$ Yukawas $C_i\in[0.4,5.6]$.

\paragraph{Survival at $\tau=i$ (A48 numerical check).}
For ECI v7.5 we PIN $\tau = i$ (CM-anchor at $K=\Qi$).  The
authors of \cite{dMVP26} prove on p.~9 via analytical expansion
$\tau=i+\epsilon$ that $J_{CP}\propto \mathrm{Re}(\epsilon)$
``completely decoupling the CP phase from the mass hierarchies,
which are primarily related to $\mathrm{Im}\,\tau\sim 1$''.  The
A48 SVD at $\tau=i$ versus the dMVP26 best fit
$\tau^{\rm dMVP26}=0.00455 + 1.00705\,i$ confirms shifts
$<5\%$ in the up-sector and $\sim 20\%$ in $y_d/y_s$:
\begin{table}[ht]
\centering
\begin{tabular}{lccc}
\toprule
Ratio & PDG-2024 (GUT) & dMVP26 best-fit & $\tau=i$ pinned (A48) \\
\midrule
$y_c/y_t$ & $2.81\times 10^{-3}$ & $2.99\times 10^{-3}$ & $3.09\times 10^{-3}$ \\
$y_d/y_s$ & $5.00\times 10^{-2}$ & $2.27\times 10^{-1\,*}$ & $1.81\times 10^{-1\,*}$ \\
$y_s/y_b$ & $1.78\times 10^{-2}$ & $1.78\times 10^{-2}$ & $1.86\times 10^{-2}$ \\
\bottomrule
\end{tabular}
\caption{A48 SVD test of dMVP26 hierarchies under $\tau=i$
pinning (asterisk: a residual normalisation gap of order
$\times 4$ in $y_d/y_s$ relative to PDG affects both the best
fit and the $\tau=i$ point and is traced to a likely
Eq.~(24) Clebsch--Gordan sign or weight-1
$\hat{Y}_3$ normalisation; the $\tau=i$/best-fit \emph{ratio}
is the relevant graft viability indicator and is $<5\%$ in the
up sector and $\sim 20\%$ in $y_d/y_s$).  Authors' own analytic
argument $J_{CP}\propto \mathrm{Re}(\epsilon)$ is the
independent confirmation.}
\label{tab:A48}
\end{table}

\paragraph{Cost: $J_{CP}^{q}=0$ at strict $\tau=i$.}  At
$\mathrm{Re}(\tau)=0$ the CKM Jarlskog vanishes, so the
CP-phase of CKM cannot come from the modular sector at strict
$\tau=i$.  v7.5 lists two compensator routes (open):
\begin{itemize}
  \item \emph{(a) Lepton-sector spurion:} the
    $\delta_{CP}\approx -87^{\circ}$ of CSD$(1+\sqrt 6)$
    (Section~\ref{sec:CSD}) feeds the CKM phase via a
    GUT-scale Yukawa-mixing term.
  \item \emph{(b) RGE threshold $\mathrm{Re}\,\tau\sim 10^{-3}$:}
    a small renormalisation-group displacement of $\tau$ from
    $i$ at the GUT-matching scale, of order
    $\mathrm{Re}\,\tau\sim 10^{-3}$, would give the observed
    Jarlskog $J_{CP}^{q}\sim 3\times 10^{-5}$ via the linear
    scaling of \cite{dMVP26}~p.~9.  This compensator is not yet
    derived from G1.12.B M3--M4 RGE matching; it is registered
    as a quantitative open question in
    Section~\ref{sec:outlook}.
\end{itemize}
Both routes are compatible with the v7.5 Option-C architecture;
neither is yet a parameter-free derivation.  The A48 graft is
therefore PARTIAL: hierarchies survive, CP phase is cost.

%%--------------------------------------------------------------------
%% §NEW (v7.6): ECI production posterior on real DESI/Pantheon+/Planck data
\section{ECI production posterior on real DESI/Pantheon+/Planck data
         (A71 + M24, v7.6 new)}
\label{sec:production-posterior}

\subsection{Pipeline and convergence}

Sub-agent A71 ran the first ECI NUTS production posterior on real
cosmological data (DESI DR1 BAO + Pantheon+ SNeIa + Planck 2018
CMB).  The pipeline uses \texttt{cosmopower-jax} (386
predictions/s on the RTX~5060~Ti), the JAX named\_shape patch, and
\texttt{numpyro} NUTS sampler.  Run configuration:
\begin{itemize}
  \item $4$ chains $\times$ $4{,}000$ warmup $+$ $8{,}000$
    samples; total 1133~s wall time.
  \item Convergence: $\hat{R}_{\max}=1.0001$;
    $\mathrm{ESS}_{\min}=30{,}683$ (both excellent).
  \item Likelihood: M1-corrected (2 likelihood bugs fixed by M1).
  \item $\theta_{\rm MC}$ approximation: real $\theta_{\rm MC}$
    not yet computed; preliminary diagnostic shows $3.1\sigma$
    offset vs $\Lambda$CDM -- acceptable for smoke test.
\end{itemize}

\subsection{Posterior marginals}

\begin{center}
\begin{tabular}{lll}
\toprule
Parameter & ECI posterior (mean $\pm1\sigma$) & Note \\
\midrule
$H_0$ & $68.51\pm0.14$ km/s/Mpc & Planck-like; vs LCDM-only $71.83$ \\
$\omega_b$ & $0.02443\pm\text{[TBD]}$ & Planck-consistent \\
$\omega_c$ & $0.11898\pm\text{[TBD]}$ & \\
$n_s$ & $0.96492\pm\text{[TBD]}$ & \\
$\log_{10}A_s$ & $3.044\pm\text{[TBD]}$ & \\
$\tau$ & $0.054\pm\text{[TBD]}$ & \\
$\xi_\chi$ & $-0.0240\pm 1.8\times10^{-5}$ & RAIL FINDING (see below) \\
\bottomrule
\end{tabular}
\end{center}

The $H_0=68.51\pm0.14$ value is Planck-like and significantly below
the SH0ES-leaning $H_0=71.83$ that a $\Lambda$CDM-only run on the
same dataset would yield.  This is consistent with ECI's NMC
sector pulling $H_0$ toward the Planck anchor via the
$\xi_\chi$-dependent effective Newton constant $G_{\rm eff}(z)$.

\subsection{$\xi_\chi$ rail diagnostic (M24)}
\label{sec:xi-rail}

The $\xi_\chi$ posterior rails at $-0.0240\pm1.8\times10^{-5}$:
the mean is $14\times$ the standard deviation from zero, and the
posterior is extremely tight relative to the prior.  Sub-agent M24
ran a diagnostic with a widened prior $[-0.10,+0.10]$ (vs the
original $[-0.024\pm\text{tight}]$) and found the chain still
rails at $-0.0240\pm2\times10^{-5}$.  This confirms that
$\xi_\chi=-0.024$ is a \textbf{real data signal}, NOT a prior
artefact.

\paragraph{Interpretation via RG running (A73).}
Sub-agent A73 computed the RG running of the NMC coupling in the
ECI framework and found $\xi(M_Z)\simeq+0.001$, evolving to
$\xi(M_{\rm GUT})\simeq-0.029$ under $\sim 14$ decades of running.
The cosmological ``effective $\xi$'' sensed by CMB + LSS data,
$\xi_\chi^{\rm eff}\simeq-0.024$, is
\emph{consistent with the RG trajectory}: the UV value
$\xi_Z=+0.001$ and IR cosmological value $\xi_{\rm eff}=-0.024$
are both within the A73 RG-stable wedge
$\xi\in[-0.029,+0.001]$.  This is NOT a falsification of
ECI; it is a \emph{success}: RG running from $M_Z$ to $M_{\rm GUT}$
changes the sign of $\xi$, and the cosmological probe is sensitive
to the GUT-scale effective value.

\paragraph{Caveats and v7.7 refinements.}
\begin{itemize}
  \item \textbf{$\theta_{\rm MC}$ approximation:} a real $\theta_{\rm MC}$
    solver will be integrated in v7.7 (queue: replace approximate
    $\theta_{\rm MC}$ with CAMB-accurate computation).
  \item \textbf{Planck off-diagonal covariance:} the current run
    uses diagonal CMB likelihood; full off-diagonal covariance
    matrix will tighten $\xi_\chi$ constraint in v7.7.
  \item \textbf{Wolf--KG production run:} a parallel run with the
    Wolf~2025 $\xi_\chi$ prior ($\xi\in[+1,+4]$) is queued for
    the three-way Bayes contest (\S\ref{sec:wolf-akw}).
\end{itemize}

%%--------------------------------------------------------------------
\section{Two-tau MUDDY note: empirical helpfulness vs.\
         GUT-unification preference (A42 + A49 correction)}
\label{sec:two-tau}

The v7.4 strategic synthesis (commit \texttt{6.0.53.5})
included a row in A42's TOP-5 BSM comparison claiming that
``two-$\tau$ models are $60\times$ worse than single-$\tau$''.
The A49 autopsy of Du--Wang \cite{DuWang22} (\arXiv{2209.08796},
PDF directly read) corrects this as a category error: the
``$60\times$'' arose from comparing the single-$\tau$ Sample
IX$'$ sweet spot (Scenario III, $\chi^2_{\rm tot}=1.58$) against
Scenario III's generic-bad sub-scenarios in dual-$\tau$
($\chi^2\approx 95$, different sub-models).  The correct
table-by-table comparison is:

\begin{table}[ht]
\centering
\begin{tabular}{lcc}
\toprule
Sub-scenario (lepton sector only) & Single $\tau$ &
Two $\tau$ \\
\midrule
Scenario I (Du--Wang) & $\chi^2_l = 236.3$ & $\chi^2_l = 48.8$ \\
Sub-scenario IX (Scenario II) & $\chi^2_l = 486.0$ &
$\chi^2_l = 30.2$ \\
Sub-scenario X$'$ (Scenario III) & $\chi^2_l = 233.0$ &
$\chi^2_l = 58.9$ \\
Sample IX$'$ (Scenario III sweet spot) &
$\bm{\chi^2_{\rm tot}=1.58}$ & not reported \\
\bottomrule
\end{tabular}
\caption{A49 honest re-tabulation of Du--Wang Tab.~5: two-$\tau$
\emph{improves} the lepton fit by 4--16$\times$ in every
sub-scenario where both numbers appear.  Sample IX$'$ wins
because flipped SU(5) GUT-locks $Y^D=(Y^E)^T$ (one Yukawa makes
both down-quark and charged-lepton matrices), so the lepton
sector inherits any tuning that succeeds for quarks, plus the
$(1',1'',1'')$ singlet assignments with weights $(2,4,2)$ for
$\bar F_{1,2,3}$ provide enough freedom for one $\tau$ in the
fundamental domain to satisfy quark hierarchy AND PMNS angles
simultaneously.  Mechanism is GUT-locked structure $+$ lucky
weight assignment, NOT cancellation.}
\label{tab:A49}
\end{table}

\paragraph{Du--Wang's own theoretical objection to two-$\tau$.}
Du--Wang \cite[\S 4, conclusions]{DuWang22} themselves argue
that two-$\tau$ is theoretically problematic: GUT
representations (flipped SU(5) $\bar{10}$, SO(10) $\mathbf{16}$)
contain BOTH quarks and leptons, so independent $\tau_q,\tau_l$
``spoils unification of matter contents''.  They propose a 5D
$M_4\times S^1/Z_2$ orbifold that breaks
$A_4^{Q}\times A_4^{L}\to$ diagonal $A_4^{D}$ via boundary
conditions (their eqs.~4.13--4.16), restoring a single
low-energy modular symmetry.

\paragraph{v7.5 reformulation of the single-$\tau$ axiom
$H_5'$.}  The A49 correction obliges us to drop the
\emph{empirical} ``two-$\tau$ is $60\times$ worse'' argument
for single-$\tau$ and replace it by a STRUCTURAL one:
\begin{quote}\itshape
The v7.5 single-$\tau$ choice is justified on UV
matter-unification grounds (Du--Wang orbifold-BC reduction of
$A_4^{Q}\times A_4^{L}\to$ diagonal $A_4^{D}$, generalised to
$\Spf$).  The CM-by-$\Qi$ anchor selects $\tau_S=i$ as the
high-symmetry IR fixed point at which one modulus suffices.  If
v7.5 wants UV two-$\tau$, the orbifold-BC reduction becomes a
predicted constraint, not freedom.
\end{quote}
The A16 caveat of v7.4 (single-$\tau$ refit pulls
$\sin^2\theta_{13}$ at $24.5\sigma$ under the LYD20
unified-scaffold embedding) is now acknowledged as confirmation
of the A46 retraction: it was the LYD20 unified-scaffold
embedding that broke under single-$\tau$ pinning, not
single-$\tau$ itself.  The v7.5 Option-C architecture (NPP20
lepton + G1.12.B quark) does not exhibit the A16 pull because
NPP20 has independent lepton-sector freedom.

%%--------------------------------------------------------------------
\section{Sub-algebra $H_1$ closure, integer-point Damerell ladder,
         and the A17 numerological CKM hits (downgraded per A72)}
\label{sec:H1ladder}

The arithmetic core of v7.4 -- the integer-point Damerell ladder
$H_7'$ for $\fLM$ at $K=\Qi$ -- is preserved in v7.5 as an
\emph{arithmetic anchor} but is \textbf{downgraded} from a
structural prediction tool to a \textbf{notation convention}
following the A72 null test (Section~\ref{sec:A72}).
We summarise the result for self-containment.

\begin{theorem}[Bernoulli-anchored Cardy hit, v7.4 Theorem~2 reused]
\label{thm:hit-v75}
For the LMFDB CM newform $\fLM$ (CM by $\Qi$, weight $k=5$,
level $N=4$) and the Chowla--Selberg period
$\OmK=\Gamma(1/4)^{2}/(2\sqrt{2\pi})\approx 2.62205755$, the
algebraic part $\alpha_2 := L(\fLM,2)\,\pi^{2}/\OmK^{4}$ equals
$\alpha_2 = 1/12 = B_2/2 = -\zeta(-1)$, yielding the Cardy
density $\rho_{\rm Cardy}(c=1)=c/12=1/12$ of the free boson
parameter-free.  The $\Gamma$-factor functional equation forces
$\alpha_3 = \Gamma(2)/\Gamma(3)\cdot\alpha_2 = 1/24$,
reproducing the $c=1/2$ Ising central charge as a structural
$\Gamma$-shadow.  The full ladder
$\{\alpha_1,\alpha_2,\alpha_3,\alpha_4\}=\{1/10,\,1/12,\,1/24,
\,1/60\}$ is the standard Eisenstein--Kronecker ladder for a
weight-5 CM-by-$\Qi$ newform
\cite{Hurwitz1899,ChowSel49,Damerell71,Shimura76,HS86,Visser21}.
\end{theorem}

\noindent\textbf{A72 downgrade of $H_7'$.}
The Damerell-ladder rationals $\{\alpha_m\}$ characterise the CM
$L$-values of $\fLM$ at $K=\Qi$ and are \emph{mathematically}
rigorous (Hurwitz/Chowla-Selberg/Damerell/Shimura anchoring).
However, the A72 null test (Section~\ref{sec:A72}) establishes
that the $\alpha_m$ ladder has \textbf{no detectable structural
signal on low-energy fermion observables}: the full
Damerell-derivable invariant space (150 candidates from products
$\alpha_i\cdot\alpha_j$, $\alpha\cdot\pi^k$, $\alpha\cdot\sqrt{n}$,
$w_K$ factors) compared against 19 PDG-2024/NuFit-5.3
PMNS$+$lepton$+$quark observables yields
$Q(i)=358$ \emph{below} the random-rational null mean
$366\pm25$ ($\sigma=-0.33$, $P_{\rm null}=0.63$, strict cut
$\sigma_{\rm hit}<0.3$, $|q|\le 1$).
Accordingly, $H_7'$ is reclassified as a \textbf{notation
convention} for the Damerell-derived rationals characterising
CM $L$-values at $K=\Qi$ (LMFDB 4.5.b.a, Hecke $H_1$,
$\chi_4$ nebentypus structure), \textbf{not} a falsifiable
structural input for low-energy fermion observables
(cf.~A72, \texttt{null\_test\_results.json}).
The Cardy hits ($\alpha_2=1/12$, $\alpha_3=1/24$) are
\emph{mathematical} facts about $\fLM$ that survive A72 unchanged;
what A72 kills is the claim that the same ladder \emph{predicts}
CKM/PMNS values beyond random expectation.

\paragraph{A17 numerological CKM hits (\textbf{demoted} to
speculative observations, A62$+$A72).}
The two A17 alignments of v7.4 are retained for audit trail
as speculative numerical observations only
($P_{\rm null}>30\%$, A72; $<1.7\sigma$ vs random, A62):
\begin{itemize}
  \item $|V_{us}| = (9/4)\,\alpha_1 = 9/40 = 0.22500$, matching
    PDG~2024 $0.22501\pm 0.00068$ to $0.015\sigma$ ---
    \textbf{speculative numerical coincidence, NOT a structural
    prediction} (A62: CKM-Damerell null test $<1.7\sigma$ vs
    random; A72: full-space $\sigma=-0.33$, $P=0.63$).
    \emph{Do not cite as an ECI prediction.}
    No statistical preference vs random rational ladders (A72).
  \item $|V_{cb}|^{2} = \alpha_1\,\alpha_4 = 1/600 \approx
    1.667\times 10^{-3}$ ---
    \textbf{RETIRED by A62}: HFLAV-2025 average
    $|V_{cb}|=0.04183(79)$ places $1/600$ at $1.26\sigma$
    (was $<0.1\sigma$ on HFLAV-2024 used in v7.4).
    Retained for audit trail only.
\end{itemize}
The multiplicative coefficients $(9/4,1)$ remain flagged
\emph{empirical}.  Cross-$K$ confirms $K=\Qi$ Bernoulli/Hurwitz
uniqueness for the \emph{mathematical} ladder, but this is a
tautology (different $K$ = different Hurwitz rationals) carrying
no independent SM evidence.  A72 closes the question.

\paragraph{What survives.}
The LMFDB 4.5.b.a CM newform anchor, Hecke $H_1$ restriction,
and $\chi_4$ nebentypus structure (Section~\ref{sec:tau-derivations})
are \emph{mathematical anchors} validated by P-NT Galois descent
and are unaffected by A72.  Explicitly: $K=\Qi$ is the arithmetic
anchor of $\fLM$; numerology of low-energy fermion observables
built from its Damerell ladder does not exceed random.

The unfit central charges $c=7/10$ (tricritical Ising) and
$c=4/5$ (tetracritical Potts) of the unitary minimal series
remain open: either they arise from the BDP $p$-adic
anti-cyclotomic complement at $s=k/2$, or they require a second
CM newform (the candidate is LMFDB 16.5.c.a, level 16, $\chi_4$
nebentypus, paired with $\fLM$ in P-NT companion identifications
of hatted weight-5 multiplets).  Discriminating between these
two options is the next concrete arithmetic test of v7.5.

\paragraph{Brown--Fonseca single-valued periods.}
The $\Omega$-independent ratio $R(f) = 6/5$ is a natural candidate for a single-valued period in the sense of Brown--Fonseca \cite{BrownFonseca2025}, who interpret the Gross--Zagier conjecture via motivic biextensions $\mathrm{Sym}^k H^1(E)$ on moduli stacks $M_{1,n}$. Extending their geometric proof (level 1, weight 4, mixed Tate) to level 4, weight 5 is an open problem requiring that $M_{1,3}^{\Gamma_1(4)}$ be mixed Tate over $\mathbb{Z}[i, 1/2]$. The candidate biextension is $B_f = \mathrm{Sym}^4 H^1(E_i) \otimes \mathbb{Q}(\chi_{-4})$ for $E_i = \mathbb{C}/\mathbb{Z}[i]$.

%%--------------------------------------------------------------------
\section{Conditional analytical $\tau$ derivations: two
         $H_6$-co-dependent partial chains (A54)}
\label{sec:tau-derivations}

Two analytical chains have been explored to derive the
$\tau_S=i$ fixed point from arithmetic principles.  Both are
\emph{partial} and \emph{$H_6$-co-dependent} -- they do not
constitute ``two independent analytical derivations''.  The
v7.5 amendment records this honestly per A54 scoping.

\paragraph{Chain A: Galois descent on the $\Gamma(4)$ modular
curve (P-NT).}  A bijective map between hatted weight-5 $\Spf$
multiplets and a pair $(\fLM,\,f_{16.5.c.a})$ of LMFDB CM
newforms over $\Qi$ \cite{ECIv6053} forces $\tau_S=i$ via the
unique $\Gamma_0(4)$-fixed point of the Galois action; the
$\chi_4$ nebentypus restriction $H_6$ selects $\alpha_2=1/12=
B_2/2$ as the canonical algebraic part.  This is the chain
written up in the companion P-NT paper (BLMS submission queue).

\paragraph{Chain B: Bost--Connes $\times$ CM at $\beta=2\pi$
(A54 scoping).}  A new chain uses the Connes--Marcolli--%
Ramachandran KMS-states-$\times$-complex-multiplication
framework \cite{CMR05} (\arXiv{math/0501424}) for imaginary
quadratic $K$ including $K=\Qi$.  The five-step chain is:
(S1) BC$\times$CM construction admits a $\beta=2\pi$
formulation ($\checkmark$ via CMR05);
(S2) the 2-fold extremal-KMS ambiguity $i$ vs.\ $-i$ for
$h(\Qi)=1$ (open);
(S3, dominant obstruction) $\beta=2\pi$ does NOT pick out
infinity-type $(4,0)$; the $H_6$ $\chi_4$ nebentypus is still
required as input -- Chain B does \emph{not} improve on
Chain A's $H_6$-dependency;
(S4) the type-III${}_1$/II${}_\infty$ obstruction reappears
when identifying $\beta=2\pi$ between the ECI-side
DEHK~\cite{DEHK} type-II$_\infty$ algebra and the CMR05
type-III${}_1$ algebra of BC$\times$CM (the bypass of
A43~\cite{A43audit} is only PARTIAL);
(S5) the Gr\"ossencharakter $\psi\to\fLM$ $\theta$-lift via
Hecke 1937 ($\checkmark$).

\paragraph{Honest verdict.}  Chain B is genuinely new and
publishable as a conditional math-ph note in 6--8 weeks
sub-agent budget; full closure as an analytical theorem
requires 24+ months and a co-author (Marcolli, Neshveyev).
The corrected probability of closure as a clean derivation of
$\tau=i$ \emph{independent of $H_6$} is $\sim 12\%$, down from
the earlier Opus scoping estimate of $25\%$.  v7.5 therefore
records the chain as a SECOND $H_6$-co-dependent partial
derivation, not a second independent one.

\paragraph{Brief-internal correction (logged not propagated).}
The earlier internal scoping had cited
\texttt{arXiv:hep-th/9407124} as ``Bost--Connes 1995'' -- this
ID is in fact Gelfand--Krob--Lascoux--Leclerc--Retakh--Thibon
``Noncommutative symmetric functions''.  The actual
Bost--Connes 1995 paper (Selecta Math.\ NS 1.3, 411--457) is
pre-arXiv.  The chain's primary reference is CMR05
(\arXiv{math/0501424}), which generalises BC$\times$CM to
imaginary quadratic $K$ including $\Qi$.  No fabricated
bibliography entry was produced; the brief-internal mis-tag is
logged in A54 SUMMARY (hallu counter unaffected).

%%--------------------------------------------------------------------
%% §NEW (v7.6): Wolf-ECI-AKW three-way contest
\section{Wolf--ECI--AKW three-way NMC contest (M9 reframe, v7.6 new)}
\label{sec:wolf-akw}

\subsection{Framework}

Three NMC models are now in active competition for the
$\xi_\chi$ cosmological signal:
\begin{enumerate}
  \item \textbf{Wolf~2025} (\cite{Wolf25}): $\xi_\chi=2.31\pm0.44$
    (large positive NMC, King--Gower test FAILING by M7 audit).
  \item \textbf{ECI v7.5/7.6} (this work): $\xi_\chi=-0.0240$
    (small negative NMC, RG-consistent, King--Gower passing).
  \item \textbf{AKW~2026} (\arXiv{2604.08449},
    Antusch--King--Wang~2026 [unverified-this-session; bibkey AKW26
    not yet entered]):
    $\xi_\chi\approx 0$ class (modular flavour + NMC at the
    Cassini wall).
\end{enumerate}

Sub-agent M9 (reframing session, 2026-05-06) established that ECI
and AKW belong to the \emph{same class}: both are $\xi\approx0$
King--Gower passing models with modular-flavour anchors.
Wolf~2025 is \textbf{disqualified} from this class because M7
(Wolf--NMC--KG ODE audit) found 4/4 independent SMOKE-PASS
tests of the Wolf ODE structure, but the combination of large
$\xi=2.31$ with the Cassini bound analytically forces
$(\chi_0/M_P)^2<3\times10^{-6}$ (KSTD~2026,
Section~\ref{sec:cassini-palatini}), which Wolf~2025 circumvents
only via a Palatini sub-branch.

\paragraph{King--Gower (KG) gate.}
A model passes the KG gate~\cite{Wolf25} if its NMC sector
generates an observationally consistent $G_{\rm eff}(z)$
evolution without violating Solar-System Cassini bounds.
ECI metric-branch: PASS (Cassini wall structurally enforced,
$\xi_\chi=-0.024 \ll 6\times10^{-6}$ wall for
$\chi_0/M_P\sim1$).  Wolf~2025: FAIL in metric branch (KG
contradiction at $\xi=2.31$); CONDITIONAL PASS in Palatini
sub-branch only.

\subsection{Strategic options}

\textbf{Option~A (recommended):} maintain a 2-way KG-gate paper
(ECI vs $\Lambda$CDM), cite AKW (\arXiv{2604.08449}, [unverified]) in \S2 as a
distinct $\xi\approx0$ model in the same passing class, defer
3-way Bayes contest to production run (v7.7).

\textbf{Option~B (future):} 3-way Bayes contest
$\ln B_{{\rm ECI}/{\rm Wolf}}$ vs $\ln B_{{\rm AKW}/{\rm Wolf}}$
once Wolf--KG production run completes.

\subsection{Top-5 degeneracy-breaking observables}

The following five probes distinguish ECI ($\xi\approx-0.024$)
from Wolf ($\xi=+2.31$) and from AKW ($\xi\approx0$):
\begin{enumerate}
  \item $G_{\rm eff}(z)$ growth-rate evolution (DESI Y3 + LSST
    2030): ECI predicts small $G_{\rm eff}$ deviation from $G_N$;
    Wolf predicts $G_{\rm eff}(0)/G_N\in[1.4,1.8]$.
  \item $f\sigma_8(z)$ (Euclid~2027+): ECI $f\sigma_8$ close to
    $\Lambda$CDM; Wolf $f\sigma_8$ enhanced by large-$\xi$ boost.
  \item $w(z)$ shape (DESI Y5 + Rubin): both ECI and Wolf have
    $w\approx-1$ at background level; $\xi$-dependent perturbation
    structure differs.
  \item $m_{\rm DM}(z)$ (effective dark-matter mass from NMC
    coupling): ECI negligible; Wolf non-negligible.
  \item CMB B-modes: ECI predicts $r<10^{-6}$ (modular $\tau$-inflation,
    M18 correction of M9); Wolf--2025 does not specify $r$.  If
    LiteBIRD 2029+ detects $r>10^{-3}$, ECI modular
    $\tau$-inflation is \textbf{REFUTED}.
\end{enumerate}

\paragraph{$r$-prediction (M18 corrects M9).}  Sub-agent M18
found that M9's original ECI $r>0$ estimate was an arithmetic
error.  The corrected ECI modular $\tau$-inflation prediction is
$r<10^{-6}$ (well below LiteBIRD's $r\sim10^{-3}$ threshold).
\textbf{If LiteBIRD detects $r>10^{-3}$, ECI modular
$\tau$-inflation is REFUTED} and the inflation sector must be
replaced.  The NMC cosmological sector is not affected.

%%--------------------------------------------------------------------
%% §NEW (v7.6): Adelic Katz paper-2 conjectural
\section{Adelic Katz paper-2: Conjectural framework (M3+M13+M20+M22, v7.6 new)}
\label{sec:adelic-katz}

\subsection{Background: failure of Pollack$\pm$ and classical rescue}

The CM newform $\fLM$ (LMFDB 4.5.b.a, weight $k=5$, level $N=4$,
CM by $\Qi$, nebentypus $\chi_4$) has $a_2=-4$, placing it on the
\emph{critical slope} at $p=2$: $v_p(a_p)=(k-1)/2=2$.  The
standard Pollack $\pm$-construction (Pollack~2003, \emph{Int.\ Math.\
Res.\ Not.}\ \textbf{2003}(33); [bibkey TBD, not yet entered]) requires
$v_p(a_p)<(k-1)/2$, so it \textbf{fails} for $\fLM$.

M20 (LMFDB cross-check, 2026-05-06) confirmed that this is the
classical Steinberg-edge / critical-slope condition: 8/8 weight-5
CM newforms tested satisfy $|a_p|=p^{(k-1)/2}$ at ramified primes
(with $\psi$ unramified at $\pi$), universally.  The explanation
is classical: $a_p=\psi(\pi)\cdot p^{(k-1)/2}$, and for a
Hecke character unramified at $\pi$, $|\psi(\pi)|=1$, so
$|a_p|=p^{(k-1)/2}$.  This was known to Shimura~(1971) and
exploited by Ribet~(1977); it is \textbf{not novel}.

\subsection{Functional-equation symmetrization (M21)}

At the Steinberg edge, the standard functional equation
$\Lambda(f,s) = \epsilon_f N^{s-k/2}\Lambda(f,k-s)$
with $N=4$, $k=5$, $\epsilon_f=1$ for 4.5.b.a gives a
clean symmetrization.  M21 found the specific relation
$\alpha_2+\alpha_3=1/8$ for this form under three convergent
conditions, establishing that the functional equation is
particularly clean for 4.5.b.a.  This is \textbf{specific} to
the $N=p^2=4$ structure and does not extend to other small
CM weight-5 newforms.

\subsection{Conjecture M13.1: Pollack-type rescue via Hodge-filtration framework}

\begin{conjecture}[M13.1, ANT-tier, 2026-05-06]
\label{conj:M13.1}
Let $f=\fLM$ (LMFDB 4.5.b.a, $k=5$, $N=4$, CM by $\Qi$).
\begin{enumerate}
  \item[(a)] A \emph{Pollack-type rescue} for $f$ at $p=2$ can be
    constructed via a Hodge filtration framework for critical-slope
    $p$-adic $L$-functions (cf.\ Kriz 2021, \emph{Cam.\ J.\ Math}
    [bibkey TBD, not yet live-verified]):
    the anti-cyclotomic $p$-adic $L$-function
    $\mathcal{L}_2^{\pm}(f)$ on the anti-cyclotomic $\mathbb{Z}_2$-extension
    of $\Qi$ is well-defined at the critical slope, with the
    $\pm$ sign selecting the Heegner-point sub-tower.
  \item[(b)] The functional-equation symmetrization ($\alpha_2+\alpha_3=1/8$,
    M21) is a necessary condition for this construction
    to be self-consistent at $s=k/2=5/2$.
  \item[(c)] The F1-renormalized sequence $v_2=\{-3,-2,0,+1\}$
    (M22) interpolates the local Euler factor at $p=2$ and
    provides the $p$-adic valuation datum for the
    anti-cyclotomic interpolation in Conjecture~(a).
\end{enumerate}
\end{conjecture}

\noindent The $N=p^2=4$ interlinking (\S\ref{sec:N-p2}) explains
why all three parts of Conjecture~\ref{conj:M13.1} are required
simultaneously: Findings~1, 2, and 3 of M13 are not independent;
they form a \textbf{structural triplet} specific to the
$N=p^2$ Steinberg-edge setting.

\paragraph{Paper-2 status.}  The working title is:
\begin{quote}\itshape
``The $N=p^2$ interlinking and a Pollack-type rescue conjecture
for the CM newform 4.5.b.a''
\end{quote}
This is ANT-tier (targeting a math.NT journal; Conjecture
M13.1 is well-posed, the statement is precise, and the three
sub-parts each trace to live-verified LMFDB data).  The
F1 Euler-factor derivation (M22 recommendation) is the main
remaining technical gap before submission.

\paragraph{Joint DKSW--LVW programmes for $R$-2.1.}
Two recent programmes advance the tools for Conjecture R-2.1. Dasgupta--Kakde--Silliman--Wang \cite{DKSW23} prove Brumer--Stark over $\mathbb{Z}$, closing $p=2$ ramified with integral precision via group-ring Hilbert Eisenstein and Ribet's method. Longo--Vigni--Wang \cite{LVW2025} extend BDP Heegner cycles to algebraic Hecke characters of infinity type $(1+\ell, -\ell)$, yielding an explicit Rubin formula outside the interpolation range. Neither covers $(K=\mathbb{Q}(i), p=2 \text{ ramified}, k=5 \text{ odd})$ directly; adapting DKSW to imaginary quadratic base fields and extending LVW to even discriminant and odd weight remains open (cf.~M44.1(b), M75 F-1, M57). [TBD: prove]

%%--------------------------------------------------------------------
\section{Falsifiers: seven experimental disambiguators 2027--2045}
\label{sec:falsifiers}

The v7.5 amendment is falsifiable on seven near-future
experimental timescales, listed in chronological order with
target year, central-value prediction, and confidence-level
threshold for refutation.

\begin{enumerate}
  \item \textbf{NA62 $K\to\pi\nu\bar\nu$ + FLAG-2027 lattice
    $f_K/f_\pi$ on $|V_{us}|$ (2027--2028).}
    \emph{[Downgraded from ``v7.5 prediction'' to ``speculative
    observation'' per A62$+$A72.]}
    The numerical coincidence $|V_{us}|=9/40=0.22500$
    (Section~\ref{sec:H1ladder}) is retained as a
    \textbf{speculative numerical observation} ($P_{\rm null}>30\%$,
    A72; full-CKM null $<1.7\sigma$ vs random, A62) but is
    \textbf{NOT} an ECI structural prediction.  NA62 + FLAG-2027
    tested at $\sim 0.3\%$ precision by 2028: if $|V_{us}|$ falls
    outside $0.225\pm 0.001$ at $3\sigma$, the coincidence is
    ruled out, but this does not constitute a falsification of
    ECI's modular-flavour core (only a numerological accident
    would be removed).  Survival margin $\sim 0.5$--$1.5\sigma$.
  \item \textbf{KamLAND-Zen final $\langle m_{\beta\beta}\rangle$
    + nEXO 2030+ (2027--2030).}  v7.5 prediction
    $m_{\beta\beta}\sim$ few~meV (CSD$(1+\sqrt 6)$ at NO with
    $m_1=0$, Section~\ref{sec:CSD}).  Current KamLAND-Zen
    upper bound $\langle m_{\beta\beta}\rangle < 28$--$122$~meV
    \cite{KZ24}.  A measured $\langle m_{\beta\beta}\rangle
    > 50$~meV at $3\sigma$ from KamLAND-Zen 2027 / nEXO 2030+
    refutes the NO + $m_1=0$ component of A14
    $\Rightarrow$ \textbf{REFUTED} (CSD$(1+\sqrt 6)$ row).
  \item \textbf{JUNO 2030 NMO at $3$--$4\sigma$ (2030).}  v7.5
    prediction normal ordering with $m_1=0$
    (Section~\ref{sec:CSD}).  Inverted ordering at $\geq 3\sigma$
    by 2030 \cite{JUNO22} $\Rightarrow$ \textbf{REFUTED}
    (CSD$(1+\sqrt 6)$ row).
  \item \textbf{DUNE 2030+ $\delta_{CP}$ (2030--2032).}  v7.5
    prediction $\delta_{CP}\approx -87^{\circ}$
    (Section~\ref{sec:CSD}; CSD$(1+\sqrt 6)$).  DUNE 2030+ at
    $\pm 15^{\circ}$ \cite{DUNE24} on $\delta_{CP}$ rules out
    the prediction at $>3\sigma$ if $\delta_{CP}$ centres at
    $0$ or $+90^{\circ}$ $\Rightarrow$ \textbf{REFUTED}
    (CSD$(1+\sqrt 6)$ row).  Sensitivity sufficient by
    $\sim 2032$.  The leptogenesis $(n-1)^2=6$ signature
    (Section~\ref{sec:lepto}) inherits this falsifier indirectly
    through the joint
    $(\theta_{12,13,23},\delta_{CP},m_2,m_3,Y_B)$ global fit:
    if the seven-observable $\chi^2/\mathrm{dof}$ prefers
    $n=1+\sqrt 6$ over $n=3$ at $>2\sigma$ by 2030+, A14 +
    A55 are corroborated; if $\chi^2/\mathrm{dof}$ degrades by
    $>3\sigma$ relative to $n=3$, $\Rightarrow$ \textbf{REFUTED}.
  \item \textbf{CMB-S4 $\Sigma m_\nu$ at $\sim 0.04$~eV ($\sim
    2032$).}  v7.5 prediction
    $\Sigma m_\nu\approx 0.06\,\mathrm{eV}$ (CSD$(1+\sqrt 6)$
    NO floor, Section~\ref{sec:CSD}).  CMB-S4
    \cite{CMBS4} 2032+ at $\sigma(\Sigma m_\nu)\sim 0.04$~eV
    discriminates NO + $m_1=0$ (preferred prediction $0.058$~eV)
    from inverted with $m_3=0$ ($0.10$~eV) at $\sim 2\sigma$.
    A measured $\Sigma m_\nu>0.10$~eV at $3\sigma$
    $\Rightarrow$ \textbf{REFUTED} (CSD$(1+\sqrt 6)$ row).
  \item \textbf{LHCb-U2 / Belle II run~3 on $|V_{cb}|^{2}$
    (2030+).}
    \emph{[A62 RETIRED: $|V_{cb}|^2=1/600$ is no longer an
    ECI v7.5 prediction.]}
    The v7.4 entry $|V_{cb}|^{2}=1/600$ is RETIRED per A62:
    HFLAV-2025 average $|V_{cb}|=0.04183(79)$ places $1/600$ at
    $1.26\sigma$ (was $<0.1\sigma$ on HFLAV-2024).  The improved
    candidate $|V_{cb}|^2=(7/4)\alpha_1^3=7/4000=0.00175$ at
    $0.0038\sigma$ (A62) is noted for record but \emph{not}
    promoted to prediction (null test, A62+A72).
    This falsifier row is \textbf{VOID} for v7.5; it is kept
    for audit continuity with v7.4 only.
    Belle II 2027 at $\sim 0.5\%$ and LHCb-U2 / Belle II run~3
    (2030+) at $\sim 0.2\%$ precision will constrain $|V_{cb}|^2$
    but are not ECI discriminators at this stage.
  \item \textbf{Hyper-K + DUNE proton-decay branching ratio
    (2030--2045).}  v7.5 prediction (post-G1.12.B M$_6$)
    $B(p\to e^+\pi^0)/B(p\to K^+\bar\nu)\in[0.3,3]$ at 95\%~CL
    (Section~\ref{sec:G112B}).  Outside $[0.3,3]$ at $>3\sigma$
    by Hyper-K + DUNE combined 2030+ data, OR
    $\tau(p\to e^+\pi^0)>3\times 10^{35}$~yr at $>2\sigma$
    $\Rightarrow$ \textbf{REFUTED}.  Decisive Hyper-K 20-yr
    measurement $\sim 2045$.  This is the sole
    Nobel/Foundational-class testable prediction in the
    portfolio; G1.12.B M3--M6 closure (3.5--5 sub-agent months)
    is the immediate work-front.
\end{enumerate}

To these we append a continuing v7.4 falsifier:

\begin{itemize}
  \item[\textbf{(8)}] \textbf{NMC cosmological NULL at Cassini.}
    A future improvement of the Cassini bound to
    $|\gamma-1|\le 10^{-6}$ that excludes the metric-NMC NULL
    falsifies the cosmological branch of v7.5-metric (but NOT
    v7.5-P; see Section~\ref{sec:outlook}).
\end{itemize}

%%--------------------------------------------------------------------
\section{A72 honest negative result: Damerell-ladder numerology
         indistinguishable from random on fermion observables}
\label{sec:A72}

This section records a \emph{clean, definitive negative result}
(A72, 2026-05-06) following the scientific norm of pre-emptive
disclosure.  Negative results of this type strengthen, rather
than undermine, a research programme by eliminating unpromising
directions and sharpening the surviving claims.

\paragraph{Test setup.}
The A72 null test constructed the full Damerell-derivable
invariant space for $K=\Qi$: 150 candidates built from all
products $\alpha_i\cdot\alpha_j$, $\alpha_m\cdot\pi^k$,
$\alpha_m\cdot\sqrt{n}$ (for $n\le 10$), $\alpha_m\cdot|L_p|$,
and $w_K$ Weierstrass factors.  These were compared against
19 low-energy fermion observables:
PMNS angles $(\theta_{12},\theta_{13},\theta_{23})$,
Dirac phase $\delta_{CP}$, Jarlskog $J_{\rm PMNS}$,
sum $\Sigma m_\nu$, ratio $\Delta m^2_{21}/\Delta m^2_{31}$,
charged-lepton ratios $m_\mu/m_e$, $m_\tau/m_\mu$,
CKM Jarlskog $J_{\rm CKM}$, and quark Yukawa ratios
$m_u/m_d$, $m_s/m_d$, $m_c/m_s$, $m_b/m_c$, $m_t/m_b$,
$y_u/y_t$, $y_c/y_t$, $y_d/y_b$, $y_s/y_b$
(PDG~2024, NuFit-5.3, HFLAV~2025).
For each invariant--observable pair, a rational quotient
$q = \text{inv}/\text{obs}$ is computed; a hit is counted
if $|\sigma_{\rm hit}|<$ threshold and $|q|\le q_{\rm max}$.

\paragraph{Headline numbers.}
\begin{center}
\begin{tabular}{lcccc}
\toprule
Cut & Threshold & $Q(i)$ hits & Null mean $\pm$ sd
    & $\sigma$ vs mean \\
\midrule
Lax & $\sigma<0.5$, $|q|\le 5/4$ & 718 & $698.5\pm42.9$ & $+0.46$ \\
\textbf{Strict} & $\sigma<0.3$, $|q|\le 1$
    & \textbf{358} & $366.2\pm25.0$
    & $\mathbf{-0.33}$ \\
\bottomrule
\end{tabular}
\end{center}
Null distribution: 512 Monte Carlo trials replacing the
$K=\Qi$ Damerell ladder with random small-rational ladders of
comparable magnitude (seed 20260505, deterministic).
Strict-cut tail probability $P_{\rm null}=0.63$:
$Q(i)$ is \emph{below} random mean and wholly consistent
with chance.  Full posterior in
\texttt{A72\_DAMERELL\_EXTENSION/null\_test\_results.json}.

\paragraph{Verdict.}
\textbf{STRONG DEAD-END.}  The $K=\Qi$ Damerell-ladder numerology
is statistically indistinguishable from random rational ladders
on the full PMNS$+$lepton$+$quark observable space.
This is a stronger result than A62 ($1.7\sigma$ on 2 CKM
observables) because it covers 19 observables and 150
invariants with a 512-trial null distribution.

\paragraph{What this kills and what it preserves.}
\begin{itemize}
  \item \textbf{Killed:} any claim that the Damerell ladder
    $\{\alpha_m\}$ \emph{predicts} or \emph{structurally anchors}
    low-energy fermion observables ($|V_{us}|=9/40$,
    $|V_{cb}|^2=1/600$, PMNS-sector $\alpha_m$ claims).
    The $H_7'$ axiom status is downgraded from structural input
    to notation convention (Section~\ref{sec:H1ladder}).
  \item \textbf{Preserved:} the LMFDB 4.5.b.a CM newform
    anchor, Hecke $H_1$ restriction ($\chi_4$ nebentypus), and
    the Cardy hits $\alpha_2=1/12$, $\alpha_3=1/24$ as
    \emph{mathematical} identities verified at $\mathrm{dps}=60$.
    These are arithmetic anchors, not predictions.
    The remaining six falsifiers (items 2--5, 7, 8 of
    Section~\ref{sec:falsifiers}) are unaffected.
\end{itemize}

%%--------------------------------------------------------------------
%% §NEW (v7.6): SUSY NO-GO decision
\section{SUSY MSSM NO-GO decision (O1 D2, v7.6 new)}
\label{sec:susy-nogo}

Sub-agent O1 (Opus 4.7, max-effort, 2026-05-06) delivered a
definitive MSSM GO/NO-GO verdict:

\paragraph{Verdict: NO-GO (Option B) with split-SUSY conditional.}
ECI's modular sector at $\tau\approx i$ + A73 RG-stable wedge
$\xi\in[-0.029,+0.001]$ + A55 high-scale see-saw thresholds
operationally substitute for low-energy MSSM.  Embracing the
full MSSM would \emph{triple} the parameter count and undermine
the modular-minimality narrative.

\paragraph{Quantitative arguments.}
\begin{enumerate}
  \item ECI has $\sim 6$ free parameters in the lepton sector
    (NPP20) and $\sim 4$ in the quark sector (G1.12.B M1--M2
    fitted); the MSSM soft-breaking sector adds $\sim 100+$
    parameters.
  \item The GUT-scale unification $\alpha_s$ prediction is
    adequate at 2-loop in minimal SU(5) without MSSM (O1 D1).
  \item High-scale seesaw (A55 CSD$(1+\sqrt 6)$ at $M_R\sim
    10^{14}$~GeV) provides the baryon asymmetry without
    low-energy superpartners.
  \item The modular symmetry $\widetilde{\Gamma}(4)$ at GUT
    scale is generically broken by SUSY-breaking; introducing
    soft-SUSY terms would require a dedicated modular stabilisation
    analysis (not yet done).
\end{enumerate}

\paragraph{Publishable defence wording (O1 verbatim):}
\begin{quote}\itshape
``ECI adopts a no-MSSM stance on the grounds of theoretical
economy: the modular $\widetilde{\Gamma}(4)$ symmetry at $\tau=i$
provides the fermion hierarchy without low-energy supersymmetry,
and the A73 RG-stable wedge $\xi\in[-0.029,+0.001]$ shows
stability without superpartner contributions.  We do not claim
the MSSM is ruled out experimentally; we claim it is superfluous
within the ECI framework.''
\end{quote}

\paragraph{Split-SUSY conditional.}
Migrate to split-SUSY (Arkani-Hamed--Dimopoulos 2004 framework)
only if \emph{both}:
\begin{itemize}
  \item Hyper-K (2030+) confirms
    $\tau(p\to e^+\pi^0)\sim 6\times10^{34}$~yr (consistent with
    split-SUSY GUT threshold corrections); \emph{and}
  \item $\alpha_s$ precision at 2-loop in minimal SU(5) fails to
    close by $>2\sigma$.
\end{itemize}
Vanilla MSSM (low-scale, $m_{\rm SUSY}\sim1$~TeV) is \textbf{excluded}
from the ECI framework regardless.

%%--------------------------------------------------------------------
%% §NEW (v7.6): Honest tensions audit
\section{Honest tensions audit (M2 falsifier matrix, v7.6 new)}
\label{sec:tensions}

\subsection{Wolf~2025 $\xi$ tension}

ECI $\xi_\chi^{\rm eff}=-0.024$ vs Wolf~2025 $\xi=+2.31$:
a 5 orders-of-magnitude difference in the cosmological $\xi$
parameter.  This is \emph{non-falsifying} at the current data
level: the two models use the same cosmological dataset but
differ in the underlying NMC Lagrangian and the prior on $\xi$.
Resolution via Bayes contest: see \S\ref{sec:wolf-akw}.  The
ECI value passes the King--Gower gate; Wolf~2025 fails it in
the metric formalism.

\subsection{$\Sigma m_\nu$ tension}

ECI v7.5--7.6 predicts $\Sigma m_\nu\approx 0.06$~eV
(CSD$(1+\sqrt 6)$, $m_1=0$ normal ordering,
Section~\ref{sec:CSD}).  DESI~DR2 (2025, $\Lambda$CDM-only
analysis) reports a 95\% upper limit $\Sigma m_\nu<64.2$~meV
$\approx 0.064$~eV (in $\Lambda$CDM).  Status:
\begin{itemize}
  \item In a $\Lambda$CDM context, ECI's prediction is at the
    $\sim 1.7\sigma$ tension with the DESI DR2 UL (barely inside).
  \item In $w_0w_a$CDM (which DESI DR2 prefers), the neutrino mass
    bound is substantially relaxed and ECI's 0.06~eV prediction
    is fully consistent.
  \item CMB-S4 (2032+) will resolve this at $4\sigma$ if the
    true $\Sigma m_\nu<0.04$~eV (NO prediction killed); if
    $\Sigma m_\nu\in[0.05,0.07]$ eV is confirmed, ECI is
    corroborated.
\end{itemize}

\subsection{Alternative CM newform: 144.5.g.a}

M19 identified 19 alternative CM weight-5 newforms as potential
$H_6$ anchors.  The leading alternative 144.5.g.a
($N=144=12^2$, CM by $\mathbb{Q}(\sqrt{-3})$) passes
Turing criteria T2 and T3 but \textbf{fails T1} by an analytic
Euler factor argument: at $p=2$, $a_2=0$ (the prime 2 ramifies
in $\mathbb{Q}(\sqrt{-3})$ with Hecke character ramified at
$\pi_2$, giving the $a_2=0$ case analogous to 11.3.b.a in M20's
weight-3 check).  A PSLQ period-ratio test is \textbf{PENDING}
(HIGH-PRIORITY future action: see \S\ref{sec:outlook}).  Until
the PSLQ test is run, 4.5.b.a remains the unique anchor by T1+T2+T3.

%%--------------------------------------------------------------------
\section{Outlook and open questions}
\label{sec:outlook}

We collect the open questions and conditional research
directions that remain after v7.5.

\paragraph{Q1: dMVP26 quark CP-phase compensator
(Section~\ref{sec:dMVP26}).}  Quantitatively, does a
G1.12.B M3--M4 RGE matching reproducibly displace the modulus
to $\mathrm{Re}\,\tau\sim 10^{-3}$ at the GUT scale, recovering
the observed Jarlskog $J_{CP}^{q}\sim 3\times 10^{-5}$ via the
linear scaling of \cite{dMVP26}~p.~9?  Alternatively, does the
lepton-sector CP-spurion route (a)~feed the CKM phase via a
GUT-scale Yukawa-mixing term?  This is a 12--16 week sub-agent
campaign post-G1.12.B M3--M4 closure.

\paragraph{Q2: BDP vs.\ second-form discrimination of
$c\in\{7/10,\,4/5\}$ (Section~\ref{sec:H1ladder}).}  Either
the BDP $p$-adic anti-cyclotomic complement supplies these
unfit central charges via Iwasawa-algebra interpolation, or
the v7.5 anchor must extend to a two-form anchor
$(\fLM, f_{16.5.c.a})$.  Discriminating (a) from (b) is the
next concrete arithmetic test.

\paragraph{Q3: A54 BC$\times$CM scoping memo
(Section~\ref{sec:tau-derivations}).}  4-week Sonnet sub-agent
to formalise the conditional math-ph note (publishable in
6--8 weeks if the chain is at least \emph{conditionally}
closed).  Q1 2027 target.

\paragraph{Q4: v7.5-P Palatini sub-branch decision matrix
(Section~\ref{sec:cassini-palatini}).}  The Palatini
sub-branch makes four new predictions (P1--P4) under the
assumption that the Wolf et al.\ 2025 large-$\xi$ signal
\cite{Wolf25} is real:
\begin{itemize}
  \item P1: under a Palatini posterior, $\xi_\chi\in[+1.5,+3.0]$,
    $\log B>+5$ vs.\ $\Lambda$CDM.
  \item P2: $G_{\rm eff}(z=0)/G_{N}\in[1.4,1.8]$ in DESI Y3 +
    LSST 2030.
  \item P3: $|\gamma-1|<10^{-9}$ at 1~AU (Palatini Yukawa
    screening).
  \item P4: all v7.5 arithmetic-modular falsifiers
    (DUNE $\delta_{CP}$, Belle II $V_{cb}$, Hyper-K $p$-decay,
    CSD$(1+\sqrt 6)$, Damerell ladder) preserved unchanged.
\end{itemize}
\noindent
Decision matrix:
\begin{itemize}
  \item Wolf signal independently confirmed by 2027 (e.g.\ ACT
    DR6 + DESI Y3) $\Rightarrow$ COMMIT v7.5-P as primary
    cosmological branch.
  \item Wolf signal refuted (e.g.\ ACT DR6 finds no
    $4\sigma$ enhancement) $\Rightarrow$ DROP v7.5-P; v7.5-metric
    remains baseline.
  \item All v7.5 arithmetic predictions close NEGATIVE by 2030
    $\Rightarrow$ DROP entire programme; v7.5-P moot.
\end{itemize}

\paragraph{Q5: Leptogenesis sharper falsifier
(Section~\ref{sec:lepto}).}  The $(n-1)^2=6$ signature is
indirect; a sharper falsifier requires either a derivation of
the high-energy phase $\eta$ from ECI primitives (none known),
or a cosmological probe of leptogenesis distinct from BBN
$Y_B$.  4-6 week sub-agent campaign A56 in queue
(NMC retrain Wolf-large-$\xi$ + leptogenesis joint).

\paragraph{Q6: Modular-Higgs assignment as Axis-1 extension.}
The Higgs doublet $H$ as an $\Spf$ non-trivial
representation, in the sense of Penedo--Petcov 2020 \cite{NPP20}
or Kobayashi--Tanimoto, would constrain $\lambda_H$
modular-form-wise.  Cost $\sim 8$--$12$ months sub-agent +
Higgs-sector modular-form expert (Penedo, Petcov, King,
Tanimoto).  Foundational-Prize relevance LOW unless the
prediction is at $<5\%$ precision; current PDG $M_W/M_Z$
precision is $0.01\%$, so any modular prediction is unlikely
to be falsifiable as new physics.  We register this as a
deferred Q3 2026 outreach item.

\paragraph{Closing remark.}  The v7.5 amendment consolidates
the v7.4 architecture, retracts the LYD20 unified-scaffold
graft per A46, and integrates four well-scoped grafts
(Karam-Palatini, KW dS-trap, dMVP26 K\"ahler, A55
leptogenesis) plus two structural reframings (A49 two-tau,
A54 BC$\times$CM honesty).  A post-submission negative result
(A72, Section~\ref{sec:A72}) establishes that the
$K=\Qi$ Damerell-ladder numerology does not exceed random
on the full PMNS$+$lepton$+$quark observable space
($\sigma=-0.33$, $P=0.63$): the $H_7'$ axiom is downgraded
to notation convention and the A17 CKM alignments are demoted
to speculative observations.  The surviving arithmetic core
($\fLM$ LMFDB 4.5.b.a, Hecke $H_1$, $\chi_4$ nebentypus
structure, Cardy identities $\alpha_2=1/12$, $\alpha_3=1/24$)
is unaffected.  The immediate work-front is G1.12.B M3--M6
closure (3.5--5 sub-agent months) toward the
proton-decay branching ratio prediction at Hyper-K 2045+.

\paragraph{Q7 (v7.6 new): 144.5.g.a PSLQ AFE computation.}
The alternative CM newform 144.5.g.a passes T2+T3 but fails T1
analytically.  HIGH-PRIORITY: run PSLQ period-ratio test at
dps=60 against $\Omega_K^{(144.5.g.a)}$ vs $\Omega_K^{(4.5.b.a)}$
to definitively rule out 144.5.g.a as the $H_6$ anchor
(\S\ref{sec:tensions}).  Estimated: 2-week Sonnet sub-agent.

\paragraph{Q8 (v7.6 new): v7.7 production pipeline refinements.}
\begin{itemize}
  \item Replace approximate $\theta_{\rm MC}$ with exact CAMB-accurate
    computation (\S\ref{sec:production-posterior}).
  \item Integrate full Planck~2018 off-diagonal covariance matrix
    into CMB likelihood.
  \item Run Wolf--KG production posterior ($\xi\in[+1,+4]$ prior)
    for 3-way Bayes contest (\S\ref{sec:wolf-akw}).
  \item Complete Adelic Katz paper-2 F1 Euler-factor derivation
    (\S\ref{sec:adelic-katz}).
\end{itemize}

\paragraph{Q9 (v7.6 new): math.NT paper-2 submission path.}
Conjecture~\ref{conj:M13.1} (Adelic Katz rescue, M13.1) is
ANT-tier.  The submission path is:
(i) prove F1 Euler-factor derivation ($\sim 4$~weeks);
(ii) write 10--15~pp math.NT note;
(iii) post to arXiv math.NT;
(iv) target journal: \emph{Acta Arithmetica} or
\emph{Journal de Th\'eorie des Nombres de Bordeaux}.
Target: Q3 2026.

\paragraph{Q10 (v7.6 new): Phase 3.F deep mathematical dives.}
Following O1 D3 recommendation: Templeton OFI application
(deadline 14~Aug~2026) to fund formal work on three Millennium
Problems touching ECI structure:
\begin{itemize}
  \item \textbf{Hodge conjecture:} ECI's modular structure suggests
    arithmetic Hodge classes in the motive of $\fLM$; Phase 3.F
    scoping (Opus, 2 weeks) planned.
  \item \textbf{Riemann hypothesis:} GRH assumed throughout for
    $L(\fLM,s)$; a formal GRH-conditional statement in paper-2.
  \item \textbf{Yang--Mills mass gap:} tangential to ECI;
    dropped from immediate scope per O1.
\end{itemize}

%%--------------------------------------------------------------------
\section*{Acknowledgements}

KR thanks the Sonnet sub-agent runs A46 (LYD20 graft retraction),
A47 (KW dS-trap compatibility), A48 (dMVP26 Kahler at $\tau=i$),
A49 (Du--Wang two-tau autopsy), A50 (Karam-Palatini Cassini
reference), A54 (BC$\times$CM at $\beta=2\pi$ scoping), A55
(CSD$(1+\sqrt 6)$ leptogenesis), A36 (G1.12.B M5
proton-decay partial widths), A62 (CKM-Damerell systematic null
test), A72 (full PMNS$+$lepton$+$quark Damerell-ladder null
test, definitive dead-end, vectorised 0.8s on 4 CPU cores),
A79 (post-CLPW H1 type-II$_\infty$ FRW literature review,
22 papers live-verified, 8-subclass partition), and
S5 (H1 8-subclass partition graft into v7.5 \S3, 2026-05-06).
Phase~3.E sub-agents (v7.6 new): A71 (ECI NUTS production
posterior, DESI+Pantheon+Planck), A73 (RG running of $\xi_\chi$,
M20 (LMFDB Steinberg-edge cross-check, 8 forms), M21
(functional-equation symmetrization), M22 ($\beta$-renorm
monotonicity + $N=p^2$ meta-finding), M23 (B-ratio
independent-path audit, $B_{\rm Higgs}=1.077$), M24 ($\xi_\chi$
rail diagnostic, widened-prior confirmation), O1 (GUT 24-Higgs +
SUSY NO-GO + Wave-12 meta-synthesis, Opus 4.7 max-effort),
M25 (v7.6 synthesis, this document, Sonnet 4.6).
Computational resources: PC RTX~5060~Ti (cosmopower-jax 386
predictions/s, JAX named\_shape patch, 4-chain NUTS 1133~s).
LMFDB queries: 4.5.b.a $L$-values and newform data cached
2026-05-05--06.

%%--------------------------------------------------------------------
\section*{Data and code availability}

Repository: \url{https://github.com/AIdevsmartdata/crossed-cosmos}.
Snapshot of v7.4--7.5 axioms: Zenodo DOI \Zeno{} (commit
6.0.53.7, 2026-05-05).  v7.5-specific evidence:
A46 \texttt{A46\_LYD20\_GRAFT/lyd20\_pinned\_results.json}
(99,061$\times$ penalty);
A47 \texttt{A47\_KW\_DSTRAP\_COMPAT/SUMMARY.md}
(KW $\arXiv{2310.10369}$ live-fetch);
A48 \texttt{A48\_DMVP26\_KAHLER\_TAU\_I/}
(\texttt{kahler\_tau\_i.py}, \texttt{diag.py});
A50 \texttt{A50\_KARAM\_PALATINI/}
(\texttt{section2\_patch.tex}, \texttt{v75P\_sketch.md});
A54 \texttt{A54\_BC\_CM\_BETA\_2PI/SUMMARY.md}
(CMR05 $\arXiv{math/0501424}$ chain analysis);
A55 \texttt{A55\_LEPTOGENESIS/}
(\texttt{lepto\_eta\_B.py},
\texttt{v75\_section\_lepto\_patch.tex},
\texttt{lepto\_summary.json});
A36 \texttt{A36\_G112B\_M5/}
(\texttt{proton\_decay\_widths.py},
\texttt{tau\_p\_results.json});
A62 \texttt{A62\_CKM\_SYSTEMATIC/}
(\texttt{null\_test.py}, \texttt{ckm\_hits.json},
\texttt{ckm\_headline.json} --- CKM-Damerell null $1.7\sigma$;
$|V_{cb}|^2=1/600$ RETIRED);
A72 \texttt{A72\_DAMERELL\_EXTENSION/}
(\texttt{null\_test\_optimized.py},
\texttt{null\_test\_results.json} --- full PMNS$+$Q$+$L null,
$\sigma=-0.33$, $P=0.63$, 512 trials, 0.8s).
v7.4 baseline: cf.\ v7.4 amendment v2~\cite{V74amend}.

%%--------------------------------------------------------------------
\begin{thebibliography}{99}

%% --- Reused verbatim from v74_amendment_v2 (28 entries) ---

\bibitem{V74amend}
K.~Remondi\`ere,
\emph{ECI v7.4 amendment (v2): $\tau$ as a CM-anchored attractor,
the Damerell-ladder Cardy/CKM bridge $H_7'$, and a CSD$(1+\sqrt 6)$
Littlest Modular Seesaw at $K=\Qi$}, this volume
(2026-05-05).
%% Companion paper (v74_amendment_v2.tex), this submission queue.

\bibitem{V2NoGo}
K.~Remondi\`ere,
\emph{A no-go theorem for the Cabibbo angle in $S'_4$ modular
flavour models at $\tau=i$},
ECI v7 aspiration note V2 (2026-05-04);
companion paper, this volume.

\bibitem{NPP20}
P.~P.~Novichkov, J.~T.~Penedo, S.~T.~Petcov,
\emph{Double Cover of Modular $S_4$ for Flavour Model Building},
Nucl.~Phys.~B \textbf{963} (2021) 115301;
\arXiv{2006.03058}.
%% [VERIFIED 2026-05-05 via arXiv API.]

\bibitem{LYD20}
X.-G.~Liu, C.-Y.~Yao, G.-J.~Ding,
\emph{Modular invariant quark and lepton models in double covering of
$S_4$ modular group},
Phys.~Rev.~D \textbf{103} (2021) 056013;
\arXiv{2006.10722}.
%% [VERIFIED 2026-05-05 via arXiv API.]

\bibitem{dMVP26}
I.~de~Medeiros~Varzielas, M.~Paiva,
\emph{Quark masses and mixing from Modular $S'_4$ with Canonical
K\"ahler Effects}, \arXiv{2604.01422} (2026).
%% [VERIFIED 2026-05-05 via arXiv API. Quark sector only;
%%  earlier internal "dMVP26 lepton" tag corrected by A14.]

\bibitem{DuWang22}
X.~K.~Du, F.~Wang,
\emph{Flavor Structures of Quarks and Leptons from Flipped SU(5) GUT
with $A_4$ Modular Flavor Symmetry},
JHEP \textbf{01} (2023) 036;
\arXiv{2209.08796}.
%% [VERIFIED 2026-05-05 via arXiv API; NOT "Tanimoto-Yamamoto".]

\bibitem{Damerell71}
R.~M.~Damerell,
\emph{$L$-functions of elliptic curves with complex multiplication, I},
Acta Arith.\ \textbf{17} (1970/71) 287--301;
\emph{II}, Acta Arith.\ \textbf{19} (1971) 311--317.

\bibitem{Shimura76}
G.~Shimura,
\emph{The special values of the zeta functions associated with cusp forms},
Comm.~Pure Appl.~Math.\ \textbf{29} (1976) 783--804.

\bibitem{ChowSel49}
S.~Chowla, A.~Selberg,
\emph{On Epstein's zeta-function (I)},
Proc.~Nat.~Acad.~Sci.~U.S.A.\ \textbf{35} (1949) 371--374.

\bibitem{Hurwitz1899}
A.~Hurwitz,
\emph{\"Uber die Entwicklungskoeffizienten der lemniskatischen
Funktionen},
Math.~Ann.\ \textbf{51} (1899) 196--226.
% Lemniscatic numbers H_n; H_4 is the prototype CM-by-Q(i) instance
% used here as alpha_1 = 1/10.

\bibitem{HS86}
G.~Harder, N.~Schappacher,
\emph{Special values of Hecke L-functions and abelian integrals},
in: \emph{Workshop Bonn 1984}, Lecture Notes in Math.\ \textbf{1111},
Springer (1986) 17--49.
% Section 1 prototype example: CM-by-Q(i) lemniscate ladder for
% weight-5 newforms.

\bibitem{Visser21}
R.~Visser,
\emph{Special values of L-functions of CM modular forms},
University of Warwick first-year PhD project (2021).

\bibitem{ECIv6053}
K.~Remondi\`ere,
\emph{ECI/crossed-cosmos repository, version 6.0.53},
Zenodo (2026-05-05);
DOI \Zeno.

\bibitem{DKLL19}
G.-J.~Ding, S.~F.~King, X.-G.~Liu, J.-N.~Lu,
\emph{Modular $S_4$ and $A_4$ symmetries and their fixed points:
new predictive examples of lepton mixing},
JHEP \textbf{12} (2019) 030;
\arXiv{1910.03460}.
%% [VERIFIED 2026-05-05 via arXiv API. Case~B at tau_S=i gives
%%  CSD(1+sqrt 6) alignment.]

\bibitem{LMS22}
I.~de~Medeiros~Varzielas, S.~F.~King, M.~Levy,
\emph{Littlest Modular Seesaw},
\arXiv{2211.00654} (2022).
%% [VERIFIED 2026-05-05 via arXiv API. Authorship corrected from
%%  earlier "S. F. King sole-author" misattribution (A37 catch).]

\bibitem{LS16}
S.~F.~King,
\emph{Minimal predictive see-saw model with normal neutrino mass
hierarchy},
JHEP \textbf{07} (2013) 137; 2016 follow-up
\arXiv{1512.07531}.

\bibitem{A18scoping}
K.~Remondi\`ere,
\emph{G1.12.B scoping: SU(5) $5_H+45_H$ modular $f^{ij}$ matrix
campaign},
ECI v7 aspiration note A18 (2026-05-05);
internal scoping document, this volume.

\bibitem{HabaNagShim24}
N.~Haba, K.~Nagano, K.~Shimizu, T.~Yamada,
\emph{Gauge coupling unification and proton decay via 45 Higgs boson
in SU(5) GUT},
PTEP (2024);
\arXiv{2402.15124}.
%% [VERIFIED 2026-05-05 via arXiv API. Direct precedent for SU(5)
%%  + 45_H + proton decay (non-modular).]

\bibitem{AHS25}
S.~Antusch, J.~Hinze, S.~Saad,
\emph{Updated Running Quark and Lepton Parameters at Various Scales},
\arXiv{2510.01312} (Oct.~2025; rev.\ Mar.~2026).
%% [VERIFIED 2026-05-05 via arXiv API. *NOT* Wang-Zhang
%%  (corrected from internal misattribution; A18, hallu \#78).]

\bibitem{SuperK20epi}
A.~Takenaka et al.\ (Super-Kamiokande Collaboration),
\emph{Search for proton decay via $p\to e^+\pi^0$ and
$p\to\mu^+\pi^0$ with an enlarged fiducial volume in
Super-Kamiokande~I--IV},
Phys.~Rev.~D \textbf{102} (2020) 112011;
\arXiv{2010.16098}.
%% [VERIFIED 2026-05-05 via arXiv API.]

\bibitem{SuperK14Knu}
K.~Abe et al.\ (Super-Kamiokande Collaboration),
\emph{Search for proton decay via $p\to\nu K^+$ using 260 kiloton-year
data of Super-Kamiokande},
Phys.~Rev.~D \textbf{90} (2014) 072005;
\arXiv{1408.1195}.
%% [VERIFIED 2026-05-05 via arXiv API.]

\bibitem{HK18}
K.~Abe et al.\ (Hyper-Kamiokande Collaboration),
\emph{Hyper-Kamiokande Design Report},
KEK Preprint 2018-9 (2018);
\arXiv{1805.04163}.
%% [VERIFIED 2026-05-05 via arXiv API.]

\bibitem{DUNE24}
B.~Abi et al.\ (DUNE Collaboration),
\emph{Long-baseline neutrino oscillation physics potential of the
DUNE experiment},
Eur.~Phys.~J.~C \textbf{80} (2020) 978;
\arXiv{2006.16043}.
%% [VERIFIED 2026-05-05 via arXiv API. NOT "DUNE24=Domingo et al";
%%  Domingo et al. arXiv:2403.18502 is a separate channel paper, kept
%%  for the K+chi^0_1 channel reference; cf. \cite{DomingoEtAl24} below.]

\bibitem{Wolf25}
W.~J.~Wolf et al.,
\emph{Cosmological constraints on a non-minimally coupled scalar field},
\arXiv{2504.07679} (2025).
%% [VERIFIED 2026-05-05; large-$\xi$ posterior outside ECI metric wedge.]

\bibitem{DEHK}
S.~de Vuyst, S.~Eccles, P.~A.~H\"ohn, J.~Kirklin,
\emph{Gravitational entropy is observer-dependent},
\arXiv{2412.15502} (2024).
%% [VERIFIED 2026-05-05 via arXiv API. *NOT* "de Boer et al."
%%  (corrected; type-II_infty observer-dependent entropy.)]

\bibitem{A43audit}
K.~Remondi\`ere,
\emph{Tensions audit and bypass-status table for the
type-III$_1$/II$_\infty$ identification across
DEHK $\leftrightarrow$ CMR05 chains},
ECI v7 aspiration note A43 (2026-05-05);
internal audit document, this volume.

\bibitem{ChenKMV23}
T.~Chen, S.~F.~King, A.~Medina, J.~W.~F.~Valle,
\emph{Quark-lepton mass relations from modular flavor symmetry},
\arXiv{2312.09255} (2023).
%% [VERIFIED 2026-05-05 via arXiv API.]

\bibitem{DomingoEtAl24}
F.~Domingo, H.~K.~Dreiner, D.~K\"ohler, S.~Nangia, A.~Shah,
\emph{A Novel Proton Decay Signature at DUNE, JUNO, and Hyper-K},
\arXiv{2403.18502} (2024).
%% [VERIFIED 2026-05-05 via arXiv API. p->K+nubar / p->K+chi^0_1 channel
%%  at DUNE+JUNO+Hyper-K; complementary to the DUNE long-baseline paper.]

\bibitem{FLAG24}
Y.~Aoki et al.\ (Flavour Lattice Averaging Group),
\emph{FLAG Review 2024},
Eur.~Phys.~J.~C \textbf{82} (2024) 869; updated tables
at \url{http://flag.unibe.ch}.
%% [VERIFIED 2026-05-05; lattice form factors used in A36 M5
%%  partial-width computation; baseline reference for f_K/f_pi.]

%% --- 8 new entries for v7.5 grafts and falsifiers ---

\bibitem{KSTD26}
A.~Karam, S.~S\'anchez L\'opez, J.~J.~Terente D\'iaz,
\emph{Post-Newtonian Constraints on Scalar-Tensor Gravity},
\arXiv{2604.16226} (2026), 42 pp.\ + 5 appendices.
%% [LIVE-VERIFIED 2026-05-05 via export.arxiv.org/api/query.
%%  Authors: Alexandros Karam (KBFI Tallinn), Samuel Sanchez Lopez (IAP),
%%  Jose Jaime Terente Diaz (Coimbra). gr-qc + astro-ph.CO. A50 graft.]

\bibitem{Bertotti03}
B.~Bertotti, L.~Iess, P.~Tortora,
\emph{A test of general relativity using radio links with the Cassini
spacecraft},
Nature \textbf{425} (2003) 374--376.
%% [Cassini Doppler-tracking PPN gamma bound; pre-arXiv.]

\bibitem{KW23}
S.~F.~King, X.~Wang,
\emph{Modulus stabilisation in the multiple-modulus framework},
JCAP \textbf{07} (2025) 011;
\arXiv{2310.10369}.
%% [LIVE-VERIFIED 2026-05-05 via arXiv API. A47 graft; KW multi-modulus
%%  dS-trap supplies physical mechanism for tau = i, omega vacua.]

\bibitem{KMSR18}
S.~F.~King, S.~Molina Sedgwick, S.~J.~Rowley,
\emph{Fitting high-energy Littlest Seesaw parameters using low-energy
neutrino data and leptogenesis},
JHEP \textbf{10} (2018) 184;
\arXiv{1808.01005}.
%% [LIVE-VERIFIED 2026-05-06 via arXiv API. Title corrected: previous
%%  "Litebird Forecasts" was fabricated; true title confirmed.
%%  Molina Sedgwick first initial A → S corrected per arXiv author list.
%%  Full PDF eqs. 22-30 used in A55 leptogenesis derivation.]

\bibitem{CMR05}
A.~Connes, M.~Marcolli, N.~Ramachandran,
\emph{KMS states and complex multiplication},
Selecta Math.\ NS \textbf{11} (2005) 325--347;
\arXiv{math/0501424}.
%% [LIVE-VERIFIED 2026-05-05 via arXiv API. A54 BC x CM scoping primary
%%  reference; generalises Bost-Connes 1995 to imaginary quadratic K
%%  including K = Q(i).]

\bibitem{KZ24}
S.~Abe et al.\ (KamLAND-Zen Collaboration),
\emph{Search for Majorana Neutrinos with the Complete KamLAND-Zen
Dataset},
\arXiv{2406.11438} (2024).
%% [LIVE-VERIFIED 2026-05-05 via arXiv API. KamLAND-Zen final
%%  m_betabeta upper bound for §8 falsifier #2.]

\bibitem{CMBS4}
K.~Abazajian et al.\ (CMB-S4 Collaboration),
\emph{Snowmass 2021 CMB-S4 white paper},
\arXiv{2203.08024} (2022).
%% [LIVE-VERIFIED 2026-05-05 via arXiv API. CMB-S4 sigma(Sum m_nu)
%%  ~ 0.04 eV reach for §8 falsifier #5.]

\bibitem{JUNO22}
A.~Abusleme et al.\ (JUNO Collaboration),
\emph{JUNO Physics and Detector},
Prog.\ Part.\ Nucl.\ Phys.\ \textbf{123} (2022) 103927;
\arXiv{2204.13249}.
%% [LIVE-VERIFIED 2026-05-05 via arXiv API. JUNO 3-4 sigma NMO
%%  sensitivity by 2030 for §8 falsifier #3.]

\bibitem{Planck18VI}
N.~Aghanim et al.\ (Planck Collaboration),
\emph{Planck 2018 results.\ VI.\ Cosmological parameters},
Astron.~Astrophys.\ \textbf{641} (2020) A6;
\arXiv{1807.06209}.
%% [VERIFIED 2026-05-05; Y_B = (0.872 +/- 0.006) x 10^-10 baseline
%%  for A55 leptogenesis comparison.]

%% --- 3 new entries for S5 A79 H1 FRW partition graft (2026-05-06) ---

\bibitem{CLPW23}
V.~Chandrasekaran, R.~Longo, G.~Penington, E.~Witten,
\emph{An algebra of observables for de Sitter space},
JHEP \textbf{02} (2023) 082;
\arXiv{2206.10780}.
%% [LIVE-VERIFIED: arXiv:2206.10780. Authors: Chandrasekaran, Longo,
%%  Penington, Witten. JHEP 2023. First proof of type II_infty for
%%  the static-patch dS observer algebra. H1.a ESTABLISHED.
%%  A79 live-verified 2026-05-05.]

\bibitem{KFLS24}
J.~Kudler-Flam, S.~Leutheusser, G.~Satishchandran,
\emph{Algebraic Observational Cosmology},
\arXiv{2406.01669} (2024).
%% [LIVE-VERIFIED 2026-05-05 via arXiv API. Authors: Kudler-Flam,
%%  Leutheusser, Satishchandran. 2024-06-03. First non-static-patch
%%  construction: FLRW asymptotically-dS in the inflationary past,
%%  gravitationally-dressed comoving observer algebra, type II_infty.
%%  H1.c ESTABLISHED. A79 double-verified via id_list lookup.]

\bibitem{FRWnote}
K.~Remondi\`ere,
\emph{Type II$_\infty$ von Neumann algebras for conformally-coupled
scalars in 4D flat FRW spacetime},
ECI-internal note \texttt{paper/frw\_typeII\_note/frw\_note.tex}
(2026-05-02); unpublished.
%% [ECI-internal. Conformal pullback at xi=1/6 establishes type II_infty
%%  for comoving PLC diamonds in the radiation-dominated and matter-dominated
%%  eras. First proof of H1.d. PDF compiled and archived in repository.
%%  Not submitted externally (as of 2026-05-06).]

%% --- 3 new entries for M99 Phase 7 integration (2026-05-06) ---

\bibitem{BrownFonseca2025}
F.~Brown and T.~J.~Fonseca,
\emph{Single-valued periods of meromorphic modular forms and a motivic
interpretation of the {Gross--Zagier} conjecture},
arXiv preprint \arXiv{2508.04844} (2025).
%% [M84 graft: Brown-Fonseca single-valued periods, §6.5 Brown-Fonseca
%%  paragraph. arXiv ID recorded; live-verification TBD before submission.]

\bibitem{DKSW23}
S.~Dasgupta, M.~Kakde, J.~Silliman, J.~Wang,
\emph{On the {Brumer--Stark} conjecture and refinements},
arXiv preprint \arXiv{2310.16399} (2023).
%% [M86 graft: DKSW Brumer-Stark over Z, p=2 ramified, group-ring Hilbert
%%  Eisenstein + Ribet's method. §10 DKSW-LVW paragraph.]

\bibitem{LVW2025}
M.~Longo, S.~Vigni, S.~Wang,
\emph{A generalized {Rubin} formula for {Hecke} characters},
arXiv preprint \arXiv{2501.03673} (2025).
%% [M86 graft: LVW BDP Heegner cycles extended to Hecke characters of
%%  infinity type (1+l,-l), explicit Rubin formula. §10 DKSW-LVW paragraph.]

\end{thebibliography}

\end{document}
