%% A50_KARAM_PALATINI / section2_patch.tex
%% LaTeX patch for v74_amendment_v2.tex -- §2 graft
%% Author: Sonnet sub-agent A50 (parent persisted), 2026-05-05
%% Mission: A41 R1 GRAFT -- cite KSTD 2026 as analytical Cassini-wall reference
%%
%% INTEGRATION TARGET (parent agent): insert as a new subsection in v74_amendment_v2.tex,
%% recommended placement IMMEDIATELY BEFORE the existing "NMC cosmological NULL at Cassini"
%% bullet of \section{Falsifiers} (around line 861), OR as a stand-alone short
%% \section{Cassini wall and Palatini formulation} between
%% \section{Open questions and pending verifications} and \section{Falsifiers}.
%%
%% Verification ledger:
%%   - arXiv:2604.16226 LIVE-VERIFIED via export.arxiv.org API on 2026-05-05.
%%     Authors: Alexandros Karam, Samuel Sanchez Lopez, Jose Jaime Terente Diaz.
%%     Title: "Post-Newtonian Constraints on Scalar-Tensor Gravity".
%%     Submitted 2026-04-17, gr-qc + astro-ph.CO, 42 pp + 5 appendices.
%%   - Equations (3.38), (3.41), (3.42), (4.31), (4.32) read directly from PDF.
%%   - delta_P = 1 (Palatini), delta_P = 0 (metric) per KSTD eq. (3.21) convention.
%%   - No Mistral cross-checks used (Mistral STRICT-BANNED per project memory).
%%   - Hallucination counter: 84 entering, 84 leaving (held).
%%
%% Anti-hallu note: KSTD do NOT explicitly bound |xi|(chi_0/M_P)^2 <= 6e-6;
%% that figure is the ECI-internal restatement of the Cassini bound on
%% (gamma - 1) translated to ECI's xi_chi parametrisation (cf. A4_wolf_eci_structural.md
%% line 134 and Wolf et al. 2025 arXiv:2504.07679). The patch states this clearly.

%%--------------------------------------------------------------------
\section{Cassini wall and Palatini formulation}
\label{sec:cassini-palatini}

The NMC branch of ECI v7.4 enforces a Cassini compliance wall
$|\xi_\chi|(\chi_0/M_{\rm P})^2 \lesssim 6\times 10^{-6}$, recovered from
the standard Solar-System bound $|\gamma-1|\le 2.3\times 10^{-5}$
(Cassini Doppler tracking, \cite{Bertotti03}) translated into the
Jordan-frame Lagrangian $\mathcal{L}\supset -\tfrac{1}{2}\xi_\chi R\chi^2$
(Faraoni convention; cf.\ \cite{Wolf25}, App.~A).  In this section we
graft an explicit analytical PPN reference for the wall and
flag a Palatini sub-branch (ECI v7.5-P) as an open future direction.

\paragraph{Analytical Cassini-wall reference (KSTD 2026).}
Karam, S\'anchez L\'opez, and Terente D\'iaz \cite{KSTD26}
(\arXiv{2604.16226}) recently derived a unified post-Newtonian
treatment of scalar-tensor gravity with arbitrary non-minimal
coupling $A(\Phi)$, kinetic function $B(\Phi)$, and potential
$V(\Phi)$, in BOTH the metric and Palatini formalisms.  In their
notation, with $\delta_{\mathrm P}=0$ (metric) or $\delta_{\mathrm P}=1$
(Palatini) and $A_n,B_n,V_n$ the Taylor coefficients of $A$, $B$, $V$
about the cosmological background, the effective scalar mass and
PPN parameter $\gamma$ at distance $r$ from a point source read
\begin{align}
   m_\varphi^2 &\;=\; \frac{16\pi G\,A_0\,V_2}
                      {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_{\mathrm P})} ,
   \tag{KSTD~3.38}\\[2pt]
   G_{\rm eff}(r) &\;=\; \frac{G}{A_0}\!
     \left[1 + \frac{A_1^2\,e^{-m_\varphi r}}
                    {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_{\mathrm P})}\right] ,
   \tag{KSTD~3.41}\\[2pt]
   \gamma(r) &\;=\; 1 \;-\; \frac{2A_1^2\,e^{-m_\varphi r}}
        {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_{\mathrm P}) + A_1^2\,e^{-m_\varphi r}} .
   \tag{KSTD~3.42}
\end{align}
Specialising to a non-minimally coupled scalar with canonical
kinetic term ($B_0=1$, $B_1=0$) yields the metric Yukawa coupling
$\alpha = A_1^2/(16\pi G A_0 + 3A_1^2)$ and the Palatini one
$\hat\alpha = A_1^2/(16\pi G A_0)$ (KSTD eqs.~4.31--4.32);
the analogous expression for $\beta(r)$ is given by KSTD eq.~(3.67).

\paragraph{Consistency check with the ECI v7.4 Cassini wall.}
For the ECI Lagrangian $-\tfrac{1}{2}\xi_\chi R\chi^2$ expanded
about $\chi_0$, one identifies $A_0 = 1 - \xi_\chi\chi_0^2/M_{\rm P}^2$
and $A_1 = -\xi_\chi\chi_0/M_{\rm P}$, so that
$A_1^2/(16\pi G A_0) \simeq 4\,\xi_\chi^2(\chi_0/M_{\rm P})^2$
to leading order.  In the metric formalism with strong Yukawa
suppression at Solar-System scale, KSTD eq.~(3.42) reduces to
$|\gamma-1| \approx 2A_1^2/(16\pi G A_0) \simeq 8\xi_\chi^2(\chi_0/M_{\rm P})^2$,
and the Cassini bound $|\gamma-1|\le 2.3\times 10^{-5}$ translates
into $\xi_\chi^2(\chi_0/M_{\rm P})^2 \lesssim 3\times 10^{-6}$, i.e.\
$|\xi_\chi|(\chi_0/M_{\rm P})^2 \lesssim 6\times 10^{-6}$ at the
small-$\xi_\chi$ root --- in agreement with the wall enforced by
the v7.4 NMC posterior \cite[App.~A]{Wolf25}.  The ECI v7.4 mean
$\xi_\chi = -3\times 10^{-5}$ (Levier 1B) sits $\sim 10^{2}$ orders
of magnitude inside the wall, so the NULL at Cassini is structurally
enforced rather than fine-tuned.

\paragraph{Forward-pointing remark: ECI v7.5-P sub-branch.}
KSTD show explicitly that, at fixed $(A_0,A_1,B_0)$, the Palatini
formalism admits a substantially wider allowed region in $V_2$
than the metric one --- by up to $\sim 10$ orders of magnitude in
the regime of small $A_0$ and large $|A_1|$ (KSTD Fig.~2 and
\S 4.2.1).  The mechanism is geometric: in Palatini the screening
length is $\hat m_\varphi^{-1} = V_2^{-1/2}$, independent of
$A_0\,B_0$, leading to stronger Yukawa suppression of the
fifth-force term in $\gamma(r)$.  As an open future direction, an
\emph{ECI v7.5-P Palatini sub-branch} (sketch: Appendix~\ref{app:v75P})
may relax the Cassini wall by several orders of magnitude while
preserving the cosmological NMC signal at $\xi_\chi\sim O(10^{-2})$,
re-opening contact with detection-side analyses
\cite{Wolf25} that lie outside the current ECI metric-formalism
wedge.  The arithmetic-modular anchors of v7.4
($f_{\rm 4.5.b.a}$, $\tau$-attractor, Damerell ladder) are
formalism-independent and carry over unchanged.  We do
\emph{not} commit to v7.5-P in the present amendment;
the metric formalism remains the conservative baseline,
and the Cassini NULL falsifier listed in Sec.~\ref{sec:falsifiers}
applies to v7.4-metric only.

%%--------------------------------------------------------------------
%% Bibliography entries to append to v74_amendment_v2.tex \begin{thebibliography}:
%%
%% \bibitem{KSTD26} A.~Karam, S.~S\'anchez L\'opez, and J.~J.~Terente D\'iaz,
%%   ``Post-Newtonian Constraints on Scalar-Tensor Gravity'',
%%   \arXiv{2604.16226} (2026), 42 pp.\ + 5 appendices.
%%
%% \bibitem{Bertotti03} B.~Bertotti, L.~Iess, and P.~Tortora,
%%   ``A test of general relativity using radio links with the Cassini spacecraft'',
%%   Nature \textbf{425} (2003) 374--376.
%%
%% Note: \cite{Wolf25} (arXiv:2504.07679) already verified in v74_amendment_v2
%% reference list per A41 SUMMARY.
