\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\begin{document}

\noindent Kevin Remondi\`ere \\
Independent Researcher \\
ORCID: \href{https://orcid.org/0009-0008-2443-7166}{0009-0008-2443-7166} \\
\href{mailto:kevin.remondiere@gmail.com}{kevin.remondiere@gmail.com} \\[1ex]
\today

\vspace{1em}
\noindent To the Editorial Board, \emph{European Physical Journal C} \\

\noindent \textbf{Re: Submission of ``ECI --- Entanglement, Complexity, Information: a framework paper threading six programs into one predictive canvas''}

\vspace{1em}

Dear Editors,

I am pleased to submit for your consideration the manuscript referenced above, authored solely by myself (Kevin Remondi\`ere, Independent Researcher, ORCID 0009-0008-2443-7166), for publication in EPJ C.

\textbf{Nature of the work.} ECI is a \emph{framework paper} --- it does not derive new results from first principles; rather, it synthesises six independently established research programs (2022--2026) into a single predictive architecture: (i) observer-dependent type-II von Neumann algebras via the QRF crossed product (CLPW 2023; De Vuyst--Eccles--H\"ohn--Kirklin 2025); (ii) the DESI DR2 dynamical-dark-energy preference addressed by non-minimally coupled thawing quintessence (Wolf 2025; Pan--Ye 2026); (iii) the $H_0$ tension reduced by Early Dark Energy (Poulin--Smith 2026); (iv) the Dark Dimension scenario (Bedroya 2025); (v) Cryptographic Censorship as a bulk-geometry selection rule; (vi) persistent-homology diagnostics of primordial non-Gaussianity (Yip 2024).

\textbf{Falsifiability.} Section~4 (``Falsifiable predictions'') lists concrete, near-term observational tests: a bound $|\xi_\chi| \lesssim 10^{-3}$ testable by DESI DR3 + Euclid, a dark-dimension signature $c' \simeq 0.05$ probed by fifth-force tests, and a persistent-homology signal in the primordial bispectrum within reach of LiteBIRD / CMB-S4.

\textbf{Reproducibility.} All derivations, numerical pipelines, and an MCMC scaffold are permanently archived at Zenodo DOI \href{https://doi.org/10.5281/zenodo.19686399}{10.5281/zenodo.19686399}, with the live repository at \url{https://github.com/AIdevsmartdata/crossed-cosmos}.

\textbf{AI-use disclosure.} Per the journal's transparency standards, AI-assisted tool usage (local LLM for LaTeX editing and code drafting) is disclosed in \texttt{AI\_USE.md} in the repository. All scientific content, derivations, and conclusions are the author's own.

\textbf{Honest caveat.} Axiom A3 (Cryptographic Censorship applied to cosmological horizons) is \emph{explicitly downgraded to a working conjecture} in v4: the original theorem is proven in AdS/CFT, and its transposition to de Sitter/FLRW is not established. All consequences drawing on A3 are flagged as conditional.

\textbf{No competing submission.} This manuscript has not been submitted elsewhere and is not under consideration by any other journal.

I believe EPJ C is the natural venue: the paper sits at the interface of theoretical cosmology, holography, and observational dark-energy physics, areas well represented in the journal, and the SCOAP3 open-access model aligns with my commitment to fully reproducible independent research.

Thank you for considering this submission.

\vspace{1em}
\noindent Sincerely, \\
Kevin Remondi\`ere

\end{document}
