% Team B draft — §1.5 "Unifying hypothesis (B): observer-dependent cosmology"
% Status: draft only. Not to be \input into eci.tex until owner selects A vs B.
% Target length: ~1 page two-column (≈ 650–750 words of physics prose).

\section{Unifying hypothesis (B): observer-dependent cosmology}
\label{sec:B-observer-dependent}

\subsection*{(i) The crossed-product framework and the generalised entropy functor}

The type-III$_1$ algebra of quantum fields restricted to a gravitational
subregion carries no trace and no density matrices; the
Chandrasekaran--Longo--Penington--Witten (CLPW) construction~\cite{CLPW2023}
and its quantum-reference-frame (QRF) completion by De~Vuyst--Eccles--H\"ohn--Kirklin
(DEHK)~\cite{DEHK2025a,DEHK2025b} repair this by crossing with the modular
automorphism of an observer clock, yielding a type-II$_1$ algebra on the static
de~Sitter patch (II$_\infty$ on Schwarzschild--de~Sitter Killing horizons) with a
well-defined trace and an entropy reproducing $S_{\rm gen} = A/(4G_N) + S_{\rm matter}$
up to a state-independent additive constant. The DEHK companion analysis makes
the observer-dependence structural: \emph{distinct QRFs give distinct crossed-product
algebras, distinct modular flows, and distinct numerical values of $S_{\rm gen}$} on the
same underlying bulk state~\cite{DEHK2025a}. The non-perturbative generalised
second law of Faulkner--Speranza~\cite{FaulknerSperanza2024} and
Kirklin~\cite{Kirklin2025} is formulated on exactly this type-II entropy.
Hypothesis~B reads this not as a technicality of the dS static patch but as the
\emph{organising principle} under which A4--A6 are to be interpreted: all
effective couplings that enter the late-time phenomenology of \S3 are labels on
the modular frame of the geodesic observer whose subregion is being probed,
not background quantities defined once and for all.

\subsection*{(ii) A4--A5--A6 as observer-frame quantities}

Under Hypothesis~B, the quantities that \S3 treats as a single numerical input
each split into a frame-labelled family.

\emph{A4 --- the NMC quintessence coupling $\xi_\chi$.} In the Jordan-frame action
of \S2, $\xi_\chi R\chi^2/2$ contributes to the graviton kinetic matrix and hence
to the type-II trace normalisation; the integration constant CLPW leave
undetermined in their trace is precisely what relates two observers' numerical
$\xi_\chi$. A geodesic observer on a different worldline, whose crossed-product
is built from a different clock Hamiltonian, measures $\xi_\chi^{(\mathcal R)}$;
the Cassini bound $|\xi_\chi|(\chi_0/M_P) \le 2.4\times 10^{-3}$ of \S3.5
constrains the \emph{solar-system-frame} value, which is the only value the
experiment is sensitive to.

\emph{A5 --- the Dark-Dimension species cutoff $c'$.} The species scale
$\Lambda_{\rm sp}(H) = M_P (H/M_P)^{c'}$ is a statement about the tower visible to
a given causal diamond. Different Hubble-patch subregions expose different
effective towers; the observer-frame reading of A5 is that $c'$ is the slope
seen in the algebra generated by a single diamond, with $c' = 1/6$ the
asymptotic quasi-de-Sitter value recovered when the diamond saturates a static
patch~\cite{Montero2022,AAL2023}.

\emph{A6 --- the persistent-homology spectrum $\mathrm{PH}_k$.} The Euler-
characteristic shift of~\eqref{eq:A6-euler} is computed on a density field
restricted to an observer-accessible region. Two observers with different
Hubble-patch subregions sample different $\mathrm{PH}_k$ distributions at the
same cosmological time; the Matsubara baseline is an ensemble average over
\emph{all} observer frames, and the diagnostic power of $\mathrm{PH}_k$ vs
the bispectrum is a frame-averaged statement.

\subsection*{(iii) A3 as the observer-frame horizon condition}

Cryptographic Censorship~\cite{CryptoCensorship} states, in its AdS/CFT proof,
that when the boundary-restricted time evolution on a code subspace is
$\varepsilon$-approximately a $k$-design, the bulk dual contains an event
horizon. Hypothesis~B re-reads this as the \emph{horizon condition bounding the
decoherence of observer-frame observables}: an observer whose modular algebra
sees an effectively pseudorandom state has lost predictive access to the
complement region, and an event horizon must cap the subregion on which
A4--A6 are defined. In this reading A3 is not an independent axiom but the
self-consistency condition for the other five: it delimits the subregion on
which the QRF crossed-product of A1 remains well-defined.

\subsection*{(iv) Caveats}

We do not claim that the QRF crossed-product is extended to generic matter-era
FLRW. CLPW is proven for the static dS patch; DEHK extend the framework to
multi-observer QRFs but still on the same Killing-horizon background. The
honest scope of Hypothesis~B is the \emph{late-time quasi-de-Sitter limit}
($\Omega_\Lambda \gtrsim 0.7$) where the comoving Hubble diamond
asymptotes to a static patch; this is exactly the regime DESI~DR2, Cassini,
and LSST~Y10 probe. Pre-recombination A4/A5 statements (the EDE axion $\phi$ at
$z_c\sim3500$, the BBN $\xi_\chi$ bound) are therefore \emph{not} covered by
the observer-dependent re-reading as of v4.6; their treatment under Hypothesis~B
requires a QRF extension to time-dependent FLRW backgrounds that is an open
problem in the algebraic-gravity literature. The paper's
\texttt{GROUND\_TRUTH.md} (Part~D.1) flags this limitation explicitly.
