% CCM_BIX_dictionary_max.tex
% Opus 4.7 (1M ctx), 2026-05-03
% Companion: /tmp/CCM_BIX_dictionary_max.{md,py}
\documentclass[11pt]{article}
\usepackage{amsmath, amssymb, amsthm, hyperref, geometry, booktabs}
\geometry{margin=1in}
\title{CCM 2511.22755 \(\leftrightarrow\) HY 2502.02661:\\
       Closure Assessment of Four Residual Gaps}
\author{Opus 4.7 (1M ctx, dictionary-draft)}
\date{2026-05-03}

\newtheorem{thm}{Theorem}
\newtheorem{prop}{Proposition}
\newtheorem{remark}{Remark}

\begin{document}
\maketitle

\section{Verbatim source quotations (PDF + HTML, fetched live)}
\paragraph{Hartnoll--Yang (HY) arXiv:2502.02661v2, JHEP 07(2025)281.}
Equation (37), \S 4.2:
\[
  D\,\psi \;=\; i\!\left(x\,\frac{d\psi}{dx} + \Delta\,\psi\right),
  \qquad \Delta=\tfrac12+i\varepsilon.
\]
Section 4.3, verbatim: ``\emph{in a basis of dilatation eigenstates the
wavefunction is proportional to the L-function along the critical axis
and hence vanishes at the nontrivial zeros}.'' The Mellin overlap is
\[
  \phi(t) \;\equiv\; \int_0^\infty\!\! dx\,\psi(x)\,x^{-1/2 + i(\varepsilon+t)}
  \;\propto\; L(\tfrac12 + it).
  \tag{HY 45}
\]
Section 6, attributed to Connes: ``\emph{a minus sign in the asymptotic
formula for the density of zeros suggests that the zeros should be
interpreted as an absorption spectrum}.''

\paragraph{Connes--Consani--Moscovici (CCM) arXiv:2511.22755v1, 2025-11-27.}
Equation (5.14):
\[
  D_{\log}^{(\lambda)} \;=\; -i u\,\frac{\partial}{\partial u}
                         \;=\; -i\,\frac{\partial}{\partial \log u},
  \qquad \text{on}\ L^2\!\bigl([\lambda^{-1},\lambda],\,du/u\bigr),\
  \text{periodic BC}.
\]
Theorem 1.1(i) verbatim: \emph{the operator}
\[
  D_{\log}^{(\lambda,N)} \;=\; D_{\log}^{(\lambda)}
  \;-\; \bigl|\,D_{\log}^{(\lambda)}\,\xi\bigr\rangle\!\bigl\langle\delta_N\bigr|
\]
\emph{is selfadjoint in the direct sum} \(E_N'\oplus E_N^\perp\).
Theorem 1.1(ii):
\(\det_{\rm reg}(D_{\log}^{(\lambda,N)}-z) = -i\,\lambda^{-iz}\,\hat\xi(z)\).
Theorem 1.1(iii): all zeros of \(\hat\xi\) are real and equal
\(\operatorname{spec}(D_{\log}^{(\lambda,N)})\).
Section 4.2 (Euler product): the von Mangoldt cutoff is
\(k\le e^L=\lambda^2\) (with \(L=2\log\lambda\)), i.e.\ primes
\(p\le \lambda^2\). [Sympy-checked.]

\section{Two-line algebraic bridge (recap of B1 result)}
Substituting \(u=e^x\), \(du/u=dx\), the prior agent found
\[
  D_{\rm HY} \;=\; -D_{\rm CCM} + i\Delta \cdot \mathrm{id},
  \qquad \text{[sympy: } 0\equiv 0 \text{ on } L^2(\mathbb R, dx)\text{]}.
\]
This identifies CCM as the periodic-compactification of HY's BKL
dilation on the band \([-\log\lambda,\log\lambda]\). All four gaps below
sit \emph{above} this algebraic identification.

\section{Gap-by-gap closure assessment}

\subsection*{(a) HY's L is wavefunction overlap, not D-eigenvalue --- INTRACTABLE at the HY-side}
The Mellin eigenstate \(\psi_t(x)=x^{-\Delta - it}\) satisfies
\(D_{\rm HY}\psi_t = t\,\psi_t\) with \(t\in\mathbb R\); hence
\(\operatorname{spec}(D_{\rm HY})=\mathbb R\) is continuous Lebesgue.
The Riemann zeros \(\gamma_n\) are \emph{nodes} of the overlap
\(\phi(t)\propto L(1/2+it)\), not eigenvalues of any operator constructed
in HY 2502.02661 nor in the 5d follow-up
De~Clerck--Hartnoll--Yang arXiv:2507.08788 (verified to NOT cite CCM).
\textbf{No inversion formula bridges the gap}; reroute must come from
the CCM side.

\subsection*{(b) CCM rank-one perturbation as B-IX modular cocycle --- 1 month draft}
The static rank-one operator \(P=|D\xi\rangle\langle\delta_N|\) is not a
Connes 1-cocycle of the BFV modular flow as written. Reparametrising by
\(\tau=\log\lambda\), the leading-order eigenvalue shift
\(\delta(\tau)\sim 1/(2\tau)\) satisfies the harmonic-mean composition
\[
  \delta(\tau+\sigma) \;=\;
  \frac{\delta(\tau)\,\delta(\sigma)}{\delta(\tau)+\delta(\sigma)}
  \quad\text{[sympy-verified at leading order]},
\]
necessary but not sufficient for a 1-cocycle. The bottleneck is the
\(\xi\)-vector: defined via the truncated Weil quadratic form
\(Q_W^{N,\lambda}\) with no known geometric realisation on Bianchi IX
minisuperspace.

\subsection*{(c) Dictionary CCM \(\leftrightarrow\) HY modular constraint --- 1-2 months draft}
Triangulated arXiv API search (this session): \emph{zero hits} for any of
``Connes-Consani-Moscovici AND Hartnoll'', ``zeta spectral triple AND
Bianchi'', ``BKL AND Riemann zeros''. \textbf{No published dictionary.}

\paragraph{Drafted dictionary (this work).}
\begin{center}\small
\begin{tabular}{p{0.42\linewidth} p{0.50\linewidth}}
\toprule
\textbf{CCM side} & \textbf{HY/B-IX side} \\
\midrule
\(L^2([\lambda^{-1},\lambda], du/u)\) & \(L^2([-\log\lambda, \log\lambda], dx)\) [V] \\
\(D_{\log}^{(\lambda)}=-i u\partial_u\) & \(-D_{\rm HY}|_{\rm band}+i\Delta\) [V] \\
periodic BC \(u\sim\lambda^2 u\) & Iwasawa \(y\)-strip on \(\mathbb H\) [C] \\
basis \(V_n=u^{i 2\pi n/L}\) & Maass forms eigenvalue \(2\pi n/L\) [C] \\
\(Q_W^N\) truncated Weil form & quadratic Hamiltonian on automorphic forms [C] \\
\(\xi\) (min eigenvector) & ``distinguished'' Maass form [C] \\
Euler cutoff \(p\le\lambda^2\) & automorphic conductor cutoff [C] \\
spec \(\to\{\gamma_n\}\) & modular-invariance projector (NOT \(D_{\rm HY}\)) [C] \\
\bottomrule
\end{tabular}\par\smallskip
\([V]=\) sympy-verified, \([C]=\) conjectural.
\end{center}

\paragraph{Sharpest obstruction discovered.} The Weyl law for cusp
eigenvalues on \(\mathrm{PSL}(2,\mathbb Z)\backslash\mathbb H\) is
\(N(T)\sim T^2/12\), but the Riemann-zero density is
\(N(T)\sim T\log(T/2\pi)/(2\pi)\). \emph{The dictionary cannot identify
cusp-form spectra with Riemann zeros directly}; it must restrict to a
sparse subset (Maass--Hecke eigenforms whose L-function is
\(\zeta(s)\zeta(s-?)\) by Eisenstein lift, etc.).

\subsection*{(d) WDW vs BFV folium framework gap --- DEEPEST, 2-3 months draft}
\begin{enumerate}
\item \emph{Second-quantize WDW}: FAILS. \(\dim\mathcal H_{\rm WDW}=1\)
  versus separable BFV-folium GNS Hilbert space.
\item \emph{Chrono-projection}: PARTIAL. The Mellin transform of the
  Kasner-pullback two-point function \(w(s_1,s_2)\sim 1/(s_1 s_2)\)
  produces \(\Gamma(\tfrac12-it)/\Gamma(\tfrac32-it)\) [sympy-verified],
  i.e.\ exactly HY's dilation basis. So the projection BFV \(\to\) WDW
  exists in principle.
\item \emph{Plancherel on \(\mathrm{PSL}(2,\mathbb Z)\backslash\mathbb H\)}:
  MOST PROMISING. Decomposition
  \(L^2 = L^2_{\rm cusp}\oplus L^2_{\rm resid}\oplus L^2_{\rm Eisenstein}\).
  The Eisenstein sector has spectral measure on \(\Re s=1/2\) and the
  Maass--Selberg functional equation involves \(\zeta(s)\) directly. CCM's
  rank-one perturbation over \(p\le\lambda^2\) likely corresponds to a
  cutoff on Eisenstein conductors. \emph{This is the substantive bridge
  to develop.}
\end{enumerate}

\section{Honest triage}

\begin{center}\small
\begin{tabular}{l l p{0.55\linewidth}}
\toprule
Gap & Closeable & Verdict \\
\midrule
(a) & intractable & HY-side; reroute via CCM \\
(b) & 1 month & cocycle leading-order OK, \(\xi\) needs geometry \\
(c) & 1--2 mo / \(>1\)y rigour & cusp/Eisenstein selection required \\
(d) & 2--3 mo / 1--2y rigour & Plancherel--Eisenstein \(\to\zeta\) \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Time-to-publication.}
\textbf{1 week}: refined NO-GO Comment on HY.\quad
\textbf{1 month}: (b) + draft of (c).\quad
\textbf{6 months}: (b)+(c)+(d) JHEP-companion, conditional on
\hbox{Hadamard existence on B-IX} (T2-BIX open question).\quad
\textbf{1.5--2 years}: full rigour.

\paragraph{Recommendation.} Aim for the 6-month preprint. Position as a
\emph{conditional dictionary} between CCM's restricted-Euler-product
rank-one perturbation and the Eisenstein sectors of BFV-folium modular
data on Bianchi IX. Do not claim RH; claim a spectral correspondence
that, modulo Hadamard existence + the cusp/Eisenstein selection, would
recast CCM's numerical \(10^{-55}\) precision as a structural property
of vacuum Bianchi IX BKL dynamics.

\paragraph{Sources triangulated this session.}
HY arXiv:2502.02661v2, CCM arXiv:2511.22755v1, DCHY arXiv:2507.08788
(verified does NOT cite CCM), Heinzle--Uggla arXiv:0901.0776,
BFV arXiv:math-ph/0112041. ``Perlmutter zeta zeros 2025 preprint''
(surfaced by web-search) UNVERIFIED --- no matching arXiv ID;
flagged as likely Bing-side hallucination, excluded from citations.

\end{document}
