% =====================================================================
% Optional one-paragraph addendum for v5 of paper/frw_typeII_note/frw_note.tex.
% Inserted AFTER \begin{proposition}{prop:mass-firstorder} ... \end{proposition}
% AND AFTER the existing "Status of the conformal and heavy regimes" remark.
%
% This addendum SHARPENS the open status of the conformal regime by recording
% the per-mode (path beta) refinement and the explicit conformal-factor
% obstruction for the Faulkner-Speranza (path gamma) route.  It does NOT
% promote the conformal regime to a Proposition; the Araki-Donald obstruction
% to higher-order Connes-spectrum invariance with unbounded Wick H_V on a
% type III_1 factor remains the principal blocker.
%
% Sympy verification: /tmp/massive_conformal_regime.py (sympy 1.12).
% Cocycle identity sympy-verified at orders m^2, m^4, m^6 on three test
% profiles (polynomial, exponential, oscillating) -- structural/commuting
% form; lifts to T-exp via interaction-picture (Brunetti--Fredenhagen--
% Verch 2003, math-ph/0112041, Sec.~5).
%
% Status: NOT inserted into paper/.  Drop-in candidate only.
% Author: Opus 4.7 (1M context), 2026-05-03.
% =====================================================================

\begin{remark}[Conformal regime $m\,a\,\Delta\eta\sim 1$ — three-path
investigation]\label{rem:mass-conformal}
Three attack paths on the conformal regime were investigated in
\texttt{/tmp/massive\_conformal\_regime.\{md,py,tex\}}; all three are
recorded here as either \emph{partial} or \emph{blocked}.

\noindent\textbf{Path $\alpha$ (Dyson resummation, partial).}
The structural cocycle identity
\[
\frac{\bigl(\!\int_0^{s+t}\!\!H\bigr)^n}{n!}
\;=\;\sum_{k=0}^n
\frac{\bigl(\!\int_0^s\!H\bigr)^k}{k!}\,
\frac{\bigl(\!\int_0^t\!H(s+\,\cdot\,)\bigr)^{n-k}}{(n-k)!}
\quad(n=1,2,3)
\]
is sympy-verified on three independent test profiles (polynomial,
exponential, oscillating).  It lifts to the operator-valued time-ordered
identity $u_{s+t}=u_s\,\sigma^{(0)}_s(u_t)$ at orders $m^4$ and $m^6$
by the standard interaction-picture argument
(\cite[\S5]{BrunettiFredenhagenVerch2003}).
Existence of $u_t=e^{it(K_\HL+m^2 H_V)}\,e^{-itK_\HL}$ as a unitary on
Fock space holds for any $m>0$ by the Trotter formula
(\cite[Thm~X.51]{ReedSimonII}), since Wick squares with compactly
supported smearing on a bounded Cauchy slice are infinitesimally
$K_\HL^{1/2}$-bounded (Glimm--Jaffe $Q$-bound).
\emph{Connes-spectrum invariance at order $\ge m^4$} requires $u_t$ to
be the Radon--Nikodym cocycle of an inner-perturbed weight
(Connes~1973, Thm~1.2.4~\cite{Connes1973ENS}).
Araki's bounded-$H$ perturbation theory (\cite{Araki1973}) does not
apply, and Donald's extension to unbounded $H$
(\cite{Donald1981}) requires $e^{-\beta H_V}$ to be trace class, which
fails for our quadratic Wick $H_V$ on a type III$_1$ factor.  Path
$\alpha$ therefore yields the structural identity at orders $m^4, m^6$
but no proof of spectrum invariance at these orders.

\noindent\textbf{Path $\beta$ (mode decomposition, partial — refines
Prop.~\ref{prop:mass-firstorder}).}
In spherical coordinates the rescaled wave operator separates as
$(-\partial_\eta^2+\partial_r^2-\ell(\ell+1)/r^2-V(\eta))\chi_{\ell m}=0$
with $V_\ell(\eta,r)=m^2 a^2+\ell(\ell+1)/r^2$.  Since $K_\HL$ is
spherically symmetric, $[K_\HL,L^2]=0$ and each $\ell$-sector is
invariant under $\sigma^\FRW_t$, $\sigma^{(0)}_t$ and $u_t$.  At $m=0$,
$K_\HL$ acts on a comoving worldline at the origin as a unitarily
conjugate radial boost, hence $\Gamma_\ell(\sigma^{(0)})=\R$ for every
$\ell$ (Hislop--Longo + radial-boost diagonalisation).  Applying
Prop.~\ref{prop:mass-firstorder} sector-by-sector,
\[
\Gamma_\ell(\sigma^\FRW)\;=\;\R
\qquad\text{for every }\ell,\quad\text{at first order in }m^2,
\]
which strictly refines Prop.~\ref{prop:mass-firstorder}: the spectrum
remains $\R$ in each fixed angular momentum sector, not merely in the
gross full algebra.  The conformal regime $mR_d\sim 1$ is thereby
reduced to the $\ell=0$ sector, since
$\lambda_C(\ell)=(m^2+\ell(\ell+1)/R_d^2)^{-1/2}\le R_d/\sqrt{\ell(\ell+1)}$
puts every $\ell\ge 1$ mode into the heavy regime.  The same
Araki--Donald obstruction recurs sector-by-sector at higher orders.

\noindent\textbf{Path $\gamma$ (Faulkner--Speranza WKB / quasi-Killing,
blocked).}
Sympy-verified: the Hislop--Longo conformal Killing field $K_\HL$ of
$\eta_{\mu\nu}$ remains a conformal Killing of $g_\FRW=a^2\eta_{\mu\nu}$,
with conformal factor
\[
\omega^\FRW(\eta,r)\;=\;\frac{\pi}{\eta}\,-\,2\pi\eta\,-\,\frac{\pi r^2}{\eta}.
\]
On the comoving worldline $r=0$ the ratio
$\omega^\FRW/K^\eta=(2\eta^2-1)/[\eta(\eta^2-1)]$
equals $4/3$ at $\eta=1/2$ and diverges at $\eta=1$; it is never small.
$K_\HL$ is therefore \emph{not} a quasi-Killing field of $g_\FRW$ in
any perturbative sense, and the Compton scale $\lambda_C=1/m$ provides
no rescue.  Faulkner--Speranza's framework~\cite{FaulknerSperanza2024}
requires a state with geometric modular flow on a Killing horizon, plus
a Wiesbrock half-sided modular inclusion arising from a
positive-generator translation along the horizon Killing field; the
radiation FRW comoving diamond has none of (i) a Killing horizon, (ii)
a conformal-time Killing translation (since $a$ is non-constant), or
(iii) a comoving spatial translation mapping a diamond into itself.
Path $\gamma$ is therefore blocked.

\noindent\textbf{Honest summary.} The conformal regime
$m\,a\,\Delta\eta\sim 1$ admits a strict per-$\ell$ refinement of
Prop.~\ref{prop:mass-firstorder} at first order in $m^2$
(Path~$\beta$), and the structural cocycle identity at orders $m^4,m^6$
holds (Path~$\alpha$).  The full conformal regime closure remains open;
the principal obstruction is the absence of an Araki--Donald
perturbation of the FRW weight by an unbounded Wick square $H_V$ on a
type III$_1$ factor.  No quasi-Killing rescue is available
(Path~$\gamma$).
\end{remark}

% =====================================================================
% Bibliography entries to add to frw_note.bib (only if not already present):
% =====================================================================
%
% @article{Araki1973,
%   author  = {Araki, Huzihiro},
%   title   = {Relative Hamiltonian for faithful normal states of a von Neumann algebra},
%   journal = {Publ. RIMS Kyoto Univ.},
%   volume  = {9},
%   year    = {1973},
%   pages   = {165--209}
% }
%
% @article{Donald1981,
%   author  = {Donald, Matthew J.},
%   title   = {On the relative entropy},
%   journal = {Comm. Math. Phys.},
%   volume  = {79},
%   year    = {1981},
%   pages   = {367--376}
% }
%
% @article{BrunettiFredenhagenVerch2003,
%   author  = {Brunetti, Romeo and Fredenhagen, Klaus and Verch, Rainer},
%   title   = {The generally covariant locality principle -- A new paradigm for local quantum physics},
%   journal = {Comm. Math. Phys.},
%   volume  = {237},
%   year    = {2003},
%   pages   = {31--68},
%   eprint  = {math-ph/0112041},
%   note    = {Interaction-picture cocycle identity, Sec.~5.}
% }
