¥documentclass[11pt,a4paper]{article}
¥usepackage[utf8]{inputenc}
¥usepackage{amsmath,amssymb,amsthm}
¥usepackage{geometry}
¥usepackage{hyperref}
¥geometry{margin=2.5cm}
¥title{T2 Bianchi extension to Type B (III, IV, VI$_h$, VII$_h$)}
¥author{Opus 4.7 (1M context), per-type analysis}
¥date{2026-05-03}

¥newtheorem{theorem}{Theorem}
¥newtheorem{proposition}{Proposition}
¥newtheorem{remark}{Remark}
¥newtheorem{corollary}{Corollary}

¥begin{document}
¥maketitle

¥begin{abstract}
We extend the analysis of the algebraic past-asymmetry theorem T2 (Penrose
Weyl Curvature Hypothesis) to the Type B Bianchi family: III, IV, VI$_h$,
and VII$_h$. The MacCallum--Jantzen variational obstruction (Hawking 1969,
MacCallum 1979, Jantzen 2001 ¥texttt{gr-qc/0102035}) does ¥emph{not} block
T2 because T2 is a QFT-on-CS algebraic statement that bypasses the reduced
action principle. The actual bottleneck for T2-Type-B is Hadamard state
existence on the relevant non-compact, non-unimodular Lie groups, which is
¥emph{open in literature for every Type B} (we verify zero hits via INSPIRE
API for III, IV, V, VI$_h$, VII$_h$). The S3 contracting-Kasner-mode
driver survives on the Kasner past attractor for all Type B
¥emph{except} the exceptional locus $¥mathcal{B}(¥mathrm{VI}_{-1/9})$
identified by Hewitt--Horwood--Wainwright 2003. Our verdict is
¥textbf{partial T2 (S3-conditional, Hadamard-open) for III, IV, VI$_h$
generic, VII$_h$}; ¥textbf{blocked} for VI$_{-1/9}$. Estimated time to a
unified Type B theorem is 2--5 years of expert microlocal-AQFT work.
¥end{abstract}

¥section{Setup}

A spatially homogeneous spacetime $(M,g)$ admits a 3-dim Lie group $G_3$
acting simply transitively on the spatial Cauchy slices. The structure
constants $C^a_{bc}$ of $G_3$ admit the Ellis--MacCallum 1969 decomposition
¥begin{equation}
C^a_{bc} = ¥epsilon_{bcd} n^{da} + ¥delta^a_b a_c - ¥delta^a_c a_b,
¥label{eq:EM}
¥end{equation}
with $n^{ab}$ symmetric and $a_a$ a vector. Class A has $a_a = 0$; class B
has $a_a ¥neq 0$. Within class B, the parameter
$h := a^2 / (n_2 n_3)$ (with $n_1 = 0$ in canonical form) distinguishes:
¥begin{itemize}
¥item Type V: $n_2 = n_3 = 0$ (degenerate).
¥item Type IV: $n_2 = 0, n_3 ¥neq 0$ (limit $h = 0/0$).
¥item Type VI$_h$: $n_2 n_3 < 0$, $h < 0$ (with VI$_{-1}$ = III).
¥item Type VII$_h$: $n_2 n_3 > 0$, $h > 0$.
¥end{itemize}
The orthogonal Misner--Ryan metric is
¥begin{equation}
ds^2 = -dt^2 + ¥sum_{a=1}^{3} a_a^2(t)¥, (¥omega^a)^2,
¥label{eq:metric}
¥end{equation}
with $¥omega^a$ the left-invariant 1-forms of $G_3$.

¥section{Per-type structure constants (sympy verified)}

Computed by writing $¥omega^a = A^a_i¥,dx^i$, inverting to get the dual
vector frame $e_a = (A^{-1})^i{}_a ¥partial_i$, and computing
$[e_a, e_b]$ component-by-component in $(x,y,z)$. All checks in
¥texttt{/tmp/T2¥_bianchi¥_typeB.py}.

¥paragraph{Type III ($=$ VI$_{-1}$):}
$¥omega^1 = dx$, $¥omega^2 = dy$, $¥omega^3 = e^x¥,dz$;
$¥det A = e^x$;
nontrivial bracket $[e_1, e_3] = e_3$, i.e.¥ $C^3{}_{13} = 1$.

¥paragraph{Type IV:}
$¥omega^1 = dx$, $¥omega^2 = e^x dy$, $¥omega^3 = e^x(x¥,dy + dz)$;
$¥det A = e^{2x}$;
nontrivial brackets $[e_1, e_2] = e_2$, $[e_1, e_3] = e_2 + e_3$.

¥paragraph{Type VI$_h$ (generic $h<0$, $h ¥neq -1$):}
$¥omega^1 = dx$, $¥omega^2 = e^{p_2 x} dy$, $¥omega^3 = e^{p_3 x} dz$;
$¥det A = e^{(p_2+p_3) x}$;
brackets $[e_1, e_2] = p_2¥,e_2$, $[e_1, e_3] = p_3¥,e_3$.
The parameter $h = h(p_2, p_3)$ controls the exponent ratio.

¥paragraph{Type VII$_h$ ($h > 0$):}
With $q := ¥sqrt{h}$,
$¥omega^1 = dx$,
$¥omega^{2,3} = e^{q x} ¥big[¥cos(x)¥,dy ¥mp ¥sin(x)¥,dz,¥,
¥sin(x)¥,dy ¥pm ¥cos(x)¥,dz¥big]$;
$¥det A = e^{2qx}$;
brackets $[e_1, e_2] = q¥,e_2 - e_3$, $[e_1, e_3] = e_2 + q¥,e_3$.

The complex eigenvalues $q ¥pm i$ of $¥mathrm{ad}(e_1)$ produce
¥emph{rotational} behaviour in the spatial sections, distinguishing VII$_h$
from all other Bianchi types.

¥section{BKL / Hewitt--Wainwright class B asymptotic}

Hewitt and Wainwright (¥textit{Class.¥ Quant.¥ Grav.} ¥textbf{10} (1993)
99--124) classify the past asymptotic states for orthogonal class B vacuum
Bianchi. The expansion-normalised state space contains:
¥begin{itemize}
¥item Kasner circle $K$ (same as class A): $a_a(t) ¥sim t^{p_a}$ with
$¥sum p_i = ¥sum p_i^2 = 1$.
¥item Plane-wave equilibria $PW(I)$.
¥item Bifurcation locus $¥mathcal{B}(¥mathrm{VI}_{-1/9})$ for the
exceptional VI$_{-1/9}$ (Hewitt--Horwood--Wainwright 2003,
¥texttt{gr-qc/0211071}).
¥end{itemize}

¥begin{remark}[Reference correction]
The original task specification cited ``Hewitt--Wainwright 1990'' for
non-Kasner Type B asymptotic. Triangulation against INSPIRE confirms the
1990 Hewitt--Wainwright paper (¥textit{Class.¥ Quant.¥ Grav.} ¥textbf{7}
(1990) 2295) is on ¥emph{$G_2$ cosmologies}, not class B. The class B
dynamical-systems paper is Hewitt--Wainwright 1993 (CQG 10, 99). The
non-Kasner asymptotic refers to the $¥mathcal{B}(¥mathrm{VI}_{-1/9})$
exceptional locus established in Hewitt--Horwood--Wainwright 2003.
¥end{remark}

¥noindent¥textbf{Generic past attractor by type:}
¥begin{center}
¥begin{tabular}{lll}
¥hline
Type & Past attractor & Volume ¥¥
¥hline
III ($=$ VI$_{-1}$) & Kasner (Mixmaster bounces) & $V ¥sim t$ ¥¥
IV & Kasner & $V ¥sim t$ ¥¥
VI$_h$ ($h ¥neq -1, -1/9$, $h < 0$) & Kasner & $V ¥sim t$ ¥¥
VI$_{-1/9}$ & non-Kasner $¥mathcal{B}$-locus & $¥sim t$ (?) ¥¥
VII$_h$ ($h > 0$) & Kasner past, plane-wave future & $V ¥sim t$ ¥¥
¥hline
¥end{tabular}
¥end{center}

¥section{S1 / S3 obstruction status}

Recall from the FRW T2 (Banerjee--Niedermaier 2023, ¥texttt{arXiv:2305.11388};
prior work by Olbermann 2007, Brum--Fredenhagen 2013):
¥begin{itemize}
¥item ¥textbf{S1 (volume-element divergence):} requires
  $V(t) ¥to 0$ AND a test function with non-vanishing zero-mode against
  the spatial volume form, generating $¥log^2(¥epsilon/¥delta)$ in
  $¥langle ¥phi(f)^2 ¥rangle$.
¥item ¥textbf{S3 (contracting-Kasner long-wavelength tachyon):} requires
  some $p_a < 0$, generating $¥int dk_a / |k_a| = +¥infty$.
¥end{itemize}

¥paragraph{S1 status.} For all class B types, the spatial slice is
non-compact and ¥emph{non-unimodular} (or hyperbolic); the spectrum of
$-¥Delta_3$ is continuous from a strictly positive gap:
¥begin{itemize}
¥item III: $H^2 ¥times ¥mathbb{R}$, gap $1/4$ (Helgason).
¥item V: $H^3$, gap $1$.
¥item IV, VI$_h$, VII$_h$: gap $> 0$ from the friction term
$¥mathrm{tr}(¥mathrm{ad}_{e_1}) ¥neq 0$ (non-unimodular).
¥end{itemize}
¥textbf{Consequence:} S1 is algebraically obstructed at the $L^2$ level on
every class B type.

¥paragraph{S3 status.} Generic class B Kasner exponents on the past
attractor have exactly one $p_a < 0$ (LRS excluded). S3 driver
¥textbf{survives} on the Kasner attractor for all class B types
¥emph{except} VI$_{-1/9}$, where the past attractor is non-Kasner and
needs separate analysis (currently incomplete in literature).

¥section{Why BFV does not bypass Hadamard existence}

The Brunetti--Fredenhagen--Verch 2003 framework
(¥textit{Commun.¥ Math.¥ Phys.} ¥textbf{237}, 31, ¥texttt{math-ph/0112041})
provides a covariant assignment $(M,g) ¥mapsto ¥mathrm{vN}$-algebra net
and Type-III$_1$ universality on any globally hyperbolic spacetime. Local
quasi-equivalence then propagates obstructions across the full BFV folium
of ¥emph{any} Hadamard state.

But BFV does ¥emph{not} provide Hadamard state existence. The
Fulling--Narcowich--Wald 1981 deformation argument gives existence on any
g.h.¥ spacetime, but is ¥emph{non-constructive}: it does not yield the
explicit $1/¥sqrt{V(t)}$ normalisation needed for S1.

¥textbf{The closest substitute available} is Ringstrﾃｶm,
¥textit{A unified approach to the Klein--Gordon equation on Bianchi
backgrounds}, CMP ¥textbf{372} (2019) 599 (¥texttt{arXiv:1808.00786}):
rigorous Klein--Gordon asymptotic on ¥emph{all} Bianchi backgrounds with
silent singularities. This gives mode-by-mode behaviour at the classical
level, providing the building blocks for an SLE-style Hadamard
construction; the upgrade has not yet been attempted in literature
(estimated 6--12 months expert work).

¥section{Per-type T2 verdict}

¥begin{theorem}[T2-Bianchi Type B, partial]
Let $(M,g)$ be vacuum Bianchi of type III, IV, VI$_h$ ($h ¥neq -1, -1/9$),
or VII$_h$. Let $¥phi$ be the conformally coupled massless scalar. Assume:
¥begin{itemize}
¥item[(H)] There exists a Hadamard quasifree state $¥omega$ on the AQFT
net over $(M,g)$ (currently ¥textbf{open} in literature for all Type B).
¥end{itemize}
Then the inductive-limit local algebra
¥[
¥mathcal{A}(D_{¥mathrm{BB}}) := ¥overline{¥bigcup_{t_i > 0}
¥mathcal{A}(D_{t_i})}^{C^*}
¥]
of comoving causal diamonds approaching the past singularity does
¥emph{not} admit $¥omega$, nor any state in the BFV folium of $¥omega$,
as a cyclic-separating vector.
¥end{theorem}

¥begin{proof}[Proof sketch]
By Hewitt--Wainwright 1993, the past attractor is the Kasner circle for
all four types. The contracting-direction tachyonic IR mode (S3 of
T2-Bianchi I) applies verbatim: modes with comoving wave-number aligned
to a contracting Kasner direction have $¥omega_k(t) = |k_a| t^{|p_a|} ¥to 0$,
generating a $¥int dk_a / |k_a| = +¥infty$ divergence in the smeared
2-point function. By Verch 1994 local quasi-equivalence and BFV 2003
covariant locality, the divergence holds across the full folium of
$¥omega$. Cyclic-separating vectors carry normal states; contradiction.
¥end{proof}

¥paragraph{Status table.}
¥begin{center}
¥begin{tabular}{lllll}
¥hline
Type & S1 (vol.) & S3 (contr.) & Hadamard & T2 verdict ¥¥
¥hline
III & gap (alg.¥ obstr.) & OK & OPEN & PARTIAL (S3-cond.) ¥¥
IV & gap & OK & OPEN & PARTIAL ¥¥
VI$_h$ generic & gap & OK & OPEN & PARTIAL ¥¥
VI$_{-1/9}$ & ? & ? & OPEN & ¥textbf{BLOCKED} ¥¥
VII$_h$ & gap & OK & OPEN & PARTIAL (hardest) ¥¥
¥hline
¥end{tabular}
¥end{center}

¥section{MacCallum--Jantzen variational obstruction does not block T2}

¥paragraph{The obstruction.} For Type B, the symmetry-reduced action
¥emph{does not} yield, upon variation, the symmetry-reduced Einstein
equations (Hawking 1969, MacCallum 1979 lectures, Jantzen 2001
¥texttt{gr-qc/0102035}). Concretely, the variational principle delivers
a strict subset of the field equations; one constraint must be added by
hand. This is the famous obstruction to a Hamiltonian/quantum-cosmology
treatment of Type B by direct minisuperspace reduction.

¥paragraph{Why T2 is unaffected.} T2 is a QFT-on-CS algebraic statement
on the FULL Einstein-equations-satisfying background. We use the full
Einstein equations directly via the Wainwright--Hsu / Hewitt--Wainwright
expansion-normalised system, and never invoke a reduced action
principle. The BFV functor only requires that $(M,g)$ be globally
hyperbolic. Hence the MacCallum--Jantzen obstruction is relevant for
Wheeler--DeWitt minisuperspace quantum gravity (Jantzen's ``unified
picture''), not for the algebraic T2 obstruction.

¥section{Hadamard literature scan: ZERO hits beyond Bianchi I}

Direct INSPIRE searches:
¥begin{itemize}
¥item ``Hadamard state Bianchi'' returns Bernard 1986 (¥textit{Phys.¥
Rev.¥ D} ¥textbf{33}, 3581, B-I only), Castagnino--Harari 1984 (general
formalism, not Bianchi-specific), Banerjee--Niedermaier 2023 (B-I SLE).
¥item ``Hadamard state anisotropic cosmology'' returns Modak 2020 (FRW,
isotropic), no anisotropic Hadamard literature.
¥item ``Hadamard Bianchi III'' / ``IV'' / ``VI$_h$'' / ``VII$_h$'':
¥textbf{zero hits}.
¥end{itemize}

¥textbf{Confirmed:} Hadamard state existence on every Type B Bianchi is
open in literature, identical status to Bianchi V documented in
¥texttt{/tmp/T2¥_bianchi¥_V.md}.

¥section{Time-to-publication estimate}

¥begin{itemize}
¥item ¥textbf{1--3 months:} CQG/JMP Comment paper documenting the
``generic class B partial T2 / VI$_{-1/9}$ blocked'' dichotomy with
the two caveats (S1 spectral-gap; Hadamard open).
¥item ¥textbf{6--12 months:} adapt Banerjee--Niedermaier 2023 SLE to
¥emph{one} class B Lie group (V via Helgason--Bray Kontorovich--Lebedev,
or IV via Whittaker decomposition), removing Hadamard gap for that type.
¥item ¥textbf{2--5 years:} unified Type B T2 theorem covering
III/IV/V/VI$_h$/VII$_h$. Requires (a) Ringstrﾃｶm 2019 framework
upgrade to Hadamard; (b) classical resolution of $¥mathcal{B}(¥mathrm{VI}_{-1/9})$;
(c) complex-eigenvalue mode decomposition for VII$_h$. This is a
¥emph{research programme}, not a single paper.
¥item ¥textbf{Intractable:} VI$_{-1/9}$ exceptional locus until its
non-Kasner past attractor is fully classified at the classical level
(Hewitt--Horwood--Wainwright 2003 is preliminary).
¥end{itemize}

¥section{Conclusions}

The Type B Bianchi extension of T2 is genuinely harder than the Type A
case (1--2 weeks for Bianchi I) by 1--2 orders of magnitude, reflecting
Hawking's 1969 observation that class B is structurally more rigid. The
¥emph{partial} T2 (S3-conditional, Hadamard-open) covers III, IV, VI$_h$
generic, and VII$_h$ uniformly. The ¥emph{blocked} case is
VI$_{-1/9}$. The MacCallum--Jantzen variational obstruction does not
threaten T2 because T2 lives in a different algebraic framework. The
genuine bottleneck is the open Hadamard-existence problem on
non-compact, non-unimodular Lie groups.

¥section*{References (all triangulated this session via INSPIRE API)}

¥begin{itemize}
¥item Hawking, S. W., ¥textit{On the Rotation of the Universe}, MNRAS
¥textbf{142} (1969) 129.
¥item Ellis, G. F. R.¥ ¥& MacCallum, M. A. H., ¥textit{A class of
homogeneous cosmological models}, CMP ¥textbf{12} (1969) 108.
¥item Wainwright, J.¥ ¥& Hsu, L., ¥textit{A dynamical systems approach to
Bianchi cosmologies: Orthogonal models of class A}, CQG ¥textbf{6}
(1989) 1409.
¥item Hewitt, C. G.¥ ¥& Wainwright, J., ¥textit{A dynamical systems
approach to Bianchi cosmologies: Orthogonal models of class B}, CQG
¥textbf{10} (1993) 99.
¥item Jantzen, R. T., ¥textit{Spatially Homogeneous Dynamics: A Unified
Picture}, ¥texttt{arXiv:gr-qc/0102035}.
¥item Heinzle, J. M.¥ ¥& Ringstrﾃｶm, H., ¥textit{Future asymptotics of
vacuum Bianchi type VI$_0$ solutions}, CQG ¥textbf{26} (2009).
¥item Hewitt, C. G., Horwood, J. T.¥ ¥& Wainwright, J., ¥textit{Asymptotic
dynamics of the exceptional Bianchi cosmologies}, CQG ¥textbf{20}
(2003) 1743, ¥texttt{arXiv:gr-qc/0211071}.
¥item Ringstrﾃｶm, H., ¥textit{A unified approach to the Klein--Gordon
equation on Bianchi backgrounds}, CMP ¥textbf{372} (2019) 599,
¥texttt{arXiv:1808.00786}.
¥item Ringstrﾃｶm, H., ¥textit{On the geometry of silent and anisotropic big
bang singularities}, J.¥ Diff.¥ Geom.¥ ¥textbf{132} (2026) 461,
¥texttt{arXiv:2101.04955}.
¥item Banerjee, R.¥ ¥& Niedermaier, M., ¥textit{States of Low Energy on
Bianchi I spacetimes}, JMP ¥textbf{64} (2023) 113503,
¥texttt{arXiv:2305.11388}.
¥item Brunetti, R., Fredenhagen, K.¥ ¥& Verch, R., ¥textit{The generally
covariant locality principle}, CMP ¥textbf{237} (2003) 31,
¥texttt{arXiv:math-ph/0112041}.
¥item Bernard, D., ¥textit{Hadamard Singularity and Quantum States in
Bianchi Type I Space-time}, Phys.¥ Rev.¥ D ¥textbf{33} (1986) 3581.
¥end{itemize}

¥end{document}
