% v6 JHEP companion draft — FORMAL track
% Title candidates (owner to choose one; delete the other two before tag):
%   (T1) An information-theoretic upper bound on observer-dependent entropy
%        growth in type-II algebras
%   (T2) Modular-time refinement of the generalised second law:
%        a dequantisation bridge
%   (T3) From Faulkner--Speranza to a complexity-bounded GSL in
%        crossed-product algebras
%
% Class choice: jheppub.cls NOT installed on build host (2026-04-22;
% SISSA download 404 on JHEP3.zip). FALLBACK: RevTeX 4-2 for the draft
% stage. Swap to jheppub once class is fetched — see v6/README.md.

\documentclass[aps,prd,superscriptaddress,nofootinbib,floatfix,10pt]{revtex4-2}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tcolorbox}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}

\newcommand{\Sgen}{S_{\mathrm{gen}}}
\newcommand{\tauR}{\tau_{R}}
\newcommand{\kR}{\kappa_{R}}
\newcommand{\Ck}{\mathcal{C}_{k}}
\newcommand{\PHk}{\mathrm{PH}_{k}}
\newcommand{\dn}{\delta n}
\newcommand{\AR}{\mathcal{A}_{R}}

\begin{document}

\title{From Faulkner--Speranza to a complexity-bounded\\
generalised second law in type-II crossed-product algebras}

\author{[author list pending]}
\affiliation{[affiliation pending]}

\date{\today}

\begin{abstract}
We extend the Eling--Guedens--Jacobson local Clausius schema
$dS = \delta Q / T + d_{i} S$ to the type-II crossed-product algebra
setting of Chandrasekaran--Longo--Penington--Witten and
De~Vuyst--Eccles--Held--Kirklin. Along the modular (Connes--Tomita--Takesaki)
flow $\tauR$ of an observer region $R$, we derive an upper bound on the
rate of generalised entropy in the form
$\, d\Sgen[R]/d\tauR \le \kR \cdot \Ck[\rho_{R}(\tauR)]
 \cdot \Theta(\PHk[\dn(\tauR)])\,$,
where $\Ck$ is the $k$-design complexity of the reduced state,
$\PHk$ is the order-$k$ persistent-homology Betti number of a
dequantised density field $\dn$, and $\Theta$ is a chameleon-type
topological activator. The dequantisation map
$\dn(x,\tauR) \equiv \mathrm{Tr}_{R}[\rho_{R}(\tauR)\hat n(x)] - \langle n\rangle$
bridges the type-II factor to classical persistent homology. The bound
recovers Wall's monotonicity inequality and the Faulkner--Speranza
modular form in the $\Theta\to 1$ limit, and saturates to the
Fan (2022) logarithmic Krylov regime under $\Ck\to\exp(S)$. We make
no cosmological claim; phenomenological consequences are deferred to a
companion paper.
\end{abstract}

\maketitle

% =============================================================
\section{Introduction}
\label{sec:intro}
% =============================================================

The local Clausius derivation of Einstein's equations by
Jacobson~\cite{Jacobson1995} and its non-equilibrium extension by
Eling--Guedens--Jacobson (EGJ, 2006) cast gravity as thermodynamics on
local causal horizons, with an internal-production term
$d_{i}S$ that encodes the departure from equilibrium. Two recent
developments invite a modular-time refinement of that schema. First,
the type-II crossed-product algebras $\AR$ of
CLPW~\cite{CLPW2023} and
DEHK~\cite{DEHK2025a,DEHK2025b} promote observer-dependent generalised
entropy from a formal pairing to a von Neumann trace. Second, the
differential GSL of Faulkner--Speranza~\cite{FaulknerSperanza2024} and
Kirklin~\cite{Kirklin2025} upgrades Wall's monotonicity to a modular
inequality for relative entropy.

The present paper combines these two lines. We propose a
differential upper bound on $d\Sgen/d\tauR$ whose source is the
$k$-design complexity $\Ck[\rho_{R}]$ of the reduced state and whose
topological activator is an exponential function of the persistent-
homology Betti number $\PHk[\dn]$ of a dequantised density field $\dn$.
The result (Eq.~\eqref{eq:main}) is a formal statement; no
cosmological prediction is claimed. Phenomenological falsifiers and
observational forecasts are developed in a companion paper
(v5.0, arXiv: to be assigned).

The novelty is threefold.
(a)~A differential modular-time bound, versus the static
type-II second law of CLPW/DEHK.
(b)~The appearance of $k$-design complexity~\cite{MaHuang2025,Haferkamp2022}
as the dominant source term, replacing geometric shear squared (EGJ) and
logarithmic Krylov complexity (Fan 2022); the logarithmic form is recovered
at saturation (Sec.~\ref{sec:relation}).
(c)~An explicit dequantisation map from $\AR$ to a classical scalar field
$\dn$, which resolves the category mismatch between quantum reduced states
and classical persistent-homology functors.

% =============================================================
\section{Setup}
\label{sec:setup}
% =============================================================

Let $\AR$ denote the type-II crossed-product algebra associated with an
observer-with-clock in region $R$, as constructed by
CLPW~\cite{CLPW2023} and extended by
DEHK~\cite{DEHK2025a,DEHK2025b}. The Connes cocycle fixes, up to an
additive constant, a faithful normal semifinite trace, and we denote by
$\sigma^{R}_{\tauR}$ the one-parameter automorphism of $\AR$ induced by
Tomita--Takesaki modular flow; $\tauR$ is the thermal (Connes--Rovelli)
time for a thermofield double state on $\AR$. The reduced state
$\rho_{R}(\tauR)\equiv\sigma^{R}_{\tauR}(\rho_{R})$ evolves
covariantly.

We adopt the modular-temperature identification
\begin{equation}
 \kR \;\equiv\; 2\pi\,T_{R},
 \label{eq:kappa}
\end{equation}
with $T_{R}$ the local Tomita--Takesaki temperature, so that $\kR$ has
units of nats per modular e-fold and reduces to the Hubble rate $H$ at
the de~Sitter Gibbons--Hawking limit.

The generalised entropy is the type-II functor
$\Sgen[R] = A[R]/(4G_{N}) + S_{\mathrm{matter}}[\rho_{R}]$, well-defined
because the crossed-product factor is semifinite.
The $k$-design complexity $\Ck[\rho_{R}]$ is defined in the
Ma--Huang pseudorandom-unitary (PRU) sense~\cite{MaHuang2025};
it is a dimensionless, monotone, state-dependent functional that
interpolates between circuit complexity~\cite{Haferkamp2022} and
information-theoretic complexity~\cite{CryptoCensorship}.
The Betti number $\PHk[\dn]$ is the order-$k$ persistent-homology count
on a classical density field $\dn$ (defined in Sec.~\ref{sec:dequant}),
following the Yip~et~al.\ cosmological pipeline~\cite{Yip2024}.

% =============================================================
\section{Main inequality}
\label{sec:main}
% =============================================================

\begin{tcolorbox}[colback=blue!3,colframe=blue!50,title=Main result (Theorem~\ref{thm:main})]
\begin{equation}
 \boxed{\;
  \frac{d\Sgen[R]}{d\tauR}
  \;\le\;
  \kR \,\cdot\, \Ck\!\left[\rho_{R}(\tauR)\right]
       \,\cdot\, \Theta\!\left(\PHk[\dn(\tauR)]\right)
 \;}
 \label{eq:main}
\end{equation}
\end{tcolorbox}

Equation~\eqref{eq:main} is an inequality, not an equality; the
equality attempt was ruled out by an internal multi-agent derivation
audit (three independent routes converged on \emph{ansatz} status;
cf.\ Appendix~\ref{app:computation}). The inequality is tight: it is
saturated in the scrambling regime where
$\Ck\to\exp(S_{R})$, and in that limit recovers the Fan~(2022)
logarithmic Krylov law as the \emph{saturating}, not \emph{derived},
rate (Sec.~\ref{sec:relation}).

The bound rests on three labelled postulates.

\begin{assumption}[M1 --- Modular complexity ansatz, POSTULATE]
\label{ass:M1}
Under modular flow on $\AR$, the internal-production part of
$d\Sgen/d\tauR$ is bounded above by $\kR\,\Ck[\rho_{R}]$. This is a
Brown--Susskind-style ``second law of complexity'' ansatz adapted to
type-II modular time, in the scaffolding proposed by
Caputa--Mag\'an (2024); no type-II theorem is claimed.
\end{assumption}

\begin{assumption}[M2 --- Semi-classical recovery of $\dn$, ANSATZ]
\label{ass:M2}
In the large-$N$ QRF coarse-graining limit, the dequantisation map
(Sec.~\ref{sec:dequant}) produces a mean-zero Gaussian random field
$\dn$ by a central-limit / law-of-large-numbers argument on the
unobserved QRF degrees of freedom. The Gaussian recovery is verified
partially on a toy type-II$_{1}$ factor
(cf.\ \texttt{V6-dequantisation-map.py}); a full proof is outside the
scope of this paper.
\end{assumption}

\begin{assumption}[M3 --- Topological activator profile, CONJECTURAL]
\label{ass:M3}
The activator $\Theta$ takes the chameleon form
$\Theta(\PHk) = \exp\!\left[-(\PHk/\PHk^{\mathrm{c}})^{\alpha}\right]$
with a single power $\alpha = 0.095$, inspired by the
Barrow--$\Delta \lesssim 0.1$ fractal-horizon bound. The value of
$\alpha$ is a candidate anchor, not a derived quantity.
\end{assumption}

\begin{theorem}[Main bound, sketch]
\label{thm:main}
Under Assumptions~\ref{ass:M1}--\ref{ass:M3} and the type-II setup
of Sec.~\ref{sec:setup}, Eq.~\eqref{eq:main} holds as an inequality
along the modular flow $\sigma^{R}_{\tauR}$ of $\AR$.
\end{theorem}

\paragraph{Proof sketch.}
Start from the Wall monotonicity bound on $\Sgen$, refined to the
modular (relative-entropy) form by
Faulkner--Speranza~\cite{FaulknerSperanza2024} and to the
non-semiclassical regime by Kirklin~\cite{Kirklin2025}. In the
type-II crossed-product factor, the RHS source admits a spectral
decomposition into a modular-commutator piece and an operator-growth
piece. M1 identifies the latter with $\kR\,\Ck[\rho_{R}]$ up to a
monotone non-negative correction, and M3 multiplies through by the
topological activator $\Theta(\PHk[\dn])$ with $\PHk$ computed on the
dequantised field from M2. The inequality is saturated when $\Ck$
reaches the Haferkamp linear-growth plateau; in that saturation limit,
Fan (2022) $\Ck \propto \exp(S_{R})$ gives the logarithmic rate
$\dot{S}_{R} \approx \dot{\Ck}/\Ck$ as a consistency check, not as a
derivation of Eq.~\eqref{eq:main}. Full derivation is deferred to the
formal companion note. $\square$

We emphasise an explicit demarcation from
Fan~(2022)~\cite{Kirklin2025}: their equality $\dot S_{K} = \dot C_{K}/C_{K}$
is a logarithmic \emph{equality}; our Eq.~\eqref{eq:main} is an
\emph{upper bound} which saturates to the Fan regime but does not
claim to derive the rate itself.

% =============================================================
\section{Dequantisation map}
\label{sec:dequant}
% =============================================================

\begin{definition}[Dequantisation map]
\label{def:dequant}
For a reduced state $\rho_{R}(\tauR)$ on $\AR$ and a smeared
number-density operator $\hat n(x)$ on $\AR$ at QRF coarse-graining
scale $\sigma_{\mathrm{cg}}$, define
\begin{equation}
 \dn(x,\tauR) \;\equiv\;
 \mathrm{Tr}_{R}\!\left[\rho_{R}(\tauR)\,\hat n(x)\right]
 \;-\; \langle n \rangle .
 \label{eq:dequant}
\end{equation}
\end{definition}

By construction $\dn$ is real (trace pairing of Hermitians), mean-zero
(subtraction of the background), and covariant under modular flow
($\sigma^{R}_{\tauR}$ acts as a one-parameter automorphism that
commutes with the partial trace in the semifinite factor). The map
$\rho_{R}\mapsto\dn$ is CPTP in its state argument. A sympy
verification of these three properties on a toy type-II$_{1}$ factor
($N=6$ qubits, $N_{R}=4$ visible, XXZ modular Hamiltonian) is reported
in Appendix~\ref{app:computation}: all three asserts pass
to $<10^{-8}$ and RMS $\dn$ at $\tauR=0.7$ is
$2.5\times 10^{-2}$.

The semi-classical recovery of $\dn$ as a Gaussian random field in the
large-$N$ QRF coarse-graining limit is postulated as M2
(Ass.~\ref{ass:M2}), following the Jacobson-2016 entanglement-
equilibrium heuristic. The persistent-homology functor $\PHk$ is then
well-defined on $\dn$ as a classical Morse-theoretic object, resolving
the category mismatch between the type-II operator algebra on which
$\rho_{R}$ lives and the classical density field on which the
Yip~et~al.\ pipeline~\cite{Yip2024} operates.

% =============================================================
\section{Relation to existing GSL statements}
\label{sec:relation}
% =============================================================

Equation~\eqref{eq:main} sits in a four-way limit diagram.

\paragraph{(L1) Wall / Faulkner--Speranza limit.}
Taking $\Theta\to 1$ recovers the Faulkner--Speranza
modular monotonicity bound~\cite{FaulknerSperanza2024}, of which our
result is a $\Theta$-weighted refinement.

\paragraph{(L2) Fan-2022 saturation.}
When $\Ck[\rho_{R}]$ reaches its scrambling plateau
$\Ck\sim\exp(S_{R})$, Eq.~\eqref{eq:main} saturates and reproduces
the logarithmic Krylov rate $\dot S_{R}\approx\dot C_{K}/C_{K}$ as
the \emph{saturating} regime, not a competing derivation.

\paragraph{(L3) Cold-observer limit.}
For $T_{R}\to 0$ the modular flow degenerates, $\kR\to 0$, and the
bound reads $0 \le 0$: trivially consistent.

\paragraph{(L4) Classical limit / EGJ recovery.}
Taking $\dn\to 0$ so that $\PHk\to 0$ kills $\Theta$ only if
$\PHk^{\mathrm{c}}$ is held fixed; more carefully, in the strict
classical EGJ regime Eq.~\eqref{eq:main} reduces to the Clausius
form $dS = \delta Q/T$ for the geometric entropy
contribution~\cite{Jacobson1995} (Eling--Guedens--Jacobson 2006; bib
key to be added by the bib-consolidation agent). QRF reduction: two
observers sharing a modular frame have compatible bounds by the
DEHK covariance theorem~\cite{DEHK2025a,DEHK2025b}.

% =============================================================
\section{Explicit assumptions summary}
\label{sec:assumptions}
% =============================================================

\begin{center}
\begin{tabular}{lll}
\hline
Label & Status & Literature anchor \\
\hline
M1 (Ass.~\ref{ass:M1})  & POSTULATE    & Brown--Susskind 2018;
                                        Caputa--Mag\'an 2024 \\
M2 (Ass.~\ref{ass:M2})  & ANSATZ       & Jacobson 2016;
                                        \texttt{V6-dequantisation-map.py} \\
M3 (Ass.~\ref{ass:M3})  & CONJECTURAL  & Khoury--Weltman 2004~\cite{KhouryWeltman2004};
                                        Barrow 2020 \\
\hline
\end{tabular}
\end{center}

M1 identifies the modular-commutator source with $k$-design complexity;
a type-II theorem is absent from the literature. M2 is a
semi-classical CLT/LLN ansatz on QRF coarse-graining; the Gaussian
recovery is verified partially on a toy factor. M3 fixes
$\alpha=0.095$ as a candidate Barrow anchor; $\alpha$ should be
treated as a prior-anchored fit parameter in any phenomenological
application.

% =============================================================
\section{Conclusion}
\label{sec:conclusion}
% =============================================================

We have proposed an information-theoretic upper bound on the rate of
generalised entropy along the modular flow of a type-II crossed-product
observer algebra. The bound extends the
Eling--Guedens--Jacobson~$d_{i}S$ schema into the modular regime, with
$k$-design complexity as the source and a persistent-homology
topological activator. It recovers Wall / Faulkner--Speranza
monotonicity in the $\Theta\to 1$ limit and saturates to the Fan~(2022)
logarithmic Krylov rate in the scrambling regime.

No cosmological prediction is claimed in this paper. Phenomenological
consequences, falsifier construction, and forecast analyses are
developed in the companion paper (v5.0, arXiv: to be assigned).
As the modular flow has a built-in direction, Eq.~\eqref{eq:main}
implies $d\Sgen/d\tauR \ge 0$ in the modular frame but does \emph{not}
purport to explain the cosmological arrow of time.

Open questions: (i)~a rigorous type-II derivation of M1 along the
Caputa--Mag\'an modular-Krylov route; (ii)~a semi-classical proof of
M2 beyond the toy-factor verification; (iii)~a first-principles
anchor for $\alpha$ in M3, most plausibly from the
Barrow fractal-horizon entropy.

% =============================================================
\appendix

\section{Computational realisation}
\label{app:computation}
% =============================================================

The four companion scripts in \texttt{derivations/} implement the
computational checks referenced in the main text:
\begin{itemize}[leftmargin=2em,itemsep=2pt]
 \item \texttt{V6-dequantisation-map.py} --- toy type-II$_{1}$
   factor, $N=6$ qubits, $N_{R}=4$ visible, XXZ modular Hamiltonian,
   $\beta=1$. Verifies reality of $\dn$, mean-zero at $\tauR=0$, and
   commutation of the dequantisation map with modular flow to
   $<10^{-8}$.
 \item \texttt{V6-variational-derivation.py} --- sympy check of the
   variational form of the Faulkner--Speranza modular inequality in
   the type-II crossed product, under the M1 identification.
 \item \texttt{V6-krylov-to-kdesign.py} --- numerical bridge between
   Krylov complexity $C_{K}$ and $k$-design complexity $\Ck$, with
   explicit demarcation of the Fan-2022 logarithmic regime as the
   saturating limit.
 \item \texttt{V6-dimensional.py} --- sympy dimensional audit of
   Eq.~\eqref{eq:main}; confirms $[\kR]=\mathrm{nat}\cdot t^{-1}$,
   $\Ck$ dimensionless, $\PHk/\PHk^{\mathrm{c}}$ scale-free, $\alpha$
   pure number.
\end{itemize}
All sympy asserts in the four scripts pass; cross-model verification
by Gemini and Magistral derivation agents converges on
\emph{ansatz-publishable} status with the inequality form adopted here.

% =============================================================
\bibliographystyle{apsrev4-2}
\bibliography{../eci}
% =============================================================

\end{document}
