% M1C_vardian_hollands.tex --- standalone obstruction note
% Theorem-attempt: does Vardian (2602.02675) + Hollands (2503.21385) upgrade M1-C
% from heuristic conjecture to theorem-grade?
%
% Author: Kevin Remondiere (drafted by Opus 4.7 sub-agent, 2026-05-02)
% Status: NEGATIVE result, four explicit obstructions O1--O4.
% Compile: pdflatex M1C_vardian_hollands

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{obstruction}[theorem]{Obstruction}
\theoremstyle{remark}
\newtheorem*{verdict}{Verdict}

\newcommand{\Ck}{\mathcal{C}_{k}}
\newcommand{\tauR}{\tau_{R}}
\newcommand{\kR}{\kappa_{R}}
\newcommand{\Sgen}{S_{\mathrm{gen}}}
\newcommand{\AR}{\mathcal{A}_R}
\newcommand{\AA}{\mathcal{A}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\OO}{\mathcal{O}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Smeas}{S_{\mathrm{meas}}}

\title{An attempted theorem-promotion of Conjecture M1-C via\\
Vardian (arXiv:2602.02675) + Hollands (arXiv:2503.21385):\\
four obstructions and a clean negative verdict}
\author{Kevin Remondi\`ere\\
\small (sub-agent draft, 2026-05-02)}
\date{2026-05-02}

\begin{document}
\maketitle

\begin{abstract}
The crypto-Sonnet 2026-Q2 lit-check identified the pairing of Vardian's
modular-Krylov boundary probe of the QES area operator (arXiv:2602.02675,
Feb 2026) with Hollands's $\lambda \le 2\pi$ chaos-bound on Rindler-time
modular flow (arXiv:2503.21385, Mar 2025) as the strongest single published
candidate for upgrading Conjecture M1-C of the ECI v6 framework
(complexity-bounded differential generalised second law on the type-II$_\infty$
crossed product $\AA(D_\OO)_{\mathrm{FRW}} \rtimes_\sigma \R$ of
\texttt{frw\_note.tex} Theorem~3.5) from heuristic conjecture to
theorem-grade status. We verify both papers via direct arXiv API query and
full-PDF reading. We construct the explicit composition chain and find
\emph{four independent obstructions}, two of which (O1, O3) are decisive on
their own. The conclusion is negative: M1-C remains a conjecture; the
Vardian + Hollands pair is the strongest currently published \emph{near-miss}
but does not close the gap.
\end{abstract}

\section{Verbatim source statements}

\subsection{Vardian, arXiv:2602.02675v1, eqs.~(26)--(27)}

Vardian's main construction (a 5-page PRL-style note with no formal
\verb|\begin{theorem}|) is the procedure (eq.~(26), p.~4):
\begin{equation}
O \in \AA \;\longrightarrow\; \{a^O_n, b^O_n\} \;\longrightarrow\;
(\mathcal{L}_S)_{mn} = (\text{Lanczos tridiagonal})
\;\longrightarrow\; P_O\, K\big|_{\AA}\, P_O
\;\longrightarrow\; \widehat{\mathcal{A}}_X / 4 G_N,
\end{equation}
giving the algebraic Ryu--Takayanagi identity (eq.~(27), p.~4):
\begin{equation}
\mathcal{L}_A \;=\; \bigoplus_\alpha S(\chi_\alpha, \mathcal{M}_A)\, P_\alpha
\;=\; K_A \otimes I_{\bar A} \;-\; K\big|_{\AA}.
\end{equation}
The setting (eqs.~(33)--(36), supplemental) is the operator-side modular
Krylov complexity of the modular Liouvillian
$\mathcal{L}_S O := [K_A \otimes I_{\bar A}, O]$.

\subsection{Hollands, arXiv:2503.21385v1, Theorem~4.1, eqs.~(39)--(40)}

\begin{theorem}[Hollands 2503.21385, Thm.~4.1, p.~9, verbatim]
Suppose that for some $a \in \R$, $\partial S(a)$ exists,
$\Phi \in \mathcal{D}(P) \cap \mathcal{D}(\log \Delta_a)$, that
$u_s \Phi \in \mathcal{D}(P)$ for all $s \in \R$, and that
$c \varphi \le \omega \le c^{-1} \varphi$ for some $c > 0$. Then the flowed
state $\varphi_s$ satisfies
\begin{equation}
\Smeas(\varphi_s \,\|\, \varphi_s \circ T \circ R)_{\mathcal{M}(a,b)}
\;\le\; (a-b)\, \partial S(a) \cdot e^{2\pi s}
\;+\; o\!\left( (b-a) e^{2\pi s} \right),
\end{equation}
for $a, b, s \in \R$ such that $(b-a) e^{2\pi s} \to 0+$. Likewise, if
$\partial \bar S(a)$ exists, $\Phi \in \mathcal{D}(P) \cap \mathcal{D}(\log
\Delta'_a)$, $u'_s \Phi \in \mathcal{D}(P)$ for all $s \in \R$, and
$c \varphi' \le \omega' \le c^{-1} \varphi'$ for some $c > 0$, then
\begin{equation}
\Smeas(\varphi'_s \,\|\, \varphi'_s \circ T \circ R)_{\mathcal{M}(b,a)}
\;\le\; (a-b)\, \partial \bar S(a) \cdot e^{-2\pi s}
\;+\; o\!\left( (a-b) e^{-2\pi s} \right),
\end{equation}
for $a, b, s \in \R$ such that $(a-b) e^{-2\pi s} \to 0+$.
\end{theorem}

The Lyapunov interpretation (eq.~(48), p.~10):
$|\psi_s(\delta_\epsilon m)| \le 2 \epsilon\, \partial \bar S(a) e^{2\pi s}
+ o(\epsilon e^{2\pi s}) + \xi$, ``In this sense, we can say that the
Lyapunov exponent is $\le 2\pi$.''

\section{The intended chain (Steps A--C)}

\textbf{Step A (Vardian).} On the FRW comoving diamond $D_\OO$, conformally
pull back to Minkowski $M_\OO$ (\texttt{frw\_note.tex} Prop.~3.4) and apply
Vardian's eq.~(26) on the boundary CFT to extract $\Ck$ and reconstruct
$K|_{\AA}$.

\textbf{Step B (Hollands).} Apply Theorem~4.1 to the modular structure of the
Hislop--Longo Möbius flow on $M_\OO$, obtaining a $2\pi$-bounded modular
Lyapunov exponent.

\textbf{Step C (composition).} Identify the modular-Lyapunov bound with the
M1-C source term $|d\langle H_{\mathrm{mod}}\rangle/d\tauR| \le \kR \cdot \Ck$
and substitute into the Faulkner--Speranza GSL (extended by Kirklin~2024 to
all perturbative orders) to obtain
$d\Sgen[R]/d\tauR \le \kR \Ck \Theta$ on $\AR$.

\section{The four obstructions}

\begin{obstruction}[Algebra-type mismatch, Vardian]\label{obs:O1}
Vardian's reconstruction (eqs.~21, 23, 27) requires the OAQEC tensor
factorisation $\HH = \HH_A \otimes \HH_{\bar A}$, which exists only for
type-I algebras. The FRW comoving-diamond local algebra is type III$_1$
(Hislop--Longo via Theorem~3.5 of \texttt{frw\_note.tex}); on type-III$_1$
there is no density matrix $\rho_A$, hence no Vardian-style
$K_A = -\log \rho_A$ to feed into eq.~(26). On the type-II$_\infty$
crossed product $\AR$ a tracial structure exists, but it is single-factor
and again does not provide an OAQEC tensor decomposition. Vardian
implicitly works in a finite-dimensional code subspace of holographic
CFT and never constructs $\Ck$ on a type-II$_\infty$ algebra.
\end{obstruction}

\begin{obstruction}[Wiesbrock structure missing, Hollands]\label{obs:O2}
Hollands Theorem~4.1 requires a half-sided modular inclusion
$(\mathcal{N} \subset \mathcal{M}, \Omega)$ in Wiesbrock's sense
(Theorem~3.1, p.~6), with positive-spectrum translation generator $P$
arising from the chiral-CFT structure $\mathcal{M} = \AA([0,\infty))$,
$\mathcal{N} = \AA([1,\infty))$. The FRW diamond carries a Hislop--Longo
Möbius flow on a \emph{bounded} causal diamond; the SO$(2,1)$ Möbius
group does not produce the required positive-spectrum translation
generator inside the diamond algebra. Hence the hypothesis of
Theorem~4.1 fails on the FRW diamond as constructed in
\texttt{frw\_note.tex} Theorem~3.5.
\end{obstruction}

\begin{obstruction}[Lyapunov is not entropy-derivative]\label{obs:O3}
Hollands eq.~(48) bounds $|\psi_s(\delta_\epsilon m)|$, an
\emph{observable-expectation flow}. M1-C wants
$d\Sgen / d\tauR$, an \emph{entropy-derivative}. Hollands has the
\emph{form} of a chaos bound but is not a bound on
$d\langle H_{\mathrm{mod}}\rangle/d\tau$ --- the Gap~2 of
\texttt{notes/m1\_audit\_2026\_05\_02.md}. The ant formula (eq.~32) gives
$\partial S(a)$ as the \emph{coefficient} on the right-hand side, not as
the LHS bounded by $\Ck$. A full-text search confirms zero occurrences of
``Krylov'', ``Lanczos'', or ``complexity'' in arXiv:2503.21385.
\end{obstruction}

\begin{obstruction}[$\kR \ne 2\pi$]\label{obs:O4}
Even granting Obstructions~\ref{obs:O1}--\ref{obs:O3}, Hollands's prefactor
$2\pi$ is the kinematic Tomita prefactor (Bisognano--Wichmann inverse
temperature). The v6 prefactor $\kR$ is a Brown--Susskind heuristic
gate-cost normalisation. There is no canonical map between them, because
they live in different conceptual layers: $2\pi$ is fixed by Tomita
modular dynamics; $\kR$ is fixed by the choice of complexity metric
(Krylov vs.\ Nielsen vs.\ spread) and gate-cost convention.
\end{obstruction}

\section{Verdict}

\begin{verdict}
\textbf{M1-C theorem-grade status: FAILS.}
The Vardian + Hollands pairing does not upgrade Conjecture M1-C to a
theorem on $\AR$. Obstructions~\ref{obs:O1} and \ref{obs:O3} are each
individually decisive; \ref{obs:O2} and \ref{obs:O4} are independent
additional failures. The pairing is the strongest currently published
near-miss and remains a useful programmatic citation for v8-bis, but it
re-localises rather than closes the audit's Gap~1 (transplant) and
Gap~2 (modular-energy-derivative lemma).
\end{verdict}

\section{Recommendation}

Reaffirm M1-C as a conjecture in v8-bis. Cite Vardian as a programmatic
reference (boundary modular-Krylov as a probe of geometric quantities,
with explicit type-I OAQEC limitation). Cite Hollands 2503.21385
(distinct from Hollands--Longo 2503.04651 already in the bib) as a
modular chaos-bound on chiral CFTs, with explicit Wiesbrock-hypothesis
limitation. Add an appendix to \texttt{notes/m1\_audit} listing the
Vardian + Hollands chain with O1--O4 enumerated, to forestall future
reviewers asking why this candidate was not tried.

A genuinely structural upgrade of M1-C would require either:
(i) extending OAQEC-style modular Krylov complexity to type-II$_\infty$
crossed products (no published construction known), \emph{or}
(ii) proving a Wiesbrock-like half-sided modular structure for FRW
comoving diamonds (no published construction known), \emph{plus}
(iii) bridging modular-Lyapunov-on-observables to
modular-energy-derivative-on-entropy in a type-II$_\infty$ algebra (no
published bridge known). Each of these is itself a substantial open
problem in mathematical physics.

\section{Triangulation log}

Both papers verified via direct arXiv API query 2026-05-03 00:21 UTC:
\begin{itemize}
\item arXiv:2602.02675v1, Vardian, ``Modular Krylov Complexity as a Boundary
Probe of Area Operator and Entanglement Islands'', published
2026-02-02T19:00:24Z, primary hep-th, ``5 pages + supplemental material
+ appendix''. PDF (412 KB) downloaded and read in full.
\item arXiv:2503.21385v1, Hollands, ``Modular Time Evolution and the QNEC'',
published 2025-03-27T11:33:35Z, primary hep-th (also math-ph, math.OA,
quant-ph), ``18 pages, no figures''. PDF (248 KB) downloaded and read in
full.
\end{itemize}
Note: \texttt{paper/eci.bib} entry \texttt{HollandsLongo2025QNEC} refers to
arXiv:2503.04651 (joint Hollands--Longo, ``A New Proof of the QNEC''),
\emph{not} the target paper 2503.21385 (Hollands solo, ``Modular Time
Evolution and the QNEC''). They are distinct papers posted 23 days apart.

\end{document}
