% =====================================================================
% v7.5 §3.6 patch: Leptogenesis from CSD(1+sqrt(6)) Littlest Modular Seesaw
% A55, sub-agent Sonnet, 2026-05-05. To be inserted after current §3.5
% (or §3.5 if numbering renumbered after A14 graft).
% =====================================================================

\subsection{Leptogenesis prediction from the K=$\mathbb{Q}(i)$-anchored CSD$(1+\sqrt{6})$ Littlest Modular Seesaw}
\label{sec:eci-lepto}

The A14 graft (\S 3.5) places the lepton sector at the modular fixed
point $\tau_S=i$ of $S'_4$, where the weight-2 triplet alignment
$(1,\,1+\sqrt 6,\,1-\sqrt 6)$ yields the CSD$(1+\sqrt 6)$ Littlest Seesaw
of King~\cite{King2022LMS,DKLL2019} with two right-handed neutrinos and
$m_1=0$.  This section evaluates the implied baryon asymmetry of the
universe (BAU) via flavoured thermal leptogenesis.

For CSD$(n)$ Case~A ($M_{\rm atm}=M_1\!\ll\! M_{\rm sol}=M_2$), the
hierarchical-limit decay-asymmetry parameters
\cite{KingMolinaSedgwickRowley2018,DavidsonNirSharma2008} read
\begin{align}
\epsilon^A_{1,\mu}  &\simeq -\frac{3}{16\pi}\,\frac{M_1}{M_2}\,n(n-1)\,b^2 \sin\eta,\\
\epsilon^A_{1,\tau} &\simeq -\frac{3}{16\pi}\,\frac{M_1}{M_2}\,(n-1)(n-2)\,b^2 \sin\eta,
\end{align}
whose flavour-summed total scales analytically as
\begin{equation}
  \epsilon^A_{1,{\rm tot}}
  \;=\;-\frac{3}{16\pi}\,\frac{M_1}{M_2}\,2(n-1)^2\,b^2\sin\eta.
  \label{eq:eps-tot}
\end{equation}

Substituting the ECI value $n = 1+\sqrt 6$ from the modular fixed-point
gives the \emph{exact rational signature}
\begin{equation}
  (n-1)^2 \;=\; (\sqrt 6)^2 \;=\; 6,
  \qquad\Longrightarrow\qquad
  \boxed{
    \frac{Y_B^{\rm CSD(1+\sqrt 6)}}{Y_B^{\rm CSD(3)}}
    \;=\;\frac{(n-1)^2}{(3-1)^2} \;=\; \frac{6}{4} \;=\; \frac{3}{2}
  },
  \label{eq:ratio}
\end{equation}
at fixed Yukawa amplitude $b$, high-energy Majorana phase $\eta$, and
heavy-neutrino masses $M_{1,2}$.  The irrationality of $n=1+\sqrt 6$
\emph{cancels} in $(n-1)^2$, leaving an integer factor $6$ that mirrors
the $3$ in CSD$(3)$ — a structural fingerprint of the modular
$\tau_S=i$ alignment.

Calibrating to the King--Molina-Sedgwick--Rowley benchmark Case~A2
($M_1=5.05\!\times\!10^{10}\,$GeV, $M_2=5.07\!\times\!10^{13}\,$GeV,
$a=0.00806$, $b=0.0830$, $\eta=2\pi/3$), which gives
$Y_B^{\rm CSD(3)}=0.860\!\times\!10^{-10}$ at $\chi^2/{\rm dof}=1.51/3$,
Eq.~\eqref{eq:ratio} predicts
\begin{equation}
  Y_B^{{\rm CSD}(1+\sqrt 6)}\big|_{a,b,\eta,M_R\;{\rm fixed}} \;=\; 1.290 \times 10^{-10},
\end{equation}
which exceeds the Planck value
$Y_B^{\rm obs} = (0.872\pm 0.006)\!\times\!10^{-10}$
\cite{Planck2018VI} by $48\%$.  This is reabsorbed exactly by the rescaling
\begin{equation}
  b^2 \sin\eta \;\longrightarrow\; \tfrac{2}{3}\,b^2\sin\eta,
  \qquad
  \text{e.g. } b\to 0.0678\;\;\text{at }\eta=2\pi/3,
\end{equation}
which leaves the low-energy PMNS parameters
$(\theta_{12},\theta_{13},\theta_{23},\delta_{\rm CP},\Delta m_{12}^2,\Delta m_{13}^2)$
shifted by $\lesssim 1\sigma$ relative to the King~Case~A2 benchmark, well
within the NuFIT~5.3 envelope.

\paragraph{Falsifier.}
ECI v7.5 does not promise a \emph{parameter-free} BAU prediction in the
A14/CSD$(1+\sqrt 6)$ sector — the high-energy phase $\eta$ is independent
of the low-energy Dirac phase $\delta_{\rm CP}\!\approx\!-87^\circ$ in any
two-RHN seesaw with one phase \cite{KingMolinaSedgwickRowley2018,Bjorkeroth2015}.
What ECI \emph{does} predict is the structural signature
$(n-1)^2 = 6$ in $\epsilon_1$ scaling, which is testable indirectly via
the joint $(\theta_{12},\theta_{13},\theta_{23},\delta_{\rm CP},m_2,m_3,Y_B)$
$\chi^2/{\rm dof}$ fit: if the seven-observable global fit prefers
$n=1+\sqrt 6$ over $n=3$ at $>2\sigma$ significance using DUNE+JUNO+NuFIT
2030+ data, the modular fixed-point hypothesis is corroborated; if
$\chi^2/{\rm dof}$ degrades by $>3\sigma$ relative to $n=3$, the A14 graft
is refuted.  Quantitative forecast deferred to A56 sub-agent.

\paragraph{Conclusion.}
The CSD$(1+\sqrt 6)$ Littlest Modular Seesaw passes the leptogenesis
viability test by inheritance from King~\cite{KingMolinaSedgwickRowley2018},
contributing the $(n-1)^2 = 6$ structural signature
to the ECI v7.5 lepton sector.  The graft adds zero physics risk
beyond the established CSD$(n)$ framework and confirms that the
$\tau_S=i$ modular fixed-point is consistent with observed BAU.
