% note.tex --- companion math.OA note to frw_note.tex
% The FRW past-light-cone type II_infty factor is hyperfinite (= R_{0,1})
%
% Author: Kevin Remondiere
% Date: 2026-05-03
% Compile: pdflatex note ; bibtex note ; pdflatex note ; pdflatex note

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{cite}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\HH}{\mathcal{H}}
\renewcommand{\AA}{\mathcal{A}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\RR}{\mathcal{R}}
\newcommand{\OO}{\mathcal{O}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\AFD}{\mathrm{AFD}}
\newcommand{\FRW}{\mathrm{FRW}}
\newcommand{\Mink}{\mathrm{M}}
\newcommand{\NN}{\mathcal{N}}

\title{The FRW past-light-cone type II$_\infty$ factor is hyperfinite\thanks{Companion to \texttt{paper/frw\_typeII\_note/frw\_note.tex} (commit \texttt{1b2d055} of \url{https://github.com/AIdevsmartdata/crossed-cosmos}). Pure expository / structural note: every step is folklore-known to the AQFT community; the contribution is to assemble them in one place for the FRW case and to record the corollary.}}
\author{K\'evin Remondi\`ere\\
\small Independent researcher\\
\small \href{mailto:kevin.remondiere@gmail.com}{kevin.remondiere@gmail.com}\\
\small ORCID: \href{https://orcid.org/0009-0008-2443-7166}{0009-0008-2443-7166}}
\date{2026-05-03}

\begin{document}
\maketitle

\begin{abstract}
The companion note \cite{FrwNote2026} constructs, for a comoving causal diamond
$D \subset $ FRW with conformally-coupled massless scalar matter in the
Bunch--Davies vacuum, a type II$_\infty$ factor
$\MM_\FRW(D) := \AA_\FRW(D) \rtimes_\sigma \R$ as the modular crossed product
of the local Hadamard von Neumann algebra. Here we observe that
$\MM_\FRW(D)$ is approximately finite-dimensional (AFD), hence isomorphic to
the unique hyperfinite type II$_\infty$ factor $\RR_{0,1}$ of Connes--Takesaki.
The argument is a routine concatenation of (i) Buchholz--D'Antoni--Fredenhagen
\cite{BDF1987}, (ii) Verch's local quasi-equivalence of Hadamard quasifree
states \cite{Verch1994}, (iii) Fewster's curved-spacetime split property
\cite{Fewster2015Split}, (iv) the Connes--Takesaki structure theorem
\cite{ConnesTakesaki1977}, and (v) Connes' uniqueness of the AFD II$_\infty$
factor \cite{Connes1976}. The corollary is that no two cosmological observers
in the BFV folium of FRW are distinguishable by the W$^*$-type of their
gravitationally-dressed crossed-product algebra alone: every such observer sees
$\RR_{0,1}$.
\end{abstract}

\section{Introduction and main theorem}\label{sec:intro}

The crossed-product construction of Witten \cite{Witten2022} and
Chandrasekaran--Longo--Penington--Witten (CLPW) \cite{CLPW2023} produces, from
a type III$_1$ local algebra and its modular automorphism group, a type II
factor of \emph{gravitationally dressed} observables. The companion note
\cite{FrwNote2026} extended the construction to a comoving causal diamond
in radiation-dominated FRW (and to dust and de Sitter cases) using
conformal pullback to Minkowski, obtaining a type II$_\infty$ factor
without invoking a Killing vector field. The output of \cite{FrwNote2026}
Theorems 3.5 and 3.6 is, schematically,
\[
\AA_\FRW(D)\;\xrightarrow[\text{Tomita--Takesaki}]{\;\sigma\;}\;
\AA_\FRW(D)\rtimes_\sigma \R \;=:\; \MM_\FRW(D),
\qquad \text{type II}_\infty.
\]
That note left open the precise W$^*$-isomorphism class of $\MM_\FRW(D)$.
The present note answers it.

\begin{theorem}[Main]\label{thm:main}
Let $\AA_\FRW(D)$ be the local von Neumann algebra of the conformally coupled
massless scalar in radiation-dominated, dust-dominated, or de Sitter FRW, in
the Bunch--Davies / conformal vacuum, associated to a doubly-bounded
comoving causal diamond $D$ with $\eta_i, \eta_f$ bounded away from the
singular conformal boundary, in the conventions of \cite{FrwNote2026}. Then
the type II$_\infty$ factor
$\MM_\FRW(D) := \AA_\FRW(D)\rtimes_\sigma \R$
of \cite[Thm.~3.5/3.6]{FrwNote2026} is approximately finite-dimensional;
\emph{a fortiori} it is isomorphic, as an abstract W$^*$-algebra equipped
with its (unique up to scaling) semifinite normal trace, to the unique
hyperfinite type II$_\infty$ factor
\[
\MM_\FRW(D)\;\cong\;\RR_{0,1}\;=\;\RR\,\overline{\otimes}\,B(\HH),
\]
where $\RR$ is the unique hyperfinite II$_1$ factor.
\end{theorem}

The proof factors through three lemmas, none of them new.

\section{Type III$_1$ structure of the local algebra}\label{sec:typeIII1}

We adopt the framework of locally covariant quantum field theory of
Brunetti--Fredenhagen--Verch \cite{BFV2003}: the conformally coupled massless
scalar on FRW is realised as a covariant functor from the category of
globally hyperbolic spacetimes to the category of $C^*$-algebras, and the
Bunch--Davies state on the conformally flat FRW pulls back to the Minkowski
vacuum on the conformally identified manifold under the rescaling
$\tilde\varphi = a(\eta)\varphi$ \cite[\S2.1, \S3]{FrwNote2026}.

\begin{lemma}[type III$_1$]\label{lem:typeIII}
$\AA_\FRW(D)$ is a factor of type III$_1$.
\end{lemma}

\begin{proof}
By \cite[Prop.~3.2 + Thm.~3.5]{FrwNote2026}, there is a unitary
$U:\HH_\FRW \to \HH_\Mink$ with $U\Omega_\FRW = \Omega_\Mink$ and
$U\AA_\FRW(D)U^* = \AA_\Mink(M_D)$, where $M_D$ is the diamond $D$
viewed as a subset of Minkowski via the conformal identification. The
Hislop--Longo theorem \cite{HislopLongo1982} states that $\AA_\Mink(M_D)$
is a type III$_1$ factor for the free massless scalar in the vacuum.
Type and factoriality are unitary invariants, hence $\AA_\FRW(D)$ is a
type III$_1$ factor. \qed
\end{proof}

\begin{remark}\label{rmk:pure-AQFT}
Lemma~\ref{lem:typeIII} can equivalently be obtained without invoking the
conformal pullback: by Verch \cite{Verch1994}, the GNS representations of all
quasifree Hadamard states on a globally hyperbolic spacetime are locally
quasi-equivalent, so the type of $\AA_\FRW(D)$ is independent of the choice
of Hadamard state in its folium. Pulling back the Minkowski vacuum gives a
Hadamard state on FRW \cite[Eq.~(2.3)]{FrwNote2026}, and the type follows
from the Hislop--Longo result. The two paths give the same answer.
\end{remark}

\section{Hyperfiniteness}\label{sec:hyperfinite}

\begin{lemma}[hyperfiniteness]\label{lem:AFD}
$\AA_\FRW(D)$ is approximately finite-dimensional.
\end{lemma}

\begin{proof}
Approximately finite-dimensional (AFD) is, for von Neumann algebras with
separable predual, equivalent to injectivity (Connes \cite{Connes1976}).
Injectivity is invariant under W$^*$-isomorphism, so by the unitary
equivalence $\AA_\FRW(D)\cong\AA_\Mink(M_D)$ from
\cite[Prop.~3.2]{FrwNote2026} it suffices to prove
$\AA_\Mink(M_D)$ is injective. By Buchholz--D'Antoni--Fredenhagen
\cite{BDF1987}, the local algebras of any Wightman quantum field theory
satisfying the standard nuclearity / split assumptions are isomorphic to
$\RR\,\overline{\otimes}\,Z$ for $\RR$ the hyperfinite type III$_1$ factor
and $Z$ the centre of the local algebra; in the present (free, factor)
case $Z=\C$, hence $\AA_\Mink(M_D)\cong\RR$, in particular injective and AFD.
The free massless scalar on Minkowski satisfies the nuclearity / split
hypotheses of \cite{BDF1987}; for the curved-spacetime extension, the
nuclearity / split property is established for free Dirac fields by
D'Antoni--Hollands \cite{DAntoniHollands2006} and for general locally
covariant theories under nuclearity by Fewster \cite{Fewster2015Split};
the scaling-algebra renormalization-group framework of Buchholz--Verch
\cite{BuchholzVerch1995} provides the broader machinery in which these
nuclearity hypotheses sit. A pedagogical overview of why all of this
works in QFT in curved spacetime is given in
\cite{Witten2022WhyQFT}.
\qed
\end{proof}

\begin{remark}\label{rmk:curved-direct}
A curved-spacetime-native proof of Lemma~\ref{lem:AFD}, without invoking
the conformal pullback, runs as follows. Verch's symplectic adjoint
continuity theorem \cite{Verch1997} establishes local primarity and
Haag duality in any quasifree Hadamard representation; Fewster's split
property \cite{Fewster2015Split} provides the funnel of type-I
intermediate factors needed for nuclearity-style arguments; combined
with Verch \cite{Verch1994} this gives, on any globally hyperbolic
spacetime where nuclearity holds, the AFD III$_1$ structure of local
algebras. For our FRW case the conformal pullback is the
shortest path, but the curved-spacetime route is in principle
self-contained.
\end{remark}

\begin{remark}\label{rmk:no-MIP}
The MIP$^*$=RE refutation of Connes' embedding conjecture
\cite{JNVWY2020} produces II$_1$ factors that do \emph{not} embed into
the ultrapower $\RR^\omega$, and \emph{a fortiori} are not AFD. The
algebra $\AA_\FRW(D)$ is well outside that pathology: it is a free-field
local algebra in a Hadamard representation, the canonical AFD-friendly
setting. The MIP$^*$=RE counterexamples involve highly non-perturbative
non-amenable group operator algebras and have no known realisation as
local algebras of a quantum field on a globally hyperbolic spacetime.
The present theorem is therefore a positive result of the form
``physically natural local algebras of free QFT on FRW remain firmly on
the AFD side of the Connes-embedding dichotomy''.
\end{remark}

\section{Crossed product is the unique AFD II$_\infty$ factor}\label{sec:CT}

\begin{lemma}[Connes--Takesaki + Connes uniqueness]\label{lem:CT}
Let $\NN$ be an AFD type III$_1$ factor with separable predual and let
$\sigma$ be a continuous one-parameter automorphism group whose
restriction to a non-trivial centraliser yields the modular flow of a
faithful normal state $\omega$. Then the crossed product
$\NN\rtimes_\sigma\R$ is an AFD type II$_\infty$ factor with full Connes
spectrum $\Gamma(\sigma)=\R$, and is therefore isomorphic to
$\RR_{0,1}=\RR\,\overline{\otimes}\,B(\HH)$, the unique AFD II$_\infty$
factor.
\end{lemma}

\begin{proof}
This is two classical results glued together. The
Connes--Takesaki structure theorem \cite{ConnesTakesaki1977} (and its
modern textbook statement, e.g.\ Takesaki II.X) states that for a
type III$_\lambda$ factor with $\lambda\in[0,1]$, the modular crossed
product $\NN\rtimes_\sigma\R$ is type II$_\infty$, semifinite, with a
canonical dual action of $\R$ scaling the trace. AFD-ness is preserved
under modular crossed products: if $\NN$ is AFD then
$\NN\rtimes_\sigma\R$ is AFD (Connes \cite[Thm.~3]{Connes1976}; see
Takesaki, \emph{Theory of Operator Algebras III}, Ch.~XII). Connes'
1976 uniqueness theorem \cite{Connes1976} shows that all AFD
II$_\infty$ factors with separable predual are isomorphic; the unique
one is denoted $\RR_{0,1}$. Combining: $\NN\rtimes_\sigma\R$ is AFD,
type II$_\infty$, separable predual, hence isomorphic to $\RR_{0,1}$.
The full Connes spectrum $\Gamma(\sigma)=\R$ is automatic for type
III$_1$ \cite{ConnesTakesaki1977}.
\qed
\end{proof}

\section{Proof of the main theorem}\label{sec:proof}

\begin{proof}[Proof of Theorem~\ref{thm:main}]
Lemma~\ref{lem:typeIII} gives type III$_1$.
Lemma~\ref{lem:AFD} gives AFD.
The modular flow $\sigma$ on $\AA_\FRW(D)$ is by definition the
Tomita--Takesaki flow of the conformal vacuum, intertwined by
$U$ to the Hislop--Longo flow on $\AA_\Mink(M_D)$
\cite[\S3.3]{FrwNote2026}; both are continuous, and the latter has full
Connes spectrum \cite{HislopLongo1982}, so by unitary intertwining the
former does too. The hypotheses of Lemma~\ref{lem:CT} are met with
$\NN=\AA_\FRW(D)$ and $\omega=$ the conformal vacuum. Lemma~\ref{lem:CT}
yields
$\AA_\FRW(D)\rtimes_\sigma\R \cong \RR_{0,1}$.
The left-hand side is, by \cite[Thm.~3.5]{FrwNote2026}, the algebra
$\MM_\FRW(D)$ of gravitationally dressed observables. \qed
\end{proof}

\begin{remark}[Separability]\label{rmk:sep}
The Bunch--Davies GNS Hilbert space $\HH_\FRW$ is separable: the one-particle
Hilbert space is $L^2$ of a second-countable Cauchy slice with respect to a
smooth weight, and the Fock space over a separable Hilbert space is
separable. So $\AA_\FRW(D)\subset B(\HH_\FRW)$ has separable predual, and
Connes' \cite{Connes1976} uniqueness applies without subtlety.
\end{remark}

\section{Corollaries}\label{sec:corollaries}

\begin{corollary}[Observer-indistinguishability by W$^*$-type alone]\label{cor:obs-indis}
Let $D, D' \subset $ FRW be two doubly-bounded comoving causal diamonds in
the conformal-vacuum Hadamard folium, in the regime of
Theorem~\ref{thm:main}. Then $\MM_\FRW(D)\cong\MM_\FRW(D')$ as W$^*$-algebras
(both $\cong\RR_{0,1}$), even though the underlying spacetime regions and
the observers' worldlines are physically distinct.
\end{corollary}

In particular, the observer-dependence in the CLPW programme cannot be
detected from the W$^*$-isomorphism class of the gravitationally dressed
algebra: it lives only in the additional structure (state, trace
normalisation, modular Hamiltonian, etc.) attached to that algebra. This
sharpens the discussion of \cite[\S6]{CLPW2023} and \cite{FaulknerSperanza2024}.

\begin{corollary}[Bianchi VI$_{-1/9}$ classification hope is closed]\label{cor:bianchi}
Conjectures in the ECI programme (R3, R8 in the project diary) that
sought to distinguish anisotropic Bianchi VI$_{-1/9}$ cosmologies from
isotropic FRW ones via the W$^*$-type of the cosmological observer's
crossed-product algebra are foreclosed in their published form: provided
the matter sector is conformally coupled and the diamond avoids the
conformal boundary, both spacetimes yield $\RR_{0,1}$.
\end{corollary}

\begin{remark}\label{rmk:negative-implication}
Corollary~\ref{cor:bianchi} is not a refutation of the wider ECI programme,
but a precise statement of what type-of-factor information \emph{alone}
cannot resolve. Distinguishing FRW from Bianchi VI$_{-1/9}$ cosmologies
requires invariants finer than the W$^*$-type --- e.g.\ trace normalisation,
modular spectrum, or expectation values of relational observables ---
none of which are foreclosed here.
\end{remark}

\section{Open questions}\label{sec:open}

\begin{enumerate}
\item \emph{Massive / non-conformally-coupled scalars.} The conformal
pullback of \cite{FrwNote2026} fails outside $\xi=1/6$, $m=0$, $d=4$.
Whether the local algebras of a massive Klein--Gordon field on FRW in a
generic Hadamard state remain AFD is not settled by the present argument.
The natural conjecture is yes, via the curved-spacetime path of
Remark~\ref{rmk:curved-direct}, contingent on nuclearity holding for
massive Hadamard FRW representations.

\item \emph{Interacting fields.} The BDF \cite{BDF1987} hyperfiniteness
result is conditional on nuclearity / split. For perturbatively
constructed interacting models (Hollands--Wald \cite{HollandsWald2001})
this is folklore but not theorem-proved on curved backgrounds.

\item \emph{Wedge-type regions and asymptotic conformal boundaries.}
The diamond restriction $\eta_i, \eta_f$ bounded away from the singularity
is essential to the conformal pullback. Conformal-boundary algebras and
past-light-cone wedges of unbounded conformal extent are excluded; they
may have different W$^*$-type behaviour.

\item \emph{Refining the indistinguishability statement.} What invariants
\emph{do} distinguish $\MM_\FRW(D)$ from $\MM_\FRW(D')$ when both equal
$\RR_{0,1}$? Candidates: the relative entropy structure on a fixed pair of
states, the trace normalisation as a function of the observer's clock,
the modular Hamiltonian spectrum on a finite-energy subspace.
\end{enumerate}

\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{FrwNote2026}
K.\ Remondi\`ere,
\emph{Type II$_\infty$ for the gravitational algebra of a comoving observer
in radiation-dominated FRW via conformal pullback},
companion note, repository \url{https://github.com/AIdevsmartdata/crossed-cosmos},
commit \texttt{1b2d055} (2026-05-02).

\bibitem{Witten2022}
E.\ Witten,
\emph{Gravity and the crossed product},
JHEP \textbf{10} (2022) 008,
arXiv:2112.12828 [hep-th].

\bibitem{CLPW2023}
V.\ Chandrasekaran, R.\ Longo, G.\ Penington, E.\ Witten,
\emph{An algebra of observables for de Sitter space},
JHEP \textbf{02} (2023) 082,
arXiv:2206.10780 [hep-th].

\bibitem{FaulknerSperanza2024}
T.\ Faulkner, A.\ J.\ Speranza,
\emph{Gravitational algebras and the generalized second law},
JHEP \textbf{11} (2024) 099,
arXiv:2405.00847 [hep-th].

\bibitem{HislopLongo1982}
P.\ D.\ Hislop, R.\ Longo,
\emph{Modular structure of the local algebras associated with the free
massless scalar field theory},
Comm.\ Math.\ Phys.\ \textbf{84} (1982) 71--85.

\bibitem{BDF1987}
D.\ Buchholz, C.\ D'Antoni, K.\ Fredenhagen,
\emph{The universal structure of local algebras},
Comm.\ Math.\ Phys.\ \textbf{111} (1987) 123--135.
DOI: 10.1007/BF01239019.

\bibitem{Verch1994}
R.\ Verch,
\emph{Local definiteness, primarity and quasiequivalence of quasifree
Hadamard quantum states in curved spacetime},
Comm.\ Math.\ Phys.\ \textbf{160} (1994) 507--536.
DOI: 10.1007/BF02173427.

\bibitem{Verch1997}
R.\ Verch,
\emph{Continuity of symplectically adjoint maps and the algebraic structure
of Hadamard vacuum representations for quantum fields on curved spacetime},
Rev.\ Math.\ Phys.\ \textbf{9} (1997) 635--674,
arXiv:funct-an/9609004.

\bibitem{BuchholzVerch1995}
D.\ Buchholz, R.\ Verch,
\emph{Scaling algebras and renormalization group in algebraic quantum field
theory},
Rev.\ Math.\ Phys.\ \textbf{7} (1995) 1195--1240,
arXiv:hep-th/9501063.

\bibitem{DAntoniHollands2006}
C.\ D'Antoni, S.\ Hollands,
\emph{Nuclearity, local quasi-equivalence and split property for Dirac
quantum fields in curved spacetime},
Comm.\ Math.\ Phys.\ \textbf{261} (2006) 133--159,
arXiv:math-ph/0106028.

\bibitem{BFV2003}
R.\ Brunetti, K.\ Fredenhagen, R.\ Verch,
\emph{The generally covariant locality principle: a new paradigm for local
quantum physics},
Comm.\ Math.\ Phys.\ \textbf{237} (2003) 31--68,
arXiv:math-ph/0112041.

\bibitem{HollandsWald2001}
S.\ Hollands, R.\ M.\ Wald,
\emph{Local Wick polynomials and time ordered products of quantum fields in
curved spacetime},
Comm.\ Math.\ Phys.\ \textbf{223} (2001) 289--326,
arXiv:gr-qc/0103074.

\bibitem{Fewster2015Split}
C.\ J.\ Fewster,
\emph{The split property for locally covariant quantum field theories in
curved spacetime},
Lett.\ Math.\ Phys.\ \textbf{105} (2015) 1633--1661,
arXiv:1501.02682.

\bibitem{ConnesTakesaki1977}
A.\ Connes, M.\ Takesaki,
\emph{The flow of weights on factors of type III},
T\^ohoku Math.\ J.\ \textbf{29} (1977) 473--575.
DOI: 10.2748/tmj/1178240493.

\bibitem{Connes1976}
A.\ Connes,
\emph{Classification of injective factors. Cases II$_1$, II$_\infty$,
III$_\lambda$, $\lambda\neq 1$},
Ann.\ of Math.\ (2) \textbf{104} (1976) 73--115.

\bibitem{Witten2022WhyQFT}
E.\ Witten,
\emph{Why does quantum field theory in curved spacetime make sense? And
what happens to the algebra of observables in the thermodynamic limit?},
arXiv:2112.11614 [hep-th] (2021).

\bibitem{JNVWY2020}
Z.\ Ji, A.\ Natarajan, T.\ Vidick, J.\ Wright, H.\ Yuen,
\emph{MIP$^*$=RE},
arXiv:2001.04383 [quant-ph] (2020).

\end{thebibliography}

\end{document}
