% =====================================================================
% A Cosmological Hayden--Preskill Protocol from the FRW Type II_infty
% Crossed-Product Algebra
%
% Author: Kevin Remondiere (synthesis with Opus 4.7 sub-agent)
% Date: 2026-05-02
%
% Standalone math.OA + quant-ph note. All algebraic claims sympy-verified
% in verify.py; all citations arXiv-API verified 2026-05-02.
% =====================================================================
\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{capt-of}
\usepackage[dvipsnames]{xcolor}

\hypersetup{colorlinks=true, linkcolor=blue!50!black, citecolor=blue!50!black,
            urlcolor=blue!50!black}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\AA}{\mathcal{A}}
\newcommand{\OO}{\mathcal{O}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\Ad}{\operatorname{Ad}}
\newcommand{\FRW}{\mathrm{FRW}}
\newcommand{\Mink}{\mathrm{M}}
\newcommand{\HL}{\mathrm{HL}}
\newcommand{\PLC}{\mathrm{PLC}}

\title{A Cosmological Hayden--Preskill Protocol from the FRW
       Type II$_\infty$ Crossed-Product Algebra}
\author{Kevin Remondi\`ere\thanks{Independent.
        \texttt{kevin.remondiere@gmail.com}.
        Algebraic verification by Opus 4.7 sub-agent;
        all sympy and arXiv-API checks reproduced in
        \texttt{verify.py} and the bibliography metadata.}}
\date{2 May 2026}

\begin{document}
\maketitle

\begin{abstract}
We combine three recent results --- (i) the type II$_\infty$ classification
of the comoving-diamond algebra of the conformally coupled massless scalar
in radiation-dominated FRW \cite{frwnote} (commit \texttt{1b2d055}),
(ii) the Krylov--Diameter correspondence
$(C_k^{-1}\,dC_k/dt = R_{\mathrm{proper}}^{-1})$ on its type II$_\infty$
crossed product \cite{piste1note},
and (iii) the Hayden--Preskill scrambling formula \cite{HaydenPreskill07}
--- to formulate a \emph{cosmological Hayden--Preskill protocol}: a
mathematical analog in which Hubble expansion plays the role of black-hole
scrambling. The main statement is:
in the past-light-cone (PLC) diamond convention, after a proper time
$\Delta t \ge H(t_0)^{-1}\,\log_2(d_{\mathrm{Hubble}}/k)$, an observer
collecting cosmologically expanded radiation can recover an arbitrary
$\log_2 k$-qubit message deposited at $t_0$ with reconstruction fidelity
$F\ge 1-\OO(k^2/d_{\mathrm{Hubble}}^2)$, conditional on (a) a rigorous
extension of modular Krylov complexity to type II$_\infty$ crossed products
and (b) a universal modular Lyapunov exponent $\lambda_L^{\mathrm{mod}}=2\pi$
on the FRW vacuum. We give a formal protocol description, a proof sketch,
explicit comparisons to the original BH scrambling protocol, the
Yoshida--Kitaev decoder, and recent strong-pseudorandom-unitary results
\cite{ChenBouland24,Schuster25}, and an honest assessment of what the
result is and is not. The primary intended contribution is conceptual: a
cosmological setting in which the Hayden--Preskill bound emerges with no
black hole present.
\end{abstract}

\tableofcontents

% =====================================================================
\section{Introduction and statement of the result}
% =====================================================================

The Hayden--Preskill scrambling protocol \cite{HaydenPreskill07} is one of
the cleanest information-theoretic statements about black holes: once a
black hole has emitted half of its initial entropy as Hawking radiation,
any further qubits thrown in re-emerge in the radiation after the
\emph{scrambling time} $t_*=(\beta/2\pi)\log S$. Two ingredients are
essential: (i) approximate Haar-random internal dynamics --- recently
sharpened to \emph{strong} pseudorandom unitaries with logarithmic depth
\cite{Schuster25,ChenBouland24} --- and (ii) the Bisognano--Wichmann
identification $\beta_{\mathrm{mod}}=2\pi$ of modular flow with thermal
flow.

In a different direction, the von Neumann algebra of observables of a
gravitational subregion was recently shown to upgrade from type
III$_1$ to type II by tensoring with a clock and crossing with the modular
group \cite{Witten22CrossedProduct,CLPW23,ChenPenington24}. For a comoving
observer in radiation-dominated FRW the resulting algebra is type II$_\infty$
\cite{frwnote}, with a canonical semifinite trace inherited from the
Connes--Takesaki dual weight \cite{ConnesTakesaki77}. On this algebra the
modular Krylov complexity \cite{CMPT24,Vardian26,AguilarGutierrez25} of the
vacuum seed satisfies the geometric identity \cite{piste1note}
\begin{equation}\label{eq:krylov-diameter}
  \frac{1}{C_k(t)}\,\frac{dC_k(t)}{dt}
   \;=\; \frac{1}{R_{\mathrm{proper}}(t)} \;+\; \OO\!\big(C_k^{-1}\big),
\end{equation}
which we call the \emph{Krylov--Diameter correspondence}. In the
past-light-cone diamond convention $R_{\mathrm{conf}}=\eta_{\mathrm{obs}}$
with $a(\eta)=a_0\eta$, one has $R_{\mathrm{proper}}=a_0\eta^2=2t$, hence
\begin{equation}\label{eq:krylov-hubble}
  \boxed{\;
    \frac{1}{C_k(t)}\,\frac{dC_k(t)}{dt}
     \;=\; H(t) \;=\; \frac{1}{2t}\quad
    \text{(radiation FRW, PLC convention)}
  \;}
\end{equation}
\emph{exactly} at leading order in $C_k^{-1}$. This is the cosmological
analog of the modular Lyapunov saturation $\lambda_L^{\mathrm{mod}}=2\pi$
of \cite{CMPT24}, with the Hubble rate playing the role of the
modular-thermal frequency.

Our main claim is that \eqref{eq:krylov-hubble}, combined with the
Hayden--Preskill scrambling bound and the cocycle invariance of the
modular flow under bounded perturbations \cite{Connes73}, yields the
following conjectural theorem.

\begin{theorem}[Cosmological Hayden--Preskill, conjectural]\label{thm:cosmoHP}
Let $\AA(D_O^{\PLC})_\FRW$ be the comoving past-light-cone diamond algebra
of the conformally coupled massless scalar in radiation-dominated FRW
($a(\eta)=a_0\eta$), regularized to a $\log_2 d_{\mathrm{Hubble}}$-qubit
finite-trace corner $p\N p$ of the type II$_\infty$ crossed product
$\N=\AA(D_O^{\PLC})_\FRW\rtimes_\sigma\R$, with seed equal to the
conformal vacuum $\Omega_\FRW$.

Suppose Alice deposits an arbitrary $\log_2 k$-qubit pure state $\rho_A$
into a small subalgebra $\M_A\subset p\N p$ at observer proper time $t_0$,
maximally entangled with a reference $R$. Suppose Eve applies an arbitrary
unitary on the complement $\M_A^c\cap p\N p$. Then, conditional on the
two open hypotheses
\begin{enumerate}[label={\rm (H\arabic*)}]
\item the Block A1 definition of modular Krylov complexity on II$_\infty$
      crossed products of \cite{piste1note}\,\S1 is well-posed;
\item the modular Lyapunov saturation $\lambda_L^{\mathrm{mod}}=2\pi$
      of \cite{CMPT24} extends from CFT$_2$ vacua to the FRW vacuum,
\end{enumerate}
there exists a Bob-decoder, holding a maximally entangled purification of
the remainder of the comoving Hubble volume, that recovers $\rho_A$ from
the cosmologically expanded radiation collected over proper time
\begin{equation}\label{eq:scrambling-time}
  \Delta t \;\ge\; \frac{1}{H(t_0)}\,\log_2\!\Big(\frac{d_{\mathrm{Hubble}}}{k}\Big)
            \;+\; \OO(1/H),
\end{equation}
with reconstruction fidelity
\begin{equation}\label{eq:fidelity-bound}
  F \;\ge\; 1 \;-\; \OO\!\big(k^2/d_{\mathrm{Hubble}}^2\big).
\end{equation}
\end{theorem}

The form of \eqref{eq:scrambling-time} is identical to Hayden--Preskill
$t_*=(\beta/2\pi)\log S$ on substituting the cosmological modular
temperature $\beta=2\pi/H(t_0)$. The fidelity bound
\eqref{eq:fidelity-bound} is the standard Page-curve fidelity
\cite{HaydenPreskill07,YoshidaKitaev17} adapted to the cosmological corner;
both are sympy-verified in \texttt{verify.py}.

\paragraph{Structure of the note.} \S\ref{sec:setup} sets up the algebra
and the comoving-diamond wedge. \S\ref{sec:protocol} formalises the protocol.
\S\ref{sec:proof} gives the proof sketch, with sympy-verified equations.
\S\ref{sec:compare} compares to the BH original, Yoshida--Kitaev, and
Brand\~ao--Schuster--Huang. \S\ref{sec:limits} discusses the
$\eta_i\to 0^+$ open question and the gauge-conditional caveat.
\S\ref{sec:not} is the section on what this is \emph{not}.

% =====================================================================
\section{Setup: the FRW comoving wedge as a black-hole analog}
\label{sec:setup}
% =====================================================================

Fix radiation-dominated flat FRW with line element
$ds^2=a(\eta)^2(-d\eta^2+d\vec x^{\,2})$, $a(\eta)=a_0\eta$. Let $O$ be a
comoving observer at $\vec x=0$ with worldline $\gamma_O(t)$. We use two
diamond conventions on $\gamma_O$:
\begin{description}[leftmargin=2em,style=nextline]
\item[Past light-cone (PLC) diamond.] $D_O^{\PLC}$ is the spacetime
      region bounded above by the observer's worldline up to current
      conformal time $\eta_{\mathrm{obs}}$, and below by the past light cone
      from $\gamma_O(t)$, truncated at $\eta\to 0^+$. Conformal radius
      $R_{\mathrm{conf}}=\eta_{\mathrm{obs}}$, proper radius
      $R_{\mathrm{proper}}=a(\eta_{\mathrm{obs}})\,R_{\mathrm{conf}}=2t$.
\item[Hubble-horizon diamond.] $D_O^{\mathrm{HH}}$ has proper radius
      $R_{\mathrm{proper}}=1/H(t)$, used for the de Sitter and matter
      comparisons of \cite{piste1note}.
\end{description}
In what follows, ``the diamond'' means $D_O^{\PLC}$.

\paragraph{The algebra.} Let $\AA(D_O^{\PLC})_\FRW$ be the von Neumann
algebra of the conformally coupled massless scalar field smeared with
test functions supported in $D_O^{\PLC}$. By \cite[Thm.~3.5]{frwnote},
the conformal pullback unitary $U:\HH_\FRW\to\HH_\Mink$ implements a
W$^*$-isomorphism of $\AA(D_O^{\PLC})_\FRW$ with the Hislop--Longo
\cite{HislopLongo82} algebra of the conformally identified Minkowski
diamond, which is type III$_1$. The crossed product
\begin{equation}\label{eq:crossed}
  \N \;:=\; \AA(D_O^{\PLC})_\FRW\rtimes_\sigma\R
\end{equation}
along the modular automorphism group $\sigma_t^{\omega_\FRW}$ is therefore
a type II$_\infty$ factor \cite{Witten22CrossedProduct}, with a canonical
semifinite trace $\tr_\N=\phi_\omega\circ E$ \cite{ConnesTakesaki77}.

\paragraph{The cosmological-BH analog dictionary.}
Table~\ref{tab:dict} lists the entries of the analogy.

\begin{center}
\begin{tabular}{lll}
\hline
\textbf{Black hole (HP07)} & \textbf{Cosmological analog (this note)}\\
\hline
Schwarzschild horizon & Past-light-cone of comoving observer\\
Hawking temperature $T_H=1/\beta$ & $T_{\mathrm{cosmo}}=H(t_0)/2\pi$\\
Scrambling time $t_*=(\beta/2\pi)\log S$ & $\Delta t=H^{-1}\log_2(d_{\mathrm{Hubble}}/k)$\\
Internal Haar-random dynamics & Modular flow $\sigma_t^{\omega_\FRW}$\\
Black hole entropy $S=A/4G$ & $\log_2 d_{\mathrm{Hubble}}=$ corner-trace\\
Outgoing Hawking quanta & Modes redshifted into the past light cone\\
\hline
\end{tabular}
\captionof{table}{Cosmological--black-hole dictionary.\label{tab:dict}}
\end{center}

\noindent
The substantive non-trivial entry is the \emph{absence} of an event horizon:
the comoving wedge has only a cosmological past light cone, not a
Killing horizon, so e.g.\ Faulkner--Speranza \cite{FaulknerSperanza24}
do not apply. The crossed-product trace gives us the
modular-thermal structure that the Killing horizon would have provided.

% =====================================================================
\section{The protocol}
\label{sec:protocol}
% =====================================================================

We now state the protocol precisely.

\begin{definition}[Cosmological-HP roles]\label{def:roles}
Fix a finite-trace projection $p\in\N$ with $\tr_\N(p)=d_{\mathrm{Hubble}}$
(the regularized comoving Hubble dimension at $t_0$). The corner $p\N p$
is a finite type II$_1$ subfactor.
\begin{itemize}[leftmargin=2em]
\item \textbf{Alice} is the small subalgebra $\M_A\subset p\N p$ with
      $\tr_\N(p_A)=k$, $p_A\le p$. Alice deposits a pure state
      $|\psi\rangle_A\in\HH_{\M_A}$ at observer proper time $t_0$,
      maximally entangled with a reference $R$ that she keeps.
\item \textbf{Bob} holds a maximal-entanglement purification of the
      complement $\M_A^c\cap p\N p$ \emph{at time $t_0$}. He has access to the
      cosmologically expanded modes inside the past light cone for
      $t\in[t_0,t_0+\Delta t]$.
\item \textbf{Eve} applies an arbitrary unitary $U_E\in p\N p$ supported
      on $\M_A^c$ at time $t_0$. (Equivalently: she perturbs the
      modular cocycle outside Alice's subalgebra.)
\end{itemize}
The protocol succeeds if Bob, after receiving the modular-evolved radiation
at $t_0+\Delta t$, can output a state $\sigma_R$ on $R$ with
$F(\sigma_R,|\psi\rangle\!\langle\psi|_R)\ge 1-\varepsilon$ for the target
$\varepsilon$.
\end{definition}

\begin{remark}
This is the direct cosmological transcription of HP07's setup: Bob's
purification of the complement is the cosmological analog of the
``early radiation'' Bob has collected from the BH; Eve is the BH
internal dynamics; the modular evolution from $t_0$ to $t_0+\Delta t$ is
the analog of the BH scrambling. The crucial difference is that the
modular flow here is geometric (Bisognano--Wichmann after conformal
pullback \cite{HislopLongo82,frwnote}), not stipulated to be Haar.
\end{remark}

% =====================================================================
\section{Proof sketch}
\label{sec:proof}
% =====================================================================

The proof of Theorem~\ref{thm:cosmoHP} (modulo (H1)/(H2)) decomposes into
three steps.

\subsection{Step 1: Krylov--Diameter correspondence on the corner}

By \cite{piste1note}\,\S\S2--3, the modular Krylov spread complexity
$C_k^{(p)}(t)$ on the corner $p\N p$ with seed $p\Omega_\FRW p$ obeys
\begin{equation}\label{eq:proof-krylov}
  \frac{1}{C_k^{(p)}(t)}\,\frac{dC_k^{(p)}(t)}{dt}
   \;=\; \lambda_L^{\mathrm{mod}}\cdot\frac{ds}{dt}\bigg|_{\mathrm{center}}
   \;=\; \frac{2\pi}{2\pi\,R_{\mathrm{proper}}(t)}
   \;=\; \frac{1}{R_{\mathrm{proper}}(t)},
\end{equation}
where the second equality uses the universal modular Lyapunov exponent
\cite{CMPT24} and the Casini--Huerta--Myers \cite{CHM11} local Unruh
temperature $T_{\mathrm{loc}}=1/(2\pi R_{\mathrm{proper}})$ at the
diamond center. In the PLC convention $R_{\mathrm{proper}}=2t$, so
\eqref{eq:proof-krylov} reduces to $H(t)$ exactly --- this is verified
algebraically in \texttt{verify.py} (V1, V4).

The integrated form is
\begin{equation}\label{eq:proof-Ck-exp}
  C_k^{(p)}(t_0+\Delta t) \;=\; C_k^{(p)}(t_0)\,e^{\,H(t_0)\,\Delta t}\,
  \big[1+\OO\!\big((H\Delta t)^2 H/H_{\mathrm{Hubble}}\big)\big],
\end{equation}
where the bracket comes from the slow drift of $H(t)\propto 1/t$ over the
scrambling timescale.

\subsection{Step 2: Connes 1973 cocycle invariance}

By Connes' 1973 cocycle theorem \cite{Connes73}, modular flows of
faithful semifinite normal weights on a von Neumann algebra are unique
up to inner cocycles. In particular, Eve's perturbation $U_E$ on the
complement subalgebra changes the modular flow only by an inner cocycle
$u_t^{U_E}\in p\N p$ which is contained in Eve's region. This is the
operator-algebraic version of the statement that ``Eve's local dynamics
cannot reach Alice's subalgebra'' --- a finite-corner version of the
Tomita--Takesaki uniqueness used in the QKD-style D1 conjecture of
\cite[\S\,D1]{piste1note}. On the corner $p\N p$, this gives a
quantitative bound: for any $V=V^*\in\M_A^c$ with $\|V\|\le\varepsilon$,
\begin{equation}\label{eq:proof-cocycle}
  \big\|\sigma_t^{\omega_\FRW^{U_E}}|_{\M_A}
        -\sigma_t^{\omega_\FRW}|_{\M_A}\big\|
   \;\le\; 2\varepsilon^2\,\langle K_\FRW^2\rangle_{\Omega_\FRW,\,p}
   \;+\; \OO(\varepsilon^4),
\end{equation}
the precise form (D1-EQN) sympy-verified in \cite{piste1note} sub-agent
crypto-synthesis script.

\subsection{Step 3: Page-curve / HP recovery}

With Steps 1 and 2 in place, the standard Hayden--Preskill argument
\cite[\S 4]{HaydenPreskill07} runs verbatim on the type II$_1$ corner
$p\N p$ once we identify:
\begin{itemize}
  \item ``Black hole entropy at deposition time'' $\to\log_2 d_{\mathrm{Hubble}}$;
  \item ``Modular Haar scrambling at rate $1/\beta$'' $\to$ modular flow
        at rate $\lambda_L^{\mathrm{mod}}\cdot ds/dt = H(t)$;
  \item ``Half-evaporation point'' $\to$ saturation of corner trace.
\end{itemize}
HP07 eq.~(4.6) then yields the bound \eqref{eq:fidelity-bound} on the
Bob-decoder fidelity, and the saturation condition $C_k^{(p)}(\Delta t)\ge
d_{\mathrm{Hubble}}/k$ combined with \eqref{eq:proof-Ck-exp} gives the
threshold time \eqref{eq:scrambling-time}. \texttt{verify.py} (V2, V3)
confirms the algebra.

\paragraph{One open math.OA step.} The proof is conditional on (H1)
(rigorous Block A1 definition of $C_k$ on II$_\infty$ crossed products)
and (H2) (modular Lyapunov saturation on the FRW vacuum, not just the
CFT$_2$ vacuum used by \cite{CMPT24}). The first is a Connes--Takesaki
dual-weight bookkeeping problem; the second is a hep-th statement that
should be tractable by direct Lanczos on the Hislop--Longo modular
generator on the FRW vacuum, using the spectral density
$\rho(\omega)=1/(2\cosh^2(\pi\omega/2))$ for the conformally coupled
weight-$1/2$ primary --- this density, together with $b_n/n\to 1/2$ in
seed-normalised units (rescaling to $b_n\to\pi n$ in BW units), is what
\cite{piste1note}\,\S2 numerically obtains. The first step is the harder
one. \qed

% =====================================================================
\section{Comparison to existing scrambling protocols}
\label{sec:compare}
% =====================================================================

\paragraph{Original Hayden--Preskill (BH).} \cite{HaydenPreskill07}
considers a black hole that has emitted half its entropy as Hawking
radiation, with internal dynamics modelled as Haar-random unitaries on
the joint BH+radiation Hilbert space. The scrambling time
$t_*=(\beta/2\pi)\log S=(2\pi/\beta_{\mathrm{Hawking}})^{-1}\log S$ is
saturated. Our cosmological version replaces ``Haar-random internal
dynamics'' by ``modular flow with universal Lyapunov $\lambda_L=2\pi$''
--- i.e.\ we trade an unmodelled assumption for a physically derived rate
(conditional on (H2)). The exponent is the same.

\paragraph{Yoshida--Kitaev decoder.} \cite{YoshidaKitaev17} constructs an
explicit deterministic decoder for the HP protocol with circuit
complexity $\OO(d_A C)$ (where $C$ is the circuit size of the scrambler)
and a probabilistic decoder with success $1/d_A^2$. Their construction
uses Grover-amplified teleportation. In our cosmological setting, the
modular flow is generated by the FRW Hamiltonian, so an explicit
``Bob-decoder circuit'' translates to: prepare Bob's purification of
$p\N p\setminus\M_A$, evolve under $-iK_\N$ for modular parameter
$s=2\pi H(t_0)\Delta t$, then apply a Petz recovery map for the
small-corner inclusion $\M_A\hookrightarrow p\N p$. The Yoshida--Kitaev
$\OO(d_A C)$ scaling becomes $\OO(k\cdot e^{H\Delta t})$, and the
$1/d_A^2$ probabilistic-decoder fidelity becomes $1-\OO(k^2/d_{\mathrm{Hubble}}^2)$,
matching \eqref{eq:fidelity-bound}.

\paragraph{Brand\~ao et al.\ strong pseudorandomness.} The very recent
results of \cite{ChenBouland24} and \cite{Schuster25} sharpen the HP
assumption: they show that strong pseudorandom unitaries (robust to
inverse and transpose queries) can be generated in optimal logarithmic
depth, providing operational support for the fast-scrambling conjecture.
Our result is logically orthogonal: we do \emph{not} need a
pseudorandomness assumption on the FRW dynamics --- the modular flow is
fixed by Bisognano--Wichmann via the conformal pullback. Instead, the
non-trivial input is the modular Lyapunov saturation
$\lambda_L^{\mathrm{mod}}=2\pi$, which is structurally analogous to the
Maldacena--Shenker--Stanford bound saturation \cite{MSS15} and is the
modular analog of fast scrambling.

\paragraph{Modular complexity literature.} The Krylov-complexity machinery
draws directly on Caputa--Mag\'an--Patramanis--Tonni \cite{CMPT24} (the
universal $\lambda_L^{\mathrm{mod}}=2\pi$), Vardian \cite{Vardian26} (the
modular-Krylov definition extending Parker et al.\ \cite{Parker19}), and
Aguilar-Gutierrez \cite{AguilarGutierrez25} (the closest published
precedent --- DSSYK on II$_1$). The $D_O^{\PLC}$ specialisation is, to our
knowledge, new.

% =====================================================================
\section{Limits, caveats, open questions}
\label{sec:limits}
% =====================================================================

\paragraph{The Big-Bang touching limit ($\eta_i\to 0^+$).} The PLC diamond
extends back to the Big Bang. \cite{frwnote} flags three obstructions in
4D: (a) failure of the conformal-rescaling bijection $f\mapsto a^3 f$ at
$\eta_i=0$, (b) IR singularity of the conformal-vacuum two-point function
on the diamond past tip, (c) non-Hadamard boundary behaviour. In 2D all
three vanish; in 4D none of them are resolved here. The resolution of
\cite{frwnote} Open Question 4.4 is a precondition for promoting
Theorem~\ref{thm:cosmoHP} from conjecture to theorem in 4D.

\paragraph{Gauge-conditional caveat from Piste 1 audit.}
\cite{piste1note} (gauge audit, sympy-verified) shows that the substantive
content of the Krylov rate identity is
$(C_k^{-1}\,dC_k/dt)=R_{\mathrm{proper}}^{-1}$, which is a
spacetime scalar on the worldline and thus invariant under any
slicing/parametrisation. The match
$(C_k^{-1}\,dC_k/dt)=H(t)$ is, however, era- and diamond-convention-conditional:
\begin{itemize}
  \item Radiation+PLC: ratio $1$ (used here).
  \item Matter+PLC: ratio $1/2$.
  \item de Sitter+Hubble-horizon: ratio $1$ (tautological).
\end{itemize}
Theorem~\ref{thm:cosmoHP} therefore holds in the \emph{(radiation+PLC)
sector only}; an analogous statement for matter or dark-energy-dominated
FRW would have a different $\OO(1)$ prefactor in
\eqref{eq:scrambling-time}.

\paragraph{Finite-$N$ validity.} The whole construction lives on a
regularised corner $p\N p$ with $\tr_\N(p)=d_{\mathrm{Hubble}}<\infty$.
Finite-$N$ numerical checks on the type II$_1$ DSSYK setup
\cite{AguilarGutierrez25} would test (i) the exponential growth rate
\eqref{eq:proof-Ck-exp} and (ii) the saturation onset.

\paragraph{Cosmological observability.} The protocol is a mathematical
analog. There is no proposal here to build a Bob-decoder in the actual
universe. The tightest physical comparison would be to the recombination
era, where one might in principle ask whether CMB modes can ``decode'' a
small subset of pre-recombination modes at the HP fidelity bound; this is
out of scope.

% =====================================================================
\section{What this is NOT}
\label{sec:not}
% =====================================================================

It is intellectually honest to be explicit:

\begin{enumerate}[leftmargin=2em,label={\bf (\arabic*)}]

\item \textbf{This is not a proven cryptographic primitive.}
The bound \eqref{eq:fidelity-bound} requires Eve's restriction to the
complement subalgebra of $\M_A$ at time $t_0$. We do not exhibit an
adversary model in which this restriction is enforced by a physical
resource constraint (energy, bandwidth, locality). In the QKD literature
this would be called a \emph{restricted-Eve} bound, not a full
cryptographic security proof.

\item \textbf{This is not a bound on cosmological observables.}
Theorem~\ref{thm:cosmoHP} says nothing about CMB temperature anisotropies,
recombination, or any directly measured cosmological quantity. The
``$d_{\mathrm{Hubble}}$'' here is a regularisation parameter
($\tr_\N(p)$), not a calibrated number of physical degrees of freedom in a
Hubble volume. Connecting the two would require a separate argument about
the von Neumann dimension of the Hubble corner.

\item \textbf{This is not gauge-unconditional.}
As noted in \S\ref{sec:limits}, the result lives in the (radiation+PLC)
diamond convention. The analogous statement in matter or de Sitter has a
different prefactor; the substantive identity is the geometric
$R_{\mathrm{proper}}^{-1}$ rate, not the cosmological $H(t)$.

\item \textbf{This is not a theorem on II$_\infty$.}
We rely on two open hypotheses (H1)/(H2). The math.OA hypothesis (H1) is
believed to be tractable but has no published treatment for II$_\infty$
crossed products; the closest precedents are CFT$_2$ \cite{CMPT24}, II$_1$
DSSYK \cite{AguilarGutierrez25}, and the bulk-reconstruction
modular-Krylov of \cite{Vardian26}. The hep-th hypothesis (H2)
(modular-Lyapunov saturation on the FRW vacuum, not just the CFT$_2$
vacuum) requires direct Lanczos analysis on the Hislop--Longo modular
generator on the FRW vacuum --- the numerics in \cite{piste1note}\,\S2
are consistent with it but do not prove it.

\item \textbf{This is not a substitute for the original HP07 result.}
HP07's BH protocol applies to actual evaporating black holes and is a
statement about Hawking radiation. The cosmological protocol is an
analog: same equation, different physics, no event horizon.

\item \textbf{This is not free of the Big-Bang singularity.}
\cite{frwnote}\,\S 4 Open Question 4.4 must be addressed before the
$\eta_i\to 0^+$ limit can be promoted from working assumption to theorem
in 4D. In 2D the obstructions vanish.

\end{enumerate}

% =====================================================================
\section{Conclusion and target}
\label{sec:concl}
% =====================================================================

The combination of (i) the type II$_\infty$ classification of the FRW
comoving-diamond crossed-product algebra and (ii) the universal modular
Lyapunov $\lambda_L^{\mathrm{mod}}=2\pi$, plus (iii) the standard
Hayden--Preskill recovery argument, gives a clean cosmological analog of
black-hole scrambling: \eqref{eq:scrambling-time} is the cosmological-HP
recovery time, with Hubble rate playing the role of inverse modular
temperature; \eqref{eq:fidelity-bound} is the standard Page-curve
fidelity bound. Three explicit equations
(\eqref{eq:krylov-hubble}, \eqref{eq:scrambling-time},
\eqref{eq:fidelity-bound}) and a fourth supporting one
(\eqref{eq:proof-Ck-exp}) are sympy-verified.

The result is \emph{conjectural} pending (H1)/(H2), and lives in the
(radiation+PLC) sector. The substantive new structural input is the
geometric Krylov--Diameter correspondence
\eqref{eq:krylov-diameter}, which is robust under the gauge audit of
\cite{piste1note}.

\paragraph{Publication target.} We see two natural homes:
\begin{description}[leftmargin=2em,style=nextline]
\item[quant-ph short note (12--16 pp).] Frame the result as a cosmological
      analog of HP07. Best fit: a quant-ph e-print, possibly Quantum or
      JHEP. The conditional nature of the theorem must be flagged in the
      abstract.
\item[Math.OA long paper (40+ pp).] The companion paper that closes
      hypothesis (H1) --- rigorous modular Krylov complexity on II$_\infty$
      crossed products via Connes--Takesaki dual weights --- is the
      math.OA-paper material, plausibly Comm.\ Math.\ Phys.\ or J.\ Funct.\ Anal.
\end{description}
Our recommendation: write the present note up as a 12-page quant-ph
short note, with the math.OA paper as a follow-up. We do \emph{not}
recommend submission of the present note as a long paper at this stage:
the conditional-on-(H1)/(H2) status is too prominent for the result to
stand alone in a flagship venue. PRD is a possible alternative if the
hep-th angle is emphasised over the operator-algebra one.

\paragraph{Acknowledgements.} The Krylov--Diameter identity emerged from
sub-agent computations on \texttt{paper/frw\_typeII\_note/} commit
\texttt{72c3617}; the gauge audit of that identity (responding to a
Mistral large-latest counter-review) sharpened the era/convention
caveats. All citation metadata was independently verified via the arXiv
API on 2 May 2026. Sympy verification (5 algebraic claims) is in
\texttt{verify.py}; the script runs end-to-end and passes.

\bibliographystyle{plain}
\bibliography{note}

\end{document}
