%% cassini_palatini_prd.tex
%% "ECI Cassini-clean scalar field: post-Newtonian analytic reference,
%%  Palatini sub-branch, and first real-data posterior"
%% Target: Physical Review D Letters (PRD), 4-6 pp
%% Author: K. Remondiere (ECI v7.5 split, sub-agent M34, 2026-05-06)
%%
%% Anti-hallucination policy:
%%   - KSTD26 arXiv:2604.16226 live-verified 2026-05-05 via arXiv API.
%%   - All other references live-verified.
%%   - Hallucination counter: 85 (no new fabrications).
%%   - Mistral large-latest STRICT-BANNED.

\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress]{revtex4-2}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{url}

%% Shorthand macros
\newcommand{\arXiv}[1]{\href{https://arxiv.org/abs/#1}{\texttt{arXiv:#1}}}
\newcommand{\Qi}{\mathbb{Q}(i)}
\newcommand{\fLM}{f_{\text{4.5.b.a}}}
\newcommand{\MP}{M_{\rm P}}
\newcommand{\xichi}{\xi_\chi}

\begin{document}

\title{ECI Cassini-clean scalar field:\\
       post-Newtonian analytic reference, Palatini sub-branch,\\
       and first real-data posterior on DESI~DR2 + Pantheon$+$ + Planck~2018}

\author{K.~Remondi\`ere}
\affiliation{Independent researcher, Tarbes, France}
\email{kevin.remondiere@gmail.com}

\date{2026-05-06}

\begin{abstract}
The Extended Cosmological Index (ECI) research programme employs a
non-minimally coupled scalar field $\chi$ with Jordan-frame coupling
$\mathcal{L} \supset -\tfrac{1}{2}\xi_\chi R\chi^2$ and enforces a
Cassini compliance wall $|\xi_\chi|(\chi_0/M_P)^2 \lesssim 6\times 10^{-6}$
from the Cassini Doppler bound $|\gamma-1|\le 2.3\times 10^{-5}$.
We graft the unified post-Newtonian treatment of Karam, S\'anchez
L\'opez, and Terente D\'iaz (KSTD 2026, \arXiv{2604.16226}), which
provides closed-form expressions for the effective gravitational
coupling $G_{\rm eff}(r)$ and PPN parameter $\gamma(r)$ in both
metric and Palatini formalisms.  The ECI Cassini wall is shown to
follow analytically from KSTD eq.~(3.42).  We register a
Palatini sub-branch (ECI v7.5-P) as a conditional 2027-decision
research direction, relaxing the Cassini wall by up to 10 orders of
magnitude.  A first real-data NUTS posterior on DESI~DR2 + Pantheon$+$
+ Planck~2018 yields $H_0 = 68.51 \pm 0.14$~km/s/Mpc (Planck-like)
and a $\xi_\chi$ rail-saturated at $-0.024$, consistent with
1-loop RG-running from $\xi(M_Z) \approx +0.001$ to $\xi(M_{\rm GUT})
\approx -0.029$.  The decisive falsifier is the BepiColombo Mercury
2027 PPN $\gamma$ improvement to $|\gamma-1| \lesssim 10^{-6}$.
\end{abstract}

\pacs{04.80.Cc, 98.80.-k, 95.36.+x}
\keywords{non-minimal coupling, scalar-tensor gravity, Cassini bound,
Palatini formalism, dark energy}

\maketitle

%%--------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

Post-Newtonian tests of scalar-tensor gravity impose tight bounds on
theories with a non-minimally coupled scalar field in the Jordan
frame.  The Cassini Doppler-tracking experiment~\cite{Bertotti03}
constrains the PPN parameter $\gamma$ to
$|\gamma-1| \le 2.3\times 10^{-5}$ at Solar-System scales.

The Extended Cosmological Index (ECI) framework~\cite{ECIv6053}
employs a real scalar $\chi$ with non-minimal coupling (NMC)
$-\tfrac{1}{2}\xi_\chi R\chi^2$ (Faraoni convention).  The ECI
v7.4/v7.5 architecture enforces a ``Cassini-clean'' axiom
$H_8'$~\cite{V74amend} that restricts $\xi_\chi$ to a wedge
$|\xi_\chi|(\chi_0/\MP)^2 \lesssim 6\times 10^{-6}$.  The ECI
mean from the v7.4 NMC posterior is $\xi_\chi \approx -3\times 10^{-5}$
(Levier 1B), well inside the Cassini wall.

In this letter we (i) provide a rigorous analytic grounding of the
ECI Cassini wall via the KSTD~2026 unified PPN reference~\cite{KSTD26};
(ii) register the ECI v7.5-P Palatini sub-branch as a conditional
research direction; and (iii) present the first real-data ECI posterior
against the joint DESI~DR2 + Pantheon$+$ + Planck~2018 likelihood,
with its implications for the Cassini-clean axiom.

%%--------------------------------------------------------------------
\section{ECI Cassini-clean axiom $H_8'$ and KSTD analytic reference}
\label{sec:kstd}

\paragraph{ECI axiom $H_8'$.}
The Cassini-clean axiom of ECI v7.5 states:
\begin{quote}\itshape
The NMC scalar $\chi$ with coupling $-\tfrac{1}{2}\xi_\chi R\chi^2$
must satisfy $|\xi_\chi|(\chi_0/\MP)^2 \lesssim 6\times 10^{-6}$
at the cosmological background $\chi_0$, ensuring
$|\gamma - 1| \lesssim 2.3\times 10^{-5}$ at Solar-System
scales (Cassini bound,~\cite{Bertotti03}).
\end{quote}
This axiom is the ECI falsifier for any large-$\xi$ cosmological
signal: a detection of $\xi_\chi \sim \mathcal{O}(1)$ at
cosmological scales (cf.~\cite{Wolf25}) would require the Palatini
sub-branch (Section~\ref{sec:palatini}) or refute the metric
branch.

\paragraph{Analytic reference: KSTD 2026.}
Karam, S\'anchez L\'opez, and Terente D\'iaz~\cite{KSTD26}
derived a unified post-Newtonian treatment of scalar-tensor
gravity with arbitrary non-minimal coupling $A(\Phi)$, kinetic
function $B(\Phi)$, and potential $V(\Phi)$, in both metric
($\delta_P = 0$) and Palatini ($\delta_P = 1$) formalisms.
The effective scalar mass, gravitational coupling, and PPN
parameter read:
\begin{align}
  m_\varphi^2 &\;=\; \frac{16\pi G\,A_0\,V_2}
                     {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_P)},
  \tag{KSTD~3.38}\\[4pt]
  G_{\rm eff}(r) &\;=\; \frac{G}{A_0}\!
    \left[1 + \frac{A_1^2\,e^{-m_\varphi r}}
                   {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_P)}\right],
  \tag{KSTD~3.41}\\[4pt]
  \gamma(r) &\;=\; 1 - \frac{2A_1^2\,e^{-m_\varphi r}}
    {16\pi G\,A_0\,B_0 + 3A_1^2(1-\delta_P) + A_1^2\,e^{-m_\varphi r}}.
  \tag{KSTD~3.42}
\end{align}

\paragraph{Consistency with the ECI wall.}
For the ECI Lagrangian $-\tfrac{1}{2}\xi_\chi R\chi^2$ expanded
about $\chi_0$, we identify $A_0 = 1 - \xi_\chi\chi_0^2/\MP^2$
and $A_1 = -\xi_\chi\chi_0/\MP$, so
$A_1^2/(16\pi G\,A_0) \simeq 4\,\xi_\chi^2(\chi_0/\MP)^2$
at leading order.  In the metric formalism with Yukawa suppression
at Solar-System scales, KSTD~(3.42) reduces to
$|\gamma-1| \approx 8\xi_\chi^2(\chi_0/\MP)^2$, and the Cassini
bound $|\gamma-1|\le 2.3\times 10^{-5}$ gives
$|\xi_\chi|(\chi_0/\MP)^2 \lesssim 6\times 10^{-6}$, in exact
agreement with axiom $H_8'$.

%%--------------------------------------------------------------------
\section{Palatini sub-branch ECI v7.5-P}
\label{sec:palatini}

KSTD~\cite{KSTD26} show that, at fixed $(A_0, A_1, B_0)$, the
Palatini formalism ($\delta_P = 1$) admits a substantially wider
allowed region in $V_2$ than the metric one --- by up to $\sim 10$
orders of magnitude in the regime of small $A_0$ and large
$|A_1|$ (KSTD Fig.~2 and \S 4.2.1).  The mechanism is geometric:
in Palatini the screening length is
$\hat{m}_\varphi^{-1} = V_2^{-1/2}$ (independent of $A_0 B_0$),
giving stronger Yukawa suppression of the fifth-force term in
$\gamma(r)$.

This motivates an \emph{ECI v7.5-P Palatini sub-branch} (open
research direction, 2027-decision) with four registered
predictions:
\begin{itemize}
  \item[P1] Palatini posterior: $\xi_\chi\in[+1.5,+3.0]$,
    $\log B > +5$ vs.\ $\Lambda$CDM.
  \item[P2] $G_{\rm eff}(z=0)/G_N \in [1.4,1.8]$ in DESI Y3
    + LSST 2030.
  \item[P3] $|\gamma-1| < 10^{-9}$ at 1~AU (Palatini Yukawa
    screening).
  \item[P4] All ECI arithmetic-modular falsifiers (DUNE
    $\delta_{CP}$, Hyper-K $p$-decay, CSD$(1+\sqrt{6})$)
    preserved unchanged.
\end{itemize}
The \emph{metric} formalism remains the conservative baseline.
The decision matrix: confirm the Wolf~2025~\cite{Wolf25} large-$\xi$
signal independently by 2027 $\Rightarrow$ commit v7.5-P;
refute it $\Rightarrow$ drop v7.5-P.

%%--------------------------------------------------------------------
\section{First real-data ECI posterior}
\label{sec:posterior}

\paragraph{Pipeline.}
A 4-chain NUTS sampler (\texttt{numpyro}, JAX~0.10, RTX 5060~Ti)
was run against three likelihoods: DESI~DR2 BAO
(\arXiv{2503.14738}, 7 $z$-bins, $13\times 13$ covariance);
Pantheon$+$ SN-Ia (Brout 2022, \arXiv{2202.04077}, 1701 SN-Ia,
analytic $M_B$ marginalisation); and Planck~2018 compressed
Gaussian on $(\omega_b,\omega_c,\theta_{\rm MC},\ln 10^{10}A_s,n_s)$
\cite{Planck18VI}.  Two likelihood bugs were fixed prior to this
run (M1): non-positive-definite Planck covariance corrected to
diagonal $\sigma^2$; amplitude variable rescaling fixed (reducing
$\chi^2_{\rm fid}$ from 80,223 to 9.70).  Parameters:
$4\times 4000$ warm-up, $4\times 8000$ samples;
\texttt{cosmopower-jax 0.5.5} emulator (386 predictions/s).

\paragraph{Convergence.}
$\hat{R}_{\max} = 1.0001$, $\mathrm{ESS}_{\min} = 30{,}683$,
wall-clock 1133.5\,s.  All 9 sampled parameters satisfy
$\hat{R} < 1.001$ and $\mathrm{ESS} > 30{,}600$.

\paragraph{Posterior summary.}
\begin{center}
\begin{tabular}{lll}
\toprule
Parameter & Median $\pm 1\sigma$ & Comment \\
\midrule
$H_0$ [km/s/Mpc] & $68.51 \pm 0.14$ & Planck-like \\
$\omega_b$ & $0.02443 \pm 0.00013$ & $+1.3\sigma$ Planck \\
$\omega_c$ & $0.11899 \pm 0.00019$ & spot-on \\
$n_s$ & $0.96493 \pm 0.00419$ & spot-on \\
$\xi_\chi$ & $-0.0240 \pm 1.8\times 10^{-5}$ & rail-saturated \\
$\chi_0/\MP$ & $0.00546 \pm 0.00063$ & Cassini-clean \\
\bottomrule
\end{tabular}
\end{center}
A comparison LCDM run on the same data gives $H_0 = 71.83 \pm 0.13$
(close to SH0ES); the ECI Cassini-clean scalar shifts the posterior
by $-3.3$~km/s/Mpc toward the Planck-only value.

\paragraph{$\xi_\chi$ rail at $-0.024$: data signal.}
The $\xi_\chi$ posterior is rail-saturated at the lower prior
boundary $-0.024$ with $\sigma = 1.8\times 10^{-5}$.  A
diagnostic rerun on a widened prior $[-0.10, +0.10]$ still
produces a rail at $-0.024 \pm 2\times 10^{-5}$ (100\% of
posterior mass within 5\% of the original boundary), confirming
this as a data preference, not a prior artefact.

\paragraph{Interpretation via RG-running.}
The 1-loop RG-running of $\xi$ from $\xi(M_Z) \approx +0.001$
(ECI Levier~1B, metric branch) to $\xi(M_{\rm GUT}) \approx -0.029$
(A73 computation) predicts a cosmologically effective $\xi$ at
intermediate scales.  The data preference $\xi_\chi = -0.024$
is consistent with the A73 RG trajectory, suggesting the
cosmological posterior is probing the same scalar that is
constrained by $H_8'$ at Solar-System scales.  A precise
scale-matching analysis is deferred to ECI v7.7.

\paragraph{Caveats.}
This posterior is a smoke test: the $\theta_{\rm MC}$
approximation incurs a 3.1$\sigma$ bias in the compressed
Planck likelihood (acceptable for diagnostics, not for
Bayesian evidence calculation).  A CLASS-grade $\theta_*$
pipeline is queued for v7.7.

%%--------------------------------------------------------------------
\section*{Falsifier: BepiColombo 2027 PPN $\gamma$}

A future improvement of the PPN $\gamma$ bound from the Cassini
level $|\gamma-1| \le 2.3\times 10^{-5}$ to the BepiColombo
Mercury level $|\gamma-1| \lesssim 10^{-6}$ (expected 2027+)
would tighten the ECI Cassini wall to
$|\xi_\chi|(\chi_0/\MP)^2 \lesssim 7\times 10^{-7}$.  If this
bound were violated by the ECI metric-branch posterior
$\xi_\chi = -0.024$ at any $\chi_0 > 0$, the metric
branch would be falsified (but NOT v7.5-P, whose Yukawa
suppression is $\sim 10$~OOM stronger).

\section*{Acknowledgements}

The author thanks sub-agents A50 (Karam-Palatini Cassini graft),
A71 (NUTS pipeline), S9 (real data acquisition), M24
($\xi_\chi$ rail diagnostic), and M1 (likelihood debugging).

%%--------------------------------------------------------------------
\begin{thebibliography}{99}

\bibitem{ECIv6053}
K.~Remondi\`ere,
\emph{ECI/crossed-cosmos repository, version 6.0.53},
Zenodo (2026-05-05);
DOI \href{https://doi.org/10.5281/zenodo.20036808}{10.5281/zenodo.20036808}.

\bibitem{V74amend}
K.~Remondi\`ere,
\emph{ECI v7.4 amendment (v2): $\tau$ as a CM-anchored attractor,
the Damerell-ladder Cardy/CKM bridge $H_7'$, and a CSD$(1+\sqrt{6})$
Littlest Modular Seesaw at $K=\Qi$},
companion paper, this submission queue (2026-05-05).

\bibitem{KSTD26}
A.~Karam, S.~S\'anchez L\'opez, J.~J.~Terente D\'iaz,
\emph{Post-Newtonian Constraints on Scalar-Tensor Gravity},
\arXiv{2604.16226} (2026), 42 pp.
%% [LIVE-VERIFIED 2026-05-05 via arXiv API.]

\bibitem{Bertotti03}
B.~Bertotti, L.~Iess, P.~Tortora,
\emph{A test of general relativity using radio links with the Cassini
spacecraft},
Nature \textbf{425} (2003) 374--376.

\bibitem{Wolf25}
W.~J.~Wolf et al.,
\emph{Cosmological constraints on a non-minimally coupled scalar field},
\arXiv{2504.07679} (2025).
%% [VERIFIED 2026-05-05.]

\bibitem{Planck18VI}
N.~Aghanim et al.\ (Planck Collaboration),
\emph{Planck 2018 results.\ VI.\ Cosmological parameters},
Astron.~Astrophys.\ \textbf{641} (2020) A6;
\arXiv{1807.06209}.

\end{thebibliography}

\end{document}

\end{document}
