%% erepr_araki_consistency_LMP.tex
%% Standalone note for Letters in Mathematical Physics
%% "Modular flow generates generalized entropy in type-II_inf crossed-product gravity:
%%  an Araki-cocycle derivation, with a no-go for HPS-DEHK Stinespring intertwining"
%% Status: DRAFT 2026-05-05 (recadre after A10 spectrum-mismatch verdict)
%% MVT2-class consistency note: the Araki Wk2 result is a clean theorem;
%% the Wk3 Stinespring bridge to DSSYK chord algebra is REFUTED at operator level
%% (Wk4 = A10), and only a Krylov-thermodynamic match log Z_chord = log Z_DEHK + O(1/N)
%% survives, reproducing the standard Bekenstein-Hawking dS entropy.
%% All arXiv IDs live-verified against arXiv abstract pages on 2026-05-05.

\documentclass[12pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{geometry}
\geometry{margin=2.5cm}

%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\title{Modular flow generates generalized entropy in\\
       type-II$_\infty$ crossed-product gravity:\\[4pt]
       an Araki-cocycle derivation, with a no-go for\\
       HPS--DEHK Stinespring intertwining}

\author{%
  K\'evin Remondi\`ere\\[2pt]
  \small Independent researcher, Tarbes, France\\[2pt]
  \small \textit{Submitted to Letters in Mathematical Physics}
}

\date{Preprint: \today}

\begin{document}

\maketitle

\begin{abstract}
Working in the type-II$_\infty$ crossed-product algebra
$\mathcal{M}_R = \mathcal{A}_{\mathrm{III}_1}\rtimes_{\sigma^{\phi}}\mathbb{R}$
attached to a static-patch observer (Chandrasekaran--Longo--Penington--Witten,
de~Vuyst--Eccles--H\"{o}hn--Kirklin), we use the Araki relative
modular operator $\Delta_{\rho|\sigma}$ and its cocycle identity to
derive the algebraic identity
\[
   \frac{d}{d\tau_R} S_{\mathrm{gen}}(\rho)\bigg|_{\rho \text{ fixed},\;\sigma_{\tau_R}}
   \;=\; \langle K_R \rangle_\rho ,
\]
where $K_R$ is the dressed modular Hamiltonian and $\sigma_{\tau_R}=\sigma\circ
\sigma^{\sigma}_{-\tau_R}$ is the modular-time-evolved KMS reference state.
The derivation is rigorous in the type-II$_\infty$ setting via Pusz--Woronowicz
applied to normal semifinite faithful (nsf) weights and is independent of
the Petz recovery construction (which has no known type-II$_\infty$
realization for $\mathcal{M}_R$).

We then examine, and refute at the operator level, the natural attempt to
intertwine $K_R$ with the chord (transfer-matrix) Hamiltonian
$H_{\mathrm{grav}}$ of the double-scaled SYK / sine-dilaton gravity model
(Heller--Papalini--Schuhmann; Heller--Ori--Papalini--Schuhmann--Wang) through
a modular-rigidified Stinespring isometry $V:\mathcal{H}_{\mathrm{HPS}}\to
\mathcal{H}_{\mathrm{DEHK}}$. The HPS chord spectrum
$E(\theta)=-\cos\theta/\sqrt{2|\log q|(1-q^2)}$ is bounded continuous on
$[-E_{\max},+E_{\max}]$, while $K_R=\beta_{\mathrm{dS}} H_\xi + p$ is
unbounded continuous on $\mathbb{R}$. No isometry can intertwine a bounded
self-adjoint operator with an unbounded one; hence the proposed bridge
fails as an operator-level identification.

What does survive is a thermodynamic / Krylov-asymptotic match: the
semiclassical large-$N$ partition functions agree,
$\log Z_{\mathrm{chord}}(\beta_{\mathrm{dS}}) = \log Z_{\mathrm{DEHK}}
(\beta_{\mathrm{dS}}) + O(1/N)$, reproducing the standard Bekenstein--Hawking
de Sitter entropy $A_{\mathrm{horizon}}/(4 G_N)$. We frame this paper as a
\emph{consistency note} (MVT2-class): the Araki theorem stands as a new
result of independent interest, while the Stinespring bridge between the
DSSYK chord algebra and the de Sitter crossed product is recorded as a
no-go obstruction for future work.
\end{abstract}

% =====================================================================
\section{Introduction}
\label{sec:intro}
% =====================================================================

\paragraph{The modular flow / generalized entropy programme.}
Algebraic methods have transformed the discussion of generalized entropy in
gravitational subregions. After Witten's observation~\cite{Witten2022} that
gravitational backreaction promotes the type-III$_1$ algebra of bulk
quantum field theory observables to a type-II crossed product, two parallel
programmes have emerged:
(i)~the de~Sitter / static-patch programme of
Chandrasekaran--Longo--Penington--Witten~\cite{CLPW2023} (CLPW), refined by
de~Vuyst--Eccles--H\"{o}hn--Kirklin~\cite{DEHK2025a} (DEHK), in which the
crossed-product trace yields finite observer-dependent entropies, and
(ii)~the generalized second law (GSL) programme of
Faulkner--Speranza~\cite{FaulknerSperanza2024} (FS24), in which monotonicity
of relative entropy under modular inclusions translates to GSL-type
statements for arbitrary cuts of Killing horizons.

Both programmes share a central structural object: the \emph{modular
Hamiltonian} $K_R$ of the dressed (crossed-product) algebra. In the
KMS-thermal de~Sitter setting at $\beta_{\mathrm{dS}} = 2\pi r_{\mathrm{dS}}$
(Bisognano--Wichmann~\cite{BisognanoWichmann1976}; Connes--Rovelli
thermal-time hypothesis~\cite{ConnesRovelli1994}), $K_R$ generates a
genuine one-parameter automorphism group, and its expectation value
$\langle K_R \rangle_\rho$ in a perturbed state $\rho$ controls the
first-law variation of the dressed entropy.

\paragraph{Two questions this note answers.}
\begin{enumerate}
\item Does the Araki relative modular operator
$\Delta_{\rho|\sigma}$~\cite{Araki1976} provide a clean type-II$_\infty$
derivation of $dS_{\mathrm{gen}}/d\tau_R = \langle K_R \rangle_\rho$,
without invoking Petz recovery? \emph{Yes} (Theorem~\ref{thm:araki-rate}).

\item Can the chord Hamiltonian $H_{\mathrm{grav}}$ of the
DSSYK / sine-dilaton holographic dictionary
of Heller--Papalini--Schuhmann~\cite{HPS2024} (HPS), refined by
Heller--Ori--Papalini--Schuhmann--Wang~\cite{HOPSW2025} (HOPSW), be
intertwined with $K_R$ via a CPTP isometry between the chord Hilbert space
$\mathcal{H}_{\mathrm{HPS}}$ and the GNS Hilbert space
$\mathcal{H}_{\mathrm{DEHK}}$? \emph{No} (Proposition~\ref{prop:no-go-spectrum}),
because their spectra have incompatible support.
\end{enumerate}

The negative answer is sharper than a generic ``algebra-type-mismatch''
argument: it pinpoints the obstruction in a concrete, falsifiable way
(bounded versus unbounded continuous spectrum). The thermodynamic match
that survives, $\log Z_{\mathrm{chord}}=\log Z_{\mathrm{DEHK}}+O(1/N)$, is
\emph{not} new (it reduces to the Bekenstein--Hawking entropy of the
cosmological horizon), and we therefore frame this paper at the level of a
consistency note: a clean lemma (Theorem~\ref{thm:araki-rate}) plus a
clean obstruction (Proposition~\ref{prop:no-go-spectrum}) plus an honest
account of what is left open.

\paragraph{Plan.}
Section~\ref{sec:araki} states and proves the Araki-cocycle identity
$dS_{\mathrm{gen}}/d\tau_R = \langle K_R \rangle_\rho$ in
type-II$_\infty$. Section~\ref{sec:stinespring} reviews the
modular-rigidified Stinespring embedding ``$C_{BS}$'' of the chord
Hilbert space into the GNS Hilbert space of $\mathcal{M}_R$, and isolates
the operator-level requirements (P1)--(P3) that any genuine bridge would
have to satisfy. Section~\ref{sec:no-go} states the spectrum-mismatch
obstruction. Section~\ref{sec:survives} records the Krylov-thermodynamic
match that does survive. Section~\ref{sec:discussion} discusses
alternative routes (Faulkner--Hollands~II type-changing recovery, modular
Krylov density-of-states matches) and open problems.

% =====================================================================
\section{Wk2: Araki cocycle in type-II$_\infty$}
\label{sec:araki}
% =====================================================================

\subsection{Setup}
Let $\mathcal{M}$ be a von Neumann algebra in standard form on a Hilbert
space $\mathcal{H}$, and let $\sigma$ be a faithful normal state with
GNS vector $|\Omega_\sigma\rangle$ in the natural positive cone
$\mathcal{P}^\sharp$. For a second faithful normal state $\rho$ with
GNS vector $|\Omega_\rho\rangle\in\mathcal{P}^\sharp$, the
\emph{Araki relative modular operator}~\cite{Araki1976} is
\begin{equation}
   \Delta_{\rho|\sigma}
   \;=\; S_{\rho|\sigma}^{\,*}\,S_{\rho|\sigma},
   \qquad
   S_{\rho|\sigma} : A|\Omega_\sigma\rangle \longmapsto
                      A^*|\Omega_\rho\rangle ,\quad A\in\mathcal{M}.
\end{equation}
Let $\sigma^\sigma_t$ denote the modular automorphism group of $\sigma$
on $\mathcal{M}$, generated by the modular Hamiltonian
$K_R = -\log\Delta_\sigma\geq 0$ (uniform sign convention, cf.~Wk2
memo~\cite{EREPR_Araki_Wk2}). The \emph{Araki relative entropy} is
\begin{equation}
   S(\rho\|\sigma)
   \;=\; -\,\langle\Omega_\rho |\,\log\Delta_{\rho|\sigma}\,|\Omega_\rho\rangle ,
   \label{eq:araki-rel-entropy}
\end{equation}
finite when $\mathrm{supp}(\rho)\subset\mathrm{supp}(\sigma)$ and $+\infty$
otherwise.

\subsection{Type-II$_\infty$ extension via Pusz--Woronowicz}
For the de~Sitter crossed product $\mathcal{M}_R=
\mathcal{A}_{\mathrm{III}_1}\rtimes_{\sigma^\phi}\mathbb{R}$ of CLPW
\cite{CLPW2023} and DEHK~\cite{DEHK2025a}, $\mathcal{M}_R$ is type-II$_\infty$
and admits a normal semifinite faithful (nsf) tracial weight $\tau_{\mathrm{II}}$
(the Haagerup trace). The Araki construction is purely algebraic and the
Tomita--Takesaki theory applies to any standard form, so $\Delta_{\rho|\sigma}$
is well-defined for every pair of nsf weights. The
Pusz--Woronowicz~\cite{PuszWoronowicz1978} variational characterization
\begin{equation}
   S(\rho\|\sigma)
   \;=\; \sup_{\substack{A\in\mathcal{M}\\ A=A^*\geq 0}}
         \bigl\{\,\rho(\log A) - \log\sigma(e^A)\,\bigr\}
\end{equation}
extends to type-II$_\infty$ weights through Haagerup's $L^p$ theory and
the Ohya--Petz~\cite{OhyaPetz1993} formulation in semifinite algebras.

\subsection{The cocycle identity and modular-time derivative}
The relative modular operator satisfies the Connes-cocycle identity (a
direct consequence of $\sigma^\sigma_t$ acting on the GNS data):
\begin{equation}
   \Delta_{\rho|\sigma\circ\sigma^\sigma_{-t}}
   \;=\; \Delta_{\rho|\sigma}\,e^{\,t K_R} ,
   \label{eq:cocycle}
\end{equation}
in the strong-resolvent sense on the joint domain. (See
Bratteli--Robinson~\cite{BR2}, Theorem~5.3.10 for the analyticity strip
$0\leq\mathrm{Im}\,t\leq\beta/2$ of $\Delta^{it}_\sigma$ on KMS states,
and Theorem~5.3.15 for the differentiability of relative entropy along
the modular flow.)

\begin{theorem}[Araki rate, Wk2]
\label{thm:araki-rate}
Let $\mathcal{M}_R$ be a type-II$_\infty$ crossed-product algebra with nsf
weight $\tau_{\mathrm{II}}$, KMS reference state $\sigma$ at inverse
temperature $\beta_{\mathrm{dS}}$ with modular Hamiltonian $K_R$, and let
$\rho$ be a normal state with $\mathrm{supp}(\rho)\subset\mathrm{supp}(\sigma)$
satisfying the regularity condition $\rho \in \mathrm{dom}\,K_R$ (i.e.\
$\langle K_R\rangle_\rho<\infty$). Define the modular-time-evolved
reference $\sigma_{\tau_R}\:=\sigma\circ\sigma^\sigma_{-\tau_R}$. Then
\begin{equation}
   \frac{d}{d\tau_R}\,S\bigl(\rho\|\sigma_{\tau_R}\bigr)\bigg|_{\tau_R=0}
   \;=\; -\,\langle K_R\rangle_\rho ,
   \label{eq:rate-relent}
\end{equation}
and consequently, with the FS24 identification
$S_{\mathrm{gen}}(\rho)=S_{\mathrm{gen}}(\sigma)-S(\rho\|\sigma)+
\text{(KMS boundary terms)}$~\cite{FaulknerSperanza2024},
\begin{equation}
   \boxed{\;\;
   \frac{dS_{\mathrm{gen}}}{d\tau_R}\bigg|_{\rho \text{ fixed}}
   \;=\; +\,\langle K_R\rangle_\rho .
   \;\;}
   \label{eq:Sgen-rate}
\end{equation}
\end{theorem}

\begin{proof}
By~\eqref{eq:cocycle},
$\log\Delta_{\rho|\sigma_{\tau_R}}=\log\Delta_{\rho|\sigma}+\tau_R K_R$ on
the joint analyticity domain. Differentiating the
formula~\eqref{eq:araki-rel-entropy}, exchanging the
$d/d\tau_R$ with the GNS expectation (justified by the dominated
convergence on the strip via the KMS analyticity of
Bratteli--Robinson~\cite{BR2}, Thm.~5.3.10), one obtains
\[
   \frac{d}{d\tau_R}S(\rho\|\sigma_{\tau_R})\bigg|_{\tau_R=0}
   \;=\; -\,\langle\Omega_\rho|\,K_R\,|\Omega_\rho\rangle
   \;=\; -\,\langle K_R\rangle_\rho ,
\]
which is~\eqref{eq:rate-relent}. The conversion to $S_{\mathrm{gen}}$
through the FS24 identification gives~\eqref{eq:Sgen-rate}; the
boundary terms vanish for the KMS reference $\sigma$ on a Killing
horizon (FS24, Sec.~3).
\end{proof}

\begin{remark}
The regularity hypothesis $\rho\in\mathrm{dom}\,K_R$ is non-trivial in
type-II$_\infty$: it requires $\rho$ to lie in the form-domain of $K_R$,
equivalent to a finite \emph{modular energy} expectation. For Boulware
or perturbed-thermal reference states in the SdS static patch, this
condition is plausible but has not been verified state-by-state in the
DEHK literature. We treat~\eqref{eq:Sgen-rate} as a conditional theorem
in this sense.
\end{remark}

\begin{remark}[Type-I sanity check]
In the type-I limit (replace $\mathcal{M}_R$ by $\mathcal{B}(\mathcal{H})$
with $\sigma=\rho_\sigma\in\mathcal{S}_1(\mathcal{H})$ a density matrix),
$\Delta_\sigma=\rho_\sigma\otimes\rho_\sigma^{-1}$ in the standard
representation, $K_R=-\log\rho_\sigma$, and~\eqref{eq:rate-relent}
reduces to the textbook formula
$d/dt\,S(\rho\|\rho_\sigma e^{-tK_R})|_{t=0}=-\mathrm{Tr}(\rho K_R)$
of Uhlmann~\cite{Uhlmann1977} and Lindblad~\cite{Lindblad1974}. The
type-II$_\infty$ statement is the natural extension to a setting where
$\rho_\sigma$ is not trace-class.
\end{remark}

\paragraph{Compatibility with the FS24 GSL.}
Equation~\eqref{eq:Sgen-rate} is consistent with the Faulkner--Speranza
GSL~\cite{FaulknerSperanza2024}: monotonicity of relative entropy under
the modular inclusions of $\mathcal{M}_R$ along the horizon implies
$d/d\tau_R\,S(\rho\|\sigma)\geq 0$ for cuts moving into the future
(modular flow as ``algebraic time''), and Theorem~\ref{thm:araki-rate}
identifies the rate with $\langle K_R\rangle_\rho$. This is the
algebraic analogue of Wall's GSL theorem~\cite{Wall2012} as recast in
the crossed-product language.

% =====================================================================
\section{Wk3: Modular-rigidified Stinespring embedding}
\label{sec:stinespring}
% =====================================================================

We now turn to the question whether $K_R$ admits a holographic dual
identification with the chord Hamiltonian of the DSSYK / sine-dilaton
holographic dictionary.

\subsection{The chord-side data (HPS / HOPSW)}
HPS~\cite{HPS2024} construct, from the chord Hilbert space
$\mathcal{H}_{\mathrm{HPS}}=\ell^2(\mathbb{N}_0)$ with chord-creation
operator $a^\dagger$ and chord-number basis $|n\rangle =
(a^\dagger)^n|0\rangle/\sqrt{[n]_{q}!}$, the gravitational transfer-matrix
Hamiltonian
\begin{equation}
   H_{\mathrm{grav}}
   \;=\; \frac{1}{\sqrt{2|\log q|(1-q^2)}}
         \Bigl[\,-\cos(\hat P) + \tfrac12 e^{i\hat P}e^{-\hat L}\,\Bigr] ,
   \label{eq:Hgrav}
\end{equation}
acting on $\mathcal{H}_{\mathrm{HPS}}$, where $\hat L = 2|\log q|\hat n$
is the chord-length operator and $\hat P$ its conjugate. The spectrum is
\begin{equation}
   \mathrm{spec}(H_{\mathrm{grav}})
   \;=\; \Bigl\{\,E(\theta)=-\cos\theta\,\bigl[2|\log q|(1-q^2)\bigr]^{-1/2}
                : \theta\in[0,\pi]\,\Bigr\} ,
   \label{eq:HPS-spec}
\end{equation}
which is bounded continuous on $[-E_{\max},+E_{\max}]$ with
$E_{\max} = [2|\log q|(1-q^2)]^{-1/2}$. The Plancherel measure is
$d\mu(\theta) = (q^2,e^{\pm 2i\theta};q^2)_\infty\,d\theta$ (HPS, eq.~28).

HOPSW~\cite{HOPSW2025} establish the holographic dictionary
$\hat L \leftrightarrow$ extremal timelike volume in the dS$_2$ geometry
anchored at past/future infinities, and identify $H_{\mathrm{grav}}$ with
the ADM Hamiltonian of sine-dilaton dS$_2$ in the high-energy
$\theta\to\pi$ limit. The Hawking temperature reads
$\beta_{\mathrm{dS}}(\theta) = -2\pi/\theta$ in the dS regime
(HOPSW, eq.~(17) and following).

\subsection{The crossed-product side (CLPW / DEHK)}
On the gravity side, CLPW~\cite{CLPW2023} and DEHK~\cite{DEHK2025a}
construct the crossed-product algebra
$\mathcal{M}_R = \mathcal{A}_R\rtimes_{\sigma^\phi}\mathbb{R}$ acting on
$\mathcal{H}_{\mathrm{DEHK}} = \mathcal{H}_{\mathrm{qft}}\otimes L^2(\mathbb{R})$,
generated by $\mathcal{A}_R$ (the bulk QFT type-III$_1$ algebra in the
static patch), the dressed observables
$e^{i pH_{\mathrm{mod}}}\,a\,e^{-ipH_{\mathrm{mod}}}$, and the clock
momentum $p$. The dressed modular Hamiltonian is (CLPW, eq.~(2.10)
footnote; DEHK, eq.~(3.6))
\begin{equation}
   K_R \;=\; \beta_{\mathrm{dS}}\,H_\xi \;+\; p ,
   \label{eq:KR}
\end{equation}
where $H_\xi$ is the boost generator (continuous unbounded spectrum on
$\mathbb{R}$ since $\mathcal{A}_R$ is type-III$_1$) and $p$ is the clock
momentum (continuous spectrum on $\mathbb{R}$ for an ideal clock). Hence
\begin{equation}
   \mathrm{spec}(K_R) \;=\; \mathbb{R} ,
\end{equation}
unbounded both above and below.

\subsection{The modular-rigidified Stinespring embedding $C_{BS}$}
The natural attempt to bridge the two sides is a CPTP isometry
$\Phi:\mathcal{B}(\mathcal{H}_{\mathrm{HPS}})\to\mathcal{M}_R$ obtained from a
Stinespring isometry
$V:\mathcal{H}_{\mathrm{HPS}}\to\mathcal{H}_{\mathrm{DEHK}}$ with $V|n\rangle =
|\xi_n\rangle$, where $\{|\xi_n\rangle\}_{n\geq 0}$ are analytic vectors
in the natural positive cone $\mathcal{P}^\sharp$ of $(\mathcal{M}_R,\sigma)$
chosen so as to be eigenstates of $\hat L$ under the HOPSW dictionary.
Berry-phase ambiguity in $V$ is rigidified by the
Caputa--Mag\'{a}n--Patramanis--Tonni~\cite{CMPT2023} modular-Lyapunov
universality $\lambda_L^{(\mathrm{mod})}=2\pi$ together with the Connes
cocycle compatibility, fixing $V$ up to a global phase.

The bridge would qualify as a \emph{genuine} algebraic identification if
it satisfied:
\begin{itemize}
\item[(P1)] (state matching) $\Phi(\rho_{\mathrm{HPS}}) = \rho_{\mathrm{DEHK}}$,
\item[(P2)] (modular-flow intertwining) $\Phi\circ\sigma^{(\mathrm{HPS})}_t
            = \sigma^{(\mathrm{DEHK})}_t\circ\Phi$ for all $t\in\mathbb{R}$,
\item[(P3)] (Connes-cocycle equivariance) the cocycle
            $[D\rho_{\mathrm{DEHK}}\!:\!D\sigma_{\mathrm{DEHK}}]_t\in\mathcal{M}_R$
            pulls back to a corresponding cocycle on
            $\mathcal{B}(\mathcal{H}_{\mathrm{HPS}})$.
\end{itemize}
The detailed construction of $V$, the Stinespring map
$\Phi(B) = \pi_R(VBV^*)$, and the Berry-rigidification by the cocycle
identity are reviewed in the predecessor memo~\cite{EREPR_Berry_Wk3}.
What concerns us here is whether (P1)--(P3) are operator-level
attainable.

% =====================================================================
\section{Wk4: Spectrum-mismatch no-go}
\label{sec:no-go}
% =====================================================================

\begin{proposition}[No-go for HPS--DEHK Stinespring intertwining]
\label{prop:no-go-spectrum}
Let $H_{\mathrm{grav}}$ be the chord (transfer-matrix) Hamiltonian on
$\mathcal{H}_{\mathrm{HPS}}$ with spectrum~\eqref{eq:HPS-spec} bounded on
$[-E_{\max},+E_{\max}]$, and let $K_R$ be the dressed modular Hamiltonian
on $\mathcal{H}_{\mathrm{DEHK}}$ with spectrum~$\mathbb{R}$. Then no
isometry $V:\mathcal{H}_{\mathrm{HPS}}\to\mathcal{H}_{\mathrm{DEHK}}$ can
intertwine $H_{\mathrm{grav}}$ with $K_R$ in the sense of
$V\,H_{\mathrm{grav}} = K_R\,V$ on $\mathrm{dom}\,H_{\mathrm{grav}}$.
Consequently, the modular-rigidified Stinespring embedding
$\Phi:\mathcal{B}(\mathcal{H}_{\mathrm{HPS}})\to\mathcal{M}_R$ of
Section~\ref{sec:stinespring} cannot satisfy (P2) in any non-trivial sense,
and (P1), (P3) fail accordingly.
\end{proposition}

\begin{proof}
Suppose, for contradiction, that $V\,H_{\mathrm{grav}} = K_R\,V$ on a
core for $H_{\mathrm{grav}}$. Since $V$ is an isometry,
$V V^*$ is the orthogonal projection onto $\mathrm{ran}\,V \subset
\mathcal{H}_{\mathrm{DEHK}}$, and $V H_{\mathrm{grav}} V^* =
K_R\,VV^*$. Then for any unit vector $\psi\in\mathrm{dom}\,H_{\mathrm{grav}}$,
\[
   \langle V\psi |\,K_R\,|V\psi\rangle
   \;=\; \langle\psi|\,H_{\mathrm{grav}}\,|\psi\rangle
   \;\in\; [-E_{\max},+E_{\max}]
\]
by~\eqref{eq:HPS-spec}. Hence $\mathrm{ran}\,V$ would lie in the
spectral subspace $\chi_{[-E_{\max},+E_{\max}]}(K_R)\,
\mathcal{H}_{\mathrm{DEHK}}$, i.e.\ $K_R$ restricted to $\mathrm{ran}\,V$
would be bounded. But $\mathrm{ran}\,V$ is infinite-dimensional (since
$V$ is an isometry from $\ell^2(\mathbb{N}_0)$), and the restriction of a
self-adjoint operator with continuous spectrum on $\mathbb{R}$ to any
infinite-dimensional invariant subspace $\mathcal{K}$ either inherits
unbounded spectrum (if $\mathcal{K}$ is not contained in a finite-spectral
band) or is unitarily equivalent to a direct sum of bounded operators on
spectral subspaces (if $\mathcal{K}$ \emph{is} contained in a bounded
spectral band of $K_R$). In the second case, $\mathrm{ran}\,V$ lives in
$\chi_{[-E_{\max},+E_{\max}]}(K_R)\,\mathcal{H}_{\mathrm{DEHK}}$, which is
proper and inconsistent with $V$ being an isometry whose image is
modular-flow invariant under $\sigma^{(\mathrm{DEHK})}_t = e^{itK_R}\,
\cdot\,e^{-itK_R}$ (since unbounded $K_R$ generates a flow that escapes
any bounded spectral band on a generic $\mathrm{ran}\,V$ direction).
The contradiction forces (P2) to fail.

(P1) fails because the $\beta_{\mathrm{dS}}$-KMS state for $K_R$ in
$\mathcal{M}_R$ is supported on the full real spectrum, while
$\rho_{\mathrm{HPS}} = \sum_n p_n(\beta)|n\rangle\langle n|$ has support
in the discrete chord basis with thermal weights determined by
$E_n^{\mathrm{chord}}$ from the bounded spectrum~\eqref{eq:HPS-spec};
$\Phi(\rho_{\mathrm{HPS}})$ then has bounded modular energy support
$[-E_{\max},+E_{\max}]$ inside $\mathcal{M}_R$, while the genuine
$\rho_{\mathrm{DEHK}}=e^{-\beta_{\mathrm{dS}}K_R}/Z_{\mathrm{DEHK}}$ does
not. (P3) is a corollary, since the Connes cocycle is built from the
relative modular operator $\Delta_{\rho_{\mathrm{DEHK}}|\sigma_{\mathrm{DEHK}}}$,
which in turn requires (P1).
\end{proof}

\begin{remark}[Categorical character of the obstruction]
The obstruction is not a small-parameter / loop-correction artifact: it is
\emph{topological} on the spectrum. Even with arbitrarily small $|\log q|$,
$E_{\max}\to\infty$ but $E(\theta)$ remains bounded for each fixed $q\in
(0,1)$, while $K_R$ is unbounded on $\mathbb{R}$ for every $\beta_{\mathrm{dS}}>0$.
The two spectra never agree as sets.
\end{remark}

\begin{remark}[On the type-I vs type-II$_\infty$ algebra mismatch]
A coarser way to state the obstruction is that
$\mathcal{B}(\mathcal{H}_{\mathrm{HPS}})$ is type-I
(the chord algebra is finite-dimensional at any chord-truncation level
$N_{\mathrm{chord}}$ and infinite-dimensional but separable in the limit;
its von Neumann completion is $\mathcal{B}(\mathcal{H}_{\mathrm{HPS}})$,
type-I), while $\mathcal{M}_R$ is type-II$_\infty$. Type-changing CPTP
maps between general von Neumann algebras do exist (Faulkner--Hollands~II
\cite{FaulknerHollandsII2020}) and even admit FH-II approximate recovery
channels under small relative-entropy perturbations. What
Proposition~\ref{prop:no-go-spectrum} shows is that the
\emph{spectral obstruction} is sharper than the algebra-type obstruction:
even within the type-changing FH-II framework, the candidate spectra
$E(\theta)$ and $\mathrm{spec}(K_R)$ are incompatible.
\end{remark}

% =====================================================================
\section{What survives: Krylov-thermodynamic match}
\label{sec:survives}
% =====================================================================

The negative content of Section~\ref{sec:no-go} is partial: while no
operator-level isometry intertwines $H_{\mathrm{grav}}$ with $K_R$, the
\emph{thermodynamic} content of the two systems agrees at semiclassical
order.

\begin{proposition}[Krylov--thermodynamic match]
\label{prop:therm-match}
In the regime $q\to 1^-$, $\theta\to\pi$ (the dS large-$N$ regime of
HPS--HOPSW), the partition functions
$Z_{\mathrm{chord}}(\beta) := \int_0^\pi (q^2,e^{\pm 2i\theta};q^2)_\infty\,
e^{-\beta E(\theta)}\,d\theta$ and $Z_{\mathrm{DEHK}}(\beta) :=
\tau_{\mathrm{II}}(e^{-\beta K_R})$ satisfy, at the de~Sitter inverse
temperature $\beta_{\mathrm{dS}}$:
\begin{equation}
   \log Z_{\mathrm{chord}}(\beta_{\mathrm{dS}})
   \;=\; \log Z_{\mathrm{DEHK}}(\beta_{\mathrm{dS}}) \;+\; O(1/N) ,
   \label{eq:therm-match}
\end{equation}
both reproducing the Bekenstein--Hawking entropy
$S_{\mathrm{BH}} = A_{\mathrm{horizon}}/(4 G_N)$ at leading order.
\end{proposition}

\begin{proof}[Sketch]
This is essentially a re-statement of (i)~the standard Bekenstein--Hawking
entropy of the cosmological horizon in CLPW/DEHK at leading
$1/G_N$~\cite{CLPW2023,DEHK2025a}, and (ii)~HOPSW eq.~(27) (chord-Krylov
disk partition function $\to$ dS$_2$ Bekenstein--Hawking
entropy)~\cite{HOPSW2025}, both regulated through the same
$\beta_{\mathrm{dS}}$-KMS thermal data on the cosmological horizon.
At higher orders, the $1/N$ corrections differ: chord-side corrections
follow from the q-Hermite Plancherel measure expansion; DEHK-side
corrections follow from the loop expansion around the saddle of
$\tau_{\mathrm{II}}(e^{-\beta K_R})$. We do not claim agreement beyond
leading order.
\end{proof}

\begin{remark}
Proposition~\ref{prop:therm-match} is \emph{not} a new result: the
Bekenstein--Hawking entropy of the cosmological horizon is recovered by
\emph{every} consistent dS holography proposal at semiclassical order.
We include it here only to record \emph{exactly} what part of the
attempted bridge survives the no-go in
Proposition~\ref{prop:no-go-spectrum}: the partition-function
thermodynamics, but not the underlying operator content.
\end{remark}

\begin{remark}[Krylov-asymptotic match]
A second weaker statement that survives is a Lanczos-asymptotic match.
The HPS chord algebra has Lanczos coefficients
$b_n = \sqrt{[n]_q(1-q^n)}\sim\alpha n + \gamma$ at large $n$ (linear
growth, signature of $\lambda_L = 2\pi T$). For $K_R$, Vardian's modular
Krylov procedure~\cite{Vardian2026} (in an AdS/CFT setting; transfer to
dS is by methodological analogy) yields modular Lanczos coefficients
$b_n^{\mathrm{mod}}$ also growing linearly at late modular time with
$\lambda_L^{(\mathrm{mod})} = 2\pi$, the
Caputa--Mag\'{a}n--Patramanis--Tonni
universality~\cite{CMPT2023}. Both spectral functions exhibit the same
asymptotic Lyapunov exponent, but as different Jacobi matrices with
different short-distance / low-$n$ structure. This is a Krylov-asymptotic
match, not an operator match, and is consistent with
Proposition~\ref{prop:no-go-spectrum}.
\end{remark}

% =====================================================================
\section{Discussion and open problems}
\label{sec:discussion}
% =====================================================================

\paragraph{What is genuinely new.} Theorem~\ref{thm:araki-rate} is, as
far as we are aware, the first explicit statement of the
Araki-cocycle identity $dS_{\mathrm{gen}}/d\tau_R = \langle K_R\rangle_\rho$
in the type-II$_\infty$ crossed-product setting of CLPW/DEHK. It
complements the FS24 GSL~\cite{FaulknerSperanza2024} by giving an
explicit \emph{rate} formula in terms of the modular-energy
expectation, and makes the modular-time first law
$\delta S_{\mathrm{gen}} = \beta_{\mathrm{dS}}\,\delta\langle K_R\rangle$
algebraically explicit on the type-II$_\infty$ trace.

\paragraph{What is genuinely refuted.} The natural attempt to identify
$\langle K_R\rangle_\rho$ with chord-Krylov complexity $C_K$ via a
Stinespring isometry between the DSSYK chord Hilbert space and the GNS
Hilbert space of $\mathcal{M}_R$ is refuted at the operator level by
Proposition~\ref{prop:no-go-spectrum}, because the spectra of the
chord Hamiltonian and $K_R$ have incompatible support (bounded vs
unbounded). What survives is the thermodynamic content of
Proposition~\ref{prop:therm-match}, i.e.\ the standard
Bekenstein--Hawking dS entropy match, and the Krylov-asymptotic
universal $\lambda_L = 2\pi$ shared between the two Lanczos sequences.

\paragraph{Alternative routes.} Three logically distinct alternatives
to the Stinespring intertwining route remain open:

\begin{enumerate}
\item \emph{Type-changing FH-II recovery.}\;
Faulkner--Hollands~II~\cite{FaulknerHollandsII2020} (and the
predecessor FHSW~\cite{FHSW2020}) construct universal approximate
recovery channels for 2-positive maps between general von~Neumann
algebras, including the type-changing case. Their bound
$S(\rho\|\sigma) - S(\Phi(\rho)\|\Phi(\sigma)) \geq -2\log F(\rho,
R\circ\Phi(\rho))$ controls relative-entropy loss without requiring
spectral matching of the modular Hamiltonians. A reformulation of
the bridge problem in FH-II language might give an
\emph{approximate-recovery} statement that escapes the
spectrum-mismatch no-go, but at the cost of giving up exact intertwining.

\item \emph{Bounded function of $K_R$.}\;
A bounded function $F(K_R)$ (e.g.\ $F(K_R) = (2/\pi)\arctan
(\beta_{\mathrm{dS}} K_R/2)$) has spectrum in $(-1,+1)$ and could
in principle match a rescaled $H_{\mathrm{grav}}$. This is an ad hoc
construction without independent physical motivation; we mention it
for completeness only.

\item \emph{Modular density-of-states matching.}\;
A genuine result at the level of the spectral measure of $K_R$, namely
that the chord-Krylov density of states $\rho(\theta)$ matches a
suitable projection of the spectral measure of $K_R$ in a controlled
high-energy / late-modular-time limit, would yield a v6.2-style consistency
note rather than a theorem. This route uses the
CMPT~\cite{CMPT2023} modular-Lyapunov universality and the modular
Krylov approach of Vardian~\cite{Vardian2026} (with the caveat that
Vardian works in AdS/CFT and the dS extension requires an independent
construction).
\end{enumerate}

\paragraph{Closing assessment.}
The ER=EPR programme in the algebraic crossed-product setting yields,
at the rigorous mathematical level, the Araki-cocycle rate formula
\eqref{eq:Sgen-rate} as a genuine new theorem (Theorem~\ref{thm:araki-rate}),
and a clean operator-level no-go for the chord--$K_R$ bridge
(Proposition~\ref{prop:no-go-spectrum}). The survival is the standard
Bekenstein--Hawking thermodynamic match (Proposition~\ref{prop:therm-match}),
which was already known from independent considerations.

We do not claim that the ER=EPR / chord-vs-modular-complexity programme
is closed, only that the most natural ``Stinespring isometry'' bridge is
ruled out, and that future work must use either type-changing approximate
recovery in the FH-II sense, a bounded-functional-calculus construction,
or a coarser density-of-states / Lanczos-asymptotic matching at the level
of spectral measures rather than operator identifications.

% =====================================================================
\section*{Acknowledgements}
% =====================================================================
We thank the algebraic gravity community for the foundational results in
\cite{Witten2022,CLPW2023,FaulknerSperanza2024,DEHK2025a,HPS2024,HOPSW2025}
and the relative-entropy / Connes-cocycle community for the foundational
results in \cite{Araki1976,PuszWoronowicz1978,BR2,Uhlmann1977,Lindblad1974}.
The framing of this note as a ``consistency note'' rather than a
full theorem reflects the obstructions identified in
Sections~\ref{sec:no-go}, and we thank the predecessor portfolio (Wk1
Petz route, Wk2 Araki memo, Wk3 Berry--Stinespring memo, Wk4 spectrum
memo) for sharpening the questions.

% =====================================================================
% Bibliography
% =====================================================================
\begin{thebibliography}{99}

\bibitem{Witten2022}
E.~Witten,
``Gravity and the crossed product,''
\textit{JHEP} \textbf{10} (2022) 008,
\href{https://doi.org/10.1007/JHEP10(2022)008}{doi:10.1007/JHEP10(2022)008},
\href{https://arxiv.org/abs/2112.12828}{arXiv:2112.12828 [hep-th]}.

\bibitem{CLPW2023}
V.~Chandrasekaran, R.~Longo, G.~Penington, and E.~Witten,
``An algebra of observables for de~Sitter space,''
\textit{JHEP} \textbf{02} (2023) 082,
\href{https://doi.org/10.1007/JHEP02(2023)082}{doi:10.1007/JHEP02(2023)082},
\href{https://arxiv.org/abs/2206.10780}{arXiv:2206.10780 [hep-th]}.

\bibitem{DEHK2025a}
J.~De~Vuyst, S.~Eccles, P.~A.~H\"{o}hn, and J.~Kirklin,
``Crossed products and quantum reference frames: on the observer-dependence
of gravitational entropy,''
\textit{JHEP} \textbf{07} (2025) 063,
\href{https://doi.org/10.1007/JHEP07(2025)063}{doi:10.1007/JHEP07(2025)063},
\href{https://arxiv.org/abs/2412.15502}{arXiv:2412.15502 [hep-th]}.

\bibitem{FaulknerSperanza2024}
T.~Faulkner and A.~J.~Speranza,
``Gravitational algebras and the generalized second law,''
\textit{JHEP} \textbf{11} (2024) 099,
\href{https://doi.org/10.1007/JHEP11(2024)099}{doi:10.1007/JHEP11(2024)099},
\href{https://arxiv.org/abs/2405.00847}{arXiv:2405.00847 [hep-th]}.

\bibitem{HPS2024}
M.~P.~Heller, J.~Papalini, and T.~Schuhmann,
``Krylov spread complexity as holographic complexity beyond JT gravity,''
\textit{Phys. Rev. Lett.} \textbf{135} (2025) 151602,
\href{https://doi.org/10.1103/PhysRevLett.135.151602}{doi:10.1103/PhysRevLett.135.151602},
\href{https://arxiv.org/abs/2412.17785}{arXiv:2412.17785 [hep-th]}.

\bibitem{HOPSW2025}
M.~P.~Heller, F.~Ori, J.~Papalini, T.~Schuhmann, and M.-T.~Wang,
``De Sitter holographic complexity from Krylov complexity in DSSYK,''
\href{https://arxiv.org/abs/2510.13986}{arXiv:2510.13986 [hep-th]} (2025).

\bibitem{Araki1976}
H.~Araki,
``Relative entropy of states of von Neumann algebras,''
\textit{Publ. RIMS, Kyoto Univ.} \textbf{11} (1976) 809--833;
follow-up: \textit{Comm. Math. Phys.} \textbf{66} (1977) 83.

\bibitem{PuszWoronowicz1978}
W.~Pusz and S.~L.~Woronowicz,
``Form convex functions and the WYDL and other inequalities,''
\textit{Lett. Math. Phys.} \textbf{2} (1978) 505--512;
related: \textit{Rep. Math. Phys.} \textbf{8} (1975) 159.
%% Pusz-Woronowicz variational characterisation of relative entropy;
%% canonical reference for the form-convex / variational formulation.

\bibitem{BR2}
O.~Bratteli and D.~W.~Robinson,
\textit{Operator Algebras and Quantum Statistical Mechanics, Vol.~2},
2nd ed., Springer, 1997.
%% Esp. Thm. 5.3.10 (KMS analyticity strip), Thm. 5.3.15
%% (modular flow differentiability of relative entropy).

\bibitem{OhyaPetz1993}
M.~Ohya and D.~Petz,
\textit{Quantum Entropy and Its Use},
Springer, 1993.

\bibitem{Uhlmann1977}
A.~Uhlmann,
``Relative entropy and the Wigner--Yanase--Dyson--Lieb concavity in an
interpolation theory,''
\textit{Comm. Math. Phys.} \textbf{54} (1977) 21.

\bibitem{Lindblad1974}
G.~Lindblad,
``Expectations and entropy inequalities for finite quantum systems,''
\textit{Comm. Math. Phys.} \textbf{39} (1974) 111.

\bibitem{Wall2012}
A.~C.~Wall,
``A proof of the generalized second law for rapidly changing fields and
arbitrary horizon slices,''
\textit{Phys. Rev. D} \textbf{85} (2012) 104049,
\href{https://doi.org/10.1103/PhysRevD.85.104049}{doi:10.1103/PhysRevD.85.104049},
\href{https://arxiv.org/abs/1105.3445}{arXiv:1105.3445 [gr-qc]}.

\bibitem{BisognanoWichmann1976}
J.~J.~Bisognano and E.~H.~Wichmann,
``On the duality condition for a Hermitian scalar field,''
\textit{J. Math. Phys.} \textbf{17} (1976) 303.

\bibitem{ConnesRovelli1994}
A.~Connes and C.~Rovelli,
``Von Neumann algebra automorphisms and time-thermodynamics relation in
generally covariant quantum theories,''
\textit{Class. Quantum Grav.} \textbf{11} (1994) 2899--2917,
\href{https://doi.org/10.1088/0264-9381/11/12/007}{doi:10.1088/0264-9381/11/12/007}.

\bibitem{CMPT2023}
P.~Caputa, J.~M.~Mag\'{a}n, D.~Patramanis, and E.~Tonni,
``Krylov complexity of modular Hamiltonian evolution,''
\textit{Phys. Rev. D} \textbf{109} (2024) 086004,
\href{https://doi.org/10.1103/PhysRevD.109.086004}{doi:10.1103/PhysRevD.109.086004},
\href{https://arxiv.org/abs/2306.14732}{arXiv:2306.14732 [hep-th]}.

\bibitem{Vardian2026}
N.~Vardian,
``Modular Krylov complexity as a boundary probe of area operator and
entanglement islands,''
\href{https://arxiv.org/abs/2602.02675}{arXiv:2602.02675 [hep-th]} (2026).
%% Note: Vardian's paper is set in AdS/CFT with QES and entanglement
%% islands; transfer to the dS / DSSYK setting of this paper is by
%% methodological analogy only.

\bibitem{FHSW2020}
T.~Faulkner, S.~Hollands, B.~Swingle, and Y.~Wang,
``Approximate recovery and relative entropy I: general von Neumann
subalgebras,''
\textit{Comm. Math. Phys.} \textbf{389} (2022) 349--397,
\href{https://arxiv.org/abs/2006.08002}{arXiv:2006.08002 [hep-th]}.

\bibitem{FaulknerHollandsII2020}
T.~Faulkner and S.~Hollands,
``Approximate recoverability and relative entropy II: 2-positive channels
of general von Neumann algebras,''
\textit{Lett. Math. Phys.} \textbf{112} (2022) 26,
\href{https://arxiv.org/abs/2010.05513}{arXiv:2010.05513 [hep-th]}.

\bibitem{EREPR_Araki_Wk2}
[ECI author(s)],
``Araki Relative Modular Operator in Type-II$_\infty$:
Week 2 Scoping Memo (ER=EPR Reopen, A13-2),''
internal portfolio memo (2026-05-04).

\bibitem{EREPR_Berry_Wk3}
[ECI author(s)],
``Berry--Stinespring channel $C_{BS}$ for HPS (type-I) $\leftrightarrow$
DEHK (type-II$_\infty$): Week 3 memo (ER=EPR Reopen, A6),''
internal portfolio memo (2026-05-05).

\end{thebibliography}

\end{document}
