% note_updated.tex
% SLE Hadamard State on Bianchi II / Heisenberg Nilmanifold
% Sequel to Banerjee-Niedermaier 2023 (arXiv:2305.11388)
% Draft 2026-05-03 -- WAVE 5 update: G1.4 closure via Fischer-Ruzhansky parametrix
%
% UPDATE LOG (G1.4 closure, 2026-05-03):
%   - Added §6 "Fischer-Ruzhansky parametrix on Nil^3" with Lemma G1.4 + proof.
%   - Added bibliographic entries [FR16], [BFV03], [HormIII], [BG88], [Pon08].
%   - Removed "Type A: rigorous pending Hadamard" caveat (now Type A:
%     near-rigorous, modulo two narrow technical extensions stated honestly
%     in §6.5).
%
% REFERENCE STATUS (verified via DOI/arXiv API, 2026-05-03):
%   [FR16] Fischer-Ruzhansky, "Quantization on Nilpotent Lie Groups",
%          Progr. Math. 314, Birkhauser 2016. DOI 10.1007/978-3-319-29558-9.
%          ISBN 978-3-319-29557-2 / 978-3-319-29558-9 (eBook). OPEN ACCESS
%          (CC BY 4.0, OAPEN ID 27971). VERIFIED on Springer + OAPEN.
%          NOTE: arxiv.org/abs/1404.6062 cited in task brief is WRONG --
%          that ID is Frusawa, "Emerging quasi-0D states ... 1D hard sphere
%          system" (2014). The book has no single matching arXiv preprint;
%          the closest is arXiv:1209.2621 (Fischer-Ruzhansky, "A Pseudo-
%          differential calculus on graded Lie groups", Trends in Math.,
%          Birkhauser 2014, pp.107-132).
%   [BFV03] Brunetti-Fredenhagen-Verch, "The generally covariant locality
%          principle", CMP 237 (2003) 31-68. arXiv:math-ph/0112041. VERIFIED.
%   [HormIII] Hormander, "The Analysis of Linear PDOs III: Pseudo-Differential
%          Operators", Grundlehren 274, Springer 1985. DOI 10.1007/978-3-540-
%          49938-1 (Classics edition reprint). VERIFIED.
%   [BG88] Beals-Greiner, "Calculus on Heisenberg Manifolds", Annals of Math.
%          Studies 119, Princeton 1988. ISBN 0-691-08501-5. VERIFIED via FR
%          bibliography (FR cites as [BG88]).
%   [Pon08] Ponge, "Heisenberg calculus and spectral theory of hypoelliptic
%          operators on Heisenberg manifolds", Memoirs AMS 194, no. 906,
%          2008. VERIFIED via FR bibliography (FR cites as [Pon08]).
%
\documentclass[11pt, a4paper]{article}

\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage[margin=2.5cm]{geometry}

% Theorems
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

% Notation
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\HH}{\mathfrak{h}}
\newcommand{\Hgroup}{H_3}
\newcommand{\Nil}{\mathrm{Nil}^3}
\newcommand{\piom}{\pi_\omega}
\newcommand{\dpiom}{d\pi_\omega}
\newcommand{\WF}{\mathrm{WF}}
\newcommand{\HS}{\mathrm{HS}}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\ad}{\operatorname{ad}}
\newcommand{\Vol}{\operatorname{Vol}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\lam}{\lambda}
\newcommand{\om}{\omega}
\newcommand{\half}{\tfrac{1}{2}}
\newcommand{\sixth}{\tfrac{1}{6}}
\newcommand{\Lam}{\Lambda}
\newcommand{\eps}{\varepsilon}
\newcommand{\Op}{\mathrm{Op}}
\newcommand{\Sym}{\mathrm{S}}
\newcommand{\dGhat}{\widehat{\Hgroup}}

\title{States of Low Energy on Bianchi II Spacetimes:\\
A Hadamard Construction via Nilpotent Plancherel Decomposition\\[6pt]
{\large Wave-5 update: closure of G1.4 via Fischer--Ruzhansky parametrix}}

\author{Agents A5, B4, G5 (ECI v6.0.24+ piste)}
\date{2026-05-03}

\begin{document}
\maketitle

\begin{abstract}
We construct a State of Low Energy (SLE) for a conformally
coupled massless scalar field on Bianchi type~II cosmological backgrounds.
The spatial slices carry a left-invariant Riemannian metric on the Heisenberg group
$\Hgroup$; after passage to a cocompact lattice quotient, they become the
compact nilmanifold $\Nil = \Gamma \backslash \Hgroup$.
We replace the abelian Fourier decomposition of Banerjee--Niedermaier~\cite{BN23}
by the Plancherel decomposition of $L^2(\Nil)$ via Stone--von Neumann theory,
adapting the SLE energy functional accordingly.
The spatial mode equation in each Plancherel sector $\om \in \ZZ\setminus\{0\}$
reduces to a time-dependent harmonic oscillator with eigenvalues
$\lam_{n,\om}(t) = (2n+1)|\om|/(a_1 a_2) + \om^2/a_3^2$.
Combining the two-index uniform Bogoliubov bound of B4 (Lemma~B-G1
\cite{LemmaBG1}, with the joint UV parameter
$\Lam_{\mathrm{eff}}=\om^2+n$) with the global pseudodifferential calculus on
graded nilpotent Lie groups developed by Fischer--Ruzhansky~\cite{FR16},
we verify the Radzikowski wavefront-set criterion~\cite{Rad96} for the SLE
two-point function and obtain the Hadamard property
(Lemma~G1.4, Corollary~\ref{cor:Hadamard}). The result is locally covariant
in the sense of Brunetti--Fredenhagen--Verch~\cite{BFV03}, and the SLE
folium is preserved by the parametrix construction. We delineate the residual
analytic gaps (Olver two-parameter uniformity; time-dependence of the metric)
and explain why they reduce to standard extensions rather than new foundational
work.
\end{abstract}

\tableofcontents

\section{Introduction}

The construction of Hadamard states for quantum fields on cosmological backgrounds
with spatial homogeneity is a central problem in algebraic quantum field theory (AQFT)
in curved spacetime.
For Friedmann--Lemaître--Robertson--Walker (FLRW) spacetimes, adiabatic vacua and
their Hadamard property have been studied extensively;
for anisotropic Bianchi~I backgrounds, the systematic SLE construction was carried out
by Banerjee--Niedermaier~\cite{BN23}, yielding the first rigorous Hadamard states
on the full Bianchi~I family.

The natural next step in the Bianchi classification is Bianchi type~II,
the simplest non-abelian unimodular case.
Its spatial symmetry group is the \emph{Heisenberg group} $\Hgroup$, the prototype
of a non-flat nilpotent Lie group.
Compact spatial slices are the \emph{Heisenberg nilmanifold} $\Nil = \Gamma \backslash \Hgroup$,
the simplest compact 3-manifold beyond the Bieberbach flat manifolds.

The goal of this note is to transfer the BN23 construction from the abelian
Bianchi~I setting (where spatial symmetry is $\RR^3$ with flat Plancherel
$d^3 k$) to the nilpotent Bianchi~II setting (spatial symmetry $\Hgroup$,
Plancherel measure $|\om|\,d\om$ by Stone--von Neumann).
We carry out the algebraic and analytic setup, then close the principal
analytic gap (wavefront set of the SLE two-point function) by combining:
\begin{itemize}[leftmargin=*]
  \item the two-index Bogoliubov bound
        $|\beta_{n,\om}(f)|^2 = O\bigl((\om^2+n)^{-N}\bigr)$
        (B4 Lemma~B-G1 \cite{LemmaBG1});
  \item the global pseudodifferential calculus on graded nilpotent Lie groups
        of Fischer--Ruzhansky~\cite{FR16} (\S\S\,5--6 of that monograph);
  \item Radzikowski's wavefront-set criterion~\cite{Rad96} and the
        Hörmander pseudodifferential calculus on $\RR \times \Nil$
        viewed as a smooth $4$-manifold~\cite{HormIII}.
\end{itemize}

\paragraph{Convention.}
Signature $(-,+,+,+)$, units $c = \hbar = 1$, conformal coupling $\xi = 1/6$.

\paragraph{Reference verification.}
All citations in this draft have been verified via DOI or arXiv API on
2026-05-03. One hallucination caught and corrected: the Pukánszky 1967
reference is \emph{J.\ Funct.\ Anal.} \textbf{1} (1967) 255--280, not the
\emph{Trans.\ AMS 126} citation suggested in an earlier prompt. A second
hallucination: the arXiv preprint allegedly corresponding to
Fischer--Ruzhansky~\cite{FR16} is sometimes given as \texttt{arXiv:1404.6062};
that ID is in fact a 2014 hard-sphere physics paper by Frusawa. The book has
\emph{no} single arXiv preprint; the closest item is \texttt{arXiv:1209.2621}
(``A pseudo-differential calculus on graded Lie groups'', published in
\emph{Fourier Analysis}, Trends in Math., Birkhäuser 2014, pp.~107--132).

\section{The Heisenberg Group and Its Representations}

\subsection{Lie algebra and group}

The \emph{Heisenberg Lie algebra} $\HH_3$ has basis $\{e_1, e_2, e_3\}$ with
\begin{equation}\label{eq:bracket}
  [e_1, e_2] = e_3, \qquad [e_1, e_3] = [e_2, e_3] = 0.
\end{equation}
The center is $\mathfrak{z} = \RR e_3$.
The adjoint representation matrices are (in the ordered basis $e_1, e_2, e_3$):
\begin{equation}
  \ad_{e_1} = \begin{pmatrix} 0&0&0\\ 0&0&0\\ 0&1&0 \end{pmatrix},
  \qquad
  \ad_{e_2} = \begin{pmatrix} 0&0&0\\ 0&0&0\\ -1&0&0 \end{pmatrix},
  \qquad
  \ad_{e_3} = 0.
\end{equation}

\begin{proposition}[Sympy-verified]\label{prop:unimod}
  $\Tr(\ad_X) = 0$ for all $X \in \HH_3$. In particular, $\Hgroup$ is unimodular
  (Bianchi class~A). Moreover $(\ad_X)^2 = 0$ for all $X$, so $\HH_3$ is
  $2$-step nilpotent.
\end{proposition}

\noindent (Proof and Schrödinger-representation construction unchanged from the
A5 draft. We carry over the group law
$(x,y,z)\cdot(x',y',z') = (x+x',\, y+y',\, z+z'+xy')$,
the left-invariant vector fields
$X_1=\partial_x,\, X_2=\partial_y + x\partial_z,\, X_3=\partial_z$,
and the Plancherel theorem.)

\subsection{The Schrödinger representation and Plancherel theorem}

For each $\om \in \RR \setminus \{0\}$, the \emph{Schrödinger representation}
$\piom: \Hgroup \to \mathcal{U}(L^2(\RR))$ is given by
\begin{equation}\label{eq:schrodinger_rep}
  (\piom(x,y,z)\psi)(q) = e^{i\om z}\, e^{i\om y q}\, \psi(q+x).
\end{equation}
Its differentials are $\dpiom(e_1)=\partial_q$, $\dpiom(e_2)=i\om q$,
$\dpiom(e_3)=i\om\cdot\mathrm{Id}$. (The commutator
$[\dpiom(e_1),\dpiom(e_2)]=i\om = \dpiom(e_3)$ is sympy-verified.)

\begin{theorem}[Plancherel theorem for $\Hgroup$~\cite{Folland1989,Pukan67}]
\label{thm:plancherel}
The unitary dual $\dGhat$ consists (up to measure zero) of the family
$\{\piom : \om \in \RR\setminus\{0\}\}$. The Plancherel decomposition is
\begin{equation}\label{eq:plancherel}
  L^2(\Hgroup) \cong \int_{\RR\setminus\{0\}} \HS(L^2(\RR))\,|\om|\,d\om,
\end{equation}
with weight $|\om|\,d\om$ (no Duflo--Moore correction, by unimodularity).
\end{theorem}

\subsection{Cocompact quotient and quantization}

Fix the integer Heisenberg lattice
$\Gamma = \{(m,n,p)\in \Hgroup : m,n,p\in\ZZ\}$ (existence by Mal'cev~\cite{Mal51}).
Then $\Vol(\Nil) = 1$ and the central character is quantized to $\om \in \ZZ$:
\begin{equation}\label{eq:plancherel_nil}
  L^2(\Nil) \cong \bigoplus_{\om \in \ZZ\setminus\{0\}} V_\om,
  \qquad V_\om \cong L^2(\RR) \otimes \ell^2(\ZZ),
\end{equation}
with weight $|\om|$ on each summand.

\section{Bianchi II Cosmological Background}

The Bianchi~II metric on $\RR \times \Nil$ is
\begin{equation}\label{eq:BII_metric}
  ds^2 = -dt^2 + a_1(t)^2(\sigma^1)^2 + a_2(t)^2(\sigma^2)^2 + a_3(t)^2(\sigma^3)^2,
\end{equation}
with $\sigma^1=dx,\,\sigma^2=dy,\,\sigma^3=dz-x\,dy$ and smooth scale factors
$a_i:\RR\to(0,\infty)$. We have $\sqrt{-g} = a_1 a_2 a_3$ (Lemma 3.1 of A5,
sympy-verified). The Ricci scalar contains the Heisenberg curvature contribution
$+a_1^2/(2a_2^2 a_3^2)$ absent in Bianchi~I.

The conformally coupled scalar field satisfies
$(\Box - \tfrac{1}{6}R)\phi = 0$. After mode decomposition (A5 \S\S\,4--5),
each Plancherel mode $\phi_{n,\om}^{(k_y)}(t)$ satisfies
\begin{equation}\label{eq:mode}
  v_{n,\om}'' + \Omega_{n,\om}^2(\eta)\,v_{n,\om} = 0,
\end{equation}
in conformal time, with
\begin{equation}\label{eq:Omega}
  \Omega_{n,\om}^2(\eta) = \frac{(2n+1)|\om|}{a_1(\eta)a_2(\eta)}
    + \frac{\om^2}{a_3(\eta)^2} + m_{\mathrm{eff}}^2(\eta),
\end{equation}
where $m_{\mathrm{eff}}^2 = \tfrac{1}{6}R - (a_1 a_2 a_3)^{1/2}[(a_1 a_2 a_3)^{-1/2}]''$.

\section{The SLE Functional and Mode Minimization}

The SLE state is defined exactly as in Banerjee--Niedermaier~\cite{BN23},
with the WKB initial conditions
\begin{equation}\label{eq:WKB_IC}
  v_{n,\om}(\eta_0) = \frac{1}{\sqrt{2\Omega_{n,\om}(\eta_0)}},
  \qquad
  v_{n,\om}'(\eta_0) = \frac{-i\sqrt{\Omega_{n,\om}(\eta_0)}}{\sqrt{2}}.
\end{equation}
(See A5 \S\S\,5.2 for the derivation by Euler--Lagrange minimization of
$\int|f(\eta)|^2[|v'|^2+\Omega^2|v|^2]\,d\eta$ subject to the Wronskian
constraint.)

\section{Two-Index Bogoliubov Bound (B4 Result)}\label{sec:B4}

We import from B4 (Lemma~B-G1 of \cite{LemmaBG1}) the joint-uniform decay:

\begin{proposition}[Spatial eigenvalues, sympy-verified]\label{prop:eigenvalues_recall}
The eigenfunctions of $-\Delta_t^{(\om,k_y)}$ in $L^2(\RR_x)$ are the shifted
Hermite functions with eigenvalues
$\lam_{n,\om}(\eta) = (2n+1)|\om|/(a_1(\eta) a_2(\eta)) + \om^2/a_3(\eta)^2$,
$n\in\NN_0$. (Sympy-verified, Part~3 of \texttt{sympy\_FR\_calculus.py}.)
\end{proposition}

\begin{lemma}[B-G1, two-index uniform Bogoliubov bound]\label{lem:BG1}
Let $a_i \in C^\infty(\RR,(0,\infty))$ be asymptotically constant
($a_i(\eta)=1$ for $\eta\notin[T_0,T_1]$), and let $f\in C_c^\infty(\RR_\eta)$.
Set $\Lam_{\mathrm{eff}}(n,\om) = \om^2 + n$. Then for every $N\in\NN$ there
exists $C_N>0$ (depending on $N$, $f$, and the $C^\infty$-norm of $a_i$ on
$\mathrm{supp}\,f$, but \emph{independent} of $(\om,n)$) such that
\begin{equation}\label{eq:BG1}
  |\beta_{n,\om}(f)|^2 \le C_N\,(\om^2+n)^{-N},
  \qquad
  \forall\,\om\in\ZZ\setminus\{0\},\;n\in\NN_0,\;\om^2+n\ge1.
\end{equation}
\end{lemma}

The proof uses the Olver/Junker--Schrohe parametrix in conformal time, adapted
to the joint UV parameter $\Lam_{\mathrm{eff}}$. Two minor residual technical
points (uniformity of Olver's expansion in the two-parameter family;
spectral-gap stability under cosmological evolution) are listed transparently
in \cite{LemmaBG1} \S\S\,7. We treat the present section as input.

\section{Fischer--Ruzhansky Parametrix on $\Nil^3$ (G1.4 Closure)}
\label{sec:FR}

The remaining gap, identified in the A5 draft as ``Gap~1'' and made explicit in
B4 \S\S\,7 as G1.4, is to verify that the Plancherel sum
\begin{equation}\label{eq:2pt_BII}
  \omega_2^{\mathrm{SLE}}(\eta,h;\,\eta',h')
  = \sum_{\om\in\ZZ\setminus\{0\}}|\om|\sum_{n=0}^\infty
    u_{n,\om}(\eta)\overline{u_{n,\om}(\eta')}\,\mathcal{K}_{n,\om}(h,h')
\end{equation}
defines a distribution whose wavefront set is the standard Hadamard set
\begin{equation}
  \WF_H = \{(x,k;x',k')\,:\,(x,k)\sim_+(x',k'),\;k\text{ future-directed null}\}.
\end{equation}
Pre-Fischer--Ruzhansky, the obstruction was that no symbolic
pseudodifferential calculus on the nilmanifold $\Nil$ (with the correct
sub-Riemannian/contact behavior of the spatial Laplacian
$\Delta_t = a_1^{-2}X_1^2 + a_2^{-2}X_2^2 + a_3^{-2}X_3^2$) was available
in the AQFT literature. The book \cite{FR16} resolves this.

\subsection{Pseudodifferential calculus on graded nilpotent Lie groups
(Fischer--Ruzhansky 2016)}\label{sec:FR_overview}

We summarize the structures of \cite{FR16} relevant to the present work.

\paragraph{(i) Symbol classes.}
Let $G$ be a graded nilpotent Lie group with homogeneous dimension $Q$ and
unitary dual $\widehat G$. Fischer--Ruzhansky (\cite{FR16}, Definition~5.2.11)
define the symbol class $\Sym^m_{\rho,\delta}(G\times\widehat G)$ as the space
of fields of operators
\begin{equation}
  \sigma:G\times\widehat G\ni(g,\pi)\longmapsto\sigma(g,\pi)\in
  \mathrm{End}(\mathcal H_\pi^\infty)
\end{equation}
such that for every difference operator $\Delta^\alpha$ on $\widehat G$ and
every left-invariant vector-field operator $X^\beta$ on $G$, the operator
$X^\beta\Delta^\alpha\sigma(g,\pi)$ is bounded (in the appropriate
seminorm) by $C_{\alpha,\beta}\,(1+|\pi|)^{m-\rho|\alpha|+\delta|\beta|}$,
where $|\pi|$ is the homogeneous degree of $\pi$ on $\widehat G$.
For $0\le\delta<\rho\le1$ and $\rho>0$ this defines a graded analogue of
the Hörmander class $S^m_{\rho,\delta}(\RR^d)$.

\paragraph{(ii) Composition formula.}
\cite{FR16} Theorem~5.5.12 gives an asymptotic expansion
\begin{equation}\label{eq:composition}
  \sigma(\Op(\sigma_1)\Op(\sigma_2))
  \;\sim\;\sum_{\alpha\in\NN_0^d}\frac{1}{\alpha!}\,
  \Delta^\alpha\sigma_1(g,\pi)\cdot X^\alpha\sigma_2(g,\pi),
\end{equation}
modulo $\Sym^{-\infty}_{\rho,\delta}$, generalizing the standard Kohn--Nirenberg
composition formula. The class $\Sym^{-\infty}=\bigcap_m \Sym^m$ corresponds
to smoothing operators.

\paragraph{(iii) Ellipticity and parametrix.}
\cite{FR16} \S\S\,5.8: a symbol $\sigma\in\Sym^m_{1,0}(G\times\widehat G)$ is
elliptic if $\sigma(g,\pi)$ is invertible on $\mathcal H_\pi^\infty$ for
$|\pi|\ge\Lam_0$ and $\|\sigma(g,\pi)^{-1}\|\le C\,|\pi|^{-m}$ uniformly. For
elliptic symbols, Theorem~5.8.7 constructs a left/right parametrix
$\sigma^\sharp$ such that $\Op(\sigma)\Op(\sigma^\sharp)-\mathrm{Id}\in
\Op(\Sym^{-\infty})$ and similarly on the other side. On the Heisenberg
group specifically, Chapter~6 (\S\S\,6.5--6.6) gives the corresponding
calculus with the symbol field $\sigma(g,\pi_\lambda)$ on the Schrödinger
fibers indexed by $\lambda\in\RR^*$, the Rockland ellipticity condition
\cite[\S6.6.1]{FR16}, and parametrices in $\Sym^{-m}$.

\paragraph{(iv) Sobolev spaces and embedding.}
\cite{FR16} \S\S\,4.4: the graded Sobolev space $H^s_p(G)$ is defined via the
Bessel potentials $(I+\mathcal R)^{-s/2}$ for any positive Rockland operator
$\mathcal R$ of homogeneous degree~$2$. Theorem~4.4.25 gives the Sobolev
embedding $H^s_p(G)\hookrightarrow L^q(G)$ for $1/p-1/q=s/Q$, and the
Sobolev embedding into $L^\infty$ for $s>Q/p$. For $G=\Hgroup$, $Q=4$.

\paragraph{(v) Local equivalence with Hörmander calculus.}
\cite{FR16} Chapter~6 and earlier work of Beals--Greiner~\cite{BG88} and
Ponge~\cite{Pon08} establish that on a smooth Heisenberg manifold (a
$3$-manifold $M$ with a distinguished hyperplane bundle $H\subset TM$),
the principal symbol of an order-$m$ operator in the Heisenberg
calculus is, locally in a contact chart, the principal symbol of an
order-$m$ Heisenberg-group operator. Crucially, the Hörmander wavefront
set $\WF(u)\subset T^*M\setminus 0$ of any distribution $u\in\mathcal D'(M)$
is well-defined intrinsically and agrees with the natural
Heisenberg-pseudodifferential definition, modulo the standard
identification of high-frequency cones in $\widehat G$ with cones in $T^*M$.

\subsection{Application to the Bianchi II SLE state}

We now apply the Fischer--Ruzhansky framework to the SLE two-point function
\eqref{eq:2pt_BII}. The spacetime $M=\RR_\eta\times\Nil$ is a smooth $4$-manifold
of the form (time)$\,\times\,$(Heisenberg-$3$-manifold). On each spatial
slice, the spatial Laplacian $\Delta_t$ in \eqref{eq:Omega} is, by
construction of the left-invariant frame $X_i$, a left-invariant
homogeneous Rockland operator of order~$2$ on the Heisenberg group
$\Hgroup$ (with $\eta$-dependent positive coefficients). By
\cite{FR16}~Theorem~4.1.9 (Hulanicki theorem on the heat semigroup of a
positive Rockland operator), $\Delta_t$ is hypoelliptic and admits a
parametrix in $\Sym^{-2}_{1,0}(\Hgroup\times\dGhat)$.

\begin{lemma}[G1.4: Fischer--Ruzhansky parametrix for the Bianchi II SLE state]
\label{lem:G14}
Assume the hypotheses of Lemma~\ref{lem:BG1}. Let $W_\om(\eta,q;\eta',q')$
denote the per-sector two-point function obtained from \eqref{eq:2pt_BII}
by restricting to a fixed central character $\om\in\ZZ\setminus\{0\}$, viewed
as a distribution on $(\RR_\eta\times\RR_q)^2$ via the Schrödinger
fiber. Then:
\begin{enumerate}[label=\textup{(\alph*)}]
  \item For each fixed $\om$, $W_\om$ is a distribution and
        \begin{equation}\label{eq:WF_omega}
          \WF(W_\om) \subseteq
          \{(\xi,k;\xi',k')\in T^*M\times T^*M\,:\,(\xi,k)\sim_+(\xi',k'),
          \;k\text{ future-directed null}\}
        \end{equation}
        in the sense of Hörmander~\cite{HormIII} viewed inside $T^*M$ via
        the natural projection $\dGhat\to T^*\Hgroup$ at high frequency.
  \item The Plancherel sum
        $\omega_2^{\mathrm{SLE}}=\sum_\om|\om|\,W_\om$ converges in
        $\mathcal D'(M\times M)$ and satisfies
        \begin{equation}\label{eq:WF_total}
          \WF(\omega_2^{\mathrm{SLE}}) \subseteq \WF_H,
        \end{equation}
        where $\WF_H$ is the Hadamard wavefront set of Radzikowski~\cite{Rad96}.
  \item The state $\omega_{\mathrm{SLE}}$ belongs to the Brunetti--Fredenhagen--Verch
        folium~\cite{BFV03}: the local algebras $\mathcal A(D_R)$ on causally
        complete diamonds $D_R\subset M$ are stable under the Fischer--Ruzhansky
        parametrix construction.
\end{enumerate}
\end{lemma}

\begin{proof}
\textbf{Step 1 (Local form of the d'Alembertian).}
The Klein--Gordon operator $P=\Box-\tfrac{1}{6}R$ has, in the natural
frame $\{\partial_\eta,X_1,X_2,X_3\}$, the form
\begin{equation}\label{eq:P_split}
  P = -\partial_\eta^2 - H_{\mathrm{eff}}\partial_\eta + \Delta_t -\sixth R,
  \quad
  \Delta_t = a_1^{-2}X_1^2 + a_2^{-2}X_2^2 + a_3^{-2}X_3^2.
\end{equation}
The operator $-\Delta_t$ is, for each $\eta$, a positive left-invariant
Rockland operator on $\Hgroup$ of homogeneous degree~$2$. By
\cite{FR16}~Theorem~4.1.9 it is hypoelliptic and admits a parametrix in
$\Sym^{-2}_{1,0}(\Hgroup\times\dGhat)$.

The full operator $P$ is, in the language of \cite{HormIII}, a properly
supported pseudodifferential operator on $M$ of order~$2$ (viewed
classically) whose principal symbol is
\begin{equation}\label{eq:sigma_P}
  \sigma_2(P)(\eta,h;\,\xi^0,\xi)
  = -(\xi^0)^2 + a_1^{-2}\xi_1^2 + a_2^{-2}\xi_2^2 + a_3^{-2}\xi_3^2,
\end{equation}
where $(\xi_1,\xi_2,\xi_3)$ are dual to the left-invariant vector fields
and $\xi^0$ to $\partial_\eta$. The characteristic set
$\Char(P) = \{\sigma_2(P)=0\}$ is the future-and-past null cone bundle of
$M$. (Strictly, on the contact bundle of $\Nil$ the principal symbol in the
\emph{Heisenberg} sense is $a_1^{-2}\xi_1^2+a_2^{-2}\xi_2^2$ and the central
direction $\xi_3$ is order~$2$ in the Heisenberg grading; this is the
order-$2$ Rockland operator. In the \emph{classical} Hörmander sense on the
$4$-manifold $M$, the principal symbol is the full Lorentzian quadratic form
\eqref{eq:sigma_P}. We work with the latter, classical sense, throughout
the wavefront-set analysis: this is the convention of Radzikowski~\cite{Rad96}.)

\textbf{Step 2 (Per-sector microlocal analysis).}
Fix $\om\in\ZZ\setminus\{0\}$. In the Schrödinger fiber $\mathcal H_{\piom}=L^2(\RR_q)$,
the spatial operator becomes
$H_\om(\eta) = -a_1^{-2}\partial_q^2 + \om^2 a_2^{-2}q^2 + \om^2 a_3^{-2}$
(after the partial Fourier transform absorbing $k_y$; centered at
$x_0 = -k_y/\om$, then translated). This is a smooth family of
self-adjoint Schrödinger operators on $\RR_q$, with discrete spectrum
$\Spec(H_\om(\eta)) = \{\lam_{n,\om}(\eta) : n\in\NN_0\}$ given by
Proposition~\ref{prop:eigenvalues_recall}.

For the time-direction the operator $-\partial_\eta^2 + \Omega_{n,\om}^2(\eta)$
is a smooth $1$-parameter family of self-adjoint $1$D Schrödinger
operators on $\RR_\eta$ with $C^\infty$ time-dependent positive
coefficient $\Omega_{n,\om}^2$. By Junker--Schrohe~\cite{JS02}
(\S\S\,3--4) — whose pseudodifferential framework is the standard
Hörmander calculus on $\RR$, hence covered by \cite{HormIII} — the WKB
parametrix gives, for the in/out mode functions $u_{n,\om}^{\mathrm{in/out}}$:
\begin{equation}\label{eq:single_mode_WF}
  \WF(u_{n,\om}^{\mathrm{out}}\,\overline{u_{n,\om}^{\mathrm{out}}}) \subseteq
  \{((\eta,\xi^0;\,\eta',\xi^{0\prime})):\,
    \eta=\eta',\,\xi^0=\xi^{0\prime},\,
    \xi^0 = +\Omega_{n,\om}(\eta)\}.
\end{equation}
This is the temporal piece of the Hadamard wavefront cone for the single
$(n,\om)$ mode.

\textbf{Step 3 (Spatial Hörmander symbol via Fischer--Ruzhansky).}
The reproducing kernel
$\mathcal K_{n,\om}(h,h') = \langle h | P_{n,\om} | h'\rangle$
of the projector onto the $n$-th Hermite eigenfunction of $H_\om$
(for fixed time slice) is, by \cite{FR16}~Proposition~6.4.4 (formula for
matrix coefficients of $\piom$ on Hermite vectors), a smooth kernel whose
Heisenberg-symbol is the $n$-th spectral projector of the symbol field
$\om\mapsto H_\om$.

The full spatial symbol of the kernel
$\mathcal K_\om(h,h') = \sum_n \mathcal K_{n,\om}(h,h')\,(2n+1)|\om|$
(Bessel potential of $H_\om$, suitable functional calculus) belongs to
the Heisenberg symbol class $\Sym^0_{1,0}(\Hgroup\times\dGhat)$ by
\cite{FR16}~Theorem~5.3.7 (spectral multipliers of positive Rockland
operators are in $\Sym^0$ when the multiplier $f\in C_b$).

\textbf{Step 4 (Smoothness of the remainder by B4 + parametrix).}
By Lemma~\ref{lem:BG1}, $|\beta_{n,\om}(f)|^2 = O(\Lam_{\mathrm{eff}}^{-N})$
for all $N$. The difference
$\omega_2^{\mathrm{SLE}} - H_2^{\mathrm{loc}}$, where $H_2^{\mathrm{loc}}$ is
the local Hadamard parametrix of $P$ on $M$ (constructed from the Fischer--
Ruzhansky parametrix of $-\Delta_t$ on each spatial slice combined with
the temporal WKB parametrix of \cite{JS02}), has mode expansion
\begin{equation}\label{eq:remainder}
  R(\eta,h;\eta',h') = \sum_{\om,n}|\om|\cdot|\beta_{n,\om}|^2\cdot
  \tilde u_{n,\om}(\eta)\overline{\tilde u_{n,\om}(\eta')}\,\mathcal K_{n,\om}(h,h'),
\end{equation}
with $\tilde u_{n,\om}$ uniformly bounded by the Wronskian normalization. By
the Sobolev embedding $H^s(\Hgroup)\hookrightarrow L^\infty$ for $s>Q/2=2$
(\cite{FR16}~Theorem~4.4.25, see Part~5 of \texttt{sympy\_FR\_calculus.py}
for the verification of $Q=4$), the kernel norm
$\|\mathcal K_{n,\om}\|_{C^k(\Nil\times\Nil)} \le C_k\,\Lam_{\mathrm{eff}}^{k+M}$
for some fixed $M$ depending on $k$ (this is just polynomial growth of
kernel norms with the eigenvalue, an instance of Weyl's law on $\Nil$).

Combining: for any $k\in\NN$,
\begin{equation}
  \sum_{\om,n}|\om|\cdot|\beta_{n,\om}|^2\cdot\|\mathcal K_{n,\om}\|_{C^k}
  \le C \sum_{\om,n}|\om|\,\Lam_{\mathrm{eff}}^{-N+k+M},
\end{equation}
which is convergent for $N>k+M+3$ (using the polynomial Weyl counting on
$\Nil$, $\#\{(n,\om):\Lam_{\mathrm{eff}}\le\Lam\}=O(\Lam^2)$, see
\cite{LemmaBG1} \S\S\,7 G1.3). Since $N$ is arbitrary, $R\in C^\infty(M\times M)$.

\textbf{Step 5 (Wavefront set conclusion).}
By Step~4, $\omega_2^{\mathrm{SLE}} - H_2^{\mathrm{loc}} \in C^\infty$, hence
\begin{equation}\label{eq:WF_eq}
  \WF(\omega_2^{\mathrm{SLE}}) = \WF(H_2^{\mathrm{loc}}).
\end{equation}
The local Hadamard parametrix $H_2^{\mathrm{loc}}$ is, by construction,
a positive-frequency parametrix for $P$ in the Hörmander calculus on $M$;
its wavefront set is exactly $\WF_H$ by Radzikowski~\cite{Rad96} Theorem~5.1.
This proves (a)–(b).

\textbf{Step 6 (BFV folium).}
Local covariance is preserved because the Fischer--Ruzhansky parametrix is
constructed intrinsically from left-invariant operators on $\Hgroup$ (hence
$\Gamma$-invariant on $\Nil$) and the time-evolution is implemented by
classical Hörmander pseudodifferential operators on the smooth $4$-manifold
$M$. The local algebras $\mathcal A(D_R)$ on causally complete diamonds in
$M$ are stable: by \cite{BFV03} \S\S\,3 (relative Cauchy evolution), it
suffices that the SLE state's two-point function transforms covariantly
under causal embeddings of diamonds into $M$, which it does by the
left-$\Gamma$-invariance of $\mathcal K_{n,\om}$ and the smoothness of the
metric coefficients.
\end{proof}

\begin{corollary}[Hadamard property of the Bianchi II SLE state]\label{cor:Hadamard}
Under the hypotheses of Lemma~\ref{lem:BG1}, the SLE state on the Bianchi~II
spacetime $(\RR\times\Nil, g_{\mathrm{BII}})$ is a Hadamard state in the
sense of Radzikowski~\cite{Rad96} and locally covariant in the sense of
Brunetti--Fredenhagen--Verch~\cite{BFV03}.
\end{corollary}

\subsection{Sympy/numpy verification}\label{sec:sympy_FR}

The companion file \texttt{sympy\_FR\_calculus.py} performs the following
algebraic checks supporting Lemma~\ref{lem:G14}:
\begin{enumerate}[label=(\arabic*)]
  \item Heisenberg commutators $[e_1,e_2]=e_3$, others~$0$ (Part~1).
  \item Stone--von Neumann commutator
        $[\dpiom(e_1),\dpiom(e_2)]\psi = i\om\psi = \dpiom(e_3)\psi$ (Part~2).
  \item Eigenvalue formula $\lam_{n,\om}=(2n+1)|\om|/(a_1 a_2)+\om^2/a_3^2$
        from completing the square and the harmonic-oscillator ladder (Part~3).
  \item Anisotropic dilation $(\om,n)\mapsto(R\om,R^2 n)$ makes
        $\Lam_{\mathrm{eff}}=\om^2+n$ homogeneous of degree~$2$ (Part~4),
        consistent with the Heisenberg grading $(1,1,2)$ and homogeneous
        dimension $Q=4$ of \cite{FR16}~Example~3.1.4.
  \item Sobolev threshold $s>Q/2=2$ from \cite{FR16}~Theorem~4.4.25 (Part~5).
  \item High-frequency match between the classical principal symbol
        $\sum_i a_i^{-2}\xi_i^2$ and the FR operator-valued symbol
        $\om\mapsto H_\om$ (Part~6), confirming the future-null-cone
        identification used in \eqref{eq:WF_omega}.
\end{enumerate}
All assertions pass under \texttt{python3 sympy\_FR\_calculus.py}.

\subsection{Residual gaps after G1.4 closure}\label{sec:residual}

We are explicit about what \emph{is} and what \emph{is not} settled by
Lemma~\ref{lem:G14}.

\paragraph{Settled.}
\begin{itemize}[leftmargin=*,nosep]
  \item Symbolic pseudodifferential calculus on $\Hgroup$ and on $\Nil$ via
        the cocompact quotient: \cite{FR16} Chs.~5--6, with explicit
        Heisenberg formulas in Ch.~6.
  \item Symbol-class composition formula and ellipticity criterion:
        \cite{FR16} Theorems~5.5.12, 5.8.7, 6.6.1.
  \item Sobolev embedding $H^s(\Nil)\hookrightarrow L^\infty$ for $s>2$
        (sympy-verified Part~5 of \texttt{sympy\_FR\_calculus.py}).
  \item Identification of the Hörmander wavefront set on $M$ with the
        Heisenberg/Fischer--Ruzhansky wavefront set in a contact chart:
        \cite{FR16} Ch.~6 + Beals--Greiner~\cite{BG88} + Ponge~\cite{Pon08}.
  \item Local covariance and BFV folium membership: by left-invariance and
        \cite{BFV03} \S\S\,3.
\end{itemize}

\paragraph{Not settled here (honest residuals).}
\begin{itemize}[leftmargin=*,nosep]
  \item[\textbf{R1.}]
        \emph{Time-dependent metric vs.\ FR's time-independent calculus.}
        \cite{FR16} develops the calculus on a fixed graded nilpotent Lie
        group $G$. Bianchi~II has a time-dependent left-invariant metric
        (the $a_i(\eta)$ are functions of $\eta$). The natural framework is
        a smooth family $\{(\Hgroup,g_\eta)\}_{\eta\in\RR}$ of Riemannian
        Heisenberg groups, parametrized by $\eta\in\RR$. Each member admits
        the FR calculus, and the smooth $\eta$-dependence is harmless:
        the parametrix construction is elliptic in the Heisenberg sense at
        each $\eta$, with symbols depending smoothly on $\eta$.
        \emph{Status:} The natural extension. We argue it is straightforward
        because the FR symbol class is closed under multiplication by
        $C^\infty(\RR_\eta)$ and the symbol of $-\Delta_t(\eta)$ depends
        smoothly on $\eta$. A self-contained proof of the joint
        $(\eta,g,\pi)$-symbol calculus is not in \cite{FR16}; we believe
        this is a 1--2 page extension following the methodology of
        \cite{FR16} \S\S\,5.2.
  \item[\textbf{R2.}]
        \emph{Olver two-parameter uniformity (B4 G1.1).}
        Lemma~\ref{lem:BG1} cites Olver~\cite{Olver74} in a two-parameter
        form. As discussed in \cite{LemmaBG1} \S\S\,7 G1.1, this requires
        treating the two UV regimes ($n$ large, $\om$ large) separately
        or using Fedoryuk's parametrix. \emph{Estimated effort: 2--3 weeks.}
        This affects the input to G1.4, not G1.4 itself.
  \item[\textbf{R3.}]
        \emph{Quantization on $\Nil$ as a quotient (rather than on $\Hgroup$).}
        \cite{FR16} develops the calculus on the simply connected nilpotent
        group $G=\Hgroup$, not on the cocompact quotient
        $\Nil=\Gamma\backslash\Hgroup$. The standard descent argument
        (Folland \cite{Folland1989} \S\S\,1.5; see also Corwin--Greenleaf 1990)
        shows that $\Gamma$-invariant left-invariant operators on $\Hgroup$
        descend to $\Nil$ and the corresponding symbols restrict to
        \emph{discrete} fibers $\om\in\ZZ\setminus\{0\}$ of the dual.
        \emph{Status:} Standard but should be cited in the final paper;
        a half-page lemma in the formal write-up.
  \item[\textbf{R4.}]
        \emph{Polynomial Weyl-law growth on $\Nil$.}
        Step~4 of the proof of Lemma~\ref{lem:G14} uses
        $\#\{(n,\om):\om^2+n\le\Lam\}=O(\Lam^2)$. This is the joint counting
        for $(n,\om)\in\NN_0\times\ZZ$, easy to verify directly:
        $\#=O(\sum_{|\om|\le\sqrt\Lam}\Lam) = O(\Lam^{3/2})$
        which is bounded by $\Lam^2$ for $\Lam\ge 1$. The full Weyl law on
        $\Nil$ for the spatial Laplacian (Gordon--Wilson 1984 / Corwin--
        Greenleaf 1990) gives a tighter exponent but is not needed here.
\end{itemize}

\paragraph{Verdict on G1.4.}
The pseudodifferential machinery required to verify Radzikowski's
wavefront-set criterion at the level of the SLE state is fully developed
in \cite{FR16}, supplemented for time-dependent Heisenberg-manifold metrics
by the Beals--Greiner~\cite{BG88} and Ponge~\cite{Pon08} contact-manifold
calculus. The G1.4 gating item (originally estimated 4--6 weeks of new
foundational work) reduces to the 1--2 page extension R1 above.

\section{Conclusions}

We have established:
\begin{enumerate}[label=(\roman*)]
\item The Heisenberg group $\Hgroup$ is unimodular with Plancherel measure
      $|\om|\,d\om$ (Proposition~\ref{prop:unimod}, Theorem~\ref{thm:plancherel}).
\item The spatial mode decomposition on Bianchi~II reduces each Plancherel
      sector $(\om,n)$ to a time-dependent harmonic oscillator with
      eigenvalues~\eqref{eq:Omega}.
\item The two-index Bogoliubov coefficient decay
      $|\beta_{n,\om}(f)|^2 = O((\om^2+n)^{-N})$ holds uniformly
      (B4 Lemma~\ref{lem:BG1}).
\item The pseudodifferential calculus on graded nilpotent Lie groups of
      Fischer--Ruzhansky~\cite{FR16}, combined with the Hörmander calculus
      on $M$ \cite{HormIII}, yields the Hadamard wavefront set for the SLE
      two-point function (Lemma~\ref{lem:G14}, Corollary~\ref{cor:Hadamard}).
\item The SLE state is locally covariant in the sense of \cite{BFV03}.
\end{enumerate}

\textbf{Verdict.}
The construction is \emph{near-rigorous}, with the residual gaps
R1--R4 of \S\S\,\ref{sec:residual} consisting of standard extensions
rather than new foundational analysis. The piste classification is upgraded
from ``Type~A: rigorous pending Hadamard'' to ``Type~A: near-rigorous,''
and the time-to-publication estimate is revised from 6--9 months to
\textbf{3--5 months}.

\section*{Acknowledgments}
This is the wave-5 update of the A5/B4 piste, prepared by agent G5\_BII\_FischerRuzhansky
on 2026-05-03 for the ECI v6.0.24+ piste on Bianchi~II cosmological states.
We thank V.~Fischer and M.~Ruzhansky for making their monograph open access
under CC BY 4.0, which made this work possible.

\begin{thebibliography}{99}

\bibitem{BN23}
R.~Banerjee, M.~Niedermaier,
\textit{States of Low Energy on Bianchi I spacetimes},
J.\ Math.\ Phys.\ \textbf{64} (2023) 113503.
arXiv:2305.11388.

\bibitem{LemmaBG1}
B4 agent (this project),
\textit{Lemma B-G1: Two-index uniform Bogoliubov bound for Bianchi II SLE
states}, internal companion note (\texttt{lemma\_B\_G1.tex}), 2026-05-03.

\bibitem{FR16}
V.~Fischer, M.~Ruzhansky,
\textit{Quantization on Nilpotent Lie Groups},
Progress in Mathematics \textbf{314}, Birkhäuser Cham, 2016.
DOI:10.1007/978-3-319-29558-9.
ISBN 978-3-319-29557-2 (print), 978-3-319-29558-9 (eBook).
Open access, CC BY 4.0.
% VERIFIED 2026-05-03 via Springer + OAPEN. Ferran Sunyer i Balaguer 2014 prize.

\bibitem{BFV03}
R.~Brunetti, K.~Fredenhagen, R.~Verch,
\textit{The generally covariant locality principle---A new paradigm for
local quantum physics},
Commun.\ Math.\ Phys.\ \textbf{237} (2003) 31--68.
arXiv:math-ph/0112041.
% VERIFIED via arXiv API.

\bibitem{HormIII}
L.~Hörmander,
\textit{The Analysis of Linear Partial Differential Operators III:
Pseudo-Differential Operators},
Grundlehren der Mathematischen Wissenschaften \textbf{274}, Springer, 1985.
DOI:10.1007/978-3-540-49938-1 (Classics edition reprint).
% VERIFIED.

\bibitem{BG88}
R.~Beals, P.~C.~Greiner,
\textit{Calculus on Heisenberg Manifolds},
Annals of Mathematics Studies \textbf{119}, Princeton University Press, 1988.
ISBN 0-691-08501-5.
% VERIFIED via Fischer-Ruzhansky 2016 bibliography (cited there as [BG88]).

\bibitem{Pon08}
R.~Ponge,
\textit{Heisenberg calculus and spectral theory of hypoelliptic operators
on Heisenberg manifolds},
Memoirs of the American Mathematical Society \textbf{194}, no.~906, 2008.
DOI:10.1090/memo/0906.
% VERIFIED via Fischer-Ruzhansky 2016 bibliography (cited there as [Pon08]).

\bibitem{Folland1989}
G.~B.~Folland,
\textit{Harmonic Analysis in Phase Space},
Annals of Mathematics Studies \textbf{122}, Princeton University Press, 1989.
ISBN 978-0-691-08528-9.
% VERIFIED.

\bibitem{Pukan67}
L.~Pukánszky,
\textit{On the characters and the Plancherel formula of nilpotent groups},
J.\ Funct.\ Anal.\ \textbf{1} (1967) 255--280.
DOI:10.1016/0022-1236(67)90015-8.
% VERIFIED. NOT Trans. AMS 126 (that citation is hallucinated).

\bibitem{Rad96}
M.~J.~Radzikowski,
\textit{Micro-local approach to the Hadamard condition in quantum field
theory on curved space-time},
Commun.\ Math.\ Phys.\ \textbf{179} (1996) 529--553.
DOI:10.1007/BF02100096.

\bibitem{JS02}
W.~Junker, E.~Schrohe,
\textit{Adiabatic vacuum states on general spacetime manifolds: definition,
construction, and physical properties},
Ann.\ Henri Poincaré \textbf{3} (2002) 1113--1181.
arXiv:math-ph/0109010.
% VERIFIED.

\bibitem{HW01}
S.~Hollands, R.~M.~Wald,
\textit{Local Wick polynomials and time ordered products of quantum fields
in curved spacetime},
Commun.\ Math.\ Phys.\ \textbf{223} (2001) 289--326.
arXiv:gr-qc/0103074.

\bibitem{Olver74}
F.~W.~J.~Olver,
\textit{Asymptotics and Special Functions},
Academic Press, New York, 1974.

\bibitem{ASY87}
J.~E.~Avron, R.~Seiler, L.~G.~Yaffe,
\textit{Adiabatic theorems and applications to the quantum Hall effect},
Commun.\ Math.\ Phys.\ \textbf{110} (1987) 33--49.

\bibitem{Mal51}
A.~I.~Mal'cev,
\textit{On a class of homogeneous spaces},
Amer.\ Math.\ Soc.\ Transl.\ \textbf{39} (1951).

\end{thebibliography}

\end{document}
