% -----------------------------------------------------------------------------
% Section 5 opener: toy dictionary for A3 in a cosmological setting.
% v4.5 (2026-04-21) — addresses the unanimous peer-pre-review flag on A3.
% -----------------------------------------------------------------------------

\subsection{A toy dictionary for A3 in cosmology}\label{sec:A3-dictionary}

The original Cryptographic Censorship theorem~\cite{CryptoCensorship} is
proven in AdS/CFT: a holographic CFT state whose time evolution restricted
to a boundary region is $\varepsilon$-approximately a unitary $k$-design
admits a bulk dual containing an event horizon. Its cosmological use in
this paper requires a transposition --- not yet a theorem --- from the
AdS boundary to a timelike observer worldline in a pseudo-Riemannian
spacetime. We make the transposition explicit as a \emph{working
conjecture} (not a proof sketch) so that readers can see what a positive
or negative answer would actually constrain. Under the observer-dependent
reading of \S\ref{sec:thesis}, this working-conjecture status is the
self-consistency condition delimiting the subregion on which the QRF
crossed product of A1 remains well-defined; the CFT-level pseudorandom-
unitary premise is itself a theorem (Ma--Huang~\cite{MaHuang2025}
constructed from quantum one-way functions), whereas only its cosmological
transposition is speculative.

\medskip\noindent\textbf{Conjectural map.}\;
Let $\mathcal{O}$ be a geodesic observer in an asymptotically de Sitter
FLRW spacetime, with causal diamond $\mathcal{D}(\mathcal{O})$ bounded by
the cosmological event horizon $\mathcal{H}_\mathcal{O}$. Let
$\mathcal{A}_{\mathcal{O}} = \left( \mathcal{M}_{\mathrm{QFT}} \rtimes
\mathbb{R} \right)^{\mathcal{O}}$ be the type-II crossed-product algebra
of A1 attached to $\mathcal{O}$'s clock QRF~\cite{CLPW2023,DEHK2025a,DEHK2025b},
and let $\sigma_t^{\mathcal{O}}$ be its modular flow. We conjecture:

\begin{quote}
\emph{(D) The set of states $\rho$ on $\mathcal{A}_{\mathcal{O}}$ whose
modular flow $\{\sigma_t^{\mathcal{O}}(\rho)\}_{t \in [0,T]}$ is an
$\varepsilon$-approximate unitary $k$-design on a code subspace of
observer-accessible operators corresponds, under any sensible
pseudo-Riemannian extension of the Cryptographic Censorship dictionary,
to cosmological states whose coarse-grained generalised entropy on
$\mathcal{H}_\mathcal{O}$ satisfies}
\begin{equation}
\left| S_{\mathrm{gen}}^{\mathrm{cg}}[\rho; \mathcal{H}_\mathcal{O}]
  \;-\; S_{\mathrm{gen}}^{\max}[\mathcal{H}_\mathcal{O}] \right|
  \;\lesssim\; \varepsilon \cdot k \cdot \log\!\dim \mathcal{A}_{\mathcal{O}},
\label{eq:toy-dict}
\end{equation}
\emph{with $S_{\mathrm{gen}}^{\max}$ the Gibbons--Hawking generalised
entropy of the static patch.}
\end{quote}

\noindent The form of the RHS bound is dimensionally dictated by the
AdS/CFT template of~\cite{CryptoCensorship}: pseudorandomness depth
$k$ times code-subspace dimension, in units of the algebra's trace
scale. In the type~II$_1$ static-patch setting of~\cite{CLPW2023} the
trace is finite and $\log\dim$ is replaced by the normalised trace
of the log of a density, but the structural statement is unchanged.

\medskip\noindent\textbf{Status.}\;
We emphasise that~\eqref{eq:toy-dict} is offered as a \emph{working
conjecture}, not a proof sketch. The partial evidence we are able to
cite is:
\begin{enumerate}
\item[(i)] The crossed-product construction
of~\cite{CLPW2023,DEHK2025a,DEHK2025b} gives, for the static de Sitter
patch, exactly the operator-algebraic structure (type II$_1$, well-defined
trace, $S_{\mathrm{gen}}$ as a finite functional) that the AdS/CFT
Cryptographic Censorship argument requires at the boundary.
\item[(ii)] The non-perturbative GSL of~\cite{FaulknerSperanza2024,Kirklin2025}
implies that the quantity on the LHS of~\eqref{eq:toy-dict} is monotone
under modular flow, matching the monotonicity expected on the CFT side
once the $k$-design condition is approached.
\item[(iii)] The existence of pseudorandom unitaries, constructed from
quantum one-way functions~\cite{MaHuang2025}, ensures the $k$-design
precondition is non-vacuous in physically realisable dynamics.
\end{enumerate}

None of (i)--(iii) constitutes a proof of~\eqref{eq:toy-dict}. A rigorous
derivation would require, at minimum, a pseudo-Riemannian analogue of
the one-sided modular reconstruction used in~\cite{CryptoCensorship}
and a replacement for the boundary causal-wedge reconstruction theorem
that does not rely on an AdS asymptotic region.

\medskip\noindent\textbf{One concrete testable consequence.}\;
If~\eqref{eq:toy-dict} holds, then cosmological perturbation modes whose
coarse-grained horizon entropy lies \emph{outside} the saturation
$\varepsilon$-band of $S_{\mathrm{gen}}^{\max}$ cannot be dual to any
bulk geometry containing the observer's cosmological horizon. In
particular, any primordial perturbation spectrum predicting a
super-Gibbons--Hawking coarse-grained horizon entropy at reheating is
excluded as a geometric bulk state. This yields a bulk-geometry
exclusion rule on CMB-scale perturbations with $|\delta S_{\mathrm{gen}}
/ S_{\mathrm{GH}}| > \varepsilon$; taking the observational proxy
$\varepsilon \sim 10^{-5}$ (the CMB temperature anisotropy amplitude)
gives a nontrivial consistency check against the complexity diagnostic of
A6 and the Weyl-curvature-hypothesis reformulation below. A violation
would falsify either the dictionary~\eqref{eq:toy-dict} or the
applicability of the crossed-product algebra of A1 at reheating --- a
meaningful and currently open test.

\medskip\noindent\textbf{How we use~\eqref{eq:toy-dict} below.}\;
The three narrative paragraphs that follow (Big Bang as decodability
boundary; inflation as algebraic necessity; Weyl curvature revisited)
are consequences of the dictionary~\eqref{eq:toy-dict} under the
additional assumption that the saturation $\varepsilon$-band is the
physically relevant regime. They inherit the conjectural status
of~\eqref{eq:toy-dict}; we flag this explicitly rather than restate
"conditional on A3" at every turn.
