% section_1_5_thesis_skeleton.tex
% Standalone scaffold for §1.5 of ECI v4.7, implementing the
% bridge-coordinator's recommendation: Hypothesis B (observer-dependent
% cosmology), scoped to the late-time quasi-de Sitter regime.
%
% NOT \input into eci.tex. Compilation not attempted. The next editing
% agent adapts this into eci.tex and re-runs the adversarial gate
% (PRINCIPLES §7).
%
% Sources:
%   - paper/team_B/draft_section_1_5_B.tex  (primary)
%   - paper/team_B/defence_memo.md          (scope restrictions)
%   - paper/team_B/observer_frame_map.md    (cross-reference table)
%   - paper/_internal_rag/GROUND_TRUTH.md Part A, D, E
%   - paper/_internal_rag/PRINCIPLES.md §§6, 8, 11, 12
%
% Three minor revisions relative to Team B's draft, per bridge §3:
%   (R1) Insert the word "heuristic" in prose for the
%        $|\dot H|/H^2$ adiabatic tolerance claim (PRINCIPLES §6).
%   (R2) Replace "organising principle" with "organising reading" to
%        stay below the framework-vs-derivation line (GROUND_TRUTH A).
%   (R3) Inline the matter-era scope exclusion in the first paragraph,
%        not only in part (iv).

\section{A unifying reading: observer-dependent cosmology in the late-time limit}
\label{sec:thesis}

% R3 inline scope exclusion up front.
The six axioms of \S1 can be given a common reading, valid in the
late-time quasi-de~Sitter regime ($\Omega_\Lambda \gtrsim 0.7$) where
every quantitative prediction of \S3 is formulated. The reading is
\emph{not} a derivation of the axioms and does not extend to the
pre-recombination sector (A4's EDE axion $\phi$ at
$z_c \sim 3500$; BBN-level constraints on $\xi_\chi$); those inputs
remain background-level inheritances from the external literature as
in v4.6.

% (i) CLPW/DEHK backbone.
\subsection*{(i) The crossed-product backbone}

The type-II$_1$ (static dS) and type-II$_\infty$ (SdS) algebras of
Chandrasekaran--Longo--Penington--Witten~\cite{CLPW2023} and the QRF
completion of De~Vuyst--Eccles--H\"ohn--Kirklin~\cite{DEHK2025a,DEHK2025b}
supply a structurally observer-dependent generalised entropy:
distinct QRFs yield distinct modular flows and numerically distinct
$S_{\rm gen}$. % citation already in eci.bib per convention audit.
% R2: organising *reading*, not principle.
The \emph{organising reading} proposed here is that the effective
couplings entering \S3 are likewise observer-frame labelled in the
late-time limit.

\subsection*{(ii) A4--A6 as observer-frame quantities}

\emph{A4 --- the NMC coupling $\xi_\chi$.} Cassini (\S3.5) bounds the
solar-system-frame value; DESI~DR2 and LSST~Y10 (\S3.1, \S3.5) bound
the late-time cosmological-frame value. Under the reading, the two
are not obliged to coincide, but in v4.6 they are identified because
the experiments' modular frames are both in the late-time quasi-dS
asymptote of a single static-patch observer.

\emph{A5 --- the species-scale exponent $c'$.} The tower visible to a
causal diamond is frame-labelled; $c' = 1/6$ is the asymptotic
quasi-dS value (Montero--Vafa--Valenzuela~\cite{Montero2022};
AAL~\cite{AAL2023}) recovered when the diamond saturates the static
patch.

\emph{A6 --- persistent homology $\mathrm{PH}_k$.} The Euler-
characteristic shift Eq.~\eqref{eq:A6-euler} is an ensemble average
over observer-frames; the Matsubara baseline is therefore the
frame-averaged diagnostic, consistent with PRINCIPLES~\S12.

\subsection*{(iii) A3 as a self-consistency condition}

Cryptographic Censorship~\cite{CryptoCensorship}, in its AdS/CFT
proof, states that boundary-restricted $\varepsilon$-approximate
$k$-design dynamics implies a bulk event horizon. Under the present
reading, A3 is not an independent axiom but the self-consistency
condition bounding the subregion on which the QRF crossed-product
of A1 remains well-defined.
% PRINCIPLES §8 + DECISIONS 2026-04-22: this is the Appendix-A-quarantined
% toy dictionary read in its minimal role, not a re-inflation.
The cosmological transposition of A3 remains a working conjecture
(Appendix~A) and the reading does not upgrade its status.

\subsection*{(iv) Caveats and scope}

% R1: heuristic label required in prose, not only in a comment.
Two scope restrictions are part of the claim, not loopholes around it.
\emph{(S1) Static-patch asymptote.} CLPW is proven on the static dS
patch; the late-time FLRW diamond approaches this asymptote with
$O(1-\Omega_\Lambda)$ corrections. The adiabatic tolerance of modular
flow to the DESI DR2 pivot $|\dot H|/H^2 \sim O(1)$ is a
\emph{heuristic} extension, at the same footing as the
Faulkner--Speranza~\cite{FaulknerSperanza2024} non-perturbative GSL.
\emph{(S2) Matter-era exclusion.} The EDE sector (\S3.2) and the BBN
$\xi_\chi$ bound (Wolf~2025) are background-level inheritances; the
observer-frame reading is not claimed for them. The paper's
separation between late-time (observer-dependent) and
pre-recombination (background-level) statements is preserved.

% Falsifier alignment, borrowed from Team A's single durable contribution
% under the weak reading (bridge §3).
A joint falsifier is available: any experimental anomaly that breaks
the consistency between the solar-system-frame $\xi_\chi$ (Cassini,
\S3.5) and the late-time cosmological-frame $\xi_\chi$
(DR3/LSST~Y10, Prediction~1b) beyond the static-patch
$O(1-\Omega_\Lambda)$ budget would damage the present reading without
touching the Lagrangian of \S2. This is not a free falsifier --- the
experimental precision on the consistency test is at least a decade
out --- but it is not zero.

% GROUND_TRUTH Part A: framework-genre declaration is the honesty anchor.
\subsection*{Status}

This section is an \emph{organising reading} of the six axioms in the
late-time quasi-dS regime, not a derivation, and does not enter the
predictive core \S3.1--\S3.7, \S4. The framework-genre declaration of
\S1 is unchanged.

% PRINCIPLES §11: when this skeleton is lifted into eci.tex for v4.7,
% a DECISIONS.md entry must land in the same commit recording:
%   (a) adoption of Hypothesis B (scoped);
%   (b) explicit rejection of Hypothesis A (UCG) per Team A's own
%       concession in its defence memo;
%   (c) the three minor revisions R1/R2/R3 listed in the header of
%       this file.
% The pre-v4.7 adversarial gate (PRINCIPLES §7) then reviews the §1.5
% text against GROUND_TRUTH.md Parts A, D, and E.

% End of skeleton.
