% section_1_5_thesis_B.tex — ECI v4.7
% Unifying reading of the six axioms: observer-dependent cosmology,
% scoped to the late-time quasi-de Sitter regime.
%
% Provenance: derived from paper/team_B/draft_section_1_5_B.tex, restructured
% to the skeleton of paper/section_1_5_thesis_skeleton.tex per the Bridge
% coordinator's verdict (paper/thesis_bridge.md, Pick B). The three minor
% revisions R1/R2/R3 listed in the skeleton header have been applied:
% (R1) "heuristic" said in prose for the |\dot H|/H^2 tolerance;
% (R2) "organising reading" replaces "organising principle";
% (R3) matter-era scope exclusion stated in the opening paragraph.
% No equation moved; no numerical value changed.

\section{A unifying reading: observer-dependent cosmology in the late-time limit}
\label{sec:thesis}

The six axioms of \S1 admit a common reading, valid in the late-time
quasi-de~Sitter regime ($\Omega_\Lambda \gtrsim 0.7$) where every
quantitative prediction of \S3 is formulated. The reading is \emph{not} a
derivation of the axioms and does not extend to the pre-recombination
sector (A4's EDE axion $\phi$ at $z_c\sim 3500$; BBN-level constraints on
$\xi_\chi$); those inputs remain background-level inheritances from the
external literature as in v4.6.

\subsection*{(i) The crossed-product backbone}

The type-III$_1$ algebra of quantum fields restricted to a gravitational
subregion carries no trace and no density matrices; the Chandrasekaran--
Longo--Penington--Witten (CLPW) construction~\cite{CLPW2023} and its
quantum-reference-frame (QRF) completion by De~Vuyst--Eccles--H\"ohn--
Kirklin (DEHK)~\cite{DEHK2025a,DEHK2025b} repair this by crossing with the
modular automorphism of an observer clock, yielding a type-II$_1$ algebra on
the static de~Sitter patch (II$_\infty$ on Schwarzschild--de~Sitter
Killing horizons) with a well-defined trace and an entropy reproducing
$S_{\rm gen} = A/(4G_N) + S_{\rm matter}$ up to a state-independent
constant. DEHK make the observer-dependence structural: distinct QRFs
yield distinct crossed-product algebras, distinct modular flows, and
distinct numerical values of $S_{\rm gen}$ on the same underlying bulk
state. The non-perturbative generalised second law of
Faulkner--Speranza~\cite{FaulknerSperanza2024} and
Kirklin~\cite{Kirklin2025} is formulated on exactly this type-II entropy.
The \emph{organising reading} proposed here is that the effective
couplings entering the late-time phenomenology of \S3 are observer-frame
labelled, not background quantities defined once and for all.

\subsection*{(ii) A4--A6 as observer-frame quantities}

\emph{A4 --- NMC coupling $\xi_\chi$.} The graviton kinetic matrix of the
Jordan-frame action of \S2 is frame-dependent via the type-II trace
normalisation; a geodesic observer whose crossed product is built from a
different clock Hamiltonian measures $\xi_\chi^{(\mathcal R)}$. The
Cassini bound $|\xi_\chi|(\chi_0/M_P)\le 2.4\times 10^{-3}$ of
\S\ref{sec:xi_constraints} constrains the \emph{solar-system-frame} value,
which is the only value the experiment is sensitive to.

\emph{A5 --- species-scale exponent $c'$.} The tower visible to a causal
diamond is frame-labelled; $c' = 1/6$ is the asymptotic quasi-dS value
(Montero--Vafa--Valenzuela~\cite{Montero2022}; AAL~\cite{AAL2023})
recovered when the diamond saturates the static patch.

\emph{A6 --- persistent homology $\mathrm{PH}_k$.} The Euler-characteristic
shift of~\eqref{eq:A6-euler} is computed on a density field restricted to
an observer-accessible region. The Matsubara baseline is the ensemble
average over observer frames; the diagnostic power of $\mathrm{PH}_k$
relative to the bispectrum is accordingly a frame-averaged statement,
consistent with PRINCIPLES~\S12 (no claim beyond derivation).

\subsection*{(iii) A3 as a self-consistency condition}

Cryptographic Censorship~\cite{CryptoCensorship}, in its AdS/CFT proof,
states that boundary-restricted $\varepsilon$-approximate $k$-design
dynamics implies a bulk event horizon. Under the present reading, A3 is
not an independent axiom but the self-consistency condition bounding the
subregion on which the QRF crossed product of A1 remains well-defined:
an observer whose modular algebra sees an effectively pseudorandom state
has lost predictive access to the complement region, and an event horizon
must cap the subregion on which A4--A6 are defined. The cosmological
transposition of A3 remains a working conjecture quarantined in
Appendix~A (PRINCIPLES~\S8); the reading does not upgrade its status.

\subsection*{(iv) Caveats and scope}

Two scope restrictions are part of the claim, not loopholes around it.

\emph{(S1) Static-patch asymptote.} CLPW is proven on the static dS
patch; DEHK extend to multi-observer QRFs on the same Killing-horizon
background. The late-time FLRW diamond approaches this asymptote with
$O(1-\Omega_\Lambda)$ corrections; the adiabatic tolerance of modular
flow to the DESI~DR2 pivot $|\dot H|/H^2 \sim O(1)$ is a \emph{heuristic}
extension, at the same footing as the Faulkner--Speranza
non-perturbative GSL.

\emph{(S2) Matter-era exclusion.} Pre-recombination A4/A5 statements
(the EDE axion $\phi$ at $z_c\sim 3500$, the BBN $\xi_\chi$ bound of
Wolf~2025~\cite{Wolf2025}) are background-level inheritances; the
observer-frame reading is not claimed for them. Their treatment under
the present reading would require a QRF extension to time-dependent FLRW
backgrounds which is an open problem in the algebraic-gravity
literature.

A joint falsifier is available: any experimental anomaly breaking the
consistency between the solar-system-frame $\xi_\chi$ (Cassini,
\S\ref{sec:xi_constraints}) and the late-time cosmological-frame
$\xi_\chi$ (DR3/LSST~Y10, Prediction~1b) beyond the static-patch
$O(1-\Omega_\Lambda)$ budget would damage the present reading without
touching the Lagrangian of \S2. The experimental precision is at least a
decade out, but the handle is not zero.

\subsection*{Status}

This section is an \emph{organising reading} of the six axioms in the
late-time quasi-dS regime, not a derivation, and does not enter the
predictive core \S3.1--\S3.7 or \S4. The framework-genre declaration of
\S1 is unchanged.

% End of section.
